### 2023-05-02 Papers/Paper ACM/paper.tex
Friedmann paper

\documentclass[manuscript]{acmart}

\renewcommand{\labelenumi}{\theenumi.}

\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[latin1]{inputenc}
\usepackage{hyperref}
\usepackage{dsfont}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{ifthen}
\usepackage{mmacells}
\usepackage[invisible]{MLaTex}

%\usepackage[displaymath,mathlines]{lineno}
%\linenumbers
%\pagewiselinenumbers
%\modulolinenumbers[1]

%permite que formulas
%se dividam por duas
%paginas diferentes
\allowdisplaybreaks

%[hypertex]

%\usepackage{showkeys}
%\usepackage[norefs]{refcheck}

\makeindex

%\usepackage{curves}

%\newcommand{\cal}{\mathcal}
\newcommand{\argmin}{\operatorname{argmin}}
\newcommand{\sech}{\operatorname{sech}}
\newcommand{\intr}{\operatorname{int}}
\newcommand{\arccosh}{\operatorname{arccosh}}
\newcommand{\arcsinh}{\operatorname{arcsinh}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\spn}{\operatorname{span}}
\newcommand{\arccotg}{\operatorname{arccotg}}
\newcommand{\inter}{\operatorname{int}}
\newcommand{\var}{\operatorname{Var}}
\newcommand{\supp}{\operatorname{supp}}
\newcommand{\car}{\operatorname{car}}
\newcommand{\cl}{\operatorname{cl}}
\newcommand{\co}{\operatorname{co}}
\newcommand{\arctanh}{\operatorname{arctanh}}
\newcommand{\esssup}{\operatorname{esssup}}
\newcommand{\essinf}{\operatorname{essinf}}
\renewcommand{\div}{\operatorname{div}}
\newcommand{\grad}{\operatorname{grad}}

\newcommand{\Rr}{{\mathbb{R}}}
\newcommand{\Cc}{{\mathbb{C}}}
\newcommand{\Nn}{{\mathbb{N}}}
\newcommand{\Zz}{{\mathbb{Z}}}
\newcommand{\Xx}{{\mathbb{X}}}
\newcommand{\Qq}{{\mathbb{Q}}}
\newcommand{\Tt}{{\mathbb{T}}}
\newcommand{\1}{{\chi}}

\newcommand{\Hh}{{\overline{H}}}
\newcommand{\ot}{{\overline{t}}}
\newcommand{\ox}{{\overline{x}}}
\newcommand{\oy}{{\overline{y}}}
\newcommand{\Ll}{{\mathcal{L}}}

\newcommand{\Aa}{{\mathcal{A}}}
\newcommand{\Cg}{{\mathcal{C}}}
\newcommand{\Ff}{{\mathcal{F}}}
\newcommand{\Gg}{{\mathcal{G}}}
\newcommand{\Ii}{{\mathcal{I}}}
\newcommand{\Dd}{{\mathcal{D}}}
\newcommand{\Uu}{{\mathcal{U}}}
\newcommand{\Vv}{{\mathcal{V}}}
\newcommand{\Ww}{{\mathcal{W}}}
\newcommand{\Mm}{{\mathcal{M}}}

\newcommand{\sfd}{{$\sigma$-field }}
\newcommand{\sfs}{{$\sigma$-fields }}
\newcommand{\cqd}{\hfill $\blacksquare$}
\newcommand{\cqda}{\hfill $\rule{1.5mm}{1.5mm}$}
\newcommand{\cqdb}{\hfill $\rule{0.75mm}{0.75mm}$}
\newcommand{\ee}{\hfill $\blacktriangleleft$}
\newcommand{\eef}{\quad\blacktriangleleft}
\newcommand{\bx}{{\bf x}}
\newcommand{\bu}{{\bf u}}
\newcommand{\bdx}{\dot{\bf x}}
\newcommand{\bp}{{\bf p}}
\newcommand{\bdp}{\dot{\bf p}}
\newcommand{\bX}{{\bf X}}
\newcommand{\bdX}{\dot{\bf X}}
\newcommand{\bP}{{\bf P}}
\newcommand{\bdP}{\dot{\bf P}}
\newcommand{\by}{{\bf y}}
\newcommand{\bdy}{\dot{\bf y}}
\newcommand{\bv}{{\bf v}}
\newcommand{\bdv}{\dot{\bf v}}
\newcommand{\bw}{{\bf w}}
\newcommand{\bdw}{\dot{\bf w}}

\newcommand{\Pf}{{\noindent \sc Proof. \ }}
\newcommand{\Rm}{{\noindent \sc Note. \ }}

\newcommand{\cws}{{\overset{*}{\rightharpoonup}}}

\newcommand{\epsi}{\varepsilon}
\def\d{{\rm d}}
\def\dx{{\rm d}x}
\def\dy{{\rm d}y}
\def\da{{\rm d}a}
\def\dt{{\rm d}t}
%\def\leq{\leqslant}
%\def\geq{\geqslant}

\def\appendixname{\empty}

\numberwithin{equation}{section}

\newtheoremstyle{thmlemcorr}{10pt}{10pt}{\itshape}{}{\bfseries}{.}{10pt}{{\thmname{#1}\thmnumber{
#2}\thmnote{ (#3)}}}
\newtheoremstyle{thmlemcorr*}{10pt}{10pt}{\itshape}{}{\bfseries}{.}\newline{{\thmname{#1}\thmnumber{
#2}\thmnote{ (#3)}}}
\newtheoremstyle{defi}{10pt}{10pt}{\itshape}{}{\bfseries}{.}{10pt}{{\thmname{#1}\thmnumber{
#2}\thmnote{ (#3)}}}
\newtheoremstyle{remexample}{10pt}{10pt}{}{}{\bfseries}{.}{10pt}{{\thmname{#1}\thmnumber{
#2}\thmnote{ (#3)}}}
\newtheoremstyle{ass}{10pt}{10pt}{}{}{\bfseries}{.}{10pt}{{\thmname{#1}\thmnumber{
A#2}\thmnote{ (#3)}}}

\theoremstyle{thmlemcorr}
\newtheorem{theorem}{Theorem}
\numberwithin{theorem}{section}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{thmlemcorr*}
\newtheorem{theorem*}{Theorem}
\newtheorem{lemma*}[theorem]{Lemma}
\newtheorem{corollary*}[theorem]{Corollary}
\newtheorem{proposition*}[theorem]{Proposition}
\newtheorem{problem*}[theorem]{Problem}
\newtheorem{conjecture*}[theorem]{Conjecture}

\theoremstyle{defi}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{hyp}{Assumption}
\newtheorem{claim}{Claim}
\newtheorem{problem}{Problem}

\theoremstyle{remexample}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{counterexample}[theorem]{Counterexample}

\newtheorem{teo}[theorem]{Theorem}
\newtheorem{lem}[theorem]{Lemma}
\newtheorem{pro}[theorem]{Proposition}
\newtheorem{cor}[theorem]{Corollary}

%\def\contentsname{}
\def\refname{References}
\def\bibname{References}
\def\chaptername{\empty}
\def\figurename{Fig.}
\def\abstractname{Abstract}

\theoremstyle{ass}

%\theoremstyle{definition}

\newtheorem{exemp}{Exemplo}

\begin{document}

\title[Detection of conserved quantities for PDE schemes]{Algorithmic detection of conserved quantities of finite-difference schemes for partial differential equations}

\author{Diogo A. Gomes}
\affiliation{%
	\institution{King Abdullah University of Science and Technology (KAUST)}
	\department[0]{AMCS/CEMSE Division}
	\city{Thuwal}
	\postcode{23955-6900}
	\country{Saudi Arabia}}
\email{diogo.gomes@kaust.edu.sa}
\author{Friedemann Krannich}
\affiliation{%
	\institution{King Abdullah University of Science and Technology (KAUST)}
	\department[0]{AMCS/CEMSE Division}
	\city{Thuwal}
	\postcode{23955-6900}
	\country{Saudi Arabia}}
\email{friedemann.krannich@kaust.edu.sa}
\author{Ricardo de Lima Ribeiro}
\affiliation{%
	\institution{King Abdullah University of Science and Technology (KAUST)}
	\department[0]{AMCS/CEMSE Division}
	\city{Thuwal}
	\postcode{23955-6900}
	\country{Saudi Arabia}}
\email{ricardo.ribeiro@kaust.edu.sa}

\keywords{Symbolic computations; Finite-difference schemes; Discrete variational derivative; Discrete partial variational derivative; Conserved quantities; Implicit schemes; Explicit schemes.}

\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10002950.10003714.10003727.10003729</concept_id>
	<concept_desc>Mathematics of computing~Partial differential equations</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10002950.10003714.10003715.10003750</concept_id>
	<concept_desc>Mathematics of computing~Discretization</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10002950.10003714.10003715.10003720.10003747</concept_id>
	<concept_desc>Mathematics of computing~Gr{\"o}bner bases and other special bases</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10010147.10010148.10010149.10010152</concept_id>
	<concept_desc>Computing methodologies~Symbolic calculus algorithms</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10010147.10010148.10010149.10010155</concept_id>
	<concept_desc>Computing methodologies~Discrete calculus algorithms</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Mathematics of computing~Partial differential equations}
\ccsdesc[500]{Mathematics of computing~Discretization}
\ccsdesc[500]{Mathematics of computing~Gr{\"o}bner bases and other special bases}
\ccsdesc[500]{Computing methodologies~Symbolic calculus algorithms}
\ccsdesc[500]{Computing methodologies~Discrete calculus algorithms}

\date{\today}

\begin{abstract}
Many partial differential equations (PDEs) admit conserved quantities like mass or energy. Those quantities are often essential to establish well-posed results. When approximating a PDE by a finite-difference scheme, it is natural to ask whether related discretized quantities remain conserved over time by the scheme. Such conservation may establish the stability of the numerical scheme. We present an algorithm for checking the preservation of a polynomial quantity under a polynomial finite-difference scheme. Our schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables. 
Additionally, we present an algorithm for, given a scheme, finding conserved quantities. 
We illustrate our algorithm with several finite-difference schemes.
\end{abstract}

\maketitle
\tableofcontents

\begin{acks}
	The authors were supported by \grantsponsor{GN01}{King Abdullah University of Science and Technology (KAUST) baseline funds}{} and 
	\grantsponsor{GN02}{KAUST OSR-CRG2021-4674}{}.
\end{acks}

\section{Introduction}
%\label{intro}
Many partial differential equations (PDEs) admit integral quantities, that are conserved in time. 
For example, the advection equation preserves energy and the heat equation conserves mass. 
When approximating PDEs by a finite-difference scheme, the question arises whether such quantities are conserved by the scheme. 
Such information can be crucial to estimate whether a scheme approximates a PDE accurately and to determine its stability.\\
Computations for determining conservation can get rather tedious.
Hence, automating them in computer algebra systems is desirable.\\
Conserved quantities correspond to conservation laws, admitted by the PDE. A conservation law \cite{cheviakov20} is an equation of the form
\begin{gather*}
	D_t\Phi[u]+D_x\Psi[u]=0
\end{gather*}
holding for all solutions $u$ of the considered PDE,
that induces the conserved quantity $\int \Phi[u]\ dx$ as
\begin{gather*}
	\frac{d}{dt}\int \Phi[u]\ dx=\int D_t\Phi[u]\ dx=-\int D_x\Psi[u]\ dx=0
\end{gather*}
assuming periodic boundary conditions in $x$.\\
An analog formulation describes conservation laws for schemes of PDEs, where the derivatives are replaced by shifts \cite{hydon01}.  The key for the construction of schemes, that admit certain conservation laws from the continuous PDE, is the use of the discrete Euler operator \cite{cheviakov20} (also called discrete variational derivative). Kupershmidt \cite[II. Theorem 31]{kupershmidt85} discovered, that an equation is a discrete conservation law if and only if it belongs to the kernel of the discrete Euler operator. Hence, standard approaches for constructing schemes with conservation are either discretizing the continuous conservation law or finding multipliers for the scheme \cite{dorodnitsyn21}, such that the result is in the kernel of the discrete Euler operator.
This strategy was, for example, recently used by Dorodnitsyn et al. to develop a scheme for the shallow water equation, that preserves energy \cite{dorodnitsyn21}. Cheviakov et al. used this idea to find schemes for the linear and nonlinear wave equation, that admit several discrete analogs of continous conservation laws \cite{cheviakov20}.
The same strategies for finding conservation laws, that are described above, work for continuous PDEs or semi-discretized PDEs:
Algorithms for finding conserved quantities, using multipliers and the Euler operator for PDEs were developed and implemented by Cheviakov \cite{cheviakov07,cheviakov10}.
Hereman et al. \cite{hereman04,hereman08}, proposed an algorithm to compute conserved densities for semi-discretized schemes for PDEs with first-order time derivative. 
Their algorithm uses the scaling symmetries of the scheme to construct conserved quantities and calculates their coefficients using the discrete Euler operator.
Gao et al. extended this algorithm to first-order time-explicit schemes \cite{gao02} and first-order time-implicit schemes \cite{gao04}.\\
Bihlo et al. \cite{bihlo12} used another approach and derived finite-difference schemes for the shallow water equations, that are symmetry-invariant on an adaptive mesh and then analyzed them regarding conserved quantities.\\
In this paper, we propose an algorithm that checks if a quantity is conserved in time under a scheme. Our approach differs from the ideas described above, as we do not try to construct discrete $\Phi$ and $\Psi$, using the scheme, but we check if a given discrete $\Phi$ is conserved in time under a given scheme.\\
Conserved quantities usually involve integral expressions or their discrete analog, sums.
Gomes et al. proposed algorithms for the simplification of sums \cite{gomes20}.
They also developed and implemented algorithms for detecting quantities conserved by PDEs and semi-discretized schemes. 
We revise these techniques in Section \ref{assp}. 
To generalize those methods to situations where we do not sum over all arguments of the functions involved, we introduce the discrete partial variational derivative (Section \ref{pvard}).
The main contribution of this paper is an algorithm for checking the conservation of a quantity under a numerical scheme.
Gerdt showed \cite{gerdt12}, that the quantity is conserved, if its discrete time derivative belongs to the difference ideal generated by the scheme. However, some quantities may add to a constant (e.g. telescopic sums) and thus be trivially preserved without belonging to the difference ideal. Moreover, Gerdt's algorithm may not terminate, as the Gr{\"o}bner basis for the difference ideal may not be finite.
We overcome these issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) with finite Gr{\"o}bner basis (Section \ref{algodetec}). 
Our algorithm works for schemes that are explicit and implicit in time and can treat schemes with several higher-order time  and space derivatives, and with several space dimensions.
Additionally, we can handle systems of equations and schemes with parameters. 
We have implemented this algorithm as part of a package in {\sc Mathematica} \cite{wolfram21} \footnote{The code is available upon request. For access contact ricardo.ribeiro@kaust.edu.sa.}.
In the examples, we show that our code finds conserved quantities and proper schemes for the time-implicit and time-explicit discretization of the Burgers equation and a system of PDEs arising in the study of mean-field games (Section \ref{eaa}).

\section{Preliminaries}
\label{assp}
Here, we follow the ideas in \cite{gomes20}.
In this paper, subscripts denote indices related to coordinates or tuples and superscripts denote indices related to sequences.
To avoid boundary terms, we work with periodic functions in $\Zz^d$.
\begin{definition}
	Let $N$, $d$ and $m$ be positive integers. The \emph{discrete torus} is $\Ii:=\{0,\ldots,N-1\}^d\subset \Zz^d$. Define the space
	\begin{gather*}
		\mathcal{P}(\Ii,\Rr^m):=\{u\ |\ u:\Ii\rightarrow\Rr^m\},
	\end{gather*}
	extended to $\Zz^d$ by periodicity.
\end{definition}
%Assuming periodicity allows us to extend the functions in $\mathcal{P}(\Ii)$ to $\Zz^d$, and we avoid discussing boundary terms.
\begin{definition}
	The space of functionals $\mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$ (not necessarily linear) on $\mathcal{P}(\Ii,\Rr^m)$ is
\begin{gather*}
	\mathcal{F}(\mathcal{P}(\Ii,\Rr^m)):= \Bigg\{\mathcal{J}:\mathcal{P}(\Ii,\Rr^m)\to \Rr\ \Bigg|\ \mathcal{J}[u]=\sum_{n\in\Ii}F_n[u],\\
	\text{ where }
	F_n[u]=G(u(n+e^1),u(n+e^2),\ldots,u(n+e^k))\\
	\text{ and }\{e^1,\ldots,e^k\}\subset\Zz^d,\ \text{G is smooth}
		\Bigg\}.
\end{gather*}
\end{definition}
In the previous definition, the $e^1,\ldots,e^k$ are not necessarily unit vectors and can vary between functionals.
\begin{example}
	\label{functional1}
	Let $d=m=1$ and
	$	\mathcal{J}[u]:=\sum_{n}(u(n+1)-u(n))^2$.
\end{example}
To simplify sums, the discrete variational derivative is a useful tool.
Let $u,v\in\mathcal{P}(\Ii,\Rr^m)$, $\epsilon 
	\in\Rr$ and $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$. Define
	\begin{gather*}
		D\mathcal{J}[u](v):=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}.
	\end{gather*}
	Below in \eqref{periodicity}, we shift the indices due to periodicity of $u$ and $v$,
	\begin{gather}
	\notag	D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\Ii}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
	\notag	=\sum_{n\in\Ii}\frac{d}{d\epsilon}G(u(n+e^1)+\epsilon v(n+e^1),\ldots,u(n+e^k)+\epsilon v(n+e^k))\bigg|_{\epsilon=0}\\
	\notag	=\sum_{n\in\Ii}D_1G(u(n+e^1),\ldots,u(n+e^k))v(n+e^1)\\
	\notag	+\ldots+D_kG(u(n+e^1),\ldots,u(n+e^k))v(n+e^k)\\
	\label{periodicity}	=\sum_{n\in\Ii}D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))v(n)\\
	\notag	+D_2G(u(n+e^1-e^2),u(n),\ldots,u(n+e^k-e^2))v(n)\\
	\notag	+\ldots+D_kG(u(n+e^1-e^k),u(n+e^2-e^k),\ldots,u(n))v(n)\\
	\notag	=\sum_{n\in\Ii}\Big(D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))\\
	\notag	+D_2G(u(n+e^1-e^2),u(n),\ldots,u(n+e^k-e^2))\\
	\notag	+\ldots+D_kG(u(n+e^1-e^k),u(n+e^2-e^k),\ldots,u(n))\Big)v(n)\\
	\notag	=:\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n).	
	\end{gather}
\begin{definition}
$\mathcal{V}(\mathcal{J})$ is the \emph{discrete variational derivative} of $\mathcal{J}$.
\end{definition}
Because we assume all related functions to be smooth, the discrete variational derivative always exists.
\begin{example}
	Let us return to Example \ref{functional1} and compute
	\begin{gather*}
		D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\Ii}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
		=\sum_{n\in\Ii}\frac{d}{d\epsilon}G[u(n+1)+\epsilon v(n+1),u(n)+\epsilon v(n)]\bigg|_{\epsilon=0}\\
		=\sum_{n\in\Ii}2(u(n+1)-u(n))v(n+1)+(-2)(u(n+1)-u(n))v(n)\\
		=\sum_{n\in\Ii}2(u(n)-u(n-1))v(n)+(-2)(u(n+1)-u(n))v(n)\\
		=\sum_{n\in\Ii}\Big(2(u(n)-u(n-1))+(-2)(u(n+1)-u(n))\Big)v(n)\\
		=\sum_{n\in\Ii}\Big(-2u(n+1)+4u(n)-2u(n-1)\Big)v(n)\\
		=\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n)
	\end{gather*}
	and hence $\mathcal{V}(\mathcal{J})[u](n)=-2u(n+1)+4u(n)-2u(n-1)$.
\end{example}
The algorithms for the simplification of sums presented in \cite{gomes20} rely on the following result:
\begin{teo}\label{theo1}
	Let $\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$. If $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[v]=0$ for all $v\in\mathcal{P}(\Ii,\Rr^m)$ and if there exists $u_0\in\mathcal{P}(\Ii,\Rr^m)$ such that $\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]$, then
		$\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$
	for all $u\in\mathcal{P}(\Ii,\Rr^m)$.
	Conversely,
		 $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u]=0$
	for all $u\in\mathcal{P}(\Ii,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
\end{teo}
\begin{proof}
	\begin{gather*}
		(\mathcal{J}-\tilde{\mathcal{J}})[u]=(\mathcal{J}-\tilde{\mathcal{J}})[u]-(\mathcal{J}-\tilde{\mathcal{J}})[u_0]\\
		=\int_{0}^{1}\frac{d}{d\lambda}(\mathcal{J}-\tilde{\mathcal{J}})[u_0+\lambda(u-u_0)]\ d\lambda\\
		=\int_{0}^{1}\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u_0+\lambda(u-u_0)](u-u_0)\ d\lambda=0
	\end{gather*}
	and hence $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
\end{proof}
We use Theorem \ref{theo1} to examine if different sums represent the same quantity.
\begin{example}
	Let $d=m=1$ and consider the functionals
	\begin{gather*}
		\mathcal{J}[u]:=\sum_{n\in\Ii}u(n)u(n+1)\text{ and }
		\tilde{\mathcal{J}}[u]:=\sum_{n\in\Ii}u(n-2)u(n-1).
	\end{gather*}
	It is clear, that $\mathcal{J}$ and $\tilde{\mathcal{J}}$ represent the same quantity (by shifting $n$). We confirm this, using the discrete variational derivative:
	\begin{gather*}
		D(\mathcal{J}-\tilde{\mathcal{J}})[u](v)=\\
		=\sum_{n\in\Ii}(u(n+1)+u(n-1)-u(n+1)-u(n-1))v(n)=0.
	\end{gather*}
	Hence, $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})=0$ and $\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]$. Therefore, both functionals sum to the same quantity.
\end{example}

\begin{example}
	Let $d=2$ and $m=1$. It may not be obvious, that
	\begin{gather*}
		\sum_{n_1,n_2}u(n_1,n_2-2)^4 - 3 u(n_1,n_2-2)^3 u(n_1+1,n_2-2) \\+ 
		u(n_1,n_2) u(n_1+1,n_2)^3 + 3 u(n_1,n_2-2)^2 u(n_1+1,n_2-2)^2\\ -
		u(n_1,n_2-2) u(n_1+1,n_2-2)^3
		- u(n_1, n_2)^4 +\\
		3 u(n_1,n_2)^3 u(n_1+1,n_2)- 3 u(n_1,n_2)^2 u(n_1+1,n_2)\\
		-3 u(n_1,n_2-2)^2 u(n_1+1,n_2-2)^2+ 3 u(n_1,n_2)^2 u(n_1+1,n_2)=0.
	\end{gather*}
	We confirm this, by computing the discrete variational derivative and noticing that the expression is $0$ for $u_0=0$.
\end{example}
\section{The discrete partial variational derivative}
\label{pvard}
Here, we generalize the discrete variational derivative for situations where we keep one (or several) of the arguments of $u$ constant.
Let $n=(n_1,\ldots,n_d)$ be the variables for the functions in $\mathcal{P}(\Ii,\Rr^m)$. We call $n$ the space variables. Let $l\in\Nn$ and call $t:=(t_1,\ldots,t_l)\in\Nn^l$ the time variables. Later in this paper, we only consider the case $l=1$.
\begin{definition}
	Let $\ \mathcal{P}(\Ii\times\Nn^l,\Rr^m):=\{u\ |\ u(\cdot,t)\in\mathcal{P}(\Ii,\Rr^m)\ \forall t\in\Nn^l \}$, extended to $\Zz^d\times\Nn^l$ by periodicity in the space variables and let $\mathcal{D}(\Nn^l):=\{u\ |\ u:\Nn^l\rightarrow\Rr\}$.
\end{definition}
When considering sums, we only sum over the space variables and not over the time variables.
\begin{definition}
	\label{quantityfunc2}
	The space of functions $\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$ is
	\begin{gather*}
		\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l)):=\\	
		\Bigg\{\mathcal{J}:\mathcal{P}(\Ii\times\Nn^l)\rightarrow\mathcal{D}(\Nn^l) \Bigg|\ \mathcal{J}[u](t)=\sum_{n\in\Ii}F_{(n,t)}[u],\\
		\text{where } F_{(n,t)}[u]=G(u((n,t)+e^1),u((n,t)+e^2),\ldots,u((n,t)+e^k)),\\
		\text{and }
		\{e^1,\ldots,e^k\}\subset\Zz^{d+l},\text{G is a polynomial in k variables}\Bigg\}.
	\end{gather*}
Here, $(n,t):=(n_1,\ldots,n_d,t_1,\ldots,t_l)\in\Ii\times\Nn^l$.
\end{definition}
Here, all quantities, whose conservation in time we check, are functions $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$. %We restrict $G$ to the set of polynomials in $k$ variables, because our algorithm below can only treat polynomial quantities.
Let $\tilde{n}:=(n_1,\dots,n_d,s_1,\dots,s_l)\in\Ii\times\Nn^l$. Using the Kronecker delta $\delta$, we rewrite
\begin{gather*}
	\mathcal{J}[u](t_1,\ldots,t_{l})
	=\sum_{\tilde{n}\in\Ii\times\Nn^l}F_{\tilde{n}}[u]\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l).
	%=\sum_{n\in\Ii}G(u(n+e^1),u(n+e^2),\ldots,u(n+e^k))\delta_{t_1,\ldots,t_{d-l+1}}(n_l,\ldots,n_d)
\end{gather*}
	Let $u,v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ and compute
	\footnotesize
	\begin{gather*}
		D^l\mathcal{J}[u](v)(t_1,\ldots,t_{l}):=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\Big|_{\epsilon=0}(t_1,\ldots,t_{l})\\
		=\sum_{\tilde{n}}D_1G(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^1)\\
		+D_2G(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^2)\\+\ldots
		+D_kG(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^k)\\
		=\sum_{\tilde{n}}D_1G(u(\tilde{n}),\ldots,u(\tilde{n}+e^k-e^1))\delta_{(t_1,\ldots,t_{l})}(s_1-e^1_{d+1},\ldots,s_l-e^1_{d+l})v(\tilde{n})\\
		+D_2G(u(\tilde{n}+e^1-e^2),\ldots,u(\tilde{n}+e^k-e^2))\delta_{(t_1,\ldots,t_{l})}(s_1-e^2_{d+1},\ldots,s_l-e^2_{d+l})v(\tilde{n})\\+\ldots
		+D_kG(u(\tilde{n}+e^1-e^k),\ldots,u(\tilde{n}))\delta_{(t_1,\ldots,t_{l})}(s_1-e^k_{d+1},\ldots,s_l-e^k_{d+l})v(\tilde{n})\\
		=\sum_{\tilde{n}}\Big(D_1G(u(\tilde{n}),\ldots,u(\tilde{n}+e^k-e^1))\delta_{(t_1+e^1_{d+1},\ldots,t_{l}+e^1_{d+l})}(s_1,\ldots,s_l)\\
		+D_2G(u(\tilde{n}+e^1-e^2),\ldots,u(\tilde{n}+e^k-e^2))\delta_{(t_1+e^2_{d+1},\ldots,t_{l}+e^2_{d+l})}(s_1,\ldots,s_l)\\+\ldots
		+D_kG(u(\tilde{n}+e^1-e^k),\ldots,u(\tilde{n}))\delta_{(t_1+e^k_{d+1},\ldots,t_{l}+e^k_{d+l})}(s_1,\ldots,s_l)\Big)v(\tilde{n})\\
		=:\sum_{\tilde{n}}\mathcal{V}^l(\mathcal{J})[u](\tilde{n},t)v(\tilde{n}).
	\end{gather*}
	\normalsize
	\begin{definition}
	We define $\mathcal{V}^l(\mathcal{J})$ as the \emph{discrete partial variational derivative} of $\mathcal{J}$.
	Note that $\mathcal{V}^0:=\mathcal{V}$.
\end{definition}
\begin{example}
	\label{dpvd1}
	The discrete partial variational derivative of 
	\begin{gather*}
		\mathcal{J}[u](t)=\sum_nu(n,t+1)-u(n,t)=\sum_{n,s}(u(n,s+1)-u(n,s))\delta_t(s)
	\end{gather*}
	equals $\delta_{t+1}(s)-\delta_t(s)$, because
	\small
	\begin{gather*}
		D^1(\mathcal{J})[u](v)(t)\\
		=\frac{d}{d\epsilon} \sum_{n,s}((u(n,s+1)+\epsilon v(n,s+1))-(u(n,s)+\epsilon v(n,s)))\delta_t(s)\\
		=\sum_{n,s}v(n,s+1)\delta_t(s)-v(n,s)\delta_t(s)=\sum_{n,s}v(n,s)\delta_t(s-1)-v(n,s)\delta_t(s)\\
		=\sum_{n,s}(\delta_{t+1}(s)-\delta_t(s))v(n,s)=\sum_{n,s}\mathcal{V}^1(\mathcal{J})(n,s,t)v(n,s).
	\end{gather*}
	\normalsize
\end{example}
\begin{example}
	In our {\sc Mathematica} implementation, Example \ref{dpvd1} is computed by
	\begin{EXE}
		LABEL="dpvd1"
		(*CODE*)
		variables = Association["indVars"->{n,t},"depVars"->{u},"timeVars"->{t}]
		expression = u[n,t+1] - u[n,t]
		PartialDVarDOperator[variables][expression]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}	
	\begin{small}
		\input{executables/dpvd1ctt.ed.tex}
		\input{executables/dpvd1ctt.result.ed.tex}
	\end{small}
\end{example}
We have a result similar to Theorem \ref{theo1} for the discrete partial variational derivative:
\begin{teo}\label{theo2}
	Let $\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$. If $\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[v]=0$ for all $v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ and if there exists $u_0\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ such that $\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]$, then
		$\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$
	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$.
	Conversely,
		$\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[u]=0$
	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
\end{teo}
The proof is similar to the proof of Theorem \ref{theo1}.
\section{Difference schemes and algebra}
\label{algodetec}
In this section, we define numerical schemes formally and introduce the tools from difference algebra, that we need for our algorithm.
During the remainder of this paper, we assume that there is only one time variable $t$ ($l=1$).
\begin{definition}
	The \emph{space of schemes} is
	\begin{gather*}
		\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr)):=\\
		\Bigg\{H:\mathcal{P}(\Ii\times\Nn,\Rr^m)\rightarrow\mathcal{P}(\Ii\times\Nn,\Rr)\ \Bigg|\\
			H[u](n,t)=H(u(n+\tilde{e}^1,t+\tilde{l}^1),\ldots,u(n+\tilde{e}^r,t+\tilde{l}^r)),\\
			\text{where }
		\{\tilde{e}^1,\ldots,\tilde{e}^r\}\subset\Zz^d,\ \{\tilde{l}^1,\ldots,\tilde{l}^r\}\subset\Zz,\\
		\text{H is a polynomial in r variables}
		\Bigg\}.
	\end{gather*}
	A \emph{scheme} is a set of functions
	\begin{gather*}
		\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))
	\end{gather*}
	that represent the equations $\{H^1[u]=0,\ldots,H^i[u]=0\}$ for $u\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$, holding pointwise for all points in $\Ii\times\Nn$.
\end{definition}
If a scheme contains a single function $H$, we call $H$ the scheme. Sometimes, we also call the expression $H[u]$ the scheme.
\begin{remark}
	In our examples, we only consider finite-difference schemes with fixed step sizes $\Delta x=\Delta t=1$. However, our algorithms and our code handle parameters. Hence, one can generalize our results to schemes with other step sizes.
\end{remark}
\begin{definition}
	A scheme $\{H^1,\ldots,H^m\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ given by
	\begin{equation*}
		H^j(u(n+\tilde{e}_j^1,t+\tilde{l}_j^1),\ldots,u(n+\tilde{e}_j^r,t+\tilde{l}_j^r))
	\end{equation*}
	is in \emph{time-explicit form} if we can rewrite the previous equation (eventually translating the scheme) as
	\begin{equation*}
		u_j(n,t+1)-\tilde{H}^j(u(n+\tilde{\tilde{e}}_j^1,t),u(n+\tilde{\tilde{e}}_j^2,t),\ldots,u(n+\tilde{\tilde{e}}_j^r,t))
	\end{equation*}
	for $1\leq j\leq m$
	with $\tilde{H}^j\in\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$. We call $\{\tilde{H}^1,\ldots,\tilde{H}^m\}$ the right-hand side of $\{H^1,\ldots,H^m\}$.
\end{definition}
\begin{example}\label{timeexplicitheateq}
	Consider the forward-difference scheme for the heat equation $u_t-u_{xx}=0$ \cite{smoller94}
	\begin{equation*}
		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t)).
	\end{equation*}
	This scheme is in time-explicit form with right-hand side
	\begin{equation*}
		u(n,t)+(u(n+1,t)-2u(n,t)+u(n-1,t)).
	\end{equation*}
\end{example}
{\bf Difference ideals.}\label{diffid}
Following Gerdt \cite{gerdt12}, we now introduce difference ideals and how we use them in our algorithm.
\begin{definition}
	Let $u=(u_1,\dots,u_m)\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$. The \emph{shift} in the $i$-th coordinate for $1\leq i\leq d+1$ by $k\in\Zz$ is
	\begin{gather*}
		\sigma_i^k\circ u_j(n,t):=u_j(n_1,\ldots,n_{i-1},n_i+k,n_{i+1},\ldots,t)
	\end{gather*}
	understanding that $\sigma_{d+1}^k$ shifts the time variable.
\end{definition}
\begin{definition}
	The \emph{set of all possible shifts} $\Theta$ is
	\begin{gather*}
		\Theta:=\{\sigma_1^{k_1}\circ\cdots\circ\sigma_{d+1}^{k_{d+1}}\ |\ k_1,\ldots,k_{d+1}\in\Zz\}.
	\end{gather*}
\end{definition}
Now we construct the difference ideal containing the scheme:
\begin{definition}
	\label{diffideal}
	Let $\ \mathcal{K}$ be the field generated by $\ \Rr$ and the variables $\{n_1,\ldots,n_d,t\}$. Let $\ \tilde{\mathcal{R}}$ be the polynomial ring over the field $\ \mathcal{K}$ and the variables $\theta\circ u_j(n,t)$ for $\theta\in\Theta$ and $1\leq j\leq m$.
	A set $\tilde{I}\subseteq\tilde{\mathcal{R}}$ is a \emph{difference ideal} if
	$p_1,p_2\in\tilde{I}$ implies $p_1+p_2\in\tilde{I}$;
	 $p_1\in\tilde{I},p_2\in\tilde{\mathcal{R}}$ implies $p_1\cdot p_2\in\tilde{I}$ and
	 $p_1\in\tilde{I},\theta\in\Theta$ implies $\theta\circ p_1\in\tilde{I}$.
\end{definition}
\begin{definition}
	$\langle H^1,\ldots,H^i\rangle$ is the smallest difference ideal containing the scheme $\{H^1,\ldots,H^i\}$.
\end{definition}
A solution of a numerical scheme is a function $u$, that makes all translations of the scheme vanish. Hence, every element of the difference ideal generated by the scheme vanishes under $u$.
Algorithm \ref{alg: discreteconservedqimplicit} seeks to determine if the polynomial associated with the discrete time derivative $\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$ belongs to the ideal $\langle H^1,\ldots,H^i\rangle$. To examine, if $\mathcal{T}\in\langle H^1,\ldots,H^i\rangle$, Gerdt proposes the notion of a standard (Gr{\"o}bner) basis for the ideal $\langle H^1,\ldots,H^i\rangle$. This standard basis may not be finite. We overcome this problem by considering polynomial instead of difference algebra and using a smaller polynomial ideal, that is contained in the difference ideal and admits a finite standard basis.
However, there may be functions $\mathcal{J}$, whose sums add to zero, even though they may fail to belong to the difference ideal. Hence, we combine the discrete partial variational derivative with polynomial ideals.
\section{The time-explicit case}
\label{time-exp}
In this section, we present Algorithm \ref{alg: discreteconservedqexplicit}  for checking the conservation of a quantity under the right-hand side of a scheme in time-explicit form. We discuss our algorithm for general schemes in the next section.
\begin{algorithm}
	\DontPrintSemicolon
	\KwIn{$\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{D}(\Nn))$ and $\{H^1,\ldots,H^m\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ right-hand side of a time-explicit scheme}
	\KwOut{ {\sc True}, if $\mathcal{J}$ is conserved in time, {\sc False} otherwise }

	\label{estep1} Build the discrete time derivative $\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$\;

	\label{estep2} Replace all instances of $u_j(n+e,t+1)$ (for $e\in\Zz^{d}$) by an appropriate translation of $\{H^1,\ldots,H^m\}$\;

	\label{estep3} Compute the discrete partial variational derivative\;

	\label{estep4} If the result is zero, conservation is {\sc True}, else {\sc False}\;

	\caption{{\sc time-explicit DiscreteConservedQ}}
	\label{alg: discreteconservedqexplicit}
\end{algorithm}
Note, that the result of Step 2 exists only at time $t$ because all instances of $t+1$ have been replaced. Thus, we do not need any computations of difference ideals or Gr{\"o}bner basis.
\begin{example}
	\label{timeexplicitheateqalgo}
	We check for conservation of $\sum_nu(n,t)$ under the scheme
	\begin{equation*}
		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t)).
	\end{equation*}
	The discrete time derivative (Step 1) is
	\begin{gather*}
		\sum_{n\in\Ii}u(n,t+1)-u(n,t).
	\end{gather*}
	Replace $u(n,t+1)$ by the right-hand side (Step 2) to get
	\begin{gather*}
		\sum_{n\in\Ii}u(n,t)+(u(n+1,t)-2u(n,t)+u(n-1,t))-u(n,t).
	\end{gather*}
	Then, we compute the discrete partial variational derivative (Step 3), which equals zero. Hence (Step 4), we have conservation.
\end{example}
\begin{example}
	\label{heatex1}
	We verify the result from Example \ref{timeexplicitheateqalgo}, using our implementation in {\sc Mathematica} of Algorithm \ref{alg: discreteconservedqexplicit}.
	\begin{EXE}
		LABEL="rhs1"
		(*CODE*)
		variables = Association["indVars"->{n},"depVars"->{u},"rhs"->{u[n]+(u[n+1]-2*u[n]+u[n-1])}]
		expression = u[n]
		DiscreteConservedQOperator[variables][expression]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/rhs1ctt.ed.tex}
		\input{executables/rhs1ctt.result.ed.tex}
	\end{small}
\end{example}
\section{The general case}
In this section, we present Algorithm \ref{alg: discreteconservedqimplicit} that deals with general, not necessarily time-explicit, schemes. We explain its steps in detail below and demonstrate the algorithm in Example \ref{heateqalgoimplicit}.\\
%In this section we explain our algorithm, that checks if a given quantity is preserved in time under a given scheme in detail.
\begin{algorithm}
	\DontPrintSemicolon
	\KwIn{$\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{D}(\Nn))$ and $\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ }
	\KwOut{ {\sc True}, if $\mathcal{J}$ is conserved in time, {\sc False} otherwise }

	\label{step1} Build the discrete time derivative $\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$\;

	\label{step2}Translate $\{H^1,\ldots,H^i\}$ by subtracting the range of the stencil of the scheme from the range of the stencil of $\mathcal{T}$\;

	\label{step3} Reduce $\mathcal{T}$ by using the Gr{\"o}bner basis generated by the translated scheme in two different ways
	\label{step3a} (a) Using the lexicographic order
	\label{step3b} (b) Using the explicit elimination order\;

	\label{step4} Choose the result that admits the least different instances of time\;

	\label{step5} Compute the discrete partial variational derivative\;

	\label{step6} Apply Steps 2 to 4 to the result\;

	\label{step7} If the result is zero, conservation is {\sc True}, else {\sc False}\;

	\caption{{\sc General DiscreteConservedQ}}
	\label{alg: discreteconservedqimplicit}
\end{algorithm}
{\bf Step 1: Build the discrete time derivative.}
We build the discrete time derivative by subtracting $\mathcal{J}[u](t)$ from $\mathcal{J}[u](t+1)$. This task is done in our code by the {\sc TimeDifference}.
	\begin{example}
		The discrete time derivative for $\sum_n u(n,t)$ is
		\begin{EXE}
			LABEL="heateq2"
			(*CODE*)
			variables = <|"indVars"->{n,t},"depVars"->{u}|>
			expression = u[n,t]
			TimeDifferenceOperator[variables][expression]
			(*CODE*)
			PUNCT={,.}
		\end{EXE}

		\begin{small}
			\input{executables/heateq2ctt.ed.tex}
			\input{executables/heateq2ctt.result.ed.tex}
		\end{small}
	\end{example}
%	Our code allows us to check not only for conservation in time but also if higher time derivatives are zero under the scheme by specifying a {\sc timederivativeorder} in the {\sc variables}.	
{\bf Step 2: Translate the scheme.}
%	Now we compute the translations of the scheme. As mentioned above, difference algebra does not guarantee termination of the algorithm.  Hence, we need to find a finite set of translations of the scheme, to guarantee termination of the algorithm. We translate the scheme such, that the range of its stencil stays exactly inside the range of the stencil of the discrete time derivative. This guarantees, that every instance of $u_j(n+e,t+l)$, occuring in $\mathcal{T}$, also occurs in the translated scheme.
We now make the step from difference algebra, with an infinite number of independent variables, to polynomial algebra, where we only have a finite number of variables to guarantee finiteness of the Gr{\"o}bner basis. Hence, we need to compute a finite number of translations of the scheme and then consider every instance $u_j(n+e,t+l)$, that appears either in the translated scheme or in the discrete time derivative as independent polynomial variable.
\begin{example}
	The algorithm treats the discrete time derivative $\sum_{n\in\Ii}u(n+1,t+1)-u(n+1,t)$
	as the polynomial $z_1-z_2$.
\end{example}
There are several possiblities which translations of the scheme to take as basis for the polynomial ideal. However, an obvious idea is to compute the minimal number of translations of the scheme, such that all instances of $u_j(n+e,t+l)$, that appear in the range of the stencil of the discrete time derivative, also appear in the translated scheme.
This is done by elementwise subtracting the range of the stencil of the scheme from the range of the stencil of the discrete time derivative. Before we define the above translations formally, we give a simple example to illustrate the idea of computing the translations:
\begin{example}
	Consider the expression
	\begin{gather*}
		u(n+3)-u(n-2)
	\end{gather*}
	which we want to reduce by the scheme $u(n+1)-u(n)=0$. We compute the translations as described above as
	\begin{gather*}
		\{u(n+3)-u(n+2),\dots,u(n-1)-u(n-2)\}.
	\end{gather*}
	and see, that the expression indeed reduces to $0$ under the scheme.
\end{example}
\begin{definition}
		Let $u=(u_1,\ldots,u_m)\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$ and
		\begin{gather*}
			F:
				\mathcal{P}(\Ii\times\Nn,\Rr^m)\rightarrow\mathcal{P}(\Ii\times\Nn,\Rr),\\
				F[u](n,t)=F(u_1((n,t)+x^1),\ldots,u_1((n,t)+x^{k_1}),\\
				u_2((n,t)+x^{k_1+1}),
				\ldots,u_2((n,t)+x^{k_2}),\\
				\ldots,
				u_m((n,t)+x^{k_{m-1}+1}),\ldots,u_m((n,t)+x^{k_m}))
		\end{gather*}
	\normalsize
	with $\{x^1,\ldots,x^{k_1},\ldots,x^{k_m}\}\subset\Zz^{d+1}$.
	The \emph{stencil} of $F$ is the $m$-tuple of sets of vectors
	\begin{gather*}
		\Big(\big\{x^1,\ldots,x^{k_1}\big\},
		\big\{x^{k_1+1},\ldots,x^{k_2}\big\},
		\ldots,
		\big\{x^{k_{m-1}+1},\ldots,x^{k_m}\big\}\Big).
	\end{gather*}
	The \emph{range} of the stencil of $F$ is
		\begin{gather*}
		\Big([x_1^{k_1 -},x_1^{k_1 +}]\times\cdots\times[x_{d+1}^{k_1 -},x_{d+1}^{k_1 +}],
		[x_1^{k_2 -},x_1^{k_2 +}]\times\cdots\times[x_{d+1}^{k_2 -},x_{d+1}^{k_2 +}],
		\\	\dots,
		[x_1^{k_m -},x_1^{k_m +}]\times\cdots\times[x_{d+1}^{k_m -},x_{d+1}^{k_m +}]
		\Big),
	\end{gather*}
	where $k_0=0$, $x^{k_l +}_i = \max_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$,
	and\\ $x^{k_l -}_i = \min_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$. Here, $[a,b]$ denotes the discrete interval in $\Zz$, i.e. $[0,2]=\{0,1,2\}$.
	\end{definition}
If for any $[a,b]$ involved $b<a$, we write $[a,b]=\emptyset$ with the convention that $\emptyset-\emptyset=\emptyset$.  Further, $\emptyset$ in the translations results in the use of the respective entry of the original scheme without translations.
	\begin{example}
		Consider the discrete time derivative
		\footnotesize
		\begin{equation*}
			u_1(n+2,t+1)+u_1(n,t+1)+u_2(n,t+1)-u_1(n+2,t)-u_1(n,t)-u_2(n,t)
		\end{equation*}
		\normalsize
		and the scheme
		\footnotesize
		$
			\{u_1(n+1,t+1)-u_1(n,t),u_2(n,t+1)-u_2(n,t)\}.
		$
		\normalsize
		Then, the stencil of the discrete time derivative equals
		\footnotesize
		\begin{equation*}
			\Big( \big\{ (2,1),(0,1),(2,0),(0,0) \big\},\big\{(0,1),(0,0)\big\} \Big)
		\end{equation*}
		\normalsize
		with range
		\footnotesize
		$
			\Big([0,2]\times[0,1],[0,0]\times [0,1]\Big).
		$
		\normalsize
		The stencil of the scheme equals the set of stencils of the equations in the scheme, i.e.
		\footnotesize
		\begin{equation*}
			\Big\{\Big(\big\{(1,1),(0,0)\big\},\emptyset\Big),\Big(\emptyset,\big\{(0,1),(0,0)\big\}\Big)\Big\}
		\end{equation*}
		\normalsize
		with range
		\footnotesize
		$
			\Big\{\Big([0,1]\times[0,1],\emptyset\Big),\Big(\emptyset,[0,0]\times [0,1]\Big)\Big\}.
		$
		\normalsize
		Hence, we get the translations for the first equation of the scheme by
		\footnotesize
		\begin{gather*}
			\Big([0,2]\times[0,1],[0,0]\times [0,1]\Big)-\Big([0,1]\times[0,1],\emptyset\Big)=\Big([0,1]\times[0,0],\emptyset\Big)
		\end{gather*}
		\normalsize
		and for the second equation of the scheme by
		\footnotesize
		\begin{gather*}
			\Big([0,2]\times[0,1],[0,0]\times [0,1]\Big)-\Big(\emptyset,[0,0]\times [0,1]\Big)=\Big(\emptyset,[0,0]\times [0,0]\Big).
		\end{gather*}
		\normalsize
		Therefore, we translate the first equation of the scheme by $(0,0)$ and $(1,0)$ and the second one by $(0,0)$ to get the translated scheme
		\footnotesize
		\begin{gather*}
			\{u_1(n+1,t+1)-u_1(n,t),
			u_1(n+2,t+1)-u_1(n+1,t),
			u_2(n,t+1)-u_2(n,t)
			\}.
		\end{gather*}
		\normalsize
	\end{example}
{\bf Step 3 and 4: Compute the Gr{\"o}bner basis and reduce the discrete time derivative.}\\
	\emph{Polynomial ideals and Gr{\"o}bner bases.}\label{polid}
	We found a finite set of polynomials (the translated scheme $\{H^1,\ldots,H^r\}$ and the discrete time derivative) with a finite number of instances of $u_j(n+e,t+l)$ in the previous step. Now we reduce $\mathcal{T}$ as an element of a polynomial ring by multivariate polynomial division with respect to the translated scheme. Here, we adapt the definitions and theorems from \cite{hibi13} to our setting.
	\begin{definition}
		Let $\mathcal{K}$ be as in Definition \ref{diffideal}. Let $\mathcal{R}$ be the polynomial ring generated by $\mathcal{K}$ and all instances of $u_j(n+e,t+l)$ that occur in $\{H^1,\ldots,H^r\}$.
		We call a set $ I\subseteq \mathcal{R}$ a \emph{(polynomial) ideal} if
		$p_1,p_2\in I$ implies $p_1+p_2\in I$ and
		 $p_1\in I,p_2\in\mathcal{R}$ implies $p_1p_2\in I$.
	\end{definition}
	\begin{definition}
		We denote by $\langle H^1,\ldots,H^r\rangle$ the smallest (polynomial) ideal containing $\{H^1,\ldots,H^r\}$.
	\end{definition}
	Given the ideal $\langle H^1,\ldots,H^r\rangle$ and the discrete time derivative $\mathcal{T}\in\mathcal{R}$, we want to determine if $\mathcal{T}\in\langle H^1,\ldots,H^r\rangle$ or if we can write $\mathcal{T}$ in a simpler form, using $\{H^1,\ldots,H^r\}$.
	Hence, using multivariate polynomial division, we search for $p^1,\ldots,p^r,p^{r+1}\in\mathcal{R}$ such that
	\begin{gather*}
		\mathcal{T}=p^1H^1+\ldots+p^rH^r+p^{r+1}.
	\end{gather*}
	Unfortunately, the resulting remainder $p^{r+1}$ may not be unique \cite[page 14, Example 1.2.3]{hibi13}, i.e. non-zero, although $\mathcal{T}$ belongs to $\langle H^1,\ldots,H^r \rangle$. 
	Replacing $\{H^1,\ldots,H^r\}$ by a Gr{\"o}bner basis for the ideal $\langle H^1,\ldots,H^r \rangle$ guarantees the uniqueness of the remainder of the polynomial division. 
	Contrary to the standard basis for the difference ideal that Gerdt's algorithm computes, a Gr{\"o}bner basis for the polynomial ideal always exists and is finite.
	A Gr{\"o}bner basis is defined up to the order of the monomials involved, so, depending on the order, we get different remainders of the polynomial division.\\
	\emph{Polynomial reduction.}
	In Algorithm \ref{alg: discreteconservedqimplicit}, we consider two monomial orders: the (a) lexicographic order and an experimental (b) elimination order.
	To reduce $\mathcal{T}$, we compute the Gr{\"o}bner basis of $\langle H^1,\ldots,H^r\rangle$ with respect to the monomial order and then calculate the remainder of the polynomial division of $\mathcal{T}$ with respect to this Gr{\"o}bner basis.
	The (a) lexicographic order is induced by the time-explicit ordering of the instances of $u_j(n+e,t+l)$, which gives more weight to instances at later times than to instances at earlier times.
	For the (b) elimination order, we first order all instances of $u_j(n+e,t+l)$ according to a pre-defined ordering. Our code uses by default the time-explicit ordering, but other orderings, like an implicit ordering or an user-defined ordering, are also possible. Then, we try to successively eliminate the instance with the highest ordering from the discrete time derivative, using the elimination order induced by a weight-matrix.
	This elimination might not always be possible, hence the reduction may leave the discrete time derivative unchanged. Then we repeat this process for the instance at the second latest time and continue until all instances of $u_j(n+e,t+l)$ have been eliminated.
	However, the elimination order is rather experimental, as its weight-matrix does not have a full rank and hence does not induce a total order on the set of monomials.\\
	Out of the two remainders from (a) and (b), we choose the result with the least number of different instances of time.\\
{\bf Step 5 to 7: Compute the discrete partial variational derivative and reduce again.}
	In the next step, we take the resulting expression and compute its discrete partial variational derivative. Then, we repeat Step 2 to Step 4 applied to the discrete partial variational derivative.
	\begin{example}
		\label{heateqalgoimplicit}
		To demonstrate Algorithm \ref{alg: discreteconservedqimplicit}, we use the setting from Example \ref{timeexplicitheateqalgo}.
		We compute the discrete time derivative (Step 1)
		\begin{gather*}
			\mathcal{T}=\sum_{n\in\Ii} u(n,t+1)-u(n,t).
		\end{gather*}
		For the translations (Step 2), the stencil of the scheme
		\begin{equation*}
			u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t))
		\end{equation*}
		\small
		is
		$
			\Big(\big\{(0,1),(0,0),(1,0),(-1,0)\big\}\Big)
		$
		\normalsize
	with range \small$\Big([-1,1]\times[0,1]\Big)$\normalsize.
	The discrete time derivative has stencil \small$\Big(\big\{(0,1),(0,0)\big\}\Big)$\normalsize 
	 and range \small$\Big([0,0]\times[0,1]\Big)$\normalsize.
	The translations are
	\small
	\begin{gather*}
	\Big([0,0]\times[0,1]\Big)-\Big([-1,1]\times[0,1]\Big)=\Big([1,-1]\times[0,0]\Big)=\Big(\emptyset\times[0,0]\Big)
	\end{gather*}
	\normalsize
	which results in no translations, as the range of the stencil of the scheme is greater than the range of the stencil of the discrete time derivative.
	Hence, the Gr{\"o}bner basis (Step 3) coincides with the original scheme for both monomial orders.
	The reduction using the lexicographic monomial order yields as remainder
	\begin{gather*}
		u(n-1,t)-2u(n,t)+u(n+1,t).
	\end{gather*}
	For the reduction using the elimination order, we first eliminate $u(n,t+1)$ and then the remaining instances of $u$.
	The elimination order yields the same remainder as the lexicographic order. Hence, we do not check for the number of instances of time (Step 4).
	We calculate the discrete partial variational derivative (Step 5) of
	\begin{gather*}
		\sum_{n\in\Ii}u(n-1,t)-2u(n,t)+u(n+1,t)
	\end{gather*}
	that equals $0$.
	Hence, repeating the above procedure (Step 6) becomes unnecessary.
	Therefore, (Step 7) we see that the scheme conserves the quantity.
	\end{example}
	\begin{example}
		We verify our result from Example \ref{heateqalgoimplicit}, using {\sc DiscreteConservedQ}, which detects automatically if we are in the time-explicit or the general setting.
		\begin{EXE}
		LABEL="heateq10"
		(*CODE*)
		variables = Association["indVars"->{n,t},"depVars"->{u},"scheme"->{u[n,t+1]-u[n,t]-(u[n+1,t]-2*u[n,t]+u[n-1,t])}]
		expression = u[n,t]
		DiscreteConservedQOperator[variables][expression]
		(*CODE*)
		PUNCT={,.}
		\end{EXE}
		\begin{small}
		\input{executables/heateq10ctt.ed.tex}
		\input{executables/heateq10ctt.result.ed.tex}
		\end{small}
	\end{example}
\begin{remark}\label{remark}
	The algorithm can only detect conservation, but can not check if something is not conserved for sure. So a resulting {\sc False} should be understood as our algorithm not detecting conservation, not as there being no conservation at all.
\end{remark}
%\begin{remark}
%	Our algorithm is also able to check for detection of time derivatives of higher order. To check for example the conservation of the first time derivative of the solution of the scheme, one can %specify {\sc timederivativeorder $\rightarrow$ 2} when calling {\sc DiscreteConservedQOperator}. In that case the algorithm will build the discrete forward second order time derivative of the %given quantity.
%\end{remark}
\section{A basis for conserved quantities}
\label{bfcq}
So far, we have discussed how to check if a quantity is preserved in time under a scheme. But it is also desirable to have a systematic way to find conserved quantities, given a scheme.
Algorithm \ref{alg: findconservedquantitybasis} finds, for a time-explicit scheme, a basis for conserved quantities that are generated by monomials up to a degree. 
\begin{algorithm}
	\DontPrintSemicolon
	\KwIn{right-hand side $\{H^1,\ldots,H^m\}\subset S(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ of time-explicit scheme, list of {\sc generators}, {\sc degree} of polynomials }
	\KwOut{basis of conserved quantities}

	\label{fstep1} Build all monomials combined from the {\sc generators} up to the total {\sc degree}\;

	\label{fstep2} Linearly combine all monomials with undetermined coefficients\;

	\label{fstep3} Compute the discrete time derivative and plug in the appropriate right-hand side(s) of the scheme\;

	Compute the discrete partial variational derivative\;

	\label{fstep4} Simplify and use the undetermined coefficients method to determine coefficients that make the result vanish\;

	\caption{{\sc FindDiscreteConservedQuantityBasis}}
	\label{alg: findconservedquantitybasis}
\end{algorithm}
We have implemented this algorithm in the {\sc FindDiscreteConservedQuantityBasis}.
\begin{example}
	We find a basis for conserved quantities that are generated by $u(n,t)$ and $n$ and that have at most degree 3, admitted by the scheme for the heat equation from Example \ref{heatex1}.
	\begin{EXE}
		LABEL="heateq11"
		(*CODE*)
		variables = <|"indVars"->{n},"depVars"->{u},"rhs"->{u[n]+u[n+1]-2*u[n]+u[n-1]}|>
		FindDiscreteConservedQuantityBasisOperator[variables][<|"degree"->3,"generators"->{u[n],n}|>]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/heateq11ctt.ed.tex}
		\input{executables/heateq11ctt.result.ed.tex}
	\end{small}
\end{example}
\section{Examples and applications}
\label{eaa}
We illustrate our code with the inviscid Burgers equation and a system of mean-field games. These only have first-order time-derivatives, but the algorithm handles PDEs with all orders of derivatives.\\
{\bf Burgers equation.}
The inviscid Burgers equation is the PDE $u_t+uu_x=0$ \cite{smoller94} in one space dimension. Any function of $u$ is a conserved quantity. We search for discretizations preserving $\int u\ dx$.
\begin{example}
	We check for conservation of mass using a forward-difference discretization:
	\begin{EXE}
		LABEL="burgers1"
		(*CODE*)
		variables = Association["indVars"->{n,t},"depVars"->{u},"scheme"->{u[n,t+1]-u[n,t]-u[n,t]*(u[n+1,t]-u[n,t])}];
		expression = u[n,t];
		DiscreteConservedQOperator[variables][expression]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/burgers1ctt.ed.tex}
		\input{executables/burgers1ctt.result.ed.tex}
	\end{small}
\end{example}
For finding a scheme that preserves mass, we check for conservation under a class of schemes with a parameter.
\begin{example}
	We discretize $u_x$ by a three-point stencil with a parameter $a$.
	\begin{EXE}
		LABEL="burgers2"
		(*CODE*)
		variables = Association["indVars"->{n,t}, "depVars"->{u}, "pars"->{a}, "scheme"->{u[n,t+1]-u[n,t]-u[n,t]*(a*(u[n+1,t]-u[n,t])+(1-a)*(u[n,t]-u[n-1,t]))}];
		expression = u[n,t];
		DiscreteConservedQOperator[variables][expression]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/burgers2ctt.ed.tex}
		\input{executables/burgers2ctt.result.ed.tex}
	\end{small}
	Conservation holds for $a=\frac{1}{2}$, the central difference case.
\end{example}
So far, we have only seen schemes that are in time-explicit form, but our algorithm allows also for general schemes:
\begin{example}
	We consider the scheme from the previous example but in the time-implicit version.
	\begin{EXE}
		LABEL="burgers3"
		(*CODE*)
		variables = Association["indVars"->{n,t}, "depVars"->{u}, "pars"->{a},"elimOrder"->"implicit", "scheme"->{u[n,t+1]-u[n,t]-u[n,t+1]*(a*(u[n+1,t+1]-u[n,t+1])+(1-a)*(u[n,t+1]-u[n-1,t+1]))}]
		DiscreteConservedQOperator[variables][u[n,t]]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/burgers3ctt.ed.tex}
		\input{executables/burgers3ctt.result.ed.tex}
	\end{small}
\end{example}
	Our results also hold for the viscous Burgers equation $u_t+uu_x-u_{xx}=0$, if we use the second-order central difference for $u_{xx}$.\\
{\bf Conserved quantities for a mean-field game.}
We study conserved quantities admitted by the discretization of a PDE system.
\begin{example}
	In \cite{gomes18}, Gomes et al. derived the following system
	\begin{gather*}
		v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0,\quad
		m_t-(vm)_x=0.
	\end{gather*}
	This system admits the conserved quantities $\int v\ dx$ and $\int m\ dx$.
	Because this system was derived from a forward-forward mean-field game, we discretize it forward in time. We check with our code if the scheme admits the same preserved quantities as the continuous system:
	\begin{EXE}
		LABEL="mfg1"
		(*CODE*)
		variables = Association[
		"indVars" -> {n, t},
		"depVars" -> {v, m},
		"display result" -> True,
		"scheme" -> {v[n, t + 1] - 
			v[n, 
			t] + (v[n + 1, t]^2/2 - m[n + 1, t]^2/2) - (v[n, t]^2/2 - 
			m[n, t]^2/2), 
			m[n, t + 1] - 
			m[n, t] - (m[n + 1, t]*v[n + 1, t]) + (m[n, t]*v[n, t])}
		];
		DiscreteConservedQOperator[variables][{v[n, t], m[n, t]}]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/mfg1ctt.ed.tex}
		\input{executables/mfg1ctt.result.ed.tex}
	\end{small}
	We reproduce this result, noting that we have a time-explicit scheme:
	\begin{EXE}
		LABEL="mfg2"
		(*CODE*)
		variables = Association[
		"indVars" -> {n},
		"depVars" -> {v, m},
		"rhs" -> { 
			v[n] -(v[n + 1]^2/2 - m[n + 1]^2/2) + (v[n]^2/2 - 
			m[n]^2/2), 
			m[n] + (m[n + 1]*v[n + 1]) - (m[n]*v[n])}
		];
		DiscreteConservedQOperator[variables][{v[n], m[n]}]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/mfg2ctt.ed.tex}
		\input{executables/mfg2ctt.result.ed.tex}
	\end{small}
	Those are the only conserved quantities for this scheme, that are polynomials up to degree 4 in $v$ and $m$.
	\begin{EXE}
		LABEL="mfg3"
		(*CODE*)
		variables = Association[
		"indVars" -> {n, t},
		"depVars" -> {v, m},
		"VarDOperator"->PartialDVarDOperator,
		"timeVars"->{t},
		"eqRhs" -> { 
			(v[n + 1, t]^2/2 - m[n + 1, t]^2/2) - (v[n, t]^2/2 - 
			m[n, t]^2/2), 
			(m[n + 1, t]*v[n + 1, t]) - (m[n, t]*v[n, t])}
		];
		FindConservedQuantityBasisOperator[variables][Association["degree"->4,"generators"->{v[n,t],m[n,t]}]]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/mfg3ctt.ed.tex}
		\input{executables/mfg3ctt.result.ed.tex}
	\end{small}
\end{example}
\begin{example}
	The related backward-forward system reads
	\begin{gather*}
		-v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0,\quad
		m_t-(vm)_x=0.
	\end{gather*}
	This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly in time for $m$ and implicitly for $v$. Our code can handle this setting, if we specify an order for the elimination of the variables, telling the code to use an explicit monomial order for $m$ and an implicit one for $v$:
	\begin{EXE}
		LABEL="mfg4"
		(*CODE*)
		variables = Association[
		"indVars" -> {n, t},
		"depVars" -> {m, v},
		"elimOrder" -> "explicitimplicit",
		"scheme" -> {-(v[n, t + 1] - 
			v[n, t]) + (v[n + 1, t+1]^2/2 - m[n + 1, t+1]^2/2) - (v[n, t+1]^2/2 - 
			m[n, t+1]^2/2), 
			m[n, t + 1] - 
			m[n, t] - (m[n + 1, t]*v[n + 1, t]) + (m[n, t]*v[n, t])}
		];
		DiscreteConservedQOperator[variables][{v[n, t], m[n, t]}]
		(*CODE*)
		PUNCT={,.}
	\end{EXE}
	\begin{small}
		\input{executables/mfg4ctt.ed.tex}
		\input{executables/mfg4ctt.result.ed.tex}
	\end{small}
\end{example}
\section{Possible extensions and concluding remarks}
\subsection{ Polynomial treatment of non-polynomial expressions}
It may be possible to extend our methods to non-polynomial PDEs and schemes. In this case, non-polynomial expressions can be handled by writing them in polynomial form.\\
{\bf Translate the scheme more accurately.}
In our elimination procedure, we work with a specific polynomial ideal that is a subset of the difference ideal generated by the scheme. However, it could be necessary to translate the scheme more than we did in our algorithm to get cancellation. Therefore, flexible methods to determine the translations of the scheme are desirable.\\
\subsection{Experimental check of conservation}
An experimental approach, that gives a hint, if a quantity is conserved, and that can refute conservation, is the following:
We fix $d$ and $N$. Then, we generate experimental initial data by randomly sampling $N\cdot d\cdot m$ values, with every value corresponding to an instance of $u_j(n,t)$ for $n\in\Ii,1\leq j\leq m$. Afterward, we calculate $u_j(n,t+1)$ for all $n\in\Ii,1\leq j\leq m$, using the scheme. With this values, we explicitly calculate the value of the discrete time derivative.
If the discrete time derivative is zero for several different samples of initial data, we might have conservation on the discrete level. Otherwise, if we find initial data such that the discrete time derivative is not zero, we can refute conservation.\\
\subsection{Discrete conserved quantities that are not conserved by the PDE}
It may arise the question if we can find conserved quantities, that are not preserved by the original PDE. Gerdt et al. \cite{gerdt10,gerdt12} define the notion of s-consistency, which guarantees that all discrete quantities are also preserved in the continuous setting. One possible extension of the code would be to check if s-consistency holds for the translated scheme.\\
\subsection{Concluding remarks}
We present an algorithm for checking the conservation of a quantity under a finite-difference scheme. Our algorithm allows for systems of equations, arbitrary time and space derivatives, and both explicit and implicit schemes. Also, we implemented a function for finding conserved quantities admitted by a scheme. We use our implementation of the algorithm to analyze the conservation properties of several schemes for PDEs that arise in applications.

%\citestyle{acmnumeric}
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}

### 2023-05-02 Papers/Paper ACM/tmp/ANTpaper.tex
Friedmann paper

\documentclass{amsart}
\usepackage{color}
\begin{document}
\section{Basic Text Statistics}
 \[\begin{array}{ll}
 \text{Language = } & \fbox{$\text{English}$} \\
 \text{Sentiment = } & \text{Neutral} \\
 \text{$\#$ Characters = } & 18321 \\
 \text{$\#$ Words = } & 3367 \\
 \text{$\#$ Sentences = } & 242 \\
 \text{$\#$ Words/Sentence = } & 13.9132 \\
 \text{$\#$ Distinct words = } & 636 \\
 \text{$\#$ Distinct words (without stop words) = } & 522 \\
 \text{Automated Readibility index = } & 11.1553 \\
 \text{Coleman-Liau index = } & 14.0676 \\
 \text{SMOG grade = } & 15.6095 \\
 \text{Glue index = } & 0.467181 \\
 \text{Word repetition index = } & 0.793288 \\
\end{array}\]

\section{Sentence Analysis}

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }

\noindent [Detection of conserved quantities for PDE schemes]Algorithmic detection of conserved quantities of finite-difference schemes for partial differential equations Diogo A.

\noindent  Gomes King Abdullah University of Science and Technology (KAUST) [0]CEMSE Division [1]AMCS Program Thuwal 23955-6900 Saudi Arabia diogo.

\noindent gomekaust.

\noindent edu.

\noindent sa Friedemann Krannich King Abdullah University of Science and Technology (KAUST) [0]CEMSE Division [1]AMCS Program Thuwal 23955-6900 Saudi Arabia friedemann.

\noindent krannickaust.

\noindent edu.

\noindent sa Ricardo de Lima Ribeiro King Abdullah University of Science and Technology (KAUST) [0]CEMSE Division [1]AMCS Program Thuwal 23955-6900 Saudi Arabia ricardo.

\noindent ribeirkaust.

\noindent edu.

\noindent sa Symbolic computations; Finite-difference schemes; Discrete variational derivative; Discrete partial variational derivative; Conserved quantities; Implicit schemes; Explicit schemes.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: like; } {\color{blue} ) }

\noindent ¡ccs2012¿ ¡concept¿ ¡concept˙id¿10002950.

\noindent 10003714.

\noindent 10003727.

\noindent 10003729¡/concept˙id¿ ¡concept˙desc¿Mathematics of computing Partial differential equations¡/concept˙desc¿ ¡concept˙significance¿500¡/concept˙significance¿ ¡/concept¿ ¡concept¿ ¡concept˙id¿10002950.

\noindent 10003714.

\noindent 10003715.

\noindent 10003750¡/concept˙id¿ ¡concept˙desc¿Mathematics of computing Discretization¡/concept˙desc¿ ¡concept˙significance¿500¡/concept˙significance¿ ¡/concept¿ ¡concept¿ ¡concept˙id¿10002950.

\noindent 10003714.

\noindent 10003715.

\noindent 10003720.

\noindent 10003747¡/concept˙id¿ ¡concept˙desc¿Mathematics of computing Grbner bases and other special bases¡/concept˙desc¿ ¡concept˙significance¿500¡/concept˙significance¿ ¡/concept¿ ¡concept¿ ¡concept˙id¿10010147.

\noindent 10010148.

\noindent 10010149.

\noindent 10010152¡/concept˙id¿ ¡concept˙desc¿Computing methodologies Symbolic calculus algorithms¡/concept˙desc¿ ¡concept˙significance¿500¡/concept˙significance¿ ¡/concept¿ ¡concept¿ ¡concept˙id¿10010147.

\noindent 10010148.

\noindent 10010149.

\noindent 10010155¡/concept˙id¿ ¡concept˙desc¿Computing methodologies Discrete calculus algorithms¡/concept˙desc¿ ¡concept˙significance¿500¡/concept˙significance¿ ¡/concept¿ ¡/ccs2012¿ [500]Mathematics of computing Partial differential equations [500]Mathematics of computing Discretization [500]Mathematics of computing Grbner bases and other special bases [500]Computing methodologies Symbolic calculus algorithms [500]Computing methodologies Discrete calculus algorithms Many partial differential equations (PDEs) admit conserved quantities like mass or energy.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: often; } {\color{blue} ) }

\noindent Those quantities are often essential to establish well-posed results.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent When approximating a PDE by a finite-difference scheme, it is natural to ask whether related discretized quantities remain conserved under the scheme.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: may; } {\color{blue} ) }

\noindent Such conservation may establish the stability of the numerical scheme.

\noindent }

\noindent {\color{green} ********************}

\noindent  We present an algorithm for checking the preservation of a polynomial quantity under a polynomial finite-difference scheme.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent In our algorithm, schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables.

\noindent }

\noindent {\color{green} ********************}

\noindent  Additionally, we present an algorithm for, given a scheme, finding conserved quantities.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }

\noindent We illustrate our algorithm by studying several finite-difference schemes.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }

\noindent The authors were supported by GN01King Abdullah University of Science and Technology (KAUST) baseline funds and GN02KAUST OSR-CRG2021-4674.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }  {\color{blue} (No verb) }

\noindent Introduction.

\noindent }

\noindent {\color{green} ********************}

\noindent  Many partial differential equations (PDEs) admit integral quantities, that are conserved in time.

\noindent   For example, the advection equation preserves energy and the heat equation conserves mass.

\noindent   When approximating PDEs by a finite-difference scheme, the question arises whether such quantities are conserved by the scheme.

\noindent   Such information can be crucial to estimate whether a scheme approximates a PDE accurately and to determine its stability.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: often, rather;  {\bf Wordiness}: more; } {\color{blue} ) }

\noindent While computations for determining conservation are often easy in simple cases, they get rather tedious for more complicated schemes or quantities.

\noindent }

\noindent {\color{green} ********************}

\noindent  Hence, automating such computations in computer algebra systems is desirable.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }

\noindent An algorithm for finding conserved quantities for (continuous) differential equations was developed and implemented in Maple by Cheviakov (see [1] and [1]).

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (No verb) }

\noindent Hereman et al.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }

\noindent (see for example [1] and [1]), proposed an algorithm to compute conserved densities for semi-discretized schemes for PDEs with first-order time derivative.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Wordiness}: which is; } {\color{blue} ) }

\noindent Their algorithm, which is implemented in a Mathematica package, uses the scaling symmetries of the scheme to construct conserved quantities and calculates their coefficients by using the discrete variational derivative.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }

\noindent Gao et al.

\noindent  extended this algorithm to first-order time-explicit schemes [1] and to first-order time-implicit schemes [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  They also implemented it in the computer algebra system Reduce.

\noindent   In this paper, we propose an algorithm that checks if a quantity is conserved in time under a scheme.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: usually; } {\color{blue} ) }

\noindent Conserved quantities usually involve integral expressions or their discrete analog, sums.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (No verb) }

\noindent Gomes et al.

\noindent }

\noindent {\color{green} ********************}

\noindent  proposed algorithms for the simplification of sums in [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }

\noindent They also developed and implemented algorithms for detecting quantities conserved by PDEs and semi-discretized schemes.

\noindent }

\noindent {\color{green} ********************}

\noindent  We revise these techniques in Section [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }

\noindent To generalize those methods to situations where we do not sum over all arguments of the functions involved, we introduce the discrete partial variational derivative (Section [1]).

\noindent }

\noindent {\color{green} ********************}

\noindent  The main contribution of this paper is an algorithm for checking the conservation of a quantity under a numerical scheme.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent Gerdt showed in [1], that if the discrete time derivative of the quantity belongs to the difference ideal generated by the scheme, the quantity is conserved.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: may; } {\color{blue} ) }

\noindent However, some quantities may add to a constant (e.

\noindent g.

\noindent  telescopic sums) and thus be trivially preserved without belonging to the difference ideal.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: may; } {\color{blue} ) }

\noindent Moreover, Gerdt’s algorithm may not terminate, as the Grbner basis for the difference ideal may not be finite.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Abstract names}: issues; } {\color{blue} ) }

\noindent We overcome this issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) with finite Grbner basis (Section [1]).

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }

\noindent Our algorithm works for schemes that are explicit and implicit in time and can treat schemes with several and higher-order time and space derivatives and with several space dimensions.

\noindent }

\noindent {\color{green} ********************}

\noindent  Additionally, we can handle systems of equations and schemes with parameters.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }

\noindent We have implemented this algorithm as part of a package in Mathematica [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }

\noindent In the examples, we show that our code finds conserved quantities and proper schemes for the time-implicit and time-explicit discretization of the Burgers equation and a system of PDEs arising in the study of mean-field games (Section [1]).

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }  {\color{blue} (No verb) }

\noindent Preliminaries.

\noindent }

\noindent {\color{green} ********************}

\noindent  Throughout this section, we follow the ideas and computations in [1].

\noindent   Here, subscripts denote indices related to coordinates or tuples while superscripts denote indices related to sequences and families.

\noindent   To avoid dealing with boundary terms, we work with periodic functions in [1].

\noindent   Let [1], [1] and [1] be natural numbers.

\noindent   The discrete torus is [1].

\noindent   Define the space [1], extended to [1] by periodicity.

\noindent   The space of functionals [1] (not necessarily linear) on [1] is [1].

\noindent   In the previous definition, the [1] are not necessarily unit vectors and can vary between functionals.

\noindent   Here, we allow for smooth [1].

\noindent   However, later we require [1] to be a polynomial function.

\noindent   Let [1], [1] and define [1].

\noindent   To simplify sums, the discrete variational derivative is a useful tool.

\noindent   Let [1], [1] and [1].

\noindent   Define [1].

\noindent   In [1], we shift the indices due to periodicity of [1] and [1], [1].

\noindent   [1] is the discrete variational derivative of [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: assume; } {\color{blue} ) }

\noindent Because we assume all related functions to be smooth, the discrete variational derivative always exists.

\noindent }

\noindent {\color{green} ********************}

\noindent  Let us return to Example [1] and compute [1] and hence [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }  {\color{blue} (No verb) }

\noindent Fundamental theorem.

\noindent }

\noindent {\color{green} ********************}

\noindent  The algorithms for the simplification of sums presented in [1] rely on the following result: Let [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }

\noindent If [1] for all [1] and if there exists [1] such that [1], then [1] for all [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }  {\color{blue} (No verb) }

\noindent Conversely, [1] for all [1], if [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (No verb) }

\noindent [1] and hence [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  We use Theorem [1] to examine if different sums represent the same quantity.

\noindent   Let [1] and consider the functionals [1].

\noindent   It is clear, that [1] and [1] represent the same quantity (by shifting [1]).

\noindent   We confirm this, using the discrete variational derivative: [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (No verb) }

\noindent Hence, [1] and [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  Therefore, both functionals sum to the same quantity.

\noindent   Let [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: may; } {\color{blue} ) }

\noindent It may not be obvious, that [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  We confirm this, by computing the discrete variational derivative and noticing that the expression is [1] for [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }  {\color{blue} (No verb) }

\noindent The discrete partial variational derivative.

\noindent }

\noindent {\color{green} ********************}

\noindent  Here, we generalize the discrete variational derivative for situations where we keep one (or several) of the arguments of [1] constant.

\noindent   Let [1] be the variables for the functions in [1].

\noindent   We call [1] the space variables.

\noindent   Let [1] and call [1] the time variables.

\noindent   Later in this paper, we only consider the case [1].

\noindent   Let [1], extended to [1] by periodicity in the space variables and let [1].

\noindent   When considering sums, we only sum over the space variables and not over the time variables.

\noindent   The space of functions [1] is [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }  {\color{blue} (No verb) }

\noindent Here, [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  Throughout this paper, all quantities, whose conservation in time we check, are functions [1].

\noindent   Define [1].

\noindent   Using the Kronecker delta [1], we rewrite [1].

\noindent   Let [1] and compute [1].

\noindent   We define [1] as the discrete partial variational derivative of [1].

\noindent   Note that [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }

\noindent The discrete partial variational derivative of [1] equals [1], because [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }

\noindent In our Mathematica implementation, Example [1] is computed by We have a result similar to Theorem [1] for the discrete partial variational derivative: Let [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }

\noindent If [1] for all [1] and if there exists [1] such that [1], then [1] for all [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }  {\color{blue} (No verb) }

\noindent Conversely, [1] for all [1], if [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  The proof is similar to the proof of Theorem [1].

\noindent   Difference schemes and algebra.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent In this section, we define numerical schemes formally and introduce the tools from difference algebra, that we need for our algorithm.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: assume; } {\color{blue} ) }

\noindent During the remainder of this paper, we assume that there is only one time variable [1] ([1]).

\noindent }

\noindent {\color{green} ********************}

\noindent  The space of schemes is [1].

\noindent   A scheme is a set of functions [1] that represent the equations [1] for [1], holding pointwise for all points in [1].

\noindent   If a scheme contains a single function [1], we call [1] the scheme.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: sometimes; } {\color{blue} ) }

\noindent Sometimes, we also call the expression [1] the scheme.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent In this work, we only consider finite-difference schemes with fixed step size [1], but one can generalize our results and algorithms to other step sizes.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent A scheme [1] given by [1] is in time-explicit form if we can rewrite the previous equation (eventually translating the scheme) as [1] for [1] and [1] with [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }

\noindent We call [1] the right-hand side of [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  Consider the forward-difference scheme for the heat equation [1] [1].

\noindent   This scheme is in time-explicit form with right-hand side [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (No verb) }

\noindent Difference ideals.

\noindent }

\noindent {\color{green} ********************}

\noindent  Following Gerdt [1], we now introduce difference ideals and how we use them in our algorithm.

\noindent   Let [1].

\noindent   The shift in the [1]-th coordinate for [1] by [1] is [1] understanding that [1] shifts the time variable.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: possible; } {\color{blue} ) }

\noindent The set of all possible shifts [1] is [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  Now we construct the difference ideal containing the scheme: Let [1] be the field generated by [1] and the variables [1].

\noindent   Let [1] be the polynomial ring over the field [1] and the variables [1] for [1] and [1].

\noindent   A set [1] is a difference ideal if [1] implies [1], [1] implies [1], [1] implies [1].

\noindent   [1] is the smallest difference ideal containing the scheme [1].

\noindent   A solution of a numerical scheme is a function [1], that makes all translations of the scheme vanish.

\noindent   Hence, every element of the difference ideal generated by the scheme vanishes under [1].

\noindent   In particular, Algorithm [1] seeks to determine if the discrete time derivative [1] belongs to the ideal [1].

\noindent   To examine, if [1], Gerdt proposes the notion of a standard (Grbner) basis for the ideal [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: may; } {\color{blue} ) }

\noindent This standard basis may not be finite.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }

\noindent We overcome this problem by considering polynomial instead of difference algebra and using a smaller polynomial ideal, that is contained in the difference ideal and admits a finite standard basis.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: may; } {\color{blue} ) }

\noindent However, there may be functions [1], whose sums add to zero, even though they may fail to belong to the difference ideal.

\noindent }

\noindent {\color{green} ********************}

\noindent  Hence, we combine the discrete partial variational derivative with polynomial ideals.

\noindent   The time-explicit case.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent In this section, we present Algorithm [1] for checking the conservation of a quantity under the right-hand side of a scheme in time-explicit form.

\noindent }

\noindent {\color{green} ********************}

\noindent  Our algorithm for general schemes can be found in the subsequent section.

\noindent   Note, that the result of step 2 exists only at time [1] because all instances of [1] have been replaced.

\noindent   Thus, we do not need any computations of difference ideals or Grbner basis.

\noindent   We check for conservation of [1] under the scheme from Example [1].

\noindent   The discrete time derivative (step 1) is [1].

\noindent   Replacing [1] by the right-hand side of the scheme (step 2) gives [1].

\noindent   Then, we compute the discrete partial variational derivative (step 3), which equals zero.

\noindent   Hence (step 4), we have conservation.

\noindent   We verify the result from Example [1], using our implementation in Mathematica of Algorithm [1].

\noindent   The general case.

\noindent   In this section, we present Algorithm [1] that deals with general, not necessarily time-explicit, schemes.

\noindent   We explain its steps in detail below and demonstrate the algorithm in Example [1].

\noindent   Step 1: Build the discrete time derivative.

\noindent   We build the discrete time derivative by subtracting [1] from [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }

\noindent This task is done in our code by the TimeDifference.

\noindent }

\noindent {\color{green} ********************}

\noindent  The discrete time derivative for [1] is Step 2: Translate the scheme.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }

\noindent To make the step from difference to polynomial algebra to guarantee termination of the algorithm, we use a standard idea in symbolic computations and consider every instance of [1] as a variable of a multivariate polynomial.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent Hence, we compute all translations of the scheme, such that all variables in the polynomial associated with [1] appear in the polynomials associated with the translated scheme.

\noindent }

\noindent {\color{green} ********************}

\noindent  The algorithm treats the discrete time derivative [1] as the polynomial equation [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }

\noindent Let [1] and [1] with [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  The stencil of [1] is the [1]-tuple of sets of vectors [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }

\noindent The range of the stencil of [1] is [1], where [1], [1], and [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  Here, [1] denotes the discrete interval in [1], i.

\noindent e.

\noindent  [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent We calculate the minimal necessary translations of the scheme, such that all instances of [1], that appear in the discrete time derivative, also appear in the translated scheme.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent This is done by elementwise subtracting the range of the stencil of the scheme from the range of the stencil of the discrete time derivative.

\noindent }

\noindent {\color{green} ********************}

\noindent  We denote the resulting translated system by [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }

\noindent If for any [1] involved [1], we write [1] with the convention that [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  Further, [1] in the translations results in the use of the respective entry of the original scheme without translations.

\noindent   Consider the discrete time derivative [1] and the scheme [1].

\noindent   Then, the stencil of the discrete time derivative equals [1] with range [1].

\noindent   The stencil of the scheme equals the set of stencils of the equations in the scheme, i.

\noindent e.

\noindent  [1] with range [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (High-glue sentence) }

\noindent Hence, we get the translations for the first equation of the scheme by [1] and for the second equation of the scheme by [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (High-glue sentence) }

\noindent Therefore, we translate the first equation of the scheme by [1] and [1] and the second one by [1] to get the translated scheme [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  Step 3 and 4: Compute the Grbner basis and reduce the discrete time derivative.

\noindent   Polynomial ideals and Grbner bases.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent We found a finite set of polynomials (the translated scheme and the discrete time derivative) with a finite number of instances of [1] in the previous step.

\noindent }

\noindent {\color{green} ********************}

\noindent  Now we reduce [1], using the translated scheme.

\noindent   Here, we adapt the definitions and theorems from [1] to our setting.

\noindent   Let [1] be as in Definition [1].

\noindent   Let [1] be the polynomial ring generated by [1] and all instances of [1] that occur in [1].

\noindent   We call a set [1] a (polynomial) ideal if [1] implies [1], [1] implies [1].

\noindent   We denote by [1] the smallest (polynomial) ideal containing [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent Given the ideal [1] and the discrete time derivative [1], we want to determine if [1] or if we can write [1] in a simpler form, using [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  Hence, using multivariate polynomial division, we search for [1] such that [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: may;  {\bf Wordiness}: unique; } {\color{blue} ) }

\noindent Unfortunately, the resulting remainder [1] may not be unique [page 14, Example 1.

\noindent 2.

\noindent 3]hibi13, i.

\noindent e.

\noindent  non-zero, although [1] belongs to [1].

\noindent }

\noindent {\color{green} ********************}

\noindent  Replacing [1] by a Grbner basis for the ideal [1] guarantees the uniqueness of the remainder of the polynomial division.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent Contrary to the standard basis for the difference ideal that Gerdt’s algorithm computes, the Grbner basis for the polynomial ideal always exists and is finite.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent A Grbner basis is defined up to the order of the monomials involved, so, depending on the order, we get different remainders of the polynomial division.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }  {\color{blue} (No verb) }

\noindent Polynomial reduction.

\noindent }

\noindent {\color{green} ********************}

\noindent  In Algorithm [1], we consider two monomial orders: (a) lexicographic and (b) explicit.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent To reduce [1], we compute the Grbner basis of [1] with respect to the monomial order and then calculate the remainder of the polynomial division of [1] with respect to this Grbner basis.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent The explicit monomial order (b) results in the Grbner basis that eliminates the instances of [1] at the latest time (with the largest [1]).

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: may, might, possible; } {\color{blue} ) }

\noindent This elimination might not always be possible, hence the reduction may leave the discrete time derivative unchanged.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent Then we repeat this process for the instance at the second latest time and continue until all instances of [1] have been eliminated.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent Out of the two remainders from (a) and (b), we choose the result with the least number of different instances of time.

\noindent }

\noindent {\color{green} ********************}

\noindent  Our code also supports other elimination orders, including user-defined ones.

\noindent   Step 5 to 7: Compute the discrete partial variational derivative and reduce again.

\noindent   In the next step, we take the resulting expression and compute its discrete partial variational derivative.

\noindent   Then, we repeat Step 2 to Step 4 applied to the discrete partial variational derivative.

\noindent   To demonstrate Algorithm [1], we use the setting from Example [1].

\noindent   We compute the discrete time derivative (step 1) [1].

\noindent   For the translations (step 2), the stencil of the scheme is [1] with range [1].

\noindent   The discrete time derivative has stencil [1] and range [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent The translations are [1] which results in no translations, as the range of the stencil of the scheme is greater than the range of the stencil of the discrete time derivative.

\noindent }

\noindent {\color{green} ********************}

\noindent  Hence, the Grbner basis (step 3) coincides with the original scheme for both monomial orders.

\noindent   The reduction using the lexicographic monomial order yields as remainder [1].

\noindent   For the reduction using the elimination order, we first eliminate [1] and then the remaining instances of [1].

\noindent   The elimination order yields the same remainder as the lexicographic order.

\noindent   Hence, we do not need to check for the number of instances of time (step 4).

\noindent   We calculate the discrete partial variational derivative (step 5) of [1] that equals [1].

\noindent   Hence, repeating the above procedure (step 6) becomes unnecessary.

\noindent   Therefore, (step 7) we see that the scheme conserves the quantity.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent We verify our result from Example [1], using DiscreteConservedQ, which detects automatically if we are in the time-explicit or the general setting.

\noindent }

\noindent {\color{green} ********************}

\noindent  The algorithm can only detect conservation, but can not check if something is not conserved for sure.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (High-glue sentence) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: should; } {\color{blue} ) }

\noindent So a resulting False should be understood as our algorithm not detecting conservation, not as there being no conservation at all.

\noindent }

\noindent {\color{green} ********************}

\noindent  A basis for conserved quantities.

\noindent   So far, we have discussed how to check if a quantity is preserved in time under a scheme.

\noindent   But it is also desirable to have a systematic way to find conserved quantities, given a scheme.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent Algorithm [1] finds, for a time-explicit scheme, a basis for conserved quantities that are generated by monomials up to a (total) degree.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }

\noindent We have implemented this algorithm in the FindDiscreteConservedQuantityBasis.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent We find a basis for conserved quantities that are generated by [1] and [1] and that have at most degree 3, admitted by the scheme for the heat equation from Example [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (No verb) }

\noindent Examples and applications.

\noindent }

\noindent {\color{green} ********************}

\noindent  We illustrate our code with the Burgers equation and a system of mean-field games.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (No verb) }

\noindent Burgers equation.

\noindent }

\noindent {\color{green} ********************}

\noindent  The Burgers equation is the PDE [1] [1] in one space dimension.

\noindent   Any function of [1] is a conserved quantity.

\noindent   We are interested in discretizations that preserve mass [1].

\noindent   We check for conservation of mass using a forward-difference discretization: We want to find a scheme that preserves this quantity.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: appropriate; } {\color{blue} ) }

\noindent Therefore we check for conservation under a class of schemes with a parameter, to find an appropriate choice for this parameter.

\noindent }

\noindent {\color{green} ********************}

\noindent  We discretize [1] using a three-point stencil with a parameter [1].

\noindent   We have conservation if we choose [1], corresponding to the standard central difference.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }

\noindent So far, we have only seen schemes that are in time-explicit form, but our algorithm allows also for general schemes: We consider the scheme from the previous example but in the time-implicit version.

\noindent }

\noindent {\color{green} ********************}

\noindent  Conserved quantities for a mean-field game.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent As a second application, we use our code to study conserved quantities admitted by the discretization of a system of PDEs.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (No verb) }

\noindent In [1], Gomes et al.

\noindent }

\noindent {\color{green} ********************}

\noindent  derived the following system of PDEs [1].

\noindent   This system admits the conserved quantities [1] and [1].

\noindent   Because this system was derived from a forward-forward mean-field game, we discretize it forward in time.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }

\noindent We check with our code if the scheme admits the same preserved quantities as the continuous system: We reproduce this result by noticing, that the scheme is in time-explicit form: Those are the only conserved quantities for this scheme, that are polynomials up to degree 4 in [1] and [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }

\noindent The related backward-forward system reads [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly (forward) in time for [1] and implicitly (backward) for [1].

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Low readability) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: possible; } {\color{blue} ) }

\noindent Our code can handle this setting, if we specify an order for the elimination of the variables, telling the code to use an explicit monomial order for [1] and an implicit one for [1]: Possible extensions and concluding remarks.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }  {\color{blue} (No verb) }

\noindent Polynomial treatment of non-polynomial expressions.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: may, may be possible, possible; } {\color{blue} ) }

\noindent It may be possible to extend our methods to non-polynomial PDEs and schemes.

\noindent }

\noindent {\color{green} ********************}

\noindent  In this case, non-polynomial expressions can be handled by writing them in polynomial form.

\noindent   For example, treat the expression [1] as the polynomial [1].

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Wordiness}: more; } {\color{blue} ) }

\noindent Translate the scheme more accurately.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent In our elimination procedure, we work with a specific polynomial ideal that is a subset of the difference ideal generated by the scheme.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (High-glue sentence) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Wordiness}: more; } {\color{blue} ) }

\noindent However, it could be necessary to translate the scheme more than we did in our algorithm to get cancellation.

\noindent }

\noindent {\color{green} ********************}

\noindent  Therefore, flexible methods to determine the translations of the scheme are desirable.

\noindent   Discrete conserved quantities that are not conserved by the PDE.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: may; } {\color{blue} ) }

\noindent It may arise the question if we can find conserved quantities, that are not preserved by the original PDE.

\noindent  Gerdt et al.

\noindent }

\noindent {\color{green} ********************}

\noindent  ([1],[1]) define the notion of s-consistency, which guarantees that all discrete quantities are also preserved in the continuous setting.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Style problems: }  {\color{green}  {\bf Hedging and Vague}: possible, would; } {\color{blue} ) }

\noindent One possible extension of our code would be to check automatically if s-consistency holds for the translated scheme.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }

\noindent Concluding remarks.

\noindent }

\noindent {\color{green} ********************}

\noindent  We present an algorithm for checking the conservation of a quantity under a finite-difference scheme.

\noindent   Our algorithm allows for systems of equations, arbitrary time and space derivatives, and both explicit and implicit schemes.

\noindent   Also, we implemented a function for finding conserved quantities admitted by a scheme.

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Long sentence) }

\noindent We use our implementation of the algorithm to analyze the conservation properties of several schemes for PDEs that arise in applications.

\noindent }

\noindent {\color{green} ********************}

\noindent

\noindent {\color{green} ********************}

\noindent {\color{red}  {\color{blue} (Low readability) }

\noindent ACM-Reference-Format references}

\noindent {\color{green} ********************}

\noindent \end{document}

### 2023-05-02 Papers/Paper ACM/tmp/CLpaper.tex
Friedmann paper

--- 2023-05-02 Papers/Paper ACM/paper.tex
@@
-\documentclass[manuscript]{acmart}
+\documentclass[sigconf,twocolumn]{acmart}
 
@@
 \usepackage{amsmath}
-%\usepackage{amssymb}
 \usepackage{amsthm}
 \usepackage[latin1]{inputenc}
+\usepackage{eurosym}
+\usepackage{epsfig}
 \usepackage{hyperref}
 \usepackage{dsfont}
+
+\usepackage[displaymath,mathlines]{lineno}
+
+\allowdisplaybreaks
+
+\usepackage{ifthen}
+
 \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
-\usepackage{ifthen}
+
 \usepackage{mmacells}
@@
 
-%\usepackage[displaymath,mathlines]{lineno}
-%\linenumbers
-%\pagewiselinenumbers
-%\modulolinenumbers[1]
-
-%permite que formulas
-%se dividam por duas
-%paginas diferentes
-\allowdisplaybreaks
-
-%[hypertex]
-
-%\usepackage{showkeys}
-%\usepackage[norefs]{refcheck}
-
 \makeindex
 
-%\usepackage{curves}
-
-%\newcommand{\cal}{\mathcal}
 \newcommand{\argmin}{\operatorname{argmin}}
@@
 \def\dt{{\rm d}t}
-%\def\leq{\leqslant}
-%\def\geq{\geqslant}
 
@@
 
-%\def\contentsname{}
 \def\refname{References}
@@
 
-%\theoremstyle{definition}
-
 \newtheorem{exemp}{Exemplo}
@@
 \author{Diogo A. Gomes}
-\affiliation{%
+\affiliation{
 	\institution{King Abdullah University of Science and Technology (KAUST)}
-	\department[0]{AMCS/CEMSE Division}
+	\department[0]{CEMSE Division}
+	\department[1]{AMCS Program}
 	\city{Thuwal}
@@
 \author{Friedemann Krannich}
-\affiliation{%
+\affiliation{
 	\institution{King Abdullah University of Science and Technology (KAUST)}
-	\department[0]{AMCS/CEMSE Division}
+	\department[0]{CEMSE Division}
+	\department[1]{AMCS Program}
 	\city{Thuwal}
@@
 \author{Ricardo de Lima Ribeiro}
-\affiliation{%
+\affiliation{
 	\institution{King Abdullah University of Science and Technology (KAUST)}
-	\department[0]{AMCS/CEMSE Division}
+	\department[0]{CEMSE Division}
+	\department[1]{AMCS Program}
 	\city{Thuwal}
@@
 Those quantities are often essential to establish well-posed results.
-When approximating a PDE by a finite-difference scheme, it is natural to ask whether related discretized quantities remain conserved over time by the scheme.
+When approximating a PDE by a finite-difference scheme, it is natural to ask whether related discretized quantities remain conserved under the scheme.
 Such conservation may establish the stability of the numerical scheme.
 We present an algorithm for checking the preservation of a polynomial quantity under a polynomial finite-difference scheme.
-Our schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables.
+In our algorithm, schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables.
 Additionally, we present an algorithm for, given a scheme, finding conserved quantities. 
-We illustrate our algorithm with several finite-difference schemes.
+We illustrate our algorithm by studying several finite-difference schemes.
 \end{abstract}
@@
 \section{Introduction}
-%\label{intro}
 Many partial differential equations (PDEs) admit integral quantities, that are conserved in time. 
@@
 Such information can be crucial to estimate whether a scheme approximates a PDE accurately and to determine its stability.\\
-Computations for determining conservation can get rather tedious.
-Hence, automating them in computer algebra systems is desirable.\\
-Conserved quantities correspond to conservation laws, admitted by the PDE. A conservation law \cite{cheviakov20} is an equation of the form
-\begin{gather*}
-	D_t\Phi[u]+D_x\Psi[u]=0
-\end{gather*}
-holding for all solutions $u$ of the considered PDE,
-that induces the conserved quantity $\int \Phi[u]\ dx$ as
-\begin{gather*}
-	\frac{d}{dt}\int \Phi[u]\ dx=\int D_t\Phi[u]\ dx=-\int D_x\Psi[u]\ dx=0
-\end{gather*}
-assuming periodic boundary conditions in $x$.\\
-An analog formulation describes conservation laws for schemes of PDEs, where the derivatives are replaced by shifts \cite{hydon01}.
-The key for the construction of schemes, that admit certain conservation laws from the continuous PDE, is the use of the discrete Euler operator \cite{cheviakov20} (also called discrete variational derivative).
-Kupershmidt \cite[II.
-Theorem 31]{kupershmidt85} discovered, that an equation is a discrete conservation law if and only if it belongs to the kernel of the discrete Euler operator.
-Hence, standard approaches for constructing schemes with conservation are either discretizing the continuous conservation law or finding multipliers for the scheme \cite{dorodnitsyn21}, such that the result is in the kernel of the discrete Euler operator.
-This strategy was, for example, recently used by Dorodnitsyn et al.
-to develop a scheme for the shallow water equation, that preserves energy \cite{dorodnitsyn21}.
-Cheviakov et al.
-used this idea to find schemes for the linear and nonlinear wave equation, that admit several discrete analogs of continous conservation laws \cite{cheviakov20}.
-The same strategies for finding conservation laws, that are described above, work for continuous PDEs or semi-discretized PDEs:
-Algorithms for finding conserved quantities, using multipliers and the Euler operator for PDEs were developed and implemented by Cheviakov \cite{cheviakov07,cheviakov10}.
-Hereman et al. \cite{hereman04,hereman08}, proposed an algorithm to compute conserved densities for semi-discretized schemes for PDEs with first-order time derivative. 
-Their algorithm uses the scaling symmetries of the scheme to construct conserved quantities and calculates their coefficients using the discrete Euler operator.
-Gao et al. extended this algorithm to first-order time-explicit schemes \cite{gao02} and first-order time-implicit schemes \cite{gao04}.\\
-Bihlo et al.
-\cite{bihlo12} used another approach and derived finite-difference schemes for the shallow water equations, that are symmetry-invariant on an adaptive mesh and then analyzed them regarding conserved quantities.\\
+While computations for determining conservation are often easy in simple cases, they get rather tedious for more complicated schemes or quantities. 
+Hence, automating such computations 
+in computer algebra systems is desirable.\\
+An algorithm for finding conserved quantities for (continuous) differential equations was developed and implemented  in {\sc Maple} by Cheviakov (see \cite{cheviakov07} and \cite{cheviakov10}).
+Hereman et al. (see for example \cite{hereman04} and \cite{hereman08}), proposed an algorithm to compute conserved densities for semi-discretized schemes for PDEs with first-order time derivative. 
+Their algorithm, which is implemented in a {\sc Mathematica} package, uses the scaling symmetries of the scheme to construct conserved quantities and calculates their coefficients by using the discrete variational derivative.
+Gao et al. extended this algorithm to first-order time-explicit schemes \cite{gao02} and to first-order time-implicit schemes \cite{gao04}. 
+They also implemented it in the computer algebra system {\sc Reduce}.\\
 In this paper, we propose an algorithm that checks if a quantity is conserved in time under a scheme.
-Our approach differs from the ideas described above, as we do not try to construct discrete $\Phi$ and $\Psi$, using the scheme, but we check if a given discrete $\Phi$ is conserved in time under a given scheme.\\
 Conserved quantities usually involve integral expressions or their discrete analog, sums.
-Gomes et al. proposed algorithms for the simplification of sums \cite{gomes20}.
+Gomes et al. proposed algorithms for the simplification of sums in \cite{gomes20}.
 They also developed and implemented algorithms for detecting quantities conserved by PDEs and semi-discretized schemes. 
@@
 The main contribution of this paper is an algorithm for checking the conservation of a quantity under a numerical scheme.
-Gerdt showed \cite{gerdt12}, that the quantity is conserved, if its discrete time derivative belongs to the difference ideal generated by the scheme.
+Gerdt showed in \cite{gerdt12}, that if the discrete time derivative of the quantity belongs to the difference ideal generated by the scheme, the quantity is conserved.
 However, some quantities may add to a constant (e.g.
@@
 Moreover, Gerdt's algorithm may not terminate, as the Gr{\"o}bner basis for the difference ideal may not be finite.
-We overcome these issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) with finite Gr{\"o}bner basis (Section \ref{algodetec}). 
-Our algorithm works for schemes that are explicit and implicit in time and can treat schemes with several higher-order time  and space derivatives, and with several space dimensions.
+We overcome this issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) with finite Gr{\"o}bner basis (Section \ref{algodetec}). 
+Our algorithm works for schemes that are explicit and implicit in time and can treat schemes with several and higher-order time  and space derivatives and with several space dimensions.
 Additionally, we can handle systems of equations and schemes with parameters. 
-We have implemented this algorithm as part of a package in {\sc Mathematica} \cite{wolfram21} \footnote{The code is available upon request. For access contact ricardo.ribeiro@kaust.edu.sa.}.
+We have implemented this algorithm as part of a package in {\sc Mathematica} \cite{wolfram21}.
 In the examples, we show that our code finds conserved quantities and proper schemes for the time-implicit and time-explicit discretization of the Burgers equation and a system of PDEs arising in the study of mean-field games (Section \ref{eaa}).
@@
 \label{assp}
-Here, we follow the ideas in \cite{gomes20}.
-In this paper, subscripts denote indices related to coordinates or tuples and superscripts denote indices related to sequences.
-To avoid boundary terms, we work with periodic functions in $\Zz^d$.
-\begin{definition}
-	Let $N$, $d$ and $m$ be positive integers. The \emph{discrete torus} is $\Ii:=\{0,\ldots,N-1\}^d\subset \Zz^d$. Define the space
+Throughout this section, we follow the ideas and computations in \cite{gomes20}.
+Here, subscripts denote indices related to coordinates or tuples while superscripts denote indices related to sequences and families.
+To avoid dealing with boundary terms, we work with periodic functions in $\Zz^d$.
+\begin{definition}
+	Let $N$, $d$ and $m$ be natural numbers. The \emph{discrete torus} is $\Ii:=\{0,\ldots,N-1\}^d\subset \Zz^d$. Define the space
 	\begin{gather*}
@@
 \end{definition}
-%Assuming periodicity allows us to extend the functions in $\mathcal{P}(\Ii)$ to $\Zz^d$, and we avoid discussing boundary terms.
 \begin{definition}
@@
 In the previous definition, the $e^1,\ldots,e^k$ are not necessarily unit vectors and can vary between functionals.
+Here, we allow for smooth $G$.
+However, later we require $G$ to be a polynomial function.
 \begin{example}
 	\label{functional1}
-	Let $d=m=1$ and
-	$	\mathcal{J}[u]:=\sum_{n}(u(n+1)-u(n))^2$.
+	Let $d=1$, $m=1$ and define
+	\begin{gather*}
+		\mathcal{J}[u]:=\sum_{n\in\Ii}(u(n+1)-u(n))^2.
+	\end{gather*}
 \end{example}
@@
 	\end{gather*}
-	Below in \eqref{periodicity}, we shift the indices due to periodicity of $u$ and $v$,
+	In \eqref{periodicity}, we shift the indices due to periodicity of $u$ and $v$,
 	\begin{gather}
@@
 	\notag	+\ldots+D_kG(u(n+e^1-e^k),u(n+e^2-e^k),\ldots,u(n))v(n)\\
-	\notag	=\sum_{n\in\Ii}\Big(D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))\\
+	\notag	=\sum_{n\in\Ii}\Big(D_1G(u(n),u(n+e^2-e^1)\ldots,u(n+e^k-e^1))\\
 	\notag	+D_2G(u(n+e^1-e^2),u(n),\ldots,u(n+e^k-e^2))\\
 	\notag	+\ldots+D_kG(u(n+e^1-e^k),u(n+e^2-e^k),\ldots,u(n))\Big)v(n)\\
-	\notag	=:\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n).	
+	\notag	=:\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u]v(n).	
 	\end{gather}
 \begin{definition}
-$\mathcal{V}(\mathcal{J})$ is the \emph{discrete variational derivative} of $\mathcal{J}$.
+$\mathcal{V}(\mathcal{J})$ is the discrete variational derivative of $\mathcal{J}$.
 \end{definition}
@@
 		=\sum_{n\in\Ii}\Big(-2u(n+1)+4u(n)-2u(n-1)\Big)v(n)\\
-		=\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n)
-	\end{gather*}
-	and hence $\mathcal{V}(\mathcal{J})[u](n)=-2u(n+1)+4u(n)-2u(n-1)$.
-\end{example}
+		=\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u]v(n)
+	\end{gather*}
+	and hence $\mathcal{V}(\mathcal{J})[u]=-2u(n+1)+4u(n)-2u(n-1)$.
+\end{example}
+
+\subsection*{Fundamental theorem}
 The algorithms for the simplification of sums presented in \cite{gomes20} rely on the following result:
@@
 If $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[v]=0$ for all $v\in\mathcal{P}(\Ii,\Rr^m)$ and if there exists $u_0\in\mathcal{P}(\Ii,\Rr^m)$ such that $\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]$, then
-		$\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$
-	for all $u\in\mathcal{P}(\Ii,\Rr^m)$.
+	\begin{equation*}
+		\mathcal{J}[u]=\tilde{\mathcal{J}}[u]
+	\end{equation*}
+	for all $u\in\mathcal{P}(\Ii,\Rr^m)$.\\
 	Conversely,
-		 $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u]=0$
+	\begin{equation*}
+		 \mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u]=0
+	\end{equation*}
 	for all $u\in\mathcal{P}(\Ii,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
@@
 	\begin{gather*}
-		\mathcal{J}[u]:=\sum_{n\in\Ii}u(n)u(n+1)\text{ and }
+		\mathcal{J}[u]:=\sum_{n\in\Ii}u(n)u(n+1),\\
 		\tilde{\mathcal{J}}[u]:=\sum_{n\in\Ii}u(n-2)u(n-1).
@@
 \begin{example}
-	Let $d=2$ and $m=1$. It may not be obvious, that
+	Let $d=2$. It may not be obvious, that
 	\begin{gather*}
@@
 We call $n$ the space variables.
-Let $l\in\Nn$ and call $t:=(t_1,\ldots,t_l)\in\Nn^l$ the time variables.
+Let $l\in\Nn$ and call $t=(t_1,\ldots,t_l)\in\Nn^l$ the time variables.
 Later in this paper, we only consider the case $l=1$.
@@
 		\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l)):=\\	
-		\Bigg\{\mathcal{J}:\mathcal{P}(\Ii\times\Nn^l)\rightarrow\mathcal{D}(\Nn^l) \Bigg|\ \mathcal{J}[u](t)=\sum_{n\in\Ii}F_{(n,t)}[u],\\
+		\Bigg\{\mathcal{J}:\mathcal{P}(\Ii\times\Nn^l)\rightarrow\mathcal{D}(\Nn^l) \Bigg|\ \mathcal{J}[u]=\sum_{n}F_{(n,t)}[u],\\
 		\text{where } F_{(n,t)}[u]=G(u((n,t)+e^1),u((n,t)+e^2),\ldots,u((n,t)+e^k)),\\
@@
 \end{definition}
-Here, all quantities, whose conservation in time we check, are functions $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$.
-%We restrict $G$ to the set of polynomials in $k$ variables, because our algorithm below can only treat polynomial quantities.
-Let $\tilde{n}:=(n_1,\dots,n_d,s_1,\dots,s_l)\in\Ii\times\Nn^l$. Using the Kronecker delta $\delta$, we rewrite
+Throughout this paper, all quantities, whose conservation in time we check, are functions $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$. 
+Define $\tilde{n}:=(n_1,...,n_d,s_1,...,s_l)\in\Ii\times\Nn^l$. Using the Kronecker delta $\delta$, we rewrite
 \begin{gather*}
@@
 	=\sum_{\tilde{n}\in\Ii\times\Nn^l}F_{\tilde{n}}[u]\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l).
-	%=\sum_{n\in\Ii}G(u(n+e^1),u(n+e^2),\ldots,u(n+e^k))\delta_{t_1,\ldots,t_{d-l+1}}(n_l,\ldots,n_d)
 \end{gather*}
@@
 		+D_kG(u(\tilde{n}+e^1-e^k),\ldots,u(\tilde{n}))\delta_{(t_1+e^k_{d+1},\ldots,t_{l}+e^k_{d+l})}(s_1,\ldots,s_l)\Big)v(\tilde{n})\\
-		=:\sum_{\tilde{n}}\mathcal{V}^l(\mathcal{J})[u](\tilde{n},t)v(\tilde{n}).
+		=:\sum_{\tilde{n}}\mathcal{V}^l(\mathcal{J})[u](t_1,\ldots,t_{l})v(\tilde{n}).
 	\end{gather*}
@@
 	\begin{definition}
-	We define $\mathcal{V}^l(\mathcal{J})$ as the \emph{discrete partial variational derivative} of $\mathcal{J}$.
+	We define $\mathcal{V}^l(\mathcal{J})$ as the discrete partial variational derivative of $\mathcal{J}$.
 	Note that $\mathcal{V}^0:=\mathcal{V}$.
@@
 		=\sum_{n,s}v(n,s+1)\delta_t(s)-v(n,s)\delta_t(s)=\sum_{n,s}v(n,s)\delta_t(s-1)-v(n,s)\delta_t(s)\\
-		=\sum_{n,s}(\delta_{t+1}(s)-\delta_t(s))v(n,s)=\sum_{n,s}\mathcal{V}^1(\mathcal{J})(n,s,t)v(n,s).
+		=\sum_{n,s}(\delta_{t+1}(s)-\delta_t(s))v(n,s)=\sum_{n,s}\mathcal{V}^1(\mathcal{J})(t)v(n,s).
 	\end{gather*}
@@
 	In our {\sc Mathematica} implementation, Example \ref{dpvd1} is computed by
-	\begin{EXE}
-		LABEL="dpvd1"
-		(*CODE*)
-		variables = Association["indVars"->{n,t},"depVars"->{u},"timeVars"->{t}]
-		expression = u[n,t+1] - u[n,t]
-		PartialDVarDOperator[variables][expression]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}	
-	\begin{small}
-		\input{executables/dpvd1ctt.ed.tex}
-		\input{executables/dpvd1ctt.result.ed.tex}
+
+	\begin{small}
+
 	\end{small}
@@
 If $\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[v]=0$ for all $v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ and if there exists $u_0\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ such that $\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]$, then
-		$\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$
-	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$.
+	\begin{equation*}
+		\mathcal{J}[u]=\tilde{\mathcal{J}}[u]
+	\end{equation*}
+	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$.\\
 	Conversely,
-		$\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[u]=0$
+	\begin{equation*}
+		\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u]=0
+	\end{equation*}
 	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
@@
 \begin{definition}
-	The \emph{space of schemes} is
+	The space of schemes is
 	\begin{gather*}
@@
 			\text{where }
-		\{\tilde{e}^1,\ldots,\tilde{e}^r\}\subset\Zz^d,\ \{\tilde{l}^1,\ldots,\tilde{l}^r\}\subset\Zz,\\
-		\text{H is a polynomial in r variables}
+		\{\tilde{e}^1,\ldots,\tilde{e}^r\}\subset\Zz^d,\ \{\tilde{l}^1,\ldots,\tilde{l}^r\}\subset\Zz
 		\Bigg\}.
 	\end{gather*}
-	A \emph{scheme} is a set of functions
+	A scheme is a set of functions
 	\begin{gather*}
@@
 If a scheme contains a single function $H$, we call $H$ the scheme. Sometimes, we also call the expression $H[u]$ the scheme.
-\begin{remark}
-	In our examples, we only consider finite-difference schemes with fixed step sizes $\Delta x=\Delta t=1$.
-However, our algorithms and our code handle parameters.
-Hence, one can generalize our results to schemes with other step sizes.
-\end{remark}
-\begin{definition}
-	A scheme $\{H^1,\ldots,H^m\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ given by
-	\begin{equation*}
-		H^j(u(n+\tilde{e}_j^1,t+\tilde{l}_j^1),\ldots,u(n+\tilde{e}_j^r,t+\tilde{l}_j^r))
-	\end{equation*}
-	is in \emph{time-explicit form} if we can rewrite the previous equation (eventually translating the scheme) as
-	\begin{equation*}
-		u_j(n,t+1)-\tilde{H}^j(u(n+\tilde{\tilde{e}}_j^1,t),u(n+\tilde{\tilde{e}}_j^2,t),\ldots,u(n+\tilde{\tilde{e}}_j^r,t))
-	\end{equation*}
-	for $1\leq j\leq m$
-	with $\tilde{H}^j\in\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$. We call $\{\tilde{H}^1,\ldots,\tilde{H}^m\}$ the right-hand side of $\{H^1,\ldots,H^m\}$.
+In this work, we only consider finite-difference schemes with fixed step size $h=1$, but one can generalize our results and algorithms to other step sizes.
+\begin{definition}
+	A scheme $\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ given by
+	\begin{equation*}
+		H^k(u(n+\tilde{e}_k^1,t+\tilde{l}_k^1),,\ldots,u(n+\tilde{e}_k^r,t+\tilde{l}_k^r))
+	\end{equation*}
+	is in time-explicit form if we can rewrite the previous equation (eventually translating the scheme) as
+	\begin{equation*}
+		u_j(n,t+1)-\tilde{H}^k(u(n+\tilde{\tilde{e}}_k^1,t),u(n+\tilde{\tilde{e}}_k^2,t),\ldots,u(n+\tilde{\tilde{e}}_k^r,t))
+	\end{equation*}
+	for $1\leq k\leq i$ and $1\leq j\leq m$
+	with $\tilde{H}^k\in\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$. We call $\{\tilde{H}^1,\ldots,\tilde{H}^i\}$ the right-hand side of $\{H^1,\ldots,H^i\}$.
 \end{definition}
 \begin{example}\label{timeexplicitheateq}
-	Consider the forward-difference scheme for the heat equation $u_t-u_{xx}=0$ \cite{smoller94}
+	Consider the forward-difference scheme for the heat equation \cite{smoller94}
 	\begin{equation*}
@@
 \end{example}
-{\bf Difference ideals.}\label{diffid}
+\subsection*{Difference ideals}\label{diffid}
 Following Gerdt \cite{gerdt12}, we now introduce difference ideals and how we use them in our algorithm.
 \begin{definition}
-	Let $u=(u_1,\dots,u_m)\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$. The \emph{shift} in the $i$-th coordinate for $1\leq i\leq d+1$ by $k\in\Zz$ is
-	\begin{gather*}
-		\sigma_i^k\circ u_j(n,t):=u_j(n_1,\ldots,n_{i-1},n_i+k,n_{i+1},\ldots,t)
+	Let $u=(u_1,\dots,u_m)\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$. The shift in the $i$-th coordinate for $1\leq i\leq d+1$ by $k\in\Zz$ is
+	\begin{gather*}
+		\sigma_i^k\circ u_j(n+e,t+l):=u_j(n_1+e_1,\ldots,n_i+e_i+k,\ldots,t+l)
 	\end{gather*}
@@
 \begin{definition}
-	The \emph{set of all possible shifts} $\Theta$ is
+	The set of all possible shifts $\Theta$ is
 	\begin{gather*}
@@
 Let $\ \tilde{\mathcal{R}}$ be the polynomial ring over the field $\ \mathcal{K}$ and the variables $\theta\circ u_j(n,t)$ for $\theta\in\Theta$ and $1\leq j\leq m$.
-	A set $\tilde{I}\subseteq\tilde{\mathcal{R}}$ is a \emph{difference ideal} if
-	$p_1,p_2\in\tilde{I}$ implies $p_1+p_2\in\tilde{I}$;
-	 $p_1\in\tilde{I},p_2\in\tilde{\mathcal{R}}$ implies $p_1\cdot p_2\in\tilde{I}$ and
-	 $p_1\in\tilde{I},\theta\in\Theta$ implies $\theta\circ p_1\in\tilde{I}$.
+	A set $\tilde{I}\subseteq\tilde{\mathcal{R}}$ is a difference ideal if
+	\begin{itemize}
+		\item $p_1,p_2\in\tilde{I}$ implies $p_1+p_2\in\tilde{I}$,
+		\item $p_1\in\tilde{I},p_2\in\tilde{\mathcal{R}}$ implies $p_1\cdot p_2\in\tilde{I}$,
+		\item $p_1\in\tilde{I},\theta\in\Theta$ implies $\theta\circ p_1\in\tilde{I}$.
+	\end{itemize}
 \end{definition}
@@
 A solution of a numerical scheme is a function $u$, that makes all translations of the scheme vanish. Hence, every element of the difference ideal generated by the scheme vanishes under $u$.
-Algorithm \ref{alg: discreteconservedqimplicit} seeks to determine if the polynomial associated with the discrete time derivative $\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$ belongs to the ideal $\langle H^1,\ldots,H^i\rangle$.
+In particular, Algorithm \ref{alg: discreteconservedqimplicit} seeks to determine if the discrete time derivative $\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$ belongs to the ideal $\langle H^1,\ldots,H^i\rangle$.
 To examine, if $\mathcal{T}\in\langle H^1,\ldots,H^i\rangle$, Gerdt proposes the notion of a standard (Gr{\"o}bner) basis for the ideal $\langle H^1,\ldots,H^i\rangle$.
 This standard basis may not be finite.
-We overcome this problem by considering polynomial instead of difference algebra and using a smaller polynomial ideal, that is contained in the difference ideal and admits a finite standard basis.
+We overcome this problem by considering polynomial instead of difference algebra and using a smaller polynomial ideal, that is contained in the difference ideal and admits a finite standard basis.\\
 However, there may be functions $\mathcal{J}$, whose sums add to zero, even though they may fail to belong to the difference ideal.
@@
 In this section, we present Algorithm \ref{alg: discreteconservedqexplicit}  for checking the conservation of a quantity under the right-hand side of a scheme in time-explicit form.
-We discuss our algorithm for general schemes in the next section.
-\begin{algorithm}
-	\DontPrintSemicolon
-	\KwIn{$\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{D}(\Nn))$ and $\{H^1,\ldots,H^m\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ right-hand side of a time-explicit scheme}
-	\KwOut{ {\sc True}, if $\mathcal{J}$ is conserved in time, {\sc False} otherwise }
-
-	\label{estep1} Build the discrete time derivative $\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$\;
-
-	\label{estep2} Replace all instances of $u_j(n+e,t+1)$ (for $e\in\Zz^{d}$) by an appropriate translation of $\{H^1,\ldots,H^m\}$\;
-
-	\label{estep3} Compute the discrete partial variational derivative\;
-
-	\label{estep4} If the result is zero, conservation is {\sc True}, else {\sc False}\;
-
-	\caption{{\sc time-explicit DiscreteConservedQ}}
-	\label{alg: discreteconservedqexplicit}
-\end{algorithm}
-Note, that the result of Step 2 exists only at time $t$ because all instances of $t+1$ have been replaced. Thus, we do not need any computations of difference ideals or Gr{\"o}bner basis.
+Our algorithm for general schemes can be found in the subsequent section.
+
+Note, that the result of step 2 exists only at time $t$ because all instances of $t+1$ have been replaced. Thus, we do not need any computations of difference ideals or Gr{\"o}bner basis.
 \begin{example}
 	\label{timeexplicitheateqalgo}
-	We check for conservation of $\sum_nu(n,t)$ under the scheme
-	\begin{equation*}
-		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t)).
-	\end{equation*}
-	The discrete time derivative (Step 1) is
-	\begin{gather*}
-		\sum_{n\in\Ii}u(n,t+1)-u(n,t).
-	\end{gather*}
-	Replace $u(n,t+1)$ by the right-hand side (Step 2) to get
-	\begin{gather*}
-		\sum_{n\in\Ii}u(n,t)+(u(n+1,t)-2u(n,t)+u(n-1,t))-u(n,t).
-	\end{gather*}
-	Then, we compute the discrete partial variational derivative (Step 3), which equals zero. Hence (Step 4), we have conservation.
+	We check for conservation of $\sum_nu(n,t)$ under the scheme from Example \ref{timeexplicitheateq}.
+	The discrete time derivative (step 1) is
+	\begin{gather*}
+		\sum_nu(n,t+1)-u(n,t).
+	\end{gather*}
+	Replacing $u(n,t+1)$ by the right-hand side of the scheme (step 2) gives
+	\begin{gather*}
+		\sum_nu(n,t)+(u(n+1,t)-2u(n,t)+u(n-1,t))-u(n,t).
+	\end{gather*}
+	Then, we compute the discrete partial variational derivative (step 3), which equals zero. Hence (step 4), we have conservation.
 \end{example}
@@
 	We verify the result from Example \ref{timeexplicitheateqalgo}, using our implementation in {\sc Mathematica} of Algorithm \ref{alg: discreteconservedqexplicit}.
-	\begin{EXE}
-		LABEL="rhs1"
-		(*CODE*)
-		variables = Association["indVars"->{n},"depVars"->{u},"rhs"->{u[n]+(u[n+1]-2*u[n]+u[n-1])}]
-		expression = u[n]
-		DiscreteConservedQOperator[variables][expression]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
-	\begin{small}
-		\input{executables/rhs1ctt.ed.tex}
-		\input{executables/rhs1ctt.result.ed.tex}
+
+	\begin{small}
+
 	\end{small}
@@
 In this section, we present Algorithm \ref{alg: discreteconservedqimplicit} that deals with general, not necessarily time-explicit, schemes.
-We explain its steps in detail below and demonstrate the algorithm in Example \ref{heateqalgoimplicit}.\\
-%In this section we explain our algorithm, that checks if a given quantity is preserved in time under a given scheme in detail.
-\begin{algorithm}
-	\DontPrintSemicolon
-	\KwIn{$\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{D}(\Nn))$ and $\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ }
-	\KwOut{ {\sc True}, if $\mathcal{J}$ is conserved in time, {\sc False} otherwise }
-
-	\label{step1} Build the discrete time derivative $\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$\;
-
-	\label{step2}Translate $\{H^1,\ldots,H^i\}$ by subtracting the range of the stencil of the scheme from the range of the stencil of $\mathcal{T}$\;
-
-	\label{step3} Reduce $\mathcal{T}$ by using the Gr{\"o}bner basis generated by the translated scheme in two different ways
-	\label{step3a} (a) Using the lexicographic order
-	\label{step3b} (b) Using the explicit elimination order\;
-
-	\label{step4} Choose the result that admits the least different instances of time\;
-
-	\label{step5} Compute the discrete partial variational derivative\;
-
-	\label{step6} Apply Steps 2 to 4 to the result\;
-
-	\label{step7} If the result is zero, conservation is {\sc True}, else {\sc False}\;
-
-	\caption{{\sc General DiscreteConservedQ}}
-	\label{alg: discreteconservedqimplicit}
-\end{algorithm}
-{\bf Step 1: Build the discrete time derivative.}
+We explain its steps in detail below and demonstrate the algorithm in Example \ref{heateqalgoimplicit}.
+
+\subsection*{Step 1: Build the discrete time derivative}
 We build the discrete time derivative by subtracting $\mathcal{J}[u](t)$ from $\mathcal{J}[u](t+1)$. This task is done in our code by the {\sc TimeDifference}.
@@
 		The discrete time derivative for $\sum_n u(n,t)$ is
-		\begin{EXE}
-			LABEL="heateq2"
-			(*CODE*)
-			variables = <|"indVars"->{n,t},"depVars"->{u}|>
-			expression = u[n,t]
-			TimeDifferenceOperator[variables][expression]
-			(*CODE*)
-			PUNCT={,.}
-		\end{EXE}
-
-		\begin{small}
-			\input{executables/heateq2ctt.ed.tex}
-			\input{executables/heateq2ctt.result.ed.tex}
-		\end{small}
-	\end{example}
-%	Our code allows us to check not only for conservation in time but also if higher time derivatives are zero under the scheme by specifying a {\sc timederivativeorder} in the {\sc variables}.	
-{\bf Step 2: Translate the scheme.}
-%	Now we compute the translations of the scheme.
-As mentioned above, difference algebra does not guarantee termination of the algorithm.
-Hence, we need to find a finite set of translations of the scheme, to guarantee termination of the algorithm.
-We translate the scheme such, that the range of its stencil stays exactly inside the range of the stencil of the discrete time derivative.
-This guarantees, that every instance of $u_j(n+e,t+l)$, occuring in $\mathcal{T}$, also occurs in the translated scheme.
-We now make the step from difference algebra, with an infinite number of independent variables, to polynomial algebra, where we only have a finite number of variables to guarantee finiteness of the Gr{\"o}bner basis.
-Hence, we need to compute a finite number of translations of the scheme and then consider every instance $u_j(n+e,t+l)$, that appears either in the translated scheme or in the discrete time derivative as independent polynomial variable.
-\begin{example}
-	The algorithm treats the discrete time derivative $\sum_{n\in\Ii}u(n+1,t+1)-u(n+1,t)$
-	as the polynomial $z_1-z_2$.
-\end{example}
-There are several possiblities which translations of the scheme to take as basis for the polynomial ideal.
-However, an obvious idea is to compute the minimal number of translations of the scheme, such that all instances of $u_j(n+e,t+l)$, that appear in the range of the stencil of the discrete time derivative, also appear in the translated scheme.
-This is done by elementwise subtracting the range of the stencil of the scheme from the range of the stencil of the discrete time derivative.
-Before we define the above translations formally, we give a simple example to illustrate the idea of computing the translations:
-\begin{example}
-	Consider the expression
-	\begin{gather*}
-		u(n+3)-u(n-2)
-	\end{gather*}
-	which we want to reduce by the scheme $u(n+1)-u(n)=0$. We compute the translations as described above as
-	\begin{gather*}
-		\{u(n+3)-u(n+2),\dots,u(n-1)-u(n-2)\}.
-	\end{gather*}
-	and see, that the expression indeed reduces to $0$ under the scheme.
+
+		\begin{small}
+
+		\end{small}
+	\end{example}
+\subsection*{Step 2: Translate the scheme}
+To make the step from difference to polynomial algebra to guarantee termination of the algorithm, we use a standard idea in symbolic computations and consider every instance of $u_j(n+e,t+l)$ as a variable of a multivariate polynomial.
+Hence, we compute all translations of the scheme, such that all variables in the polynomial associated with $\mathcal{T}$ appear in the polynomials associated with the translated scheme.
+\begin{example}
+	The algorithm treats the discrete time derivative $u(n+1,t+1)-u(n+1,t)$
+	as the polynomial equation $x_1-x_2$.
 \end{example}
@@
 				\mathcal{P}(\Ii\times\Nn,\Rr^m)\rightarrow\mathcal{P}(\Ii\times\Nn,\Rr),\\
-				F[u](n,t)=F(u_1((n,t)+x^1),\ldots,u_1((n,t)+x^{k_1}),\\
+				F[u]=F(u_1((n,t)+x^1),\ldots,u_1((n,t)+x^{k_1}),\\
 				u_2((n,t)+x^{k_1+1}),
@@
 	\end{gather*}
-	where $k_0=0$, $x^{k_l +}_i = \max_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$,
-	and\\ $x^{k_l -}_i = \min_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$. Here, $[a,b]$ denotes the discrete interval in $\Zz$, i.e. $[0,2]=\{0,1,2\}$.
-	\end{definition}
+	where $k_0=0$, $x^{k_l +}_i = \max_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$,\\
+	and $x^{k_l -}_i = \min_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$. Here, $[a,b]$ denotes the discrete interval in $\Zz$, i.e. $[0,2]=\{0,1,2\}$.
+	\end{definition}
+We calculate the minimal necessary translations of the scheme, such that all instances of $u_j(n+e,t+l)$, that appear in the discrete time derivative, also appear in the translated scheme.
+This is done by elementwise subtracting the range of the stencil of the scheme from the range of the stencil of the discrete time derivative.
+We denote the resulting translated system by $\{H^1,\ldots,H^r\}$.
 If for any $[a,b]$ involved $b<a$, we write $[a,b]=\emptyset$ with the convention that $\emptyset-\emptyset=\emptyset$.
@@
 		Consider the discrete time derivative
-		\footnotesize
 		\begin{equation*}
@@
 		\end{equation*}
-		\normalsize
 		and the scheme
-		\footnotesize
-		$
+		\begin{equation*}
 			\{u_1(n+1,t+1)-u_1(n,t),u_2(n,t+1)-u_2(n,t)\}.
-		$
-		\normalsize
+		\end{equation*}
 		Then, the stencil of the discrete time derivative equals 
-		\footnotesize
 		\begin{equation*}
@@
 		\end{equation*}
-		\normalsize
 		with range
-		\footnotesize
-		$
+		\begin{equation*}
 			\Big([0,2]\times[0,1],[0,0]\times [0,1]\Big).
-		$
-		\normalsize
+		\end{equation*}
 		The stencil of the scheme equals the set of stencils of the equations in the scheme, i.e.
-		\footnotesize
-		\begin{equation*}
-			\Big\{\Big(\big\{(1,1),(0,0)\big\},\emptyset\Big),\Big(\emptyset,\big\{(0,1),(0,0)\big\}\Big)\Big\}
-		\end{equation*}
-		\normalsize
+		\begin{equation*}
+			\Bigg\{\Big(\big\{(1,1),(0,0)\big\},\emptyset\Big),\Big(\emptyset,\big\{(0,1),(0,0)\big\}\Big)\Bigg\}
+		\end{equation*}
 		with range
-		\footnotesize
-		$
-			\Big\{\Big([0,1]\times[0,1],\emptyset\Big),\Big(\emptyset,[0,0]\times [0,1]\Big)\Big\}.
-		$
-		\normalsize
+		\begin{equation*}
+			\Bigg\{\Big([0,1]\times[0,1],\emptyset\Big),\Big(\emptyset,[0,0]\times [0,1]\Big)\Bigg\}.
+		\end{equation*}
 		Hence, we get the translations for the first equation of the scheme by
@@
 		Therefore, we translate the first equation of the scheme by $(0,0)$ and $(1,0)$ and the second one by $(0,0)$ to get the translated scheme
-		\footnotesize
 		\begin{gather*}
 			\{u_1(n+1,t+1)-u_1(n,t),
-			u_1(n+2,t+1)-u_1(n+1,t),
+			u_1(n+2,t+1)-u_1(n+1,t),\\
 			u_2(n,t+1)-u_2(n,t)
@@
 		\end{gather*}
-		\normalsize
-	\end{example}
-{\bf Step 3 and 4: Compute the Gr{\"o}bner basis and reduce the discrete time derivative.}\\
-	\emph{Polynomial ideals and Gr{\"o}bner bases.}\label{polid}
-	We found a finite set of polynomials (the translated scheme $\{H^1,\ldots,H^r\}$ and the discrete time derivative) with a finite number of instances of $u_j(n+e,t+l)$ in the previous step.
-Now we reduce $\mathcal{T}$ as an element of a polynomial ring by multivariate polynomial division with respect to the translated scheme.
+	\end{example}
+\subsection*{Step 3 and 4: Compute the Gr{\"o}bner basis and reduce the discrete time derivative}
+	\subsubsection*{Polynomial ideals and Gr{\"o}bner bases}\label{polid}
+	We found a finite set of polynomials (the translated scheme and the discrete time derivative) with a finite number of instances of $u_j(\tilde{n}+e,t+l)$ in the previous step.
+Now we reduce $\mathcal{T}$, using the translated scheme.
 Here, we adapt the definitions and theorems from \cite{hibi13} to our setting.
 	\begin{definition}
-		Let $\mathcal{K}$ be as in Definition \ref{diffideal}. Let $\mathcal{R}$ be the polynomial ring generated by $\mathcal{K}$ and all instances of $u_j(n+e,t+l)$ that occur in $\{H^1,\ldots,H^r\}$.
-		We call a set $ I\subseteq \mathcal{R}$ a \emph{(polynomial) ideal} if
-		$p_1,p_2\in I$ implies $p_1+p_2\in I$ and
-		 $p_1\in I,p_2\in\mathcal{R}$ implies $p_1p_2\in I$.
+		Let $\mathcal{K}$ be as in Definition \ref{diffideal}.
+Let $\mathcal{R}$ be the polynomial ring generated by $\mathcal{K}$ and all instances of $u_j(n+e,t+l)$ that occur in $\{H^1,\ldots,H^r,\mathcal{T}\}$.
+		We call a set $ I\subseteq \mathcal{R}$ a (polynomial) ideal if
+		\begin{itemize}
+			\item $p_1,p_2\in I$ implies $p_1+p_2\in I$,
+			\item $p_1\in I,p_2\in\mathcal{R}$ implies $p_1p_2\in I$.
+		\end{itemize}
 	\end{definition}
@@
 	Replacing $\{H^1,\ldots,H^r\}$ by a Gr{\"o}bner basis for the ideal $\langle H^1,\ldots,H^r \rangle$ guarantees the uniqueness of the remainder of the polynomial division. 
-	Contrary to the standard basis for the difference ideal that Gerdt's algorithm computes, a Gr{\"o}bner basis for the polynomial ideal always exists and is finite.
-	A Gr{\"o}bner basis is defined up to the order of the monomials involved, so, depending on the order, we get different remainders of the polynomial division.\\
-	\emph{Polynomial reduction.}
-	In Algorithm \ref{alg: discreteconservedqimplicit}, we consider two monomial orders: the (a) lexicographic order and an experimental (b) elimination order.
-	To reduce $\mathcal{T}$, we compute the Gr{\"o}bner basis of $\langle H^1,\ldots,H^r\rangle$ with respect to the monomial order and then calculate the remainder of the polynomial division of $\mathcal{T}$ with respect to this Gr{\"o}bner basis.
-	The (a) lexicographic order is induced by the time-explicit ordering of the instances of $u_j(n+e,t+l)$, which gives more weight to instances at later times than to instances at earlier times.
-	For the (b) elimination order, we first order all instances of $u_j(n+e,t+l)$ according to a pre-defined ordering.
-Our code uses by default the time-explicit ordering, but other orderings, like an implicit ordering or an user-defined ordering, are also possible.
-Then, we try to successively eliminate the instance with the highest ordering from the discrete time derivative, using the elimination order induced by a weight-matrix.
+	Contrary to the standard basis for the difference ideal that Gerdt's algorithm computes, the Gr{\"o}bner basis for the polynomial ideal always exists and is finite.\\
+	A Gr{\"o}bner basis is defined up to the order of the monomials involved, so, depending on the order, we get different remainders of the polynomial division.
+	\subsubsection*{Polynomial reduction}
+	In Algorithm \ref{alg: discreteconservedqimplicit}, we consider two monomial orders: (a) lexicographic and (b) explicit.
+	To reduce $\mathcal{T}$, we compute the Gr{\"o}bner basis of $\langle H^1,\ldots,H^r\rangle$ with respect to the monomial order and then calculate the remainder of the polynomial division of $\mathcal{T}$ with respect to this Gr{\"o}bner basis.\\
+	The explicit monomial order (b) results in the Gr{\"o}bner basis that eliminates the instances of $u_j(n+e,t+l)$ at the latest time (with the largest $l$).
 This elimination might not always be possible, hence the reduction may leave the discrete time derivative unchanged.
-Then we repeat this process for the instance at the second latest time and continue until all instances of $u_j(n+e,t+l)$ have been eliminated.
-	However, the elimination order is rather experimental, as its weight-matrix does not have a full rank and hence does not induce a total order on the set of monomials.\\
-	Out of the two remainders from (a) and (b), we choose the result with the least number of different instances of time.\\
-{\bf Step 5 to 7: Compute the discrete partial variational derivative and reduce again.}
+Then we repeat this process for the instance at the second latest time and continue until all instances of $u$ have been eliminated.\\
+	Out of the two remainders from (a) and (b), we choose the result with the least number of different instances of time. Our code also supports other elimination orders, including user-defined ones.
+\subsection*{Step 5 to 7: Compute the discrete partial variational derivative and reduce again}
 	In the next step, we take the resulting expression and compute its discrete partial variational derivative. Then, we repeat Step 2 to Step 4 applied to the discrete partial variational derivative.
@@
 		To demonstrate Algorithm \ref{alg: discreteconservedqimplicit}, we use the setting from Example \ref{timeexplicitheateqalgo}.
-		We compute the discrete time derivative (Step 1)
-		\begin{gather*}
-			\mathcal{T}=\sum_{n\in\Ii} u(n,t+1)-u(n,t).
-		\end{gather*}
-		For the translations (Step 2), the stencil of the scheme
-		\begin{equation*}
-			u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t))
-		\end{equation*}
-		\small
-		is
-		$
-			\Big(\big\{(0,1),(0,0),(1,0),(-1,0)\big\}\Big)
-		$
-		\normalsize
-	with range \small$\Big([-1,1]\times[0,1]\Big)$\normalsize.
-	The discrete time derivative has stencil \small$\Big(\big\{(0,1),(0,0)\big\}\Big)$\normalsize 
-	 and range \small$\Big([0,0]\times[0,1]\Big)$\normalsize.
+		We compute the discrete time derivative (step 1)
+		\begin{gather*}
+			\mathcal{T}=\sum_n u(n,t+1)-u(n,t).
+		\end{gather*}
+		For the translations (step 2), the stencil of the scheme is
+		\begin{gather*}
+			\Big(\big\{(0,1),(0,0),(-1,0),(1,0)\big\}\Big)
+		\end{gather*}
+	with range $\Big([-1,1]\times[0,1]\Big)$.
+	The discrete time derivative has stencil $\Big(\big\{(0,1),(0,0)\big\}\Big)$
+	and range $\Big([0,0]\times[0,1]\Big)$.
 	The translations are
@@
 	which results in no translations, as the range of the stencil of the scheme is greater than the range of the stencil of the discrete time derivative.
-	Hence, the Gr{\"o}bner basis (Step 3) coincides with the original scheme for both monomial orders.
+	Hence, the Gr{\"o}bner basis (step 3) coincides with the original scheme for both monomial orders.
 	The reduction using the lexicographic monomial order yields as remainder
@@
 	For the reduction using the elimination order, we first eliminate $u(n,t+1)$ and then the remaining instances of $u$.
-	The elimination order yields the same remainder as the lexicographic order. Hence, we do not check for the number of instances of time (Step 4).
-	We calculate the discrete partial variational derivative (Step 5) of
-	\begin{gather*}
-		\sum_{n\in\Ii}u(n-1,t)-2u(n,t)+u(n+1,t)
+	The elimination order yields the same remainder as the lexicographic order. Hence, we do not need to check for the number of instances of time (step 4).
+	We calculate the discrete partial variational derivative (step 5) of
+	\begin{gather*}
+		\sum_nu(n-1,t)-2u(n,t)+u(n+1,t)
 	\end{gather*}
 	that equals $0$.
-	Hence, repeating the above procedure (Step 6) becomes unnecessary.
-	Therefore, (Step 7) we see that the scheme conserves the quantity.
+	Hence, repeating the above procedure (step 6) becomes unnecessary.
+	Therefore, (step 7) we see that the scheme conserves the quantity.
 	\end{example}
@@
 		We verify our result from Example \ref{heateqalgoimplicit}, using {\sc DiscreteConservedQ}, which detects automatically if we are in the time-explicit or the general setting.
-		\begin{EXE}
-		LABEL="heateq10"
-		(*CODE*)
-		variables = Association["indVars"->{n,t},"depVars"->{u},"scheme"->{u[n,t+1]-u[n,t]-(u[n+1,t]-2*u[n,t]+u[n-1,t])}]
-		expression = u[n,t]
-		DiscreteConservedQOperator[variables][expression]
-		(*CODE*)
-		PUNCT={,.}
-		\end{EXE}
-		\begin{small}
-		\input{executables/heateq10ctt.ed.tex}
-		\input{executables/heateq10ctt.result.ed.tex}
-		\end{small}
-	\end{example}
-\begin{remark}\label{remark}
+
+		\begin{small}
+
+		\end{small}
+	\end{example}
+\begin{remark}
 	The algorithm can only detect conservation, but can not check if something is not conserved for sure.
@@
 \end{remark}
-%\begin{remark}
-%	Our algorithm is also able to check for detection of time derivatives of higher order.
-To check for example the conservation of the first time derivative of the solution of the scheme, one can %specify {\sc timederivativeorder $\rightarrow$ 2} when calling {\sc DiscreteConservedQOperator}.
-In that case the algorithm will build the discrete forward second order time derivative of the %given quantity.
-%\end{remark}
 \section{A basis for conserved quantities}
@@
 So far, we have discussed how to check if a quantity is preserved in time under a scheme. But it is also desirable to have a systematic way to find conserved quantities, given a scheme.
-Algorithm \ref{alg: findconservedquantitybasis} finds, for a time-explicit scheme, a basis for conserved quantities that are generated by monomials up to a degree. 
-\begin{algorithm}
-	\DontPrintSemicolon
-	\KwIn{right-hand side $\{H^1,\ldots,H^m\}\subset S(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ of time-explicit scheme, list of {\sc generators}, {\sc degree} of polynomials }
-	\KwOut{basis of conserved quantities}
-
-	\label{fstep1} Build all monomials combined from the {\sc generators} up to the total {\sc degree}\;
-
-	\label{fstep2} Linearly combine all monomials with undetermined coefficients\;
-
-	\label{fstep3} Compute the discrete time derivative and plug in the appropriate right-hand side(s) of the scheme\;
-
-	Compute the discrete partial variational derivative\;
-
-	\label{fstep4} Simplify and use the undetermined coefficients method to determine coefficients that make the result vanish\;
-
-	\caption{{\sc FindDiscreteConservedQuantityBasis}}
-	\label{alg: findconservedquantitybasis}
-\end{algorithm}
+Algorithm \ref{alg: findconservedquantitybasis} finds, for a time-explicit scheme, a basis for conserved quantities that are generated by monomials up to a (total) degree.
+
 We have implemented this algorithm in the {\sc FindDiscreteConservedQuantityBasis}.
@@
 	We find a basis for conserved quantities that are generated by $u(n,t)$ and $n$ and that have at most degree 3, admitted by the scheme for the heat equation from Example \ref{heatex1}.
-	\begin{EXE}
-		LABEL="heateq11"
-		(*CODE*)
-		variables = <|"indVars"->{n},"depVars"->{u},"rhs"->{u[n]+u[n+1]-2*u[n]+u[n-1]}|>
-		FindDiscreteConservedQuantityBasisOperator[variables][<|"degree"->3,"generators"->{u[n],n}|>]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
-	\begin{small}
-		\input{executables/heateq11ctt.ed.tex}
-		\input{executables/heateq11ctt.result.ed.tex}
+
+	\begin{small}
+
 	\end{small}
@@
 \label{eaa}
-We illustrate our code with the inviscid Burgers equation and a system of mean-field games.
-These only have first-order time-derivatives, but the algorithm handles PDEs with all orders of derivatives.\\
-{\bf Burgers equation.}
-The inviscid Burgers equation is the PDE $u_t+uu_x=0$ \cite{smoller94} in one space dimension. Any function of $u$ is a conserved quantity. We search for discretizations preserving $\int u\ dx$.
+We illustrate our code with the Burgers equation and a system of mean-field games.
+\subsection*{Burgers equation}
+The Burgers equation is the PDE $u_t+uu_x=0$ \cite{smoller94} in one space dimension. Any function of $u$ is a conserved quantity. We are interested in discretizations that preserve mass $\int u\ dx$.
 \begin{example}
 	We check for conservation of mass using a forward-difference discretization:
-	\begin{EXE}
-		LABEL="burgers1"
-		(*CODE*)
-		variables = Association["indVars"->{n,t},"depVars"->{u},"scheme"->{u[n,t+1]-u[n,t]-u[n,t]*(u[n+1,t]-u[n,t])}];
-		expression = u[n,t];
-		DiscreteConservedQOperator[variables][expression]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
-	\begin{small}
-		\input{executables/burgers1ctt.ed.tex}
-		\input{executables/burgers1ctt.result.ed.tex}
-	\end{small}
-\end{example}
-For finding a scheme that preserves mass, we check for conservation under a class of schemes with a parameter.
-\begin{example}
-	We discretize $u_x$ by a three-point stencil with a parameter $a$.
-	\begin{EXE}
-		LABEL="burgers2"
-		(*CODE*)
-		variables = Association["indVars"->{n,t}, "depVars"->{u}, "pars"->{a}, "scheme"->{u[n,t+1]-u[n,t]-u[n,t]*(a*(u[n+1,t]-u[n,t])+(1-a)*(u[n,t]-u[n-1,t]))}];
-		expression = u[n,t];
-		DiscreteConservedQOperator[variables][expression]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
-	\begin{small}
-		\input{executables/burgers2ctt.ed.tex}
-		\input{executables/burgers2ctt.result.ed.tex}
-	\end{small}
-	Conservation holds for $a=\frac{1}{2}$, the central difference case.
+
+	\begin{small}
+
+	\end{small}
+\end{example}
+We want to find a scheme that preserves this quantity. Therefore we check for conservation under a class of schemes with a parameter, to find an appropriate choice for this parameter.
+\begin{example}
+	We discretize $u_x$ using a three-point stencil with a parameter $a$.
+
+	\begin{small}
+
+	\end{small}
+	We have conservation if we choose $a=\frac{1}{2}$, corresponding to the standard central difference.
 \end{example}
@@
 	We consider the scheme from the previous example but in the time-implicit version.
-	\begin{EXE}
-		LABEL="burgers3"
-		(*CODE*)
-		variables = Association["indVars"->{n,t}, "depVars"->{u}, "pars"->{a},"elimOrder"->"implicit", "scheme"->{u[n,t+1]-u[n,t]-u[n,t+1]*(a*(u[n+1,t+1]-u[n,t+1])+(1-a)*(u[n,t+1]-u[n-1,t+1]))}]
-		DiscreteConservedQOperator[variables][u[n,t]]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
-	\begin{small}
-		\input{executables/burgers3ctt.ed.tex}
-		\input{executables/burgers3ctt.result.ed.tex}
-	\end{small}
-\end{example}
-	Our results also hold for the viscous Burgers equation $u_t+uu_x-u_{xx}=0$, if we use the second-order central difference for $u_{xx}$.\\
-{\bf Conserved quantities for a mean-field game.}
-We study conserved quantities admitted by the discretization of a PDE system.
-\begin{example}
-	In \cite{gomes18}, Gomes et al. derived the following system
-	\begin{gather*}
-		v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0,\quad
+
+	\begin{small}
+
+	\end{small}
+\end{example}
+\subsection*{Conserved quantities for a mean-field game}
+As a second application, we use our code to study conserved quantities admitted by the discretization of a system of PDEs.
+\begin{example}
+	In \cite{gomes18}, Gomes et al. derived the following system of PDEs
+	\begin{gather*}
+		v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0\\
 		m_t-(vm)_x=0.
@@
 We check with our code if the scheme admits the same preserved quantities as the continuous system:
-	\begin{EXE}
-		LABEL="mfg1"
-		(*CODE*)
-		variables = Association[
-		"indVars" -> {n, t},
-		"depVars" -> {v, m},
-		"display result" -> True,
-		"scheme" -> {v[n, t + 1] - 
-			v[n, 
-			t] + (v[n + 1, t]^2/2 - m[n + 1, t]^2/2) - (v[n, t]^2/2 - 
-			m[n, t]^2/2), 
-			m[n, t + 1] - 
-			m[n, t] - (m[n + 1, t]*v[n + 1, t]) + (m[n, t]*v[n, t])}
-		];
-		DiscreteConservedQOperator[variables][{v[n, t], m[n, t]}]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
-	\begin{small}
-		\input{executables/mfg1ctt.ed.tex}
-		\input{executables/mfg1ctt.result.ed.tex}
-	\end{small}
-	We reproduce this result, noting that we have a time-explicit scheme:
-	\begin{EXE}
-		LABEL="mfg2"
-		(*CODE*)
-		variables = Association[
-		"indVars" -> {n},
-		"depVars" -> {v, m},
-		"rhs" -> { 
-			v[n] -(v[n + 1]^2/2 - m[n + 1]^2/2) + (v[n]^2/2 - 
-			m[n]^2/2), 
-			m[n] + (m[n + 1]*v[n + 1]) - (m[n]*v[n])}
-		];
-		DiscreteConservedQOperator[variables][{v[n], m[n]}]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
-	\begin{small}
-		\input{executables/mfg2ctt.ed.tex}
-		\input{executables/mfg2ctt.result.ed.tex}
+
+	\begin{small}
+
+	\end{small}
+	We reproduce this result by noticing, that the scheme is in time-explicit form:
+
+	\begin{small}
+
 	\end{small}
 	Those are the only conserved quantities for this scheme, that are polynomials up to degree 4 in $v$ and $m$.
-	\begin{EXE}
-		LABEL="mfg3"
-		(*CODE*)
-		variables = Association[
-		"indVars" -> {n, t},
-		"depVars" -> {v, m},
-		"VarDOperator"->PartialDVarDOperator,
-		"timeVars"->{t},
-		"eqRhs" -> { 
-			(v[n + 1, t]^2/2 - m[n + 1, t]^2/2) - (v[n, t]^2/2 - 
-			m[n, t]^2/2), 
-			(m[n + 1, t]*v[n + 1, t]) - (m[n, t]*v[n, t])}
-		];
-		FindConservedQuantityBasisOperator[variables][Association["degree"->4,"generators"->{v[n,t],m[n,t]}]]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
-	\begin{small}
-		\input{executables/mfg3ctt.ed.tex}
-		\input{executables/mfg3ctt.result.ed.tex}
+
+	\begin{small}
+
 	\end{small}
@@
 	\begin{gather*}
-		-v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0,\quad
+		-v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0\\
 		m_t-(vm)_x=0.
 	\end{gather*}
-	This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly in time for $m$ and implicitly for $v$.
+	This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly (forward) in time for $m$ and implicitly (backward) for $v$.
 Our code can handle this setting, if we specify an order for the elimination of the variables, telling the code to use an explicit monomial order for $m$ and an implicit one for $v$:
-	\begin{EXE}
-		LABEL="mfg4"
-		(*CODE*)
-		variables = Association[
-		"indVars" -> {n, t},
-		"depVars" -> {m, v},
-		"elimOrder" -> "explicitimplicit",
-		"scheme" -> {-(v[n, t + 1] - 
-			v[n, t]) + (v[n + 1, t+1]^2/2 - m[n + 1, t+1]^2/2) - (v[n, t+1]^2/2 - 
-			m[n, t+1]^2/2), 
-			m[n, t + 1] - 
-			m[n, t] - (m[n + 1, t]*v[n + 1, t]) + (m[n, t]*v[n, t])}
-		];
-		DiscreteConservedQOperator[variables][{v[n, t], m[n, t]}]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
-	\begin{small}
-		\input{executables/mfg4ctt.ed.tex}
-		\input{executables/mfg4ctt.result.ed.tex}
+
+	\begin{small}
+
 	\end{small}
@@
 \section{Possible extensions and concluding remarks}
-\subsection{ Polynomial treatment of non-polynomial expressions}
-It may be possible to extend our methods to non-polynomial PDEs and schemes. In this case, non-polynomial expressions can be handled by writing them in polynomial form.\\
-{\bf Translate the scheme more accurately.}
+\subsection*{Polynomial treatment of non-polynomial expressions}
+It may be possible to extend our methods to non-polynomial PDEs and schemes.
+In this case, non-polynomial expressions can be handled by writing them in polynomial form.
+For example, treat the expression $2\cos(u(n+e^1))e^{u(n+e^2)}+\frac{1}{u(n+e^3)}$
+as the polynomial $2x_1x_2+x_3$.
+\subsection*{Translate the scheme more accurately}
 In our elimination procedure, we work with a specific polynomial ideal that is a subset of the difference ideal generated by the scheme.
 However, it could be necessary to translate the scheme more than we did in our algorithm to get cancellation.
-Therefore, flexible methods to determine the translations of the scheme are desirable.\\
-\subsection{Experimental check of conservation}
-An experimental approach, that gives a hint, if a quantity is conserved, and that can refute conservation, is the following:
-We fix $d$ and $N$.
-Then, we generate experimental initial data by randomly sampling $N\cdot d\cdot m$ values, with every value corresponding to an instance of $u_j(n,t)$ for $n\in\Ii,1\leq j\leq m$.
-Afterward, we calculate $u_j(n,t+1)$ for all $n\in\Ii,1\leq j\leq m$, using the scheme.
-With this values, we explicitly calculate the value of the discrete time derivative.
-If the discrete time derivative is zero for several different samples of initial data, we might have conservation on the discrete level.
-Otherwise, if we find initial data such that the discrete time derivative is not zero, we can refute conservation.\\
-\subsection{Discrete conserved quantities that are not conserved by the PDE}
+Therefore, flexible methods to determine the translations of the scheme are desirable.
+\subsection*{Discrete conserved quantities that are not conserved by the PDE}
 It may arise the question if we can find conserved quantities, that are not preserved by the original PDE.
 Gerdt et al.
-\cite{gerdt10,gerdt12} define the notion of s-consistency, which guarantees that all discrete quantities are also preserved in the continuous setting.
-One possible extension of the code would be to check if s-consistency holds for the translated scheme.\\
-\subsection{Concluding remarks}
+(\cite{gerdt10},\cite{gerdt12}) define the notion of s-consistency, which guarantees that all discrete quantities are also preserved in the continuous setting.
+One possible extension of our code would be to check automatically if s-consistency holds for the translated scheme.
+\subsection*{Concluding remarks}
 We present an algorithm for checking the conservation of a quantity under a finite-difference scheme.
@@
 
-%\citestyle{acmnumeric}
 \bibliographystyle{ACM-Reference-Format}

### 2023-05-02 Papers/Paper ACM/tmp/CPpaper.tex
Friedmann paper

--- 2023-05-02 Papers/Paper ACM/tmp/CLpaper.tex
@@
 	In our {\sc Mathematica} implementation, Example \ref{dpvd1} is computed by
-
+	\begin{EXE}
+		LABEL="dpvd1"
+		(*CODE*)
+		variables = Association["indVars"->{n,t},"depVars"->{u},"timeVars"->{t}]
+		expression = u[n,t+1] - u[n,t]
+		PartialDVarDOperator[variables][expression]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}	
 	\begin{small}
@@
 	We verify the result from Example \ref{timeexplicitheateqalgo}, using our implementation in {\sc Mathematica} of Algorithm \ref{alg: discreteconservedqexplicit}.
-
+	\begin{EXE}
+		LABEL="rhs1"
+		(*CODE*)
+		variables = Association["indVars"->{n},"depVars"->{u},"rhs"->{u[n]+(u[n+1]-2*u[n]+u[n-1])}]
+		expression = u[n]
+		DiscreteConservedQOperator[variables][expression]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 		The discrete time derivative for $\sum_n u(n,t)$ is
+		\begin{EXE}
+			LABEL="heateq2"
+			(*CODE*)
+			variables = <|"indVars"->{n,t},"depVars"->{u}|>
+			expression = u[n,t]
+			TimeDifferenceOperator[variables][expression]
+			(*CODE*)
+			PUNCT={,.}
+		\end{EXE}
 
@@
 		We verify our result from Example \ref{heateqalgoimplicit}, using {\sc DiscreteConservedQ}, which detects automatically if we are in the time-explicit or the general setting.
-
+		\begin{EXE}
+		LABEL="heateq10"
+		(*CODE*)
+		variables = Association["indVars"->{n,t},"depVars"->{u},"scheme"->{u[n,t+1]-u[n,t]-(u[n+1,t]-2*u[n,t]+u[n-1,t])}]
+		expression = u[n,t]
+		DiscreteConservedQOperator[variables][expression]
+		(*CODE*)
+		PUNCT={,.}
+		\end{EXE}
 		\begin{small}
@@
 	We find a basis for conserved quantities that are generated by $u(n,t)$ and $n$ and that have at most degree 3, admitted by the scheme for the heat equation from Example \ref{heatex1}.
-
+	\begin{EXE}
+		LABEL="heateq11"
+		(*CODE*)
+		variables = <|"indVars"->{n},"depVars"->{u},"rhs"->{u[n]+u[n+1]-2*u[n]+u[n-1]}|>
+		FindDiscreteConservedQuantityBasisOperator[variables][<|"degree"->3,"generators"->{u[n],n}|>]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 	We check for conservation of mass using a forward-difference discretization:
-
+	\begin{EXE}
+		LABEL="burgers1"
+		(*CODE*)
+		variables = Association["indVars"->{n,t},"depVars"->{u},"scheme"->{u[n,t+1]-u[n,t]-u[n,t]*(u[n+1,t]-u[n,t])}];
+		expression = u[n,t];
+		DiscreteConservedQOperator[variables][expression]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 	We discretize $u_x$ using a three-point stencil with a parameter $a$.
-
+	\begin{EXE}
+		LABEL="burgers2"
+		(*CODE*)
+		variables = Association["indVars"->{n,t}, "depVars"->{u}, "pars"->{a}, "scheme"->{u[n,t+1]-u[n,t]-u[n,t]*(a*(u[n+1,t]-u[n,t])+(1-a)*(u[n,t]-u[n-1,t]))}];
+		expression = u[n,t];
+		DiscreteConservedQOperator[variables][expression]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 	We consider the scheme from the previous example but in the time-implicit version.
-
+	\begin{EXE}
+		LABEL="burgers3"
+		(*CODE*)
+		variables = Association["indVars"->{n,t}, "depVars"->{u}, "pars"->{a},"elimOrder"->"implicit", "scheme"->{u[n,t+1]-u[n,t]-u[n,t+1]*(a*(u[n+1,t+1]-u[n,t+1])+(1-a)*(u[n,t+1]-u[n-1,t+1]))}]
+		DiscreteConservedQOperator[variables][u[n,t]]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 We check with our code if the scheme admits the same preserved quantities as the continuous system:
-
+	\begin{EXE}
+		LABEL="mfg1"
+		(*CODE*)
+		variables = Association[
+		"indVars" -> {n, t},
+		"depVars" -> {v, m},
+		"display result" -> True,
+		"scheme" -> {v[n, t + 1] - 
+			v[n, 
+			t] + (v[n + 1, t]^2/2 - m[n + 1, t]^2/2) - (v[n, t]^2/2 - 
+			m[n, t]^2/2), 
+			m[n, t + 1] - 
+			m[n, t] - (m[n + 1, t]*v[n + 1, t]) + (m[n, t]*v[n, t])}
+		];
+		DiscreteConservedQOperator[variables][{v[n, t], m[n, t]}]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 	We reproduce this result by noticing, that the scheme is in time-explicit form:
-
+	\begin{EXE}
+		LABEL="mfg2"
+		(*CODE*)
+		variables = Association[
+		"indVars" -> {n},
+		"depVars" -> {v, m},
+		"rhs" -> { 
+			v[n] -(v[n + 1]^2/2 - m[n + 1]^2/2) + (v[n]^2/2 - 
+			m[n]^2/2), 
+			m[n] + (m[n + 1]*v[n + 1]) - (m[n]*v[n])}
+		];
+		DiscreteConservedQOperator[variables][{v[n], m[n]}]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 	Those are the only conserved quantities for this scheme, that are polynomials up to degree 4 in $v$ and $m$.
-
+	\begin{EXE}
+		LABEL="mfg3"
+		(*CODE*)
+		variables = Association[
+		"indVars" -> {n, t},
+		"depVars" -> {v, m},
+		"VarDOperator"->PartialDVarDOperator,
+		"timeVars"->{t},
+		"eqRhs" -> { 
+			(v[n + 1, t]^2/2 - m[n + 1, t]^2/2) - (v[n, t]^2/2 - 
+			m[n, t]^2/2), 
+			(m[n + 1, t]*v[n + 1, t]) - (m[n, t]*v[n, t])}
+		];
+		FindConservedQuantityBasisOperator[variables][Association["degree"->4,"generators"->{v[n,t],m[n,t]}]]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 Our code can handle this setting, if we specify an order for the elimination of the variables, telling the code to use an explicit monomial order for $m$ and an implicit one for $v$:
-
+	\begin{EXE}
+		LABEL="mfg4"
+		(*CODE*)
+		variables = Association[
+		"indVars" -> {n, t},
+		"depVars" -> {m, v},
+		"elimOrder" -> "explicitimplicit",
+		"scheme" -> {-(v[n, t + 1] - 
+			v[n, t]) + (v[n + 1, t+1]^2/2 - m[n + 1, t+1]^2/2) - (v[n, t+1]^2/2 - 
+			m[n, t+1]^2/2), 
+			m[n, t + 1] - 
+			m[n, t] - (m[n + 1, t]*v[n + 1, t]) + (m[n, t]*v[n, t])}
+		];
+		DiscreteConservedQOperator[variables][{v[n, t], m[n, t]}]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}

### 2023-05-02 Papers/Paper ACM/tmp/FCpaper.tex
Friedmann paper

--- 2023-05-02 Papers/Paper ACM/tmp/CLpaper.tex
@@
 
-\begin{document}
+\usepackage{color}\begin{document}
 
@@
 They also developed and implemented algorithms for detecting quantities conserved by PDEs and semi-discretized schemes. 
-We revise these techniques in Section \ref{assp}. 
-To generalize those methods to situations where we do not sum over all arguments of the functions involved, we introduce the discrete partial variational derivative (Section \ref{pvard}).
+We revise these techniques in Section \ref{assp}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{assp}\end{verbatim}.}
+. 
+To generalize those methods to situations where we do not sum over all arguments of the functions involved, we introduce the discrete partial variational derivative (Section \ref{pvard}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{pvard}\end{verbatim}.}
+).
 The main contribution of this paper is an algorithm for checking the conservation of a quantity under a numerical scheme.
@@
 Moreover, Gerdt's algorithm may not terminate, as the Gr{\"o}bner basis for the difference ideal may not be finite.
-We overcome this issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) with finite Gr{\"o}bner basis (Section \ref{algodetec}). 
+We overcome this issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) with finite Gr{\"o}bner basis (Section \ref{algodetec}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{algodetec}\end{verbatim}.}
+). 
 Our algorithm works for schemes that are explicit and implicit in time and can treat schemes with several and higher-order time  and space derivatives and with several space dimensions.
@@
 We have implemented this algorithm as part of a package in {\sc Mathematica} \cite{wolfram21}.
-In the examples, we show that our code finds conserved quantities and proper schemes for the time-implicit and time-explicit discretization of the Burgers equation and a system of PDEs arising in the study of mean-field games (Section \ref{eaa}).
+In the examples, we show that our code finds conserved quantities and proper schemes for the time-implicit and time-explicit discretization of the Burgers equation and a system of PDEs arising in the study of mean-field games (Section \ref{eaa}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{eaa}\end{verbatim}.}
+).
 
@@
 Here, subscripts denote indices related to coordinates or tuples while superscripts denote indices related to sequences and families.
-To avoid dealing with boundary terms, we work with periodic functions in $\Zz^d$.
+To avoid dealing with boundary terms, we work with periodic functions in \({\mathbb{Z}}^d\)
+{\color{blue} New variable(s): \(d\), \(Z\),  }
+.
 \begin{definition}
-	Let $N$, $d$ and $m$ be natural numbers. The \emph{discrete torus} is $\Ii:=\{0,\ldots,N-1\}^d\subset \Zz^d$. Define the space
+	Let \(N\)
+{\color{blue} New variable(s): \(N\),  }
+, \(d\) and \(m\)
+{\color{blue} New variable(s): \(m\),  }
+ be natural numbers. The \emph{discrete torus} is \({\mathcal{I}}:=\{0,\ldots,N-1\}^d\subset {\mathbb{Z}}^d\). Define the space
 	\begin{gather*}
-		\mathcal{P}(\Ii,\Rr^m):=\{u\ |\ u:\Ii\rightarrow\Rr^m\},
+		\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m):=\{u\ |\ u:{\mathcal{I}}\rightarrow{\mathbb{R}}^m\},
 	\end{gather*}
-	extended to $\Zz^d$ by periodicity.
+	extended to \({\mathbb{Z}}^d\) by periodicity.
 \end{definition}
 \begin{definition}
-	The space of functionals $\mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$ (not necessarily linear) on $\mathcal{P}(\Ii,\Rr^m)$ is
+	The space of functionals \(\mathcal{F}(\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m))\)
+{\color{blue} New variable(s): \(F\), \(P\), \(R\),  }
+ (not necessarily linear) on \(\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m)\) is
 \begin{gather*}
-	\mathcal{F}(\mathcal{P}(\Ii,\Rr^m)):= \Bigg\{\mathcal{J}:\mathcal{P}(\Ii,\Rr^m)\to \Rr\ \Bigg|\ \mathcal{J}[u]=\sum_{n\in\Ii}F_n[u],\\
+	\mathcal{F}(\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m)):= \Bigg\{\mathcal{J}:\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m)\to {\mathbb{R}}\ \Bigg|\ \mathcal{J}[u]=\sum_{n\in{\mathcal{I}}}F_n[u],\\
 	\text{ where }
 	F_n[u]=G(u(n+e^1),u(n+e^2),\ldots,u(n+e^k))\\
-	\text{ and }\{e^1,\ldots,e^k\}\subset\Zz^d,\ \text{G is smooth}
+	\text{ and }\{e^1,\ldots,e^k\}\subset{\mathbb{Z}}^d,\ \text{G is smooth}
 		\Bigg\}.
@@
 \end{definition}
-In the previous definition, the $e^1,\ldots,e^k$ are not necessarily unit vectors and can vary between functionals.
-Here, we allow for smooth $G$.
-However, later we require $G$ to be a polynomial function.
+In the previous definition, the \(e^1,\ldots,e^k\)
+{\color{blue} New variable(s): \(k\),  }
+ are not necessarily unit vectors and can vary between functionals. Here, we allow for smooth \(G\)
+{\color{blue} New variable(s): \(G\),  }
+. However, later we require \(G\) to be a polynomial function.
 \begin{example}
 	\label{functional1}
-	Let $d=1$, $m=1$ and define
+	Let \(d=1\), \(m=1\) and define
 	\begin{gather*}
-		\mathcal{J}[u]:=\sum_{n\in\Ii}(u(n+1)-u(n))^2.
+		\mathcal{J}[u]:=\sum_{n\in{\mathcal{I}}}(u(n+1)-u(n))^2.
 	\end{gather*}
@@
 To simplify sums, the discrete variational derivative is a useful tool.
-Let $u,v\in\mathcal{P}(\Ii,\Rr^m)$, $\epsilon 
-	\in\Rr$ and $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$. Define
+Let \(u,v\in\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m)\)
+{\color{blue} New variable(s): \(u\), \(v\),  }
+, \(\epsilon 
+	\in{\mathbb{R}}\)
+{\color{blue} New variable(s): \(\epsilon\),  }
+ and \(\mathcal{J}\in\mathcal{F}(\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m))\)
+{\color{blue} New variable(s): \(J\),  }
+. Define
 	\begin{gather*}
@@
 	\end{gather*}
-	In \eqref{periodicity}, we shift the indices due to periodicity of $u$ and $v$,
+	In \eqref{periodicity}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\eqref{periodicity}\end{verbatim}.}
+, we shift the indices due to periodicity of \(u\) and \(v\),
 	\begin{gather}
-	\notag	D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\Ii}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
-	\notag	=\sum_{n\in\Ii}\frac{d}{d\epsilon}G(u(n+e^1)+\epsilon v(n+e^1),\ldots,u(n+e^k)+\epsilon v(n+e^k))\bigg|_{\epsilon=0}\\
-	\notag	=\sum_{n\in\Ii}D_1G(u(n+e^1),\ldots,u(n+e^k))v(n+e^1)\\
+	\notag	D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in{\mathcal{I}}}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
+	\notag	=\sum_{n\in{\mathcal{I}}}\frac{d}{d\epsilon}G(u(n+e^1)+\epsilon v(n+e^1),\ldots,u(n+e^k)+\epsilon v(n+e^k))\bigg|_{\epsilon=0}\\
+	\notag	=\sum_{n\in{\mathcal{I}}}D_1G(u(n+e^1),\ldots,u(n+e^k))v(n+e^1)\\
 	\notag	+\ldots+D_kG(u(n+e^1),\ldots,u(n+e^k))v(n+e^k)\\
-	\label{periodicity}	=\sum_{n\in\Ii}D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))v(n)\\
+	\label{periodicity}	=\sum_{n\in{\mathcal{I}}}D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))v(n)\\
 	\notag	+D_2G(u(n+e^1-e^2),u(n),\ldots,u(n+e^k-e^2))v(n)\\
 	\notag	+\ldots+D_kG(u(n+e^1-e^k),u(n+e^2-e^k),\ldots,u(n))v(n)\\
-	\notag	=\sum_{n\in\Ii}\Big(D_1G(u(n),u(n+e^2-e^1)\ldots,u(n+e^k-e^1))\\
+	\notag	=\sum_{n\in{\mathcal{I}}}\Big(D_1G(u(n),u(n+e^2-e^1)\ldots,u(n+e^k-e^1))\\
 	\notag	+D_2G(u(n+e^1-e^2),u(n),\ldots,u(n+e^k-e^2))\\
 	\notag	+\ldots+D_kG(u(n+e^1-e^k),u(n+e^2-e^k),\ldots,u(n))\Big)v(n)\\
-	\notag	=:\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u]v(n).	
+	\notag	=:\sum_{n\in{\mathcal{I}}}\mathcal{V}(\mathcal{J})[u]v(n).	
 	\end{gather}
 \begin{definition}
-$\mathcal{V}(\mathcal{J})$ is the discrete variational derivative of $\mathcal{J}$.
+\(\mathcal{V}(\mathcal{J})\)
+{\color{blue} New variable(s): \(V\),  }
+ is the discrete variational derivative of \(\mathcal{J}\).
 \end{definition}
@@
 	\begin{gather*}
-		D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\Ii}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
-		=\sum_{n\in\Ii}\frac{d}{d\epsilon}G[u(n+1)+\epsilon v(n+1),u(n)+\epsilon v(n)]\bigg|_{\epsilon=0}\\
-		=\sum_{n\in\Ii}2(u(n+1)-u(n))v(n+1)+(-2)(u(n+1)-u(n))v(n)\\
-		=\sum_{n\in\Ii}2(u(n)-u(n-1))v(n)+(-2)(u(n+1)-u(n))v(n)\\
-		=\sum_{n\in\Ii}\Big(2(u(n)-u(n-1))+(-2)(u(n+1)-u(n))\Big)v(n)\\
-		=\sum_{n\in\Ii}\Big(-2u(n+1)+4u(n)-2u(n-1)\Big)v(n)\\
-		=\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u]v(n)
+		D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in{\mathcal{I}}}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
+		=\sum_{n\in{\mathcal{I}}}\frac{d}{d\epsilon}G[u(n+1)+\epsilon v(n+1),u(n)+\epsilon v(n)]\bigg|_{\epsilon=0}\\
+		=\sum_{n\in{\mathcal{I}}}2(u(n+1)-u(n))v(n+1)+(-2)(u(n+1)-u(n))v(n)\\
+		=\sum_{n\in{\mathcal{I}}}2(u(n)-u(n-1))v(n)+(-2)(u(n+1)-u(n))v(n)\\
+		=\sum_{n\in{\mathcal{I}}}\Big(2(u(n)-u(n-1))+(-2)(u(n+1)-u(n))\Big)v(n)\\
+		=\sum_{n\in{\mathcal{I}}}\Big(-2u(n+1)+4u(n)-2u(n-1)\Big)v(n)\\
+		=\sum_{n\in{\mathcal{I}}}\mathcal{V}(\mathcal{J})[u]v(n)
 	\end{gather*}
-	and hence $\mathcal{V}(\mathcal{J})[u]=-2u(n+1)+4u(n)-2u(n-1)$.
+	and hence \(\mathcal{V}(\mathcal{J})[u]=-2u(n+1)+4u(n)-2u(n-1)\).
 \end{example}
@@
 \begin{teo}\label{theo1}
-	Let $\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$.
-If $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[v]=0$ for all $v\in\mathcal{P}(\Ii,\Rr^m)$ and if there exists $u_0\in\mathcal{P}(\Ii,\Rr^m)$ such that $\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]$, then
+	Let \(\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m))\).
+If \(\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[v]=0\) for all \(v\in\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m)\) and if there exists \(u_0\in\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m)\)
+{\color{blue} New variable(s): \(u_0\),  }
+ such that \(\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]\), then
 	\begin{equation*}
@@
 	\end{equation*}
-	for all $u\in\mathcal{P}(\Ii,\Rr^m)$.\\
+	for all \(u\in\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m)\).\\
 	Conversely,
@@
 	\end{equation*}
-	for all $u\in\mathcal{P}(\Ii,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
+	for all \(u\in\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m)\), if \(\mathcal{J}[u]=\tilde{\mathcal{J}}[u]\).
 \end{teo}
@@
 	\end{gather*}
-	and hence $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
+	and hence \(\mathcal{J}[u]=\tilde{\mathcal{J}}[u]\).
 \end{proof}
@@
 \begin{example}
-	Let $d=m=1$ and consider the functionals
+	Let \(d=m=1\) and consider the functionals
 	\begin{gather*}
-		\mathcal{J}[u]:=\sum_{n\in\Ii}u(n)u(n+1),\\
-		\tilde{\mathcal{J}}[u]:=\sum_{n\in\Ii}u(n-2)u(n-1).
+		\mathcal{J}[u]:=\sum_{n\in{\mathcal{I}}}u(n)u(n+1),\\
+		\tilde{\mathcal{J}}[u]:=\sum_{n\in{\mathcal{I}}}u(n-2)u(n-1).
 	\end{gather*}
-	It is clear, that $\mathcal{J}$ and $\tilde{\mathcal{J}}$ represent the same quantity (by shifting $n$). We confirm this, using the discrete variational derivative:
+	It is clear, that \(\mathcal{J}\) and \(\tilde{\mathcal{J}}\) represent the same quantity (by shifting \(n\)
+{\color{blue} New variable(s): \(n\),  }
+). We confirm this, using the discrete variational derivative:
 	\begin{gather*}
 		D(\mathcal{J}-\tilde{\mathcal{J}})[u](v)=\\
-		=\sum_{n\in\Ii}(u(n+1)+u(n-1)-u(n+1)-u(n-1))v(n)=0.
+		=\sum_{n\in{\mathcal{I}}}(u(n+1)+u(n-1)-u(n+1)-u(n-1))v(n)=0.
 	\end{gather*}
-	Hence, $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})=0$ and $\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]$. Therefore, both functionals sum to the same quantity.
+	Hence, \(\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})=0\) and \(\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]\). Therefore, both functionals sum to the same quantity.
 \end{example}
@@
 \begin{example}
-	Let $d=2$. It may not be obvious, that
+	Let \(d=2\). It may not be obvious, that
 	\begin{gather*}
@@
 	\end{gather*}
-	We confirm this, by computing the discrete variational derivative and noticing that the expression is $0$ for $u_0=0$.
+	We confirm this, by computing the discrete variational derivative and noticing that the expression is \(0\)
+{\color{blue} New variable(s): \(\text{Null}\),  }
+ for \(u_0=0\).
 \end{example}
@@
 \label{pvard}
-Here, we generalize the discrete variational derivative for situations where we keep one (or several) of the arguments of $u$ constant.
-Let $n=(n_1,\ldots,n_d)$ be the variables for the functions in $\mathcal{P}(\Ii,\Rr^m)$.
-We call $n$ the space variables.
-Let $l\in\Nn$ and call $t=(t_1,\ldots,t_l)\in\Nn^l$ the time variables.
-Later in this paper, we only consider the case $l=1$.
+Here, we generalize the discrete variational derivative for situations where we keep one (or several) of the arguments of \(u\) constant.
+Let \(n=(n_1,\ldots,n_d)\)
+{\color{blue} New variable(s): \(n_1\), \(n_d\),  }
+ be the variables for the functions in \(\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m)\). We call \(n\) the space variables. Let \(l\in{\mathbb{N}}\)
+{\color{blue} New variable(s): \(l\),  }
+ and call \(t=(t_1,\ldots,t_l)\in{\mathbb{N}}^l\)
+{\color{blue} New variable(s): \(t\), \(t_1\), \(t_l\),  }
+ the time variables. Later in this paper, we only consider the case \(l=1\).
 \begin{definition}
-	Let $\ \mathcal{P}(\Ii\times\Nn^l,\Rr^m):=\{u\ |\ u(\cdot,t)\in\mathcal{P}(\Ii,\Rr^m)\ \forall t\in\Nn^l \}$, extended to $\Zz^d\times\Nn^l$ by periodicity in the space variables and let $\mathcal{D}(\Nn^l):=\{u\ |\ u:\Nn^l\rightarrow\Rr\}$.
+	Let {\color{red} \(\ \mathcal{P}({\mathcal{I}}\times{\mathbb{N}}^l,{\mathbb{R}}^m):=\{u\ |\ u(\cdot,t)\in\mathcal{P}({\mathcal{I}},{\mathbb{R}}^m)\ \forall t\in{\mathbb{N}}^l \}\)
+
+  Equation errors and warnings:
+
+ Unbalanced $|\hdots|$}
+
+, extended to \({\mathbb{Z}}^d\times{\mathbb{N}}^l\) by periodicity in the space variables and let {\color{red} \(\mathcal{D}({\mathbb{N}}^l):=\{u\ |\ u:{\mathbb{N}}^l\rightarrow{\mathbb{R}}\}\)
+
+  Equation errors and warnings:
+
+ Unbalanced $|\hdots|$}
+
+{\color{blue} New variable(s): \(D\),  }
+.
 \end{definition}
@@
 	\label{quantityfunc2}
-	The space of functions $\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$ is
+	The space of functions \(\mathcal{F}(\mathcal{P}({\mathcal{I}}\times{\mathbb{N}}^l,{\mathbb{R}}^m),\mathcal{D}({\mathbb{N}}^l))\) is
 	\begin{gather*}
-		\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l)):=\\	
-		\Bigg\{\mathcal{J}:\mathcal{P}(\Ii\times\Nn^l)\rightarrow\mathcal{D}(\Nn^l) \Bigg|\ \mathcal{J}[u]=\sum_{n}F_{(n,t)}[u],\\
+		\mathcal{F}(\mathcal{P}({\mathcal{I}}\times{\mathbb{N}}^l,{\mathbb{R}}^m),\mathcal{D}({\mathbb{N}}^l)):=\\	
+		\Bigg\{\mathcal{J}:\mathcal{P}({\mathcal{I}}\times{\mathbb{N}}^l)\rightarrow\mathcal{D}({\mathbb{N}}^l) \Bigg|\ \mathcal{J}[u]=\sum_{n}F_{(n,t)}[u],\\
 		\text{where } F_{(n,t)}[u]=G(u((n,t)+e^1),u((n,t)+e^2),\ldots,u((n,t)+e^k)),\\
 		\text{and }
-		\{e^1,\ldots,e^k\}\subset\Zz^{d+l},\text{G is a polynomial in k variables}\Bigg\}.
+		\{e^1,\ldots,e^k\}\subset{\mathbb{Z}}^{d+l},\text{G is a polynomial in k variables}\Bigg\}.
 	\end{gather*}
-Here, $(n,t):=(n_1,\ldots,n_d,t_1,\ldots,t_l)\in\Ii\times\Nn^l$.
+Here, \((n,t):=(n_1,\ldots,n_d,t_1,\ldots,t_l)\in{\mathcal{I}}\times{\mathbb{N}}^l\).
 \end{definition}
-Throughout this paper, all quantities, whose conservation in time we check, are functions $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$. 
-Define $\tilde{n}:=(n_1,...,n_d,s_1,...,s_l)\in\Ii\times\Nn^l$. Using the Kronecker delta $\delta$, we rewrite
+Throughout this paper, all quantities, whose conservation in time we check, are functions \(\mathcal{J}\in\mathcal{F}(\mathcal{P}({\mathcal{I}}\times{\mathbb{N}}^l,{\mathbb{R}}^m),\mathcal{D}({\mathbb{N}}^l))\).
+Define \(\tilde{n}:=(n_1,...,n_d,s_1,...,s_l)\in{\mathcal{I}}\times{\mathbb{N}}^l\)
+{\color{blue} New variable(s): \(s_1\), \(s_l\),  }
+. Using the Kronecker delta \(\delta\)
+{\color{blue} New variable(s): \(\delta\),  }
+, we rewrite
 \begin{gather*}
 	\mathcal{J}[u](t_1,\ldots,t_{l})
-	=\sum_{\tilde{n}\in\Ii\times\Nn^l}F_{\tilde{n}}[u]\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l).
+	=\sum_{\tilde{n}\in{\mathcal{I}}\times{\mathbb{N}}^l}F_{\tilde{n}}[u]\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l).
 \end{gather*}
-	Let $u,v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ and compute
+	Let \(u,v\in\mathcal{P}({\mathcal{I}}\times{\mathbb{N}}^l,{\mathbb{R}}^m)\) and compute
 	\footnotesize
@@
 	\begin{definition}
-	We define $\mathcal{V}^l(\mathcal{J})$ as the discrete partial variational derivative of $\mathcal{J}$.
-	Note that $\mathcal{V}^0:=\mathcal{V}$.
+	We define \(\mathcal{V}^l(\mathcal{J})\) as the discrete partial variational derivative of \(\mathcal{J}\).
+	Note that \(\mathcal{V}^0:=\mathcal{V}\).
 \end{definition}
@@
 	\end{gather*}
-	equals $\delta_{t+1}(s)-\delta_t(s)$, because
+	equals \(\delta_{t+1}(s)-\delta_t(s)\)
+{\color{blue} New variable(s): \(s\), \(\delta _t\),  }
+, because
 	\small
@@
 \begin{teo}\label{theo2}
-	Let $\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$.
-If $\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[v]=0$ for all $v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ and if there exists $u_0\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ such that $\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]$, then
+	Let \(\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}({\mathcal{I}}\times{\mathbb{N}}^l,{\mathbb{R}}^m),\mathcal{D}({\mathbb{N}}^l))\).
+If \(\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[v]=0\) for all \(v\in\mathcal{P}({\mathcal{I}}\times{\mathbb{N}}^l,{\mathbb{R}}^m)\) and if there exists \(u_0\in\mathcal{P}({\mathcal{I}}\times{\mathbb{N}}^l,{\mathbb{R}}^m)\) such that \(\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]\), then
 	\begin{equation*}
@@
 	\end{equation*}
-	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$.\\
+	for all \(u\in\mathcal{P}({\mathcal{I}}\times{\mathbb{N}}^l,{\mathbb{R}}^m)\).\\
 	Conversely,
@@
 	\end{equation*}
-	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
+	for all \(u\in\mathcal{P}({\mathcal{I}}\times{\mathbb{N}}^l,{\mathbb{R}}^m)\), if \(\mathcal{J}[u]=\tilde{\mathcal{J}}[u]\).
 \end{teo}
@@
 In this section, we define numerical schemes formally and introduce the tools from difference algebra, that we need for our algorithm.
-During the remainder of this paper, we assume that there is only one time variable $t$ ($l=1$).
+During the remainder of this paper, we assume that there is only one time variable \(t\) (\(l=1\)).
 \begin{definition}
@@
 	\begin{gather*}
-		\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr)):=\\
-		\Bigg\{H:\mathcal{P}(\Ii\times\Nn,\Rr^m)\rightarrow\mathcal{P}(\Ii\times\Nn,\Rr)\ \Bigg|\\
+		\mathcal{S}(\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}^m),\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}})):=\\
+		\Bigg\{H:\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}^m)\rightarrow\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}})\ \Bigg|\\
 			H[u](n,t)=H(u(n+\tilde{e}^1,t+\tilde{l}^1),\ldots,u(n+\tilde{e}^r,t+\tilde{l}^r)),\\
 			\text{where }
-		\{\tilde{e}^1,\ldots,\tilde{e}^r\}\subset\Zz^d,\ \{\tilde{l}^1,\ldots,\tilde{l}^r\}\subset\Zz
+		\{\tilde{e}^1,\ldots,\tilde{e}^r\}\subset{\mathbb{Z}}^d,\ \{\tilde{l}^1,\ldots,\tilde{l}^r\}\subset{\mathbb{Z}}
 		\Bigg\}.
@@
 	\begin{gather*}
-		\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))
+		\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}^m),\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}))
 	\end{gather*}
-	that represent the equations $\{H^1[u]=0,\ldots,H^i[u]=0\}$ for $u\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$, holding pointwise for all points in $\Ii\times\Nn$.
+	that represent the equations \(\{H^1[u]=0,\ldots,H^i[u]=0\}\)
+{\color{blue} New variable(s): \(H\), \(i\),  }
+ for \(u\in\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}^m)\), holding pointwise for all points in \({\mathcal{I}}\times{\mathbb{N}}\).
 \end{definition}
-If a scheme contains a single function $H$, we call $H$ the scheme. Sometimes, we also call the expression $H[u]$ the scheme.
-In this work, we only consider finite-difference schemes with fixed step size $h=1$, but one can generalize our results and algorithms to other step sizes.
+If a scheme contains a single function \(H\), we call \(H\) the scheme. Sometimes, we also call the expression \(H[u]\) the scheme.
+In this work, we only consider finite-difference schemes with fixed step size \(h=1\)
+{\color{blue} New variable(s): \(h\),  }
+, but one can generalize our results and algorithms to other step sizes.
 \begin{definition}
-	A scheme $\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ given by
+	A scheme \(\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}^m),\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}))\)
+{\color{blue} New variable(s): \(S\),  }
+ given by
 	\begin{equation*}
@@
 	\end{equation*}
+{\color{blue} New variable(s): \(e\), \(r\),  }
+
 	is in time-explicit form if we can rewrite the previous equation (eventually translating the scheme) as
@@
 	\end{equation*}
-	for $1\leq k\leq i$ and $1\leq j\leq m$
-	with $\tilde{H}^k\in\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$. We call $\{\tilde{H}^1,\ldots,\tilde{H}^i\}$ the right-hand side of $\{H^1,\ldots,H^i\}$.
+{\color{blue} New variable(s): \(u_j\),  }
+
+	for \(1\leq k\leq i\) and \(1\leq j\leq m\)
+{\color{blue} New variable(s): \(j\),  }
+
+	with \(\tilde{H}^k\in\mathcal{S}(\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}^m),\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}))\).
+We call \(\{\tilde{H}^1,\ldots,\tilde{H}^i\}\) the right-hand side of \(\{H^1,\ldots,H^i\}\).
 \end{definition}
@@
 \begin{definition}
-	Let $u=(u_1,\dots,u_m)\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$. The shift in the $i$-th coordinate for $1\leq i\leq d+1$ by $k\in\Zz$ is
+	Let \(u=(u_1,\dots,u_m)\in\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}^m)\)
+{\color{blue} New variable(s): \(u_1\), \(u_m\),  }
+. The shift in the \(i\)-th coordinate for \(1\leq i\leq d+1\) by \(k\in{\mathbb{Z}}\) is
 	\begin{gather*}
@@
 	\end{gather*}
-	understanding that $\sigma_{d+1}^k$ shifts the time variable.
+	understanding that \(\sigma_{d+1}^k\)
+{\color{blue} New variable(s): \(\sigma\),  }
+ shifts the time variable.
 \end{definition}
 \begin{definition}
-	The set of all possible shifts $\Theta$ is
+	The set of all possible shifts \(\Theta\)
+{\color{blue} New variable(s): \(\Theta\),  }
+ is
 	\begin{gather*}
-		\Theta:=\{\sigma_1^{k_1}\circ\cdots\circ\sigma_{d+1}^{k_{d+1}}\ |\ k_1,\ldots,k_{d+1}\in\Zz\}.
+		\Theta:=\{\sigma_1^{k_1}\circ\cdots\circ\sigma_{d+1}^{k_{d+1}}\ |\ k_1,\ldots,k_{d+1}\in{\mathbb{Z}}\}.
 	\end{gather*}
@@
 	\label{diffideal}
-	Let $\ \mathcal{K}$ be the field generated by $\ \Rr$ and the variables $\{n_1,\ldots,n_d,t\}$.
-Let $\ \tilde{\mathcal{R}}$ be the polynomial ring over the field $\ \mathcal{K}$ and the variables $\theta\circ u_j(n,t)$ for $\theta\in\Theta$ and $1\leq j\leq m$.
-	A set $\tilde{I}\subseteq\tilde{\mathcal{R}}$ is a difference ideal if
+	Let \(\ \mathcal{K}\)
+{\color{blue} New variable(s): \(K\),  }
+ be the field generated by \(\ {\mathbb{R}}\) and the variables \(\{n_1,\ldots,n_d,t\}\).
+Let \(\ \tilde{\mathcal{R}}\) be the polynomial ring over the field \(\ \mathcal{K}\) and the variables \(\theta\circ u_j(n,t)\)
+{\color{blue} New variable(s): \(\theta\),  }
+ for \(\theta\in\Theta\) and \(1\leq j\leq m\).
+	A set \(\tilde{I}\subseteq\tilde{\mathcal{R}}\) is a difference ideal if
 	\begin{itemize}
-		\item $p_1,p_2\in\tilde{I}$ implies $p_1+p_2\in\tilde{I}$,
-		\item $p_1\in\tilde{I},p_2\in\tilde{\mathcal{R}}$ implies $p_1\cdot p_2\in\tilde{I}$,
-		\item $p_1\in\tilde{I},\theta\in\Theta$ implies $\theta\circ p_1\in\tilde{I}$.
+		\item \(p_1,p_2\in\tilde{I}\)
+{\color{blue} New variable(s): \(p_1\), \(p_2\),  }
+ implies \(p_1+p_2\in\tilde{I}\),
+		\item \(p_1\in\tilde{I},p_2\in\tilde{\mathcal{R}}\) implies \(p_1\cdot p_2\in\tilde{I}\),
+		\item \(p_1\in\tilde{I},\theta\in\Theta\) implies \(\theta\circ p_1\in\tilde{I}\).
 	\end{itemize}
@@
 \begin{definition}
-	$\langle H^1,\ldots,H^i\rangle$ is the smallest difference ideal containing the scheme $\{H^1,\ldots,H^i\}$.
+	\(\langle H^1,\ldots,H^i\rangle\) is the smallest difference ideal containing the scheme \(\{H^1,\ldots,H^i\}\).
 \end{definition}
-A solution of a numerical scheme is a function $u$, that makes all translations of the scheme vanish. Hence, every element of the difference ideal generated by the scheme vanishes under $u$.
-In particular, Algorithm \ref{alg: discreteconservedqimplicit} seeks to determine if the discrete time derivative $\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$ belongs to the ideal $\langle H^1,\ldots,H^i\rangle$.
-To examine, if $\mathcal{T}\in\langle H^1,\ldots,H^i\rangle$, Gerdt proposes the notion of a standard (Gr{\"o}bner) basis for the ideal $\langle H^1,\ldots,H^i\rangle$.
+A solution of a numerical scheme is a function \(u\), that makes all translations of the scheme vanish. Hence, every element of the difference ideal generated by the scheme vanishes under \(u\).
+In particular, Algorithm \ref{alg: discreteconservedqimplicit}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{alg: discreteconservedqimplicit}\end{verbatim}.}
+ seeks to determine if the discrete time derivative \(\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)\)
+{\color{blue} New variable(s): \(T\),  }
+ belongs to the ideal \(\langle H^1,\ldots,H^i\rangle\).
+To examine, if \(\mathcal{T}\in\langle H^1,\ldots,H^i\rangle\), Gerdt proposes the notion of a standard (Gr{\"o}bner) basis for the ideal \(\langle H^1,\ldots,H^i\rangle\).
 This standard basis may not be finite.
 We overcome this problem by considering polynomial instead of difference algebra and using a smaller polynomial ideal, that is contained in the difference ideal and admits a finite standard basis.\\
-However, there may be functions $\mathcal{J}$, whose sums add to zero, even though they may fail to belong to the difference ideal.
+However, there may be functions \(\mathcal{J}\), whose sums add to zero, even though they may fail to belong to the difference ideal.
 Hence, we combine the discrete partial variational derivative with polynomial ideals.
@@
 \label{time-exp}
-In this section, we present Algorithm \ref{alg: discreteconservedqexplicit}  for checking the conservation of a quantity under the right-hand side of a scheme in time-explicit form.
-Our algorithm for general schemes can be found in the subsequent section.
+In this section, we present Algorithm \ref{alg: discreteconservedqexplicit}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{alg: discreteconservedqexplicit}\end{verbatim}.}
+  for checking the conservation of a quantity under the right-hand side of a scheme in time-explicit form. Our algorithm for general schemes can be found in the subsequent section.
 
-Note, that the result of step 2 exists only at time $t$ because all instances of $t+1$ have been replaced. Thus, we do not need any computations of difference ideals or Gr{\"o}bner basis.
+Note, that the result of step 2 exists only at time \(t\) because all instances of \(t+1\) have been replaced. Thus, we do not need any computations of difference ideals or Gr{\"o}bner basis.
 \begin{example}
 	\label{timeexplicitheateqalgo}
-	We check for conservation of $\sum_nu(n,t)$ under the scheme from Example \ref{timeexplicitheateq}.
+	We check for conservation of \(\sum_nu(n,t)\) under the scheme from Example \ref{timeexplicitheateq}.
 	The discrete time derivative (step 1) is
@@
 	\end{gather*}
-	Replacing $u(n,t+1)$ by the right-hand side of the scheme (step 2) gives
+	Replacing \(u(n,t+1)\) by the right-hand side of the scheme (step 2) gives
 	\begin{gather*}
@@
 	\label{heatex1}
-	We verify the result from Example \ref{timeexplicitheateqalgo}, using our implementation in {\sc Mathematica} of Algorithm \ref{alg: discreteconservedqexplicit}.
+	We verify the result from Example \ref{timeexplicitheateqalgo}, using our implementation in {\sc Mathematica} of Algorithm \ref{alg: discreteconservedqexplicit}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{alg: discreteconservedqexplicit}\end{verbatim}.}
+.
 
@@
 \section{The general case}
-In this section, we present Algorithm \ref{alg: discreteconservedqimplicit} that deals with general, not necessarily time-explicit, schemes.
-We explain its steps in detail below and demonstrate the algorithm in Example \ref{heateqalgoimplicit}.
+In this section, we present Algorithm \ref{alg: discreteconservedqimplicit}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{alg: discreteconservedqimplicit}\end{verbatim}.}
+ that deals with general, not necessarily time-explicit, schemes. We explain its steps in detail below and demonstrate the algorithm in Example \ref{heateqalgoimplicit}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{heateqalgoimplicit}\end{verbatim}.}
+.
 
 \subsection*{Step 1: Build the discrete time derivative}
-We build the discrete time derivative by subtracting $\mathcal{J}[u](t)$ from $\mathcal{J}[u](t+1)$. This task is done in our code by the {\sc TimeDifference}.
+We build the discrete time derivative by subtracting \(\mathcal{J}[u](t)\) from \(\mathcal{J}[u](t+1)\). This task is done in our code by the {\sc TimeDifference}.
 	\begin{example}
-		The discrete time derivative for $\sum_n u(n,t)$ is
+		The discrete time derivative for \(\sum_n u(n,t)\) is
 
@@
 \subsection*{Step 2: Translate the scheme}
-To make the step from difference to polynomial algebra to guarantee termination of the algorithm, we use a standard idea in symbolic computations and consider every instance of $u_j(n+e,t+l)$ as a variable of a multivariate polynomial.
-Hence, we compute all translations of the scheme, such that all variables in the polynomial associated with $\mathcal{T}$ appear in the polynomials associated with the translated scheme.
+To make the step from difference to polynomial algebra to guarantee termination of the algorithm, we use a standard idea in symbolic computations and consider every instance of \(u_j(n+e,t+l)\) as a variable of a multivariate polynomial.
+Hence, we compute all translations of the scheme, such that all variables in the polynomial associated with \(\mathcal{T}\) appear in the polynomials associated with the translated scheme.
 \begin{example}
-	The algorithm treats the discrete time derivative $u(n+1,t+1)-u(n+1,t)$
-	as the polynomial equation $x_1-x_2$.
+	The algorithm treats the discrete time derivative \(u(n+1,t+1)-u(n+1,t)\)
+	as the polynomial equation \(x_1-x_2\)
+{\color{blue} New variable(s): \(x_1\), \(x_2\),  }
+.
 \end{example}
 	\begin{definition}
-		Let $u=(u_1,\ldots,u_m)\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$ and
+		Let \(u=(u_1,\ldots,u_m)\in\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}^m)\) and
 		\begin{gather*}
 			F:
-				\mathcal{P}(\Ii\times\Nn,\Rr^m)\rightarrow\mathcal{P}(\Ii\times\Nn,\Rr),\\
+				\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}^m)\rightarrow\mathcal{P}({\mathcal{I}}\times{\mathbb{N}},{\mathbb{R}}),\\
 				F[u]=F(u_1((n,t)+x^1),\ldots,u_1((n,t)+x^{k_1}),\\
@@
 	\normalsize
-	with $\{x^1,\ldots,x^{k_1},\ldots,x^{k_m}\}\subset\Zz^{d+1}$.
-	The \emph{stencil} of $F$ is the $m$-tuple of sets of vectors
+	with \(\{x^1,\ldots,x^{k_1},\ldots,x^{k_m}\}\subset{\mathbb{Z}}^{d+1}\)
+{\color{blue} New variable(s): \(x\), \(k_1\), \(k_m\),  }
+.
+	The \emph{stencil} of \(F\) is the \(m\)-tuple of sets of vectors
 	\begin{gather*}
@@
 	\end{gather*}
-	The \emph{range} of the stencil of $F$ is
+	The \emph{range} of the stencil of \(F\) is
 		\begin{gather*}
@@
 	\end{gather*}
-	where $k_0=0$, $x^{k_l +}_i = \max_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$,\\
-	and $x^{k_l -}_i = \min_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$. Here, $[a,b]$ denotes the discrete interval in $\Zz$, i.e. $[0,2]=\{0,1,2\}$.
+	where \(k_0=0\)
+{\color{blue} New variable(s): \(k_0\),  }
+, \(x^{k_l +}_i = \max_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}\)
+{\color{blue} New variable(s): \(k_l\), \(x_i\),  }
+,\\
+	and \(x^{k_l -}_i = \min_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}\). Here, \([a,b]\)
+{\color{blue} New variable(s): \(a\), \(b\),  }
+ denotes the discrete interval in \({\mathbb{Z}}\), i.e. \([0,2]=\{0,1,2\}\).
 	\end{definition}
-We calculate the minimal necessary translations of the scheme, such that all instances of $u_j(n+e,t+l)$, that appear in the discrete time derivative, also appear in the translated scheme.
+We calculate the minimal necessary translations of the scheme, such that all instances of \(u_j(n+e,t+l)\), that appear in the discrete time derivative, also appear in the translated scheme.
 This is done by elementwise subtracting the range of the stencil of the scheme from the range of the stencil of the discrete time derivative.
-We denote the resulting translated system by $\{H^1,\ldots,H^r\}$.
-If for any $[a,b]$ involved $b<a$, we write $[a,b]=\emptyset$ with the convention that $\emptyset-\emptyset=\emptyset$.
-Further, $\emptyset$ in the translations results in the use of the respective entry of the original scheme without translations.
+We denote the resulting translated system by \(\{H^1,\ldots,H^r\}\).
+If for any \([a,b]\) involved \(b<a\), we write \([a,b]=\emptyset\) with the convention that \(\emptyset-\emptyset=\emptyset\).
+Further, \(\emptyset\) in the translations results in the use of the respective entry of the original scheme without translations.
 	\begin{example}
@@
 		\end{equation*}
+{\color{blue} New variable(s): \(u_2\),  }
+
 		Then, the stencil of the discrete time derivative equals 
@@
 		\normalsize
-		Therefore, we translate the first equation of the scheme by $(0,0)$ and $(1,0)$ and the second one by $(0,0)$ to get the translated scheme
+		Therefore, we translate the first equation of the scheme by \((0,0)\) and \((1,0)\) and the second one by \((0,0)\) to get the translated scheme
 		\begin{gather*}
@@
 	\subsubsection*{Polynomial ideals and Gr{\"o}bner bases}\label{polid}
-	We found a finite set of polynomials (the translated scheme and the discrete time derivative) with a finite number of instances of $u_j(\tilde{n}+e,t+l)$ in the previous step.
-Now we reduce $\mathcal{T}$, using the translated scheme.
+	We found a finite set of polynomials (the translated scheme and the discrete time derivative) with a finite number of instances of \(u_j(\tilde{n}+e,t+l)\) in the previous step.
+Now we reduce \(\mathcal{T}\), using the translated scheme.
 Here, we adapt the definitions and theorems from \cite{hibi13} to our setting.
 	\begin{definition}
-		Let $\mathcal{K}$ be as in Definition \ref{diffideal}.
-Let $\mathcal{R}$ be the polynomial ring generated by $\mathcal{K}$ and all instances of $u_j(n+e,t+l)$ that occur in $\{H^1,\ldots,H^r,\mathcal{T}\}$.
-		We call a set $ I\subseteq \mathcal{R}$ a (polynomial) ideal if
+		Let \(\mathcal{K}\) be as in Definition \ref{diffideal}.
+Let \(\mathcal{R}\) be the polynomial ring generated by \(\mathcal{K}\) and all instances of \(u_j(n+e,t+l)\) that occur in \(\{H^1,\ldots,H^r,\mathcal{T}\}\).
+		We call a set \( I\subseteq \mathcal{R}\) a (polynomial) ideal if
 		\begin{itemize}
-			\item $p_1,p_2\in I$ implies $p_1+p_2\in I$,
-			\item $p_1\in I,p_2\in\mathcal{R}$ implies $p_1p_2\in I$.
+			\item \(p_1,p_2\in I\) implies \(p_1+p_2\in I\),
+			\item \(p_1\in I,p_2\in\mathcal{R}\) implies \(p_1p_2\in I\).
 		\end{itemize}
@@
 	\begin{definition}
-		We denote by $\langle H^1,\ldots,H^r\rangle$ the smallest (polynomial) ideal containing $\{H^1,\ldots,H^r\}$.
+		We denote by \(\langle H^1,\ldots,H^r\rangle\) the smallest (polynomial) ideal containing \(\{H^1,\ldots,H^r\}\).
 	\end{definition}
-	Given the ideal $\langle H^1,\ldots,H^r\rangle$ and the discrete time derivative $\mathcal{T}\in\mathcal{R}$, we want to determine if $\mathcal{T}\in\langle H^1,\ldots,H^r\rangle$ or if we can write $\mathcal{T}$ in a simpler form, using $\{H^1,\ldots,H^r\}$.
-	Hence, using multivariate polynomial division, we search for $p^1,\ldots,p^r,p^{r+1}\in\mathcal{R}$ such that
+	Given the ideal \(\langle H^1,\ldots,H^r\rangle\) and the discrete time derivative \(\mathcal{T}\in\mathcal{R}\), we want to determine if \(\mathcal{T}\in\langle H^1,\ldots,H^r\rangle\) or if we can write \(\mathcal{T}\) in a simpler form, using \(\{H^1,\ldots,H^r\}\).
+	Hence, using multivariate polynomial division, we search for \(p^1,\ldots,p^r,p^{r+1}\in\mathcal{R}\)
+{\color{blue} New variable(s): \(p\),  }
+ such that
 	\begin{gather*}
@@
 	\end{gather*}
-	Unfortunately, the resulting remainder $p^{r+1}$ may not be unique \cite[page 14, Example 1.2.3]{hibi13}, i.e. non-zero, although $\mathcal{T}$ belongs to $\langle H^1,\ldots,H^r \rangle$. 
-	Replacing $\{H^1,\ldots,H^r\}$ by a Gr{\"o}bner basis for the ideal $\langle H^1,\ldots,H^r \rangle$ guarantees the uniqueness of the remainder of the polynomial division. 
+	Unfortunately, the resulting remainder \(p^{r+1}\) may not be unique \cite[page 14, Example 1.2.3]{hibi13}, i.e. non-zero, although \(\mathcal{T}\) belongs to \(\langle H^1,\ldots,H^r \rangle\). 
+	Replacing \(\{H^1,\ldots,H^r\}\) by a Gr{\"o}bner basis for the ideal \(\langle H^1,\ldots,H^r \rangle\) guarantees the uniqueness of the remainder of the polynomial division. 
 	Contrary to the standard basis for the difference ideal that Gerdt's algorithm computes, the Gr{\"o}bner basis for the polynomial ideal always exists and is finite.\\
@@
 	\subsubsection*{Polynomial reduction}
-	In Algorithm \ref{alg: discreteconservedqimplicit}, we consider two monomial orders: (a) lexicographic and (b) explicit.
-	To reduce $\mathcal{T}$, we compute the Gr{\"o}bner basis of $\langle H^1,\ldots,H^r\rangle$ with respect to the monomial order and then calculate the remainder of the polynomial division of $\mathcal{T}$ with respect to this Gr{\"o}bner basis.\\
-	The explicit monomial order (b) results in the Gr{\"o}bner basis that eliminates the instances of $u_j(n+e,t+l)$ at the latest time (with the largest $l$).
+	In Algorithm \ref{alg: discreteconservedqimplicit}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{alg: discreteconservedqimplicit}\end{verbatim}.}
+, we consider two monomial orders: (a) lexicographic and (b) explicit.
+	To reduce \(\mathcal{T}\), we compute the Gr{\"o}bner basis of \(\langle H^1,\ldots,H^r\rangle\) with respect to the monomial order and then calculate the remainder of the polynomial division of \(\mathcal{T}\) with respect to this Gr{\"o}bner basis.\\
+	The explicit monomial order (b) results in the Gr{\"o}bner basis that eliminates the instances of \(u_j(n+e,t+l)\) at the latest time (with the largest \(l\)).
 This elimination might not always be possible, hence the reduction may leave the discrete time derivative unchanged.
-Then we repeat this process for the instance at the second latest time and continue until all instances of $u$ have been eliminated.\\
+Then we repeat this process for the instance at the second latest time and continue until all instances of \(u\) have been eliminated.\\
 	Out of the two remainders from (a) and (b), we choose the result with the least number of different instances of time. Our code also supports other elimination orders, including user-defined ones.
@@
 		\label{heateqalgoimplicit}
-		To demonstrate Algorithm \ref{alg: discreteconservedqimplicit}, we use the setting from Example \ref{timeexplicitheateqalgo}.
+		To demonstrate Algorithm \ref{alg: discreteconservedqimplicit}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{alg: discreteconservedqimplicit}\end{verbatim}.}
+, we use the setting from Example \ref{timeexplicitheateqalgo}.
 		We compute the discrete time derivative (step 1)
@@
 		\end{gather*}
-	with range $\Big([-1,1]\times[0,1]\Big)$.
-	The discrete time derivative has stencil $\Big(\big\{(0,1),(0,0)\big\}\Big)$
-	and range $\Big([0,0]\times[0,1]\Big)$.
+	with range \(\Big([-1,1]\times[0,1]\Big)\).
+	The discrete time derivative has stencil \(\Big(\big\{(0,1),(0,0)\big\}\Big)\)
+	and range \(\Big([0,0]\times[0,1]\Big)\).
 	The translations are
@@
 	\end{gather*}
-	For the reduction using the elimination order, we first eliminate $u(n,t+1)$ and then the remaining instances of $u$.
+	For the reduction using the elimination order, we first eliminate \(u(n,t+1)\) and then the remaining instances of \(u\).
 	The elimination order yields the same remainder as the lexicographic order. Hence, we do not need to check for the number of instances of time (step 4).
@@
 	\end{gather*}
-	that equals $0$.
+	that equals \(0\).
 	Hence, repeating the above procedure (step 6) becomes unnecessary.
@@
 So far, we have discussed how to check if a quantity is preserved in time under a scheme. But it is also desirable to have a systematic way to find conserved quantities, given a scheme.
-Algorithm \ref{alg: findconservedquantitybasis} finds, for a time-explicit scheme, a basis for conserved quantities that are generated by monomials up to a (total) degree.
+Algorithm \ref{alg: findconservedquantitybasis}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{alg: findconservedquantitybasis}\end{verbatim}.}
+ finds, for a time-explicit scheme, a basis for conserved quantities that are generated by monomials up to a (total) degree.
 
@@
 \begin{example}
-	We find a basis for conserved quantities that are generated by $u(n,t)$ and $n$ and that have at most degree 3, admitted by the scheme for the heat equation from Example \ref{heatex1}.
+	We find a basis for conserved quantities that are generated by \(u(n,t)\) and \(n\) and that have at most degree 3, admitted by the scheme for the heat equation from Example \ref{heatex1}.
 
@@
 \subsection*{Burgers equation}
-The Burgers equation is the PDE $u_t+uu_x=0$ \cite{smoller94} in one space dimension. Any function of $u$ is a conserved quantity. We are interested in discretizations that preserve mass $\int u\ dx$.
+The Burgers equation is the PDE \(u_t+uu_x=0\)
+{\color{blue} New variable(s): \(u_t\), \(u_x\),  }
+ \cite{smoller94} in one space dimension. Any function of \(u\) is a conserved quantity. We are interested in discretizations that preserve mass \(\int u\ dx\).
 \begin{example}
@@
 \begin{example}
-	We discretize $u_x$ using a three-point stencil with a parameter $a$.
+	We discretize \(u_x\) using a three-point stencil with a parameter \(a\).
 
@@
 	\end{small}
-	We have conservation if we choose $a=\frac{1}{2}$, corresponding to the standard central difference.
+	We have conservation if we choose \(a=\frac{1}{2}\), corresponding to the standard central difference.
 \end{example}
@@
 	\end{gather*}
-	This system admits the conserved quantities $\int v\ dx$ and $\int m\ dx$.
+	This system admits the conserved quantities \(\int v\ dx\) and \(\int m\ dx\).
 	Because this system was derived from a forward-forward mean-field game, we discretize it forward in time.
@@
 	\end{small}
-	Those are the only conserved quantities for this scheme, that are polynomials up to degree 4 in $v$ and $m$.
+	Those are the only conserved quantities for this scheme, that are polynomials up to degree 4 in \(v\) and \(m\).
 
@@
 	\end{gather*}
-	This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly (forward) in time for $m$ and implicitly (backward) for $v$.
-Our code can handle this setting, if we specify an order for the elimination of the variables, telling the code to use an explicit monomial order for $m$ and an implicit one for $v$:
+	This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly (forward) in time for \(m\) and implicitly (backward) for \(v\).
+Our code can handle this setting, if we specify an order for the elimination of the variables, telling the code to use an explicit monomial order for \(m\) and an implicit one for \(v\):
 
@@
 In this case, non-polynomial expressions can be handled by writing them in polynomial form.
-For example, treat the expression $2\cos(u(n+e^1))e^{u(n+e^2)}+\frac{1}{u(n+e^3)}$
-as the polynomial $2x_1x_2+x_3$.
+For example, treat the expression \(2\cos(u(n+e^1))e^{u(n+e^2)}+\frac{1}{u(n+e^3)}\)
+as the polynomial \(2x_1x_2+x_3\)
+{\color{blue} New variable(s): \(x_3\),  }
+.
 \subsection*{Translate the scheme more accurately}

### 2023-05-02 Papers/Paper ACM/tmp/GCpaper.tex
Friedmann paper

\documentclass{article}
\begin{document}

[Detection of conserved quantities for PDE schemes]{Algorithmic detection of conserved quantities of finite-difference schemes for partial differential equations}

{Diogo A. Gomes}
{
	{King Abdullah University of Science and Technology (KAUST)}
	[0]{CEMSE Division}
	[1]{AMCS Program}
	{Thuwal}
	{23955-6900}
	{Saudi Arabia}}
{diogo.gomekaust.edu.sa}
{Friedemann Krannich}
{
	{King Abdullah University of Science and Technology (KAUST)}
	[0]{CEMSE Division}
	[1]{AMCS Program}
	{Thuwal}
	{23955-6900}
	{Saudi Arabia}}
{friedemann.krannickaust.edu.sa}
{Ricardo de Lima Ribeiro}
{
	{King Abdullah University of Science and Technology (KAUST)}
	[0]{CEMSE Division}
	[1]{AMCS Program}
	{Thuwal}
	{23955-6900}
	{Saudi Arabia}}
{ricardo.ribeirkaust.edu.sa}

{Symbolic computations; Finite-difference schemes; Discrete variational derivative; Discrete partial variational derivative; Conserved quantities; Implicit schemes; Explicit schemes.}

	<ccs2012>
	<concept>
	<concept_id>10002950.10003714.10003727.10003729</concept_id>
	<concept_desc>Mathematics of computing~Partial differential equations</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10002950.10003714.10003715.10003750</concept_id>
	<concept_desc>Mathematics of computing~Discretization</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10002950.10003714.10003715.10003720.10003747</concept_id>
	<concept_desc>Mathematics of computing~Gr{}bner bases and other special bases</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10010147.10010148.10010149.10010152</concept_id>
	<concept_desc>Computing methodologies~Symbolic calculus algorithms</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10010147.10010148.10010149.10010155</concept_id>
	<concept_desc>Computing methodologies~Discrete calculus algorithms</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	</ccs2012>

[500]{Mathematics of computing~Partial differential equations}
[500]{Mathematics of computing~Discretization}
[500]{Mathematics of computing~Gr{}bner bases and other special bases}
[500]{Computing methodologies~Symbolic calculus algorithms}
[500]{Computing methodologies~Discrete calculus algorithms}

{}

Many partial differential equations (PDEs) admit conserved quantities like mass or energy. Those quantities are often essential to establish well-posed results. When approximating a PDE by a finite-difference scheme, it is natural to ask whether related discretized quantities remain conserved under the scheme. Such conservation may establish the stability of the numerical scheme. We present an algorithm for checking the preservation of a polynomial quantity under a polynomial finite-difference scheme. In our algorithm, schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables. 
Additionally, we present an algorithm for, given a scheme, finding conserved quantities. 
We illustrate our algorithm by studying several finite-difference schemes.

	The authors were supported by {GN01}{King Abdullah University of Science and Technology (KAUST) baseline funds}{} and 
	{GN02}{KAUST OSR-CRG2021-4674}{}.

Introduction.

Many partial differential equations (PDEs) admit integral quantities, that are conserved in time. 
For example, the advection equation preserves energy and the heat equation conserves mass. 
When approximating PDEs by a finite-difference scheme, the question arises whether such quantities are conserved by the scheme. 
Such information can be crucial to estimate whether a scheme approximates a PDE accurately and to determine its stability.
While computations for determining conservation are often easy in simple cases, they get rather tedious for more complicated schemes or quantities. 
Hence, automating such computations 
in computer algebra systems is desirable.
An algorithm for finding conserved quantities for (continuous) differential equations was developed and implemented  in {Maple} by Cheviakov (see [1] and [1]).
Hereman et al. (see for example [1] and [1]), proposed an algorithm to compute conserved densities for semi-discretized schemes for PDEs with first-order time derivative. 
Their algorithm, which is implemented in a {Mathematica} package, uses the scaling symmetries of the scheme to construct conserved quantities and calculates their coefficients by using the discrete variational derivative.
Gao et al. extended this algorithm to first-order time-explicit schemes [1] and to first-order time-implicit schemes [1]. 
They also implemented it in the computer algebra system {Reduce}.
In this paper, we propose an algorithm that checks if a quantity is conserved in time under a scheme.
Conserved quantities usually involve integral expressions or their discrete analog, sums.
Gomes et al. proposed algorithms for the simplification of sums in [1].
They also developed and implemented algorithms for detecting quantities conserved by PDEs and semi-discretized schemes. 
We revise these techniques in Section [1]. 
To generalize those methods to situations where we do not sum over all arguments of the functions involved, we introduce the discrete partial variational derivative (Section [1]).
The main contribution of this paper is an algorithm for checking the conservation of a quantity under a numerical scheme.
Gerdt showed in [1], that if the discrete time derivative of the quantity belongs to the difference ideal generated by the scheme, the quantity is conserved. However, some quantities may add to a constant (e.g. telescopic sums) and thus be trivially preserved without belonging to the difference ideal. Moreover, Gerdt's algorithm may not terminate, as the Gr{}bner basis for the difference ideal may not be finite.
We overcome this issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) with finite Gr{}bner basis (Section [1]). 
Our algorithm works for schemes that are explicit and implicit in time and can treat schemes with several and higher-order time  and space derivatives and with several space dimensions.
Additionally, we can handle systems of equations and schemes with parameters. 
We have implemented this algorithm as part of a package in {Mathematica} [1].
In the examples, we show that our code finds conserved quantities and proper schemes for the time-implicit and time-explicit discretization of the Burgers equation and a system of PDEs arising in the study of mean-field games (Section [1]).

Preliminaries.

Throughout this section, we follow the ideas and computations in [1].
Here, subscripts denote indices related to coordinates or tuples while superscripts denote indices related to sequences and families.
To avoid dealing with boundary terms, we work with periodic functions in [1].

	Let [1], [1] and [1] be natural numbers. The {discrete torus} is [1]. Define the space
	[1],
	extended to [1] by periodicity.

	The space of functionals [1] (not necessarily linear) on [1] is
[1].

In the previous definition, the [1] are not necessarily unit vectors and can vary between functionals. Here, we allow for smooth [1]. However, later we require [1] to be a polynomial function.

	Let [1], [1] and define
	[1].

To simplify sums, the discrete variational derivative is a useful tool.
Let [1], [1] and [1]. Define
	[1].
	In [1], we shift the indices due to periodicity of [1] and [1],
	[1].

[1] is the discrete variational derivative of [1].

Because we assume all related functions to be smooth, the discrete variational derivative always exists.

	Let us return to Example [1] and compute
	[1]
	and hence [1].

Fundamental theorem.

The algorithms for the simplification of sums presented in [1] rely on the following result:

	Let [1]. If [1] for all [1] and if there exists [1] such that [1], then
	[1]
	for all [1].
	Conversely,
	[1]
	for all [1], if [1].

	[1]
	and hence [1].

We use Theorem [1] to examine if different sums represent the same quantity.

	Let [1] and consider the functionals
	[1].
	It is clear, that [1] and [1] represent the same quantity (by shifting [1]). We confirm this, using the discrete variational derivative:
	[1].
	Hence, [1] and [1]. Therefore, both functionals sum to the same quantity.

	Let [1]. It may not be obvious, that
	[1].
	We confirm this, by computing the discrete variational derivative and noticing that the expression is [1] for [1].

The discrete partial variational derivative.

Here, we generalize the discrete variational derivative for situations where we keep one (or several) of the arguments of [1] constant.
Let [1] be the variables for the functions in [1]. We call [1] the space variables. Let [1] and call [1] the time variables. Later in this paper, we only consider the case [1].

	Let [1], extended to [1] by periodicity in the space variables and let [1].

When considering sums, we only sum over the space variables and not over the time variables.

	The space of functions [1] is
	[1].
Here, [1].

Throughout this paper, all quantities, whose conservation in time we check, are functions [1]. 
Define [1]. Using the Kronecker delta [1], we rewrite
[1].
	Let [1] and compute

	[1].

	We define [1] as the discrete partial variational derivative of [1].
	Note that [1].

	The discrete partial variational derivative of 
	[1]
	equals [1], because

	[1].

	In our {Mathematica} implementation, Example [1] is computed by

We have a result similar to Theorem [1] for the discrete partial variational derivative:

	Let [1]. If [1] for all [1] and if there exists [1] such that [1], then
	[1]
	for all [1].
	Conversely,
	[1]
	for all [1], if [1].

The proof is similar to the proof of Theorem [1].
Difference schemes and algebra.

In this section, we define numerical schemes formally and introduce the tools from difference algebra, that we need for our algorithm.
During the remainder of this paper, we assume that there is only one time variable [1] ([1]).

	The space of schemes is
	[1].
	A scheme is a set of functions
	[1]
	that represent the equations [1] for [1], holding pointwise for all points in [1].

If a scheme contains a single function [1], we call [1] the scheme. Sometimes, we also call the expression [1] the scheme.
In this work, we only consider finite-difference schemes with fixed step size [1], but one can generalize our results and algorithms to other step sizes.

	A scheme [1] given by
	[1]
	is in time-explicit form if we can rewrite the previous equation (eventually translating the scheme) as
	[1]
	for [1] and [1]
	with [1]. We call [1] the right-hand side of [1].

	Consider the forward-difference scheme for the heat equation [1]
	[1].
	This scheme is in time-explicit form with right-hand side
	[1].

Difference ideals.

Following Gerdt [1], we now introduce difference ideals and how we use them in our algorithm.

	Let [1]. The shift in the [1]-th coordinate for [1] by [1] is
	[1]
	understanding that [1] shifts the time variable.

	The set of all possible shifts [1] is
	[1].

Now we construct the difference ideal containing the scheme:

	Let [1] be the field generated by [1] and the variables [1]. Let [1] be the polynomial ring over the field [1] and the variables [1] for [1] and [1].
	A set [1] is a difference ideal if

		 [1] implies [1],
		 [1] implies [1],
		 [1] implies [1].

	[1] is the smallest difference ideal containing the scheme [1].

A solution of a numerical scheme is a function [1], that makes all translations of the scheme vanish. Hence, every element of the difference ideal generated by the scheme vanishes under [1].
In particular, Algorithm [1] seeks to determine if the discrete time derivative [1] belongs to the ideal [1]. To examine, if [1], Gerdt proposes the notion of a standard (Gr{}bner) basis for the ideal [1]. This standard basis may not be finite. We overcome this problem by considering polynomial instead of difference algebra and using a smaller polynomial ideal, that is contained in the difference ideal and admits a finite standard basis.
However, there may be functions [1], whose sums add to zero, even though they may fail to belong to the difference ideal. Hence, we combine the discrete partial variational derivative with polynomial ideals.
The time-explicit case.

In this section, we present Algorithm [1]  for checking the conservation of a quantity under the right-hand side of a scheme in time-explicit form. Our algorithm for general schemes can be found in the subsequent section.

Note, that the result of step 2 exists only at time [1] because all instances of [1] have been replaced. Thus, we do not need any computations of difference ideals or Gr{}bner basis.

	We check for conservation of [1] under the scheme from Example [1].
	The discrete time derivative (step 1) is
	[1].
	Replacing [1] by the right-hand side of the scheme (step 2) gives
	[1].
	Then, we compute the discrete partial variational derivative (step 3), which equals zero. Hence (step 4), we have conservation.

	We verify the result from Example [1], using our implementation in {Mathematica} of Algorithm [1].

The general case.

In this section, we present Algorithm [1] that deals with general, not necessarily time-explicit, schemes. We explain its steps in detail below and demonstrate the algorithm in Example [1].

Step 1: Build the discrete time derivative.

We build the discrete time derivative by subtracting [1] from [1]. This task is done in our code by the {TimeDifference}.

		The discrete time derivative for [1] is

Step 2: Translate the scheme.

To make the step from difference to polynomial algebra to guarantee termination of the algorithm, we use a standard idea in symbolic computations and consider every instance of [1] as a variable of a multivariate polynomial.
Hence, we compute all translations of the scheme, such that all variables in the polynomial associated with [1] appear in the polynomials associated with the translated scheme.

	The algorithm treats the discrete time derivative [1]
	as the polynomial equation [1].

		Let [1] and
		[1]

	with [1].
	The {stencil} of [1] is the [1]-tuple of sets of vectors
	[1].
	The {range} of the stencil of [1] is
		[1],
	where [1], [1],
	and [1]. Here, [1] denotes the discrete interval in [1], i.e. [1].

We calculate the minimal necessary translations of the scheme, such that all instances of [1], that appear in the discrete time derivative, also appear in the translated scheme. This is done by elementwise subtracting the range of the stencil of the scheme from the range of the stencil of the discrete time derivative. We denote the resulting translated system by [1].
If for any [1] involved [1], we write [1] with the convention that [1].  Further, [1] in the translations results in the use of the respective entry of the original scheme without translations.

		Consider the discrete time derivative
		[1]
		and the scheme
		[1].
		Then, the stencil of the discrete time derivative equals 
		[1]
		with range
		[1].
		The stencil of the scheme equals the set of stencils of the equations in the scheme, i.e.
		[1]
		with range
		[1].
		Hence, we get the translations for the first equation of the scheme by

		[1]

		and for the second equation of the scheme by

		[1].

		Therefore, we translate the first equation of the scheme by [1] and [1] and the second one by [1] to get the translated scheme
		[1].

Step 3 and 4: Compute the Gr{}bner basis and reduce the discrete time derivative.

	Polynomial ideals and Gr{}bner bases.

	We found a finite set of polynomials (the translated scheme and the discrete time derivative) with a finite number of instances of [1] in the previous step. Now we reduce [1], using the translated scheme. Here, we adapt the definitions and theorems from [1] to our setting.

		Let [1] be as in Definition [1]. Let [1] be the polynomial ring generated by [1] and all instances of [1] that occur in [1].
		We call a set [1] a (polynomial) ideal if

			 [1] implies [1],
			 [1] implies [1].

		We denote by [1] the smallest (polynomial) ideal containing [1].

	Given the ideal [1] and the discrete time derivative [1], we want to determine if [1] or if we can write [1] in a simpler form, using [1].
	Hence, using multivariate polynomial division, we search for [1] such that
	[1].
	Unfortunately, the resulting remainder [1] may not be unique [page 14, Example 1.2.3]{hibi13}, i.e. non-zero, although [1] belongs to [1]. 
	Replacing [1] by a Gr{}bner basis for the ideal [1] guarantees the uniqueness of the remainder of the polynomial division. 
	Contrary to the standard basis for the difference ideal that Gerdt's algorithm computes, the Gr{}bner basis for the polynomial ideal always exists and is finite.
	A Gr{}bner basis is defined up to the order of the monomials involved, so, depending on the order, we get different remainders of the polynomial division.
	Polynomial reduction.

	In Algorithm [1], we consider two monomial orders: (a) lexicographic and (b) explicit.
	To reduce [1], we compute the Gr{}bner basis of [1] with respect to the monomial order and then calculate the remainder of the polynomial division of [1] with respect to this Gr{}bner basis.
	The explicit monomial order (b) results in the Gr{}bner basis that eliminates the instances of [1] at the latest time (with the largest [1]). This elimination might not always be possible, hence the reduction may leave the discrete time derivative unchanged. Then we repeat this process for the instance at the second latest time and continue until all instances of [1] have been eliminated.
	Out of the two remainders from (a) and (b), we choose the result with the least number of different instances of time. Our code also supports other elimination orders, including user-defined ones.
Step 5 to 7: Compute the discrete partial variational derivative and reduce again.

	In the next step, we take the resulting expression and compute its discrete partial variational derivative. Then, we repeat Step 2 to Step 4 applied to the discrete partial variational derivative.

		To demonstrate Algorithm [1], we use the setting from Example [1].
		We compute the discrete time derivative (step 1)
		[1].
		For the translations (step 2), the stencil of the scheme is
		[1]
	with range [1].
	The discrete time derivative has stencil [1]
	and range [1].
	The translations are

	[1]

	which results in no translations, as the range of the stencil of the scheme is greater than the range of the stencil of the discrete time derivative.
	Hence, the Gr{}bner basis (step 3) coincides with the original scheme for both monomial orders.
	The reduction using the lexicographic monomial order yields as remainder
	[1].
	For the reduction using the elimination order, we first eliminate [1] and then the remaining instances of [1].
	The elimination order yields the same remainder as the lexicographic order. Hence, we do not need to check for the number of instances of time (step 4).
	We calculate the discrete partial variational derivative (step 5) of
	[1]
	that equals [1].
	Hence, repeating the above procedure (step 6) becomes unnecessary.
	Therefore, (step 7) we see that the scheme conserves the quantity.

		We verify our result from Example [1], using {DiscreteConservedQ}, which detects automatically if we are in the time-explicit or the general setting.

	The algorithm can only detect conservation, but can not check if something is not conserved for sure. So a resulting {False} should be understood as our algorithm not detecting conservation, not as there being no conservation at all.

A basis for conserved quantities.

So far, we have discussed how to check if a quantity is preserved in time under a scheme. But it is also desirable to have a systematic way to find conserved quantities, given a scheme.
Algorithm [1] finds, for a time-explicit scheme, a basis for conserved quantities that are generated by monomials up to a (total) degree.

We have implemented this algorithm in the {FindDiscreteConservedQuantityBasis}.

	We find a basis for conserved quantities that are generated by [1] and [1] and that have at most degree 3, admitted by the scheme for the heat equation from Example [1].

Examples and applications.

We illustrate our code with the Burgers equation and a system of mean-field games.
Burgers equation.

The Burgers equation is the PDE [1] [1] in one space dimension. Any function of [1] is a conserved quantity. We are interested in discretizations that preserve mass [1].

	We check for conservation of mass using a forward-difference discretization:

We want to find a scheme that preserves this quantity. Therefore we check for conservation under a class of schemes with a parameter, to find an appropriate choice for this parameter.

	We discretize [1] using a three-point stencil with a parameter [1].

	We have conservation if we choose [1], corresponding to the standard central difference.

So far, we have only seen schemes that are in time-explicit form, but our algorithm allows also for general schemes:

	We consider the scheme from the previous example but in the time-implicit version.

Conserved quantities for a mean-field game.

As a second application, we use our code to study conserved quantities admitted by the discretization of a system of PDEs.

	In [1], Gomes et al. derived the following system of PDEs
	[1].
	This system admits the conserved quantities [1] and [1].
	Because this system was derived from a forward-forward mean-field game, we discretize it forward in time. We check with our code if the scheme admits the same preserved quantities as the continuous system:

	We reproduce this result by noticing, that the scheme is in time-explicit form:

	Those are the only conserved quantities for this scheme, that are polynomials up to degree 4 in [1] and [1].

	The related backward-forward system reads
	[1].
	This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly (forward) in time for [1] and implicitly (backward) for [1]. Our code can handle this setting, if we specify an order for the elimination of the variables, telling the code to use an explicit monomial order for [1] and an implicit one for [1]:

Possible extensions and concluding remarks.

Polynomial treatment of non-polynomial expressions.

It may be possible to extend our methods to non-polynomial PDEs and schemes. In this case, non-polynomial expressions can be handled by writing them in polynomial form. For example, treat the expression [1]
as the polynomial [1].
Translate the scheme more accurately.

In our elimination procedure, we work with a specific polynomial ideal that is a subset of the difference ideal generated by the scheme. However, it could be necessary to translate the scheme more than we did in our algorithm to get cancellation. Therefore, flexible methods to determine the translations of the scheme are desirable.
Discrete conserved quantities that are not conserved by the PDE.

It may arise the question if we can find conserved quantities, that are not preserved by the original PDE. Gerdt et al. ([1],[1]) define the notion of s-consistency, which guarantees that all discrete quantities are also preserved in the continuous setting. One possible extension of our code would be to check automatically if s-consistency holds for the translated scheme.
Concluding remarks.

We present an algorithm for checking the conservation of a quantity under a finite-difference scheme. Our algorithm allows for systems of equations, arbitrary time and space derivatives, and both explicit and implicit schemes. Also, we implemented a function for finding conserved quantities admitted by a scheme. We use our implementation of the algorithm to analyze the conservation properties of several schemes for PDEs that arise in applications.

{ACM-Reference-Format}
{references}

\end{document}

### 2023-05-02 Papers/Paper ACM/MLaTeX.m

@@ -1,0 +1,1777 @@
+(* ::Package:: *)
+
+BeginPackage["MLaTeX`"]
+Needs["CellsToTeX`"]
+
+
+MLaTeX::usage="Processes the file"
+
+(* Basic LaTeX check *)
+
+LATEXSANITYCHECK::usage="Minimal check on latex";
+BASICLATEXVALIDATION::usage="Checks if all environments open and closes, checks if
+braces are balanced in text and if all formulas open and close";
+
+(* clean up and includes handling *)
+
+BASICTEXTCLEAN::usage="Takes a LaTeX string and returns the file without comments both % and {comment} environment. Also replaces \\n\\r by \\n. ";
+BASICTEXTCLEANA::usage="just like BASICTEXTCLEAN but does not clean mmacells and algorithms";
+BASICINCLUDE::usage="Scans string for \\input and replaces the appropriate file";
+BASICINCLUDEREMOVER::usage="Scans string for \\input and if the file does not exist removes \\input";
+BEGINENDDOCCLEANER::usage="Replaces macros for begin{document} and end{document}";
+INCLUDEANDCLEAN::usage="Applies BASICTEXTCLEAN followed by BASICINCLUDE until a fixed point is reached";
+INCLUDEANDCLEANA::usage="Applies BASICTEXTCLEANA followed by BASICINCLUDE until a fixed point is reached";
+MLaTeXMARKUPREMOVER::usage="Takes a string, removes MLATEX PATTERNS, replaces mequation by equation, removes MLaTex Package";
+FORMULAFIXER::usage="Replaces in a string $$...$$ by \\[...\\] and $...$ by \\(...\\)";
+
+(* TO BE REMOVED IF REPLACED BY NEW METHODS*)
+
+LATEXFORMULACLEANER::usage="Removes formulas and images and replaces formulas by [1]"; 
+LATEXONLYFORMULACLEANER::usage="Removes formulas them by [1]"; 
[…]

### 2023-05-02 PDESymbolics2022/Discrete.m
comments

@@ -362,6 +362,9 @@ Reduction[variables_Association][schemeexpression_Association] :=
       schexp["exp"] // Variables // Flatten] // Flatten, 
     Lookup[var, "pars", {}]];
  polyvars = Sort[polyvars, TimeOrderedQOperator[Append[variables,"elimOrder"->"explicit"]]];
+ 
+ (*from here *)
+ 
   scheme = 
    GroebnerBasis[polylist, polyvars, 
     CoefficientDomain -> RationalFunctions];
@@ -373,12 +376,16 @@ Reduction[variables_Association][schemeexpression_Association] :=
   order of variables in which they are eliminated , 
   default is permutations, 
   specify function that does that by function name*)
+ 
   eliminationlist = 
    Lookup[var, "EliminationListOperator", EliminationListOperator][
      Append[var, "scheme" -> polylist]][
     schexp["exp"]]; (*list of lists*)
   reducelist = VariableEliminationOperator[Append[var,"elimOrder"->Lookup[var,"elimOrder","explicit"]]][#, schexp] & /@ eliminationlist;
    Append[reducelist, normalred] // PiecewiseExpand // PiecewiseListClean
+   
+   (* to here there are some picewise objects *)
+   
 	];
 
 drlMatrix[n_] := 
@@ -423,6 +430,9 @@ VariableEliminationOperator[variables_Association][
       	n1 = Position[polyvars,elem]//First//First;
       	n2 = Length[polyvars];
       	ordermatrix = elimMatrix[n1,n2];
+      	
+      (* from here *)	
+      	
       Gbasis = GroebnerBasis[schexp["scheme"], polyvars, CoefficientDomain -> RationalFunctions, MonomialOrder -> ordermatrix];
       schexp["exp"] = 
        PolynomialReduce[schexp["exp"], Gbasis, polyvars, CoefficientDomain -> RationalFunctions, MonomialOrder -> ordermatrix] // Last;
@@ -432,9 +442,12 @@ VariableEliminationOperator[variables_Association][
        ];
       ];
       VariableEliminationOperator[var][Drop[elimlist, 1],schexp],
-      schexp
+      schexp      
       ]
      ]
+     
+     (* up to here there are some picewise objects *)
+     
     ]
    ];
 

### 2023-05-04 PDESymbolics2022/Discrete.m
Update Discrete.m
There was an output that needed PiecewiseExpand.

@@ -707,7 +707,7 @@ FindDiscreteConservedQuantityBasisOperator[variables_Association][problem_] := W
           Cases[#, Subscript[\[FormalA], _], {0, Infinity}]]},
        Map[Function[coeff, conservationlaws /. coeff -> 1], 
          coefficients] /. (Function[c, Rule[c, 0]] /@ 
-          coefficients)] &, conservationlaws] // PiecewiseBeautify]; 
+          coefficients)] &, conservationlaws] //PiecewiseExpand// PiecewiseBeautifyOperator[variables]]; 
           
 End[] (* End Private Context *)
 (*EndPackage[]*)

### 2023-05-04 PDESymbolics2022/ParametricGroebnerBasis.m
Update ParametricGroebnerBasis.m
Fixed an issue with AutoReduceOperator: It now looks into all the cases of parameters (refine...)

@@ -26,7 +26,7 @@ AutoReduceOperator::usage =
 "AutoReduceOperator[variables][polylist] returns the reduced (smallest set generating the same ideal) polylist.";
 
 PiecewisePolynomialReduceRemainderOperator::usage =
-"PiecewisePolynomialReduceRemainderOperator[variables][p, polylist] returns a Piecewise remainder for the redeuctions.";
+"PiecewisePolynomialReduceRemainderOperator[variables][p, polylist] returns a Piecewise remainder for the reductions.";
 
 PiecewiseApplyConditionOperator::usage =
 "PiecewiseApplyConditionOperator[variables][px] Simplifies the expressions with conditions as Assumptions";
@@ -44,7 +44,7 @@ Begin["`Private`"]
 InferGeneratorsOperator[variables_][xplist_List] :=
     With[ {glc = First@EchoLabel["InferGenerators: gLc"]@GenericLinearCombinationOperator[variables][EchoLabel["InferGenerators: input"]@xplist]},
         InferGeneratorsOperator[variables][glc]
-    ];
+    ]//QuietEcho;
 
 InferGeneratorsOperator[variables_Association][xp_] :=
     Kleisli[InferGenerators][variables][xp];
@@ -206,16 +206,16 @@ AutoReduceStepOperator[variables_][polylist_] :=
 AutoReduceStep[variables_][polylist_List] :=
     Module[ {generators},
         generators = 
-         Lookup[variables, "generators", 
-          InferGeneratorsOperator[variables][polylist]];
+         EchoLabel["AutoReduceStep: generators"]@Lookup[variables, "generators", 
+          InferGeneratorsOperator[variables][EchoLabel["AutoReduceStep: input"]@polylist]];
         With[ {ps = 
-           ReverseSortBy[First[MonomialList[#, generators]] &]@polylist},
+           EchoLabel["AutoReduceStep: ps"]@ReverseSortBy[First[MonomialList[#, generators]] &]@polylist},
             Module[ {unrefined},
                 unrefined = 
-                 DeleteCases[0]@
                   Table[
-                   PiecewisePolynomialReduceRemainderOperator[variables][ps[[i]],
+                   PiecewisePolynomialReduceRemainderOperator[Append["generators"->generators]@variables][ps[[i]],
                      ps[[i + 1 ;;]]], {i, 1, Length@ps}];
[…]

### 2023-05-04 PDESymbolics2022/PiecewiseTools.m
Update PiecewiseTools.m
We can pass the reduce/resolve/first to PiecewiseBeautify through PiecewiseBeautifyOperator

@@ -158,7 +158,7 @@ ConditionalPiecewiseMap[F_,XP_] :=
     ]
 
 Clear[PiecewiseBeautify];
-Options[PiecewiseBeautify] = {"domain"->Complex};
+Options[PiecewiseBeautify] = {"domain"->Complex, "reduce"->Resolve};
 
 
 (*keeping to check one alteration
@@ -178,9 +178,9 @@ PiecewiseBeautify[P_,OptionsPattern[]] :=
 PiecewiseBeautify[P_,OptionsPattern[]] :=
     Which[
         Head[P] === Piecewise,
-        With[ {pp = PiecewiseLastCaseClean[P]},
+        With[ {pp = PiecewiseLastCaseClean[P],reduce = OptionValue["reduce"]},
             Module[ {qq},
-                qq = Simplify@EchoLabel["PiecewiseBeautify:  qq"]@MapThread[BooleanConvert@(*Reduce*)Resolve[!#1&&#2,OptionValue["domain"]]&, {FoldList[Or,False, pp[[All,2]]][[;;-2]],pp[[All,2]]}];
+                qq = Simplify@EchoLabel["PiecewiseBeautify:  qq"]@MapThread[BooleanConvert@reduce[!#1&&#2,OptionValue["domain"]]&, {FoldList[Or,False, pp[[All,2]]][[;;-2]],pp[[All,2]]}];
                 Piecewise[Select[Transpose[{pp[[All,1]], qq}], #[[1]]=!=$Failed &], $Failed]
             ]
         ],
@@ -192,7 +192,7 @@ PiecewiseBeautifyOperator[variables_][xp_] :=
     Module[ {facts = Lookup[variables, "facts", True]}, 
         (*Print["piecewiseBeautifyOperator domain: ", Lookup[variables,"domain",Complex]]*)
         Assuming[facts,
-            PiecewiseBeautify[xp,"domain"->Lookup[variables,"domain",Complex]]
+            PiecewiseBeautify[xp,"domain"->Lookup[variables,"domain",Complex], "reduce"->Lookup[variables, "reduce", Reduce]]
         ]
     ];
 

### 2023-05-04 PDESymbolics2022/ParametricGroebnerBasis.m
Update ParametricGroebnerBasis.m
GropOp accepts reduce as Reduce,Resolve or First.

ComprehensiveGroebnerSystemOperator is the Kleisli of GrobOp.

@@ -31,6 +31,9 @@ PiecewisePolynomialReduceRemainderOperator::usage =
 PiecewiseApplyConditionOperator::usage =
 "PiecewiseApplyConditionOperator[variables][px] Simplifies the expressions with conditions as Assumptions";
 
+ComprehensiveGroebnerSystemOperator::usage =
+"ComprehensiveGroebnerSystemOperator[variables][preGrobner] returns a piecewise expression with groebner basis for each set of conditions.";
+
 GrobOp::usage =
 "GrobOp[variables][preGrobner] returns a piecewise expression with groebner basis for each set of conditions.";
 
@@ -281,6 +284,9 @@ PiecewiseApplyConditionOperator[variables_][px_Piecewise] :=
    
 Clear[GrobOp];
 
+ComprehensiveGroebnerSystemOperator[variables_][ideal_]:=
+Kleisli[GrobOp][variables][ideal];
+
 hasFailed[x_] :=
     MemberQ[$Failed][x];
 
@@ -394,14 +400,15 @@ GrobOp[variables_][preGrobner_List, sPolynomials_List] :=
     ];
 
 GrobOp[variables_][preGrobner_List] :=
-    Module[ {newPreGrobner, sPolynomials, newArgs, facts ,newVariables, generators},
+    Module[ {newPreGrobner, sPolynomials, newArgs, facts ,newVariables, generators,reduce},
         facts = Lookup[variables, "facts", True];
+        reduce = Lookup[variables, "reduce", Resolve];
         If[Lookup[variables,"VarDOperator",VarDOperator] === VarDOperator,
         generators = Lookup[variables, "generators", InferGeneratorsOperator[variables][preGrobner]],
         (*if "VarDOperator" is DVarDOperator*)
         generators = Lookup[variables, "generators", DiscreteInferGeneratorsOperator[variables][preGrobner]]
         ];
-        newVariables = Append["generators"->generators]@variables;
+        newVariables = Append["reduce"->reduce]@Append["generators"->generators]@variables;
         If[ facts === False,
             $Failed,
             newPreGrobner = EchoLabel["monic preGrobner"]@ MonicOperator[newVariables][preGrobner];

### 2023-05-05 PDESymbolics2022/Discrete.m
Updated grobner operator in discrete.m

@@ -37,6 +37,14 @@ FindDiscreteConservedQuantityOperator::usage =
 FindDiscreteConservedQuantityBasisOperator::usage =
 	"FindDiscreteConservedQuantityBasisOperator[variables][<|\"degree\"->degree,\"generators\"->generators|>] finds a monomial basis for conserved quantities made from the generators up to the total degree";
 
+
+(* debug *)
+
+VariableEliminationOperator::usaga ="";
+
+
+
+
 Begin["`Private`"] (* Begin Private Context *) 
 
 RangeSchemeTranslationsOperator[variables_Association][masterstencil_,stencil_] :=
@@ -365,12 +373,48 @@ Reduction[variables_Association][schemeexpression_Association] :=
  
  (*from here *)
  
+ (* Friedmann code *)
+ (*
   scheme = 
    GroebnerBasis[polylist, polyvars, 
     CoefficientDomain -> RationalFunctions];
+    
+    *)
+    
+  (* DG code *)  
+    (*
+    Print["original scheme = ", schexp["scheme"]];
+    Print["polylist = ",polylist];
+    Print["polyvars = ",polyvars];
+    *)
+  scheme=GrobOp[Append[variables, "generators"->polyvars]][polylist];  
+    
+   (*
+    Print["polyscheme = ", scheme]; 
+    *)
+    
+  (* Friedmann code *)
+ (*  
   normalred = 
    PolynomialReduce[schexp["exp"], scheme, polyvars, 
      CoefficientDomain -> RationalFunctions] // Last;
+     
+      *)
+    
+  (* DG code *)  
+    
+    
+    normalred= PiecewisePolynomialReduceRemainderOperator[Append[variables, "generators"->polyvars]][schexp["exp"],scheme];
+     
+     (*
+    Print["original scheme = ", schexp["scheme"]];
+    Print["polylist = ",polylist];
+    Print["polyvars = ",polyvars];
+    Print["polyscheme = ", scheme]; 
+    Print["normalred = ", normalred]; 
+     *)
+     
+     
    normalred = Association["exp"->normalred,"scheme"->scheme]//PiecewiseExpandAssociation;
    (* set up list of lists with elimination order, i.e. 
   order of variables in which they are eliminated , 
@@ -432,10 +476,18 @@ VariableEliminationOperator[variables_Association][
       	ordermatrix = elimMatrix[n1,n2];
       	
       (* from here *)	
-      	
[…]

### 2023-05-07 Papers/Paper ACM/paper.tex
paper changes

@@
 \usepackage{amsmath}
-%\usepackage{amssymb}
+\usepackage{amsfonts}
 \usepackage{amsthm}
@@
 Conserved quantities correspond to conservation laws, admitted by the PDE. A conservation law \cite{cheviakov20} is an equation of the form
+
 \begin{gather*}
@@
 \section{The general case}
+\label{secgen}
 In this section, we present Algorithm \ref{alg: discreteconservedqimplicit} that deals with general, not necessarily time-explicit, schemes.

### 2023-05-07 Papers/Paper CASC format/paper.tex
paper changes

--- 2023-05-07 Papers/Paper ACM/paper.tex
@@
-\documentclass[manuscript]{acmart}
-
-\renewcommand{\labelenumi}{\theenumi.}
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.21 of 2022/01/12
+%
+\documentclass[runningheads]{llncs}
+%
+\usepackage[T1]{fontenc}
+% T1 fonts will be used to generate the final print and online PDFs,
+% so please use T1 fonts in your manuscript whenever possible.
+% Other font encondings may result in incorrect characters.
+%
+\usepackage{graphicx}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following two lines
+% to display URLs in blue roman font according to Springer's eBook style:
+%\usepackage{color}
+%\renewcommand\UrlFont{\color{blue}\rmfamily}
+%
 
@@
 \usepackage{amsfonts}
-\usepackage{amsthm}
 \usepackage[latin1]{inputenc}
@@
 \usepackage{mmacells}
-\usepackage[invisible]{MLaTex}
-
-%\usepackage[displaymath,mathlines]{lineno}
-%\linenumbers
-%\pagewiselinenumbers
-%\modulolinenumbers[1]
-
-%permite que formulas
-%se dividam por duas
-%paginas diferentes
+\usepackage[invisible]{MLaTeX}
+
 \allowdisplaybreaks
 
-%[hypertex]
-
-%\usepackage{showkeys}
-%\usepackage[norefs]{refcheck}
-
-\makeindex
-
-%\usepackage{curves}
-
-%\newcommand{\cal}{\mathcal}
 \newcommand{\argmin}{\operatorname{argmin}}
@@
 
-\newcommand{\Rr}{{\mathbb{R}}}
-\newcommand{\Cc}{{\mathbb{C}}}
-\newcommand{\Nn}{{\mathbb{N}}}
+\newcommand{\Rr}{\mathbb{R}}
+\newcommand{\Cc}{\mathbb{C}}
+\newcommand{\Nn}{\mathbb{N}}
 \newcommand{\Zz}{{\mathbb{Z}}}
-\newcommand{\Xx}{{\mathbb{X}}}
-\newcommand{\Qq}{{\mathbb{Q}}}
-\newcommand{\Tt}{{\mathbb{T}}}
-\newcommand{\1}{{\chi}}
-
-\newcommand{\Hh}{{\overline{H}}}
-\newcommand{\ot}{{\overline{t}}}
-\newcommand{\ox}{{\overline{x}}}
-\newcommand{\oy}{{\overline{y}}}
-\newcommand{\Ll}{{\mathcal{L}}}
-
-\newcommand{\Aa}{{\mathcal{A}}}
-\newcommand{\Cg}{{\mathcal{C}}}
-\newcommand{\Ff}{{\mathcal{F}}}
-\newcommand{\Gg}{{\mathcal{G}}}
-\newcommand{\Ii}{{\mathcal{I}}}
-\newcommand{\Dd}{{\mathcal{D}}}
-\newcommand{\Uu}{{\mathcal{U}}}
-\newcommand{\Vv}{{\mathcal{V}}}
-\newcommand{\Ww}{{\mathcal{W}}}
-\newcommand{\Mm}{{\mathcal{M}}}
-
-\newcommand{\sfd}{{$\sigma$-field }}
-\newcommand{\sfs}{{$\sigma$-fields }}
-\newcommand{\cqd}{\hfill $\blacksquare$}
-\newcommand{\cqda}{\hfill $\rule{1.5mm}{1.5mm}$}
-\newcommand{\cqdb}{\hfill $\rule{0.75mm}{0.75mm}$}
-\newcommand{\ee}{\hfill $\blacktriangleleft$}
+\newcommand{\Xx}{\mathbb{X}}
+\newcommand{\Qq}{\mathbb{Q}}
+\newcommand{\Tt}{\mathbb{T}}
+\newcommand{\1}{\chi}
+
+\newcommand{\Hh}{\overline{H}}
+\newcommand{\ot}{\overline{t}}
+\newcommand{\ox}{\overline{x}}
+\newcommand{\oy}{\overline{y}}
+\newcommand{\Ll}{\mathcal{L}}
+
+\newcommand{\Aa}{\mathcal{A}}
+\newcommand{\Cg}{\mathcal{C}}
+\newcommand{\Ff}{\mathcal{F}}
+\newcommand{\Gg}{\mathcal{G}}
+\newcommand{\Ii}{\mathcal{I}}
+\newcommand{\Dd}{\mathcal{D}}
+\newcommand{\Uu}{\mathcal{U}}
+\newcommand{\Vv}{\mathcal{V}}
+\newcommand{\Ww}{\mathcal{W}}
+\newcommand{\Mm}{\mathcal{M}}
+
 \newcommand{\eef}{\quad\blacktriangleleft}
-\newcommand{\bx}{{\bf x}}
-\newcommand{\bu}{{\bf u}}
-\newcommand{\bdx}{\dot{\bf x}}
-\newcommand{\bp}{{\bf p}}
-\newcommand{\bdp}{\dot{\bf p}}
-\newcommand{\bX}{{\bf X}}
-\newcommand{\bdX}{\dot{\bf X}}
-\newcommand{\bP}{{\bf P}}
-\newcommand{\bdP}{\dot{\bf P}}
-\newcommand{\by}{{\bf y}}
-\newcommand{\bdy}{\dot{\bf y}}
-\newcommand{\bv}{{\bf v}}
-\newcommand{\bdv}{\dot{\bf v}}
-\newcommand{\bw}{{\bf w}}
-\newcommand{\bdw}{\dot{\bf w}}
-
-\newcommand{\Pf}{{\noindent \sc Proof. \ }}
-\newcommand{\Rm}{{\noindent \sc Note. \ }}
-
-\newcommand{\cws}{{\overset{*}{\rightharpoonup}}}
+\newcommand{\bx}{\boldsymbol{x}}
+\newcommand{\bu}{\boldsymbol{u}}
+\newcommand{\bdx}{\dot{\boldsymbol{x}}}
+\newcommand{\bp}{\boldsymbol{p}}
+\newcommand{\bdp}{\dot{\boldsymbol{p}}}
+\newcommand{\bX}{\boldsymbol{X}}
+\newcommand{\bdX}{\dot{\boldsymbol{X}}}
+\newcommand{\bP}{\boldsymbol{P}}
+\newcommand{\bdP}{\dot{\boldsymbol{P}}}
+\newcommand{\by}{\boldsymbol{y}}
+\newcommand{\bdy}{\dot{\boldsymbol{y}}}
+\newcommand{\bv}{\boldsymbol{v}}
+\newcommand{\bdv}{\dot{\boldsymbol{v}}}
+\newcommand{\bw}{\boldsymbol{w}}
+\newcommand{\bdw}{\dot{\boldsymbol{w}}}
+
+\newcommand{\cws}{\overset{*}{\rightharpoonup}}
 
 \newcommand{\epsi}{\varepsilon}
-\def\d{{\rm d}}
-\def\dx{{\rm d}x}
-\def\dy{{\rm d}y}
-\def\da{{\rm d}a}
-\def\dt{{\rm d}t}
-%\def\leq{\leqslant}
-%\def\geq{\geqslant}
-
-\def\appendixname{\empty}
-
-\numberwithin{equation}{section}
-
-\newtheoremstyle{thmlemcorr}{10pt}{10pt}{\itshape}{}{\bfseries}{.}{10pt}{{\thmname{#1}\thmnumber{
-#2}\thmnote{ (#3)}}}
-\newtheoremstyle{thmlemcorr*}{10pt}{10pt}{\itshape}{}{\bfseries}{.}\newline{{\thmname{#1}\thmnumber{
-#2}\thmnote{ (#3)}}}
-\newtheoremstyle{defi}{10pt}{10pt}{\itshape}{}{\bfseries}{.}{10pt}{{\thmname{#1}\thmnumber{
-#2}\thmnote{ (#3)}}}
-\newtheoremstyle{remexample}{10pt}{10pt}{}{}{\bfseries}{.}{10pt}{{\thmname{#1}\thmnumber{
-#2}\thmnote{ (#3)}}}
-\newtheoremstyle{ass}{10pt}{10pt}{}{}{\bfseries}{.}{10pt}{{\thmname{#1}\thmnumber{
-A#2}\thmnote{ (#3)}}}
-
-\theoremstyle{thmlemcorr}
-\newtheorem{theorem}{Theorem}
-\numberwithin{theorem}{section}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{conjecture}[theorem]{Conjecture}
-
-\theoremstyle{thmlemcorr*}
-\newtheorem{theorem*}{Theorem}
-\newtheorem{lemma*}[theorem]{Lemma}
-\newtheorem{corollary*}[theorem]{Corollary}
-\newtheorem{proposition*}[theorem]{Proposition}
-\newtheorem{problem*}[theorem]{Problem}
-\newtheorem{conjecture*}[theorem]{Conjecture}
-
-\theoremstyle{defi}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{hyp}{Assumption}
-\newtheorem{claim}{Claim}
-\newtheorem{problem}{Problem}
-
-\theoremstyle{remexample}
-\newtheorem{remark}[theorem]{Remark}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{counterexample}[theorem]{Counterexample}
-
-\newtheorem{teo}[theorem]{Theorem}
-\newtheorem{lem}[theorem]{Lemma}
-\newtheorem{pro}[theorem]{Proposition}
-\newtheorem{cor}[theorem]{Corollary}
-
-%\def\contentsname{}
-\def\refname{References}
-\def\bibname{References}
-\def\chaptername{\empty}
-\def\figurename{Fig.}
-\def\abstractname{Abstract}
-
-\theoremstyle{ass}
-
-%\theoremstyle{definition}
-
-\newtheorem{exemp}{Exemplo}
+\newcommand{\dx}{\mathrm{d}x}
+\newcommand{\dy}{\mathrm{d}y}
+\newcommand{\da}{\mathrm{d}a}
+\newcommand{\dt}{\mathrm{d}t}
 
 \begin{document}
-
-\title[Detection of conserved quantities for PDE schemes]{Algorithmic detection of conserved quantities of finite-difference schemes for partial differential equations}
-
-\author{Diogo A. Gomes}
-\affiliation{%
-	\institution{King Abdullah University of Science and Technology (KAUST)}
-	\department[0]{AMCS/CEMSE Division}
-	\city{Thuwal}
-	\postcode{23955-6900}
-	\country{Saudi Arabia}}
-\email{diogo.gomes@kaust.edu.sa}
-\author{Friedemann Krannich}
-\affiliation{%
-	\institution{King Abdullah University of Science and Technology (KAUST)}
-	\department[0]{AMCS/CEMSE Division}
-	\city{Thuwal}
-	\postcode{23955-6900}
-	\country{Saudi Arabia}}
-\email{friedemann.krannich@kaust.edu.sa}
-\author{Ricardo de Lima Ribeiro}
-\affiliation{%
-	\institution{King Abdullah University of Science and Technology (KAUST)}
-	\department[0]{AMCS/CEMSE Division}
-	\city{Thuwal}
-	\postcode{23955-6900}
-	\country{Saudi Arabia}}
-\email{ricardo.ribeiro@kaust.edu.sa}
-
-\keywords{Symbolic computations; Finite-difference schemes; Discrete variational derivative; Discrete partial variational derivative; Conserved quantities; Implicit schemes; Explicit schemes.}
-
-\begin{CCSXML}
-	<ccs2012>
-	<concept>
-	<concept_id>10002950.10003714.10003727.10003729</concept_id>
-	<concept_desc>Mathematics of computing~Partial differential equations</concept_desc>
-	<concept_significance>500</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10002950.10003714.10003715.10003750</concept_id>
-	<concept_desc>Mathematics of computing~Discretization</concept_desc>
-	<concept_significance>500</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10002950.10003714.10003715.10003720.10003747</concept_id>
-	<concept_desc>Mathematics of computing~Gr{\"o}bner bases and other special bases</concept_desc>
-	<concept_significance>500</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10010147.10010148.10010149.10010152</concept_id>
-	<concept_desc>Computing methodologies~Symbolic calculus algorithms</concept_desc>
-	<concept_significance>500</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10010147.10010148.10010149.10010155</concept_id>
-	<concept_desc>Computing methodologies~Discrete calculus algorithms</concept_desc>
-	<concept_significance>500</concept_significance>
-	</concept>
-	</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Mathematics of computing~Partial differential equations}
-\ccsdesc[500]{Mathematics of computing~Discretization}
-\ccsdesc[500]{Mathematics of computing~Gr{\"o}bner bases and other special bases}
-\ccsdesc[500]{Computing methodologies~Symbolic calculus algorithms}
-\ccsdesc[500]{Computing methodologies~Discrete calculus algorithms}
-
-\date{\today}
-
+%
+\title{Algorithmic detection of conserved quantities for finite-difference schemes\thanks{Supported by King Abdullah University of Science and Technology (KAUST) baseline funds and KAUST OSR-CRG2021-4674.}}
+%
+%\titlerunning{Abbreviated paper title}
+%
+\author{Diogo A. Gomes\inst{1}\orcidID{0000-0002-3129-3956} \and
+	Friedemann Krannich\inst{1}\orcidID{1111-2222-3333-4444} \and
+	Bashayer Majrashi\inst{1}\orcidID{0009-0008-3677-8064 } \and
+	Ricardo de Lima Ribeiro\inst{1}\orcidID{0000-0002-2707-2210}}
+%
+\authorrunning{D.A. Gomes et al.}
+%
+\institute{King Abdullah University of Science and Technology (KAUST), AMCS/CEMSE Division, Thuwal 23955-6900, Saudi Arabia\
+	\email{{diogo.gomes,friedemann.krannich,bashayer.majrashi, ricardo.ribeiro}@kaust.edu.sa}}
+%
+\maketitle
+%
 \begin{abstract}
 	Many partial differential equations (PDEs) admit conserved quantities like mass or energy.
-Those quantities are often essential to establish well-posed results.
-When approximating a PDE by a finite-difference scheme, it is natural to ask whether related discretized quantities remain conserved over time by the scheme.
-Such conservation may establish the stability of the numerical scheme.
-We present an algorithm for checking the preservation of a polynomial quantity under a polynomial finite-difference scheme.
+These quantities are often essential for establishing well-posedness results.
+When approximating a PDE with a finite-difference scheme, it is crucial to determine whether related discretized quantities remain conserved over time by the scheme.
+Such conservation may ensure the stability of the numerical scheme.
+We present an algorithm for verifying the preservation of a polynomial quantity under a polynomial finite-difference scheme.
 Our schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables.
-Additionally, we present an algorithm for, given a scheme, finding conserved quantities. 
+Additionally, we introduce an algorithm for finding conserved quantities, given a scheme.
 We illustrate our algorithm with several finite-difference schemes.
+Our approach incorporates parametric G\"obner bases to handle parameters, ensuring accurate computation of conserved quantities.
+
+	\keywords{Symbolic computations \and Finite-difference schemes \and Discrete variational derivative \and Discrete partial variational derivative \and Conserved quantities \and Implicit schemes \and Explicit schemes \and Parametric Groebner basis.}
 \end{abstract}
-
-\maketitle
-\tableofcontents
-
-\begin{acks}
-	The authors were supported by \grantsponsor{GN01}{King Abdullah University of Science and Technology (KAUST) baseline funds}{} and 
-	\grantsponsor{GN02}{KAUST OSR-CRG2021-4674}{}.
-\end{acks}
 
@@
 %\label{intro}
-Many partial differential equations (PDEs) admit integral quantities, that are conserved in time. 
+In many partial differential equations (PDEs), conserved integral quantities play a crucial role, as they often represent properties of physical significance, such as energy or mass.
 For example, the advection equation preserves energy and the heat equation conserves mass.
-When approximating PDEs by a finite-difference scheme, the question arises whether such quantities are conserved by the scheme. 
-Such information can be crucial to estimate whether a scheme approximates a PDE accurately and to determine its stability.\\
-Computations for determining conservation can get rather tedious.
-Hence, automating them in computer algebra systems is desirable.\\
+Since closed-form solutions for these PDEs are often unavailable, finite-difference schemes serve as a practical numerical alternative.
+Hence, it becomes imperative to determine whether the chosen scheme preserves these conserved quantities.
+This information is not only valuable for gauging the accuracy of the approximation, but also vital for evaluating its stability.
+Automating conservation computations in computer algebra systems streamlines the process, saves time, enhances accuracy, and may create new research opportunities.
+
+{\color{red} $\Phi$ and $\Psi$ need domains and ranges, also explain that divergence is only in $x$}
+
 Conserved quantities correspond to conservation laws, admitted by the PDE. A conservation law \cite{cheviakov20} is an equation of the form
-
-\begin{gather*}
-	D_t\Phi[u]+D_x\Psi[u]=0
+\begin{gather*}
+	\frac{\partial }{\partial t}\Phi[u]+\div \Psi[u]=0
 \end{gather*}
@@
 \end{gather*}
-assuming periodic boundary conditions in $x$.\\
-An analog formulation describes conservation laws for schemes of PDEs, where the derivatives are replaced by shifts \cite{hydon01}.
-The key for the construction of schemes, that admit certain conservation laws from the continuous PDE, is the use of the discrete Euler operator \cite{cheviakov20} (also called discrete variational derivative).
+assuming periodic boundary conditions in $x$.
+
+There is a similar formulation for
+for conservation laws for finite difference schemes of a PDE  where the derivatives in the preceding identity are replaced by shifts \cite{hydon01}.
+.
+As a result, standard approaches for constructing schemes with conservation involve either discretizing the continuous conservation law or finding multipliers for the scheme \cite{dorodnitsyn21}.
+Alternatively, 
+to verify that a particular scheme admits conservation laws, we can use the  discrete Euler operator \cite{cheviakov20}.
 Kupershmidt \cite[II.
-Theorem 31]{kupershmidt85} discovered, that an equation is a discrete conservation law if and only if it belongs to the kernel of the discrete Euler operator.
-Hence, standard approaches for constructing schemes with conservation are either discretizing the continuous conservation law or finding multipliers for the scheme \cite{dorodnitsyn21}, such that the result is in the kernel of the discrete Euler operator.
-This strategy was, for example, recently used by Dorodnitsyn et al.
-to develop a scheme for the shallow water equation, that preserves energy \cite{dorodnitsyn21}.
-Cheviakov et al.
-used this idea to find schemes for the linear and nonlinear wave equation, that admit several discrete analogs of continous conservation laws \cite{cheviakov20}.
-The same strategies for finding conservation laws, that are described above, work for continuous PDEs or semi-discretized PDEs:
-Algorithms for finding conserved quantities, using multipliers and the Euler operator for PDEs were developed and implemented by Cheviakov \cite{cheviakov07,cheviakov10}.
-Hereman et al. \cite{hereman04,hereman08}, proposed an algorithm to compute conserved densities for semi-discretized schemes for PDEs with first-order time derivative. 
-Their algorithm uses the scaling symmetries of the scheme to construct conserved quantities and calculates their coefficients using the discrete Euler operator.
-Gao et al. extended this algorithm to first-order time-explicit schemes \cite{gao02} and first-order time-implicit schemes \cite{gao04}.\\
+Theorem 31]{kupershmidt85} discovered that an equation represents a discrete conservation law if and only if it is in the kernel of the discrete Euler operator
+This strategy has been employed in recent studies, such as by Dorodnitsyn et al.
+for developing a shallow water equation scheme that preserves energy \cite{dorodnitsyn21}, and by Cheviakov et al.
+for creating linear and nonlinear wave equation schemes that admit discrete analogs of continuous conservation laws \cite{cheviakov20}.
+These strategies also apply to continuous or semi-discretized PDEs.
+Algorithms using multipliers and the Euler operator for PDEs were developed and implemented by Cheviakov \cite{cheviakov07,cheviakov10}.
+Hereman et al.
+\cite{hereman04,hereman08} proposed an algorithm for computing conserved densities for semi-discretized PDE schemes with first-order time derivatives, using scaling symmetries and the discrete Euler operator.
+Gao et al.
+extended this algorithm to first-order time-explicit \cite{gao02} and time-implicit schemes \cite{gao04}.
 Bihlo et al.
-\cite{bihlo12} used another approach and derived finite-difference schemes for the shallow water equations, that are symmetry-invariant on an adaptive mesh and then analyzed them regarding conserved quantities.\\
-In this paper, we propose an algorithm that checks if a quantity is conserved in time under a scheme.
-Our approach differs from the ideas described above, as we do not try to construct discrete $\Phi$ and $\Psi$, using the scheme, but we check if a given discrete $\Phi$ is conserved in time under a given scheme.\\
-Conserved quantities usually involve integral expressions or their discrete analog, sums.
-Gomes et al. proposed algorithms for the simplification of sums \cite{gomes20}.
-They also developed and implemented algorithms for detecting quantities conserved by PDEs and semi-discretized schemes. 
-We revise these techniques in Section \ref{assp}. 
-To generalize those methods to situations where we do not sum over all arguments of the functions involved, we introduce the discrete partial variational derivative (Section \ref{pvard}).
-The main contribution of this paper is an algorithm for checking the conservation of a quantity under a numerical scheme.
-Gerdt showed \cite{gerdt12}, that the quantity is conserved, if its discrete time derivative belongs to the difference ideal generated by the scheme.
-However, some quantities may add to a constant (e.g.
-telescopic sums) and thus be trivially preserved without belonging to the difference ideal.
-Moreover, Gerdt's algorithm may not terminate, as the Gr{\"o}bner basis for the difference ideal may not be finite.
-We overcome these issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) with finite Gr{\"o}bner basis (Section \ref{algodetec}). 
-Our algorithm works for schemes that are explicit and implicit in time and can treat schemes with several higher-order time  and space derivatives, and with several space dimensions.
-Additionally, we can handle systems of equations and schemes with parameters. 
-We have implemented this algorithm as part of a package in {\sc Mathematica} \cite{wolfram21} \footnote{The code is available upon request. For access contact ricardo.ribeiro@kaust.edu.sa.}.
-In the examples, we show that our code finds conserved quantities and proper schemes for the time-implicit and time-explicit discretization of the Burgers equation and a system of PDEs arising in the study of mean-field games (Section \ref{eaa}).
+\cite{bihlo12} used another approach and derived finite-difference schemes for the shallow water equations, that are symmetry-invariant on an adaptive mesh and then analyzed them regarding conserved quantities.
+
+In this paper, we present an algorithm that verifies if a quantity is conserved over time under a given scheme.
+Our approach diverges from previous methods, as we do not attempt to construct discrete $\Phi$ and $\Psi$ using the scheme, but rather check if a given discrete $\Phi$ is conserved over time under the scheme.
+Typically, conserved quantities involve integral expressions or their discrete analogs, such as sums.
+Gomes et al.
+proposed algorithms for simplifying sums \cite{gomes20} and developed algorithms for detecting conserved quantities in PDEs and semi-discretized schemes \cite{gomes20}.
+We review these techniques in Section \ref{assp}.
+To generalize these methods for cases without summing over all arguments of the involved functions, we introduce the discrete partial variational derivative (Section \ref{pvard}).
+Our main contribution is an algorithm for verifying the conservation of a quantity under a numerical scheme.
+Gerdt \cite{gerdt12} demonstrated that the quantity is conserved if its discrete time derivative belongs to the difference ideal generated by the scheme.
+However, some quantities may add only to a constant and be trivially preserved without being part of the difference ideal.
+Furthermore, Gerdt's algorithm may not always terminate, as the Gr{\"o}bner basis for the difference ideal might be infinite.
+We address these issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) having a finite Gr{\"o}bner basis (Section \ref{algodetec}).
+Our algorithm is applicable to schemes that are explicit and implicit in time, can handle schemes with multiple higher-order time and space derivatives, multiple space dimensions, systems of equations, and schemes with parameters.
+As the process of finding conserved quantities involves polynomial division, special care must be taken to ensure that the division is performed correctly with respect to the parameters.
+We employ parametric Gr{\"o}bner  bases, which are crucial for the proper treatment of parameters in our algorithm, see Section \ref{pgb}
+ We have implemented this algorithm as a part of a {\sc Mathematica} package \cite{wolfram21}\footnote{The code is available upon request.
+For access, contact ricardo.ribeiro@kaust.edu.sa.}.
+Our examples include Burgers equation and a PDE system related to mean-field games (Section \ref{eaa}).
+In these examples,
+ we show that our code detects conserved quantities and can be used to select suitable schemes for time-implicit and time-explicit discretizations.
 
@@
 \label{assp}
-Here, we follow the ideas in \cite{gomes20}.
-In this paper, subscripts denote indices related to coordinates or tuples and superscripts denote indices related to sequences.
+
+In this section, we introduce the concept of discrete torus and discrete variational derivatives, and discuss their properties and applications in simplifying sums. 
+These definitions are essential for the analysis of methods presented in the rest of the paper.
+Here, we 
+discuss the connection between discrete functionals and discrete variational derivatives and present examples that illustrate how to simplify sums using this approach.
+We follow the ideas in \cite{gomes20}.
+In this paper, subscripts denote coordinate or tuple indices, and superscripts denote sequence indices.
 To avoid boundary terms, we work with periodic functions in $\Zz^d$.
-\begin{definition}
-	Let $N$, $d$ and $m$ be positive integers. The \emph{discrete torus} is $\Ii:=\{0,\ldots,N-1\}^d\subset \Zz^d$. Define the space
-	\begin{gather*}
-		\mathcal{P}(\Ii,\Rr^m):=\{u\ |\ u:\Ii\rightarrow\Rr^m\},
-	\end{gather*}
-	extended to $\Zz^d$ by periodicity.
-\end{definition}
-%Assuming periodicity allows us to extend the functions in $\mathcal{P}(\Ii)$ to $\Zz^d$, and we avoid discussing boundary terms.
-\begin{definition}
+
+When discretizing PDEs using finite difference schemes, we need to work with functions defined on discrete sets.
+For instance, periodic functions in $\Rr^d$ can be replaced by functions defined on the discrete torus, which will be introduced in the following discussion.
+\begin{definition}
+	\label{d1}
+	Let $N$, $d$ and $m$ be positive integers. The \emph{discrete torus} is the set $\Ii:=\{0,\ldots,N-1\}^d\subset \Zz^d$. Define the space
+	\begin{gather*}
+		\mathcal{P}(\Ii,\Rr^m):=\{u\ |\ u:\Ii\rightarrow\Rr^m\}.
+	\end{gather*}
+Functions in $\mathcal{P}(\Ii,\Rr^m)$ extend periodically to $\Zz^d$.
+\end{definition}
+We introduce discrete functionals in the previous function space. These functionals are the analog of integral functionals for discrete approximations of functions.
+\begin{definition}
+	\label{d2}
 	The space of functionals $\mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$ (not necessarily linear) on $\mathcal{P}(\Ii,\Rr^m)$ is
@@
 \end{definition}
-In the previous definition, the $e^1,\ldots,e^k$ are not necessarily unit vectors and can vary between functionals.
+In the previous definition, $e^1,\ldots,e^k$ are not necessarily unit vectors and may vary between functionals.
 \begin{example}
@@
 \end{example}
-To simplify sums, the discrete variational derivative is a useful tool.
+The discrete variational derivative is a useful tool for simplifying sums.
 Let $u,v\in\mathcal{P}(\Ii,\Rr^m)$, $\epsilon 
@@
 \end{gather*}
-	Below in \eqref{periodicity}, we shift the indices due to periodicity of $u$ and $v$,
+In \eqref{periodicity}, we shift the indices because of the periodicity of $u$ and $v$,
 \begin{gather}
@@
 \end{gather}
+The operator $\mathcal{V}(\mathcal{J})$ can be viewed as a representation of the derivative $D\mathcal{J}u$ with respect to the $L^2$ inner product $\langle w, v\rangle=\sum_{n\in \Ii} w_n v_n$, leading to the following definition.
 \begin{definition}
@@
 \end{definition}
-Because we assume all related functions to be smooth, the discrete variational derivative always exists.
-\begin{example}
-	Let us return to Example \ref{functional1} and compute
+Because we assume all related functions are smooth, the discrete variational derivative always exists.
+\begin{example}
+Now, we return to Example \ref{functional1} and compute
 	\begin{gather*}
@@
 		=\sum_{n\in\Ii}\Big(-2u(n+1)+4u(n)-2u(n-1)\Big)v(n)\\
-		=\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n)
-	\end{gather*}
-	and hence $\mathcal{V}(\mathcal{J})[u](n)=-2u(n+1)+4u(n)-2u(n-1)$.
-\end{example}
-The algorithms for the simplification of sums presented in \cite{gomes20} rely on the following result:
-\begin{teo}\label{theo1}
+		=\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n).
+	\end{gather*}
+Thus, $\mathcal{V}(\mathcal{J})[u](n)=-2u(n+1)+4u(n)-2u(n-1)$.
+\end{example}
+The algorithms for the simplification of sums presented in \cite{gomes20} relies on the following result, which 
+provides a criterion to compare functionals and evaluate if they represent the same quantity.
+
+\begin{theorem}\label{theo1}
 	Let $\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$.
@@
 	for all $u\in\mathcal{P}(\Ii,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
-\end{teo}
+\end{theorem}
 \begin{proof}
+	We have
 	\begin{gather*}
@@
 \end{proof}
-We use Theorem \ref{theo1} to examine if different sums represent the same quantity.
+We use Theorem \ref{theo1} to determine whether different sums represent the same quantity.
 \begin{example}
@@
 	\end{gather*}
-	It is clear, that $\mathcal{J}$ and $\tilde{\mathcal{J}}$ represent the same quantity (by shifting $n$). We confirm this, using the discrete variational derivative:
+	It is clear that $\mathcal{J}$ and $\tilde{\mathcal{J}}$ represent the same quantity (by shifting $n$). We confirm this, using the discrete variational derivative:
 	\begin{gather*}
@@
 	\end{gather*}
-	Hence, $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})=0$ and $\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]$. Therefore, both functionals sum to the same quantity.
+	Hence, $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})=0$ and $\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]$. Therefore, both functionals represent the same quantity.
 \end{example}
@@
 	\end{gather*}
-	We confirm this, by computing the discrete variational derivative and noticing that the expression is $0$ for $u_0=0$.
-\end{example}
+	We confirm this by computing the discrete variational derivative and noticing that the expression is $0$ for $u=0$.
+\end{example}
+
 \section{The discrete partial variational derivative}
 \label{pvard}
-Here, we generalize the discrete variational derivative for situations where we keep one (or several) of the arguments of $u$ constant.
+
+In this section, we introduce the concept of the discrete partial variational derivative, a generalization of the discrete variational derivative for cases where one or several arguments of the indices are kept constant.
+This generalization provides a more flexible approach to handling various problems in conservation laws for finite-difference schemes, addressing challenges and limitations of the traditional discrete variational derivative.
+The main result of this section is Theorem \ref{theo2}, which gives a criterion for equality of two functionals based on their discrete partial variational derivatives."
+
 Let $n=(n_1,\ldots,n_d)$ be the variables for the functions in $\mathcal{P}(\Ii,\Rr^m)$.
@@
 Let $l\in\Nn$ and call $t:=(t_1,\ldots,t_l)\in\Nn^l$ the time variables.
-Later in this paper, we only consider the case $l=1$.
-\begin{definition}
-	Let $\ \mathcal{P}(\Ii\times\Nn^l,\Rr^m):=\{u\ |\ u(\cdot,t)\in\mathcal{P}(\Ii,\Rr^m)\ \forall t\in\Nn^l \}$, extended to $\Zz^d\times\Nn^l$ by periodicity in the space variables and let $\mathcal{D}(\Nn^l):=\{u\ |\ u:\Nn^l\rightarrow\Rr\}$.
-\end{definition}
-When considering sums, we only sum over the space variables and not over the time variables.
+For generality, we consider the vectorial time case, while later in this paper, we focus specifically on the case when $l=1$.
+The analog to Definition \ref{d1} is the following. 
+\begin{definition}
+	Let $\mathcal{P}(\Ii\times\Nn^l,\Rr^m):=\{u\ |\ u(\cdot,t)\in\mathcal{P}(\Ii,\Rr^m)\ \forall t\in\Nn^l \}$.
+	Functions in $\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ extend to $\Zz^d\times\Nn^l$ through  periodicity in the space variables.
+	Let $\mathcal{D}(\Nn^l):=\{u\ |\ u:\Nn^l\rightarrow\Rr\}$.
+\end{definition}
+We can regard $\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ as the set of functions depending in space and time and $\mathcal{D}(\Nn^l)$ as the functions depending only on time.
+
+Next, we examine functions that involve summation only over spatial variables, extending the concept presented in Definition \ref{d2}.
 \begin{definition}
@@
 \end{definition}
+
 Here, all quantities, whose conservation in time we check, are functions $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$.
@@
 \end{gather*}
-	Let $u,v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ and compute
+Let $u,v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$, then
 \footnotesize
@@
 \begin{definition}
-	We define $\mathcal{V}^l(\mathcal{J})$ as the \emph{discrete partial variational derivative} of $\mathcal{J}$.
-	Note that $\mathcal{V}^0:=\mathcal{V}$.
+$\mathcal{V}^l(\mathcal{J})$ is the \emph{discrete partial variational derivative} of $\mathcal{J}$.
+	Note that $\mathcal{V}^0=\mathcal{V}$.
 \end{definition}
@@
 	\end{gather*}
-	equals $\delta_{t+1}(s)-\delta_t(s)$, because
+	is $\delta_{t+1}(s)-\delta_t(s)$, because
 	\small
@@
 We have a result similar to Theorem \ref{theo1} for the discrete partial variational derivative:
-\begin{teo}\label{theo2}
+\begin{theorem}\label{theo2}
 	Let $\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$.
@@
 	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
-\end{teo}
+\end{theorem}
 The proof is similar to the proof of Theorem \ref{theo1}.
+
 \section{Difference schemes and algebra}
 \label{algodetec}
-In this section, we define numerical schemes formally and introduce the tools from difference algebra, that we need for our algorithm.
-During the remainder of this paper, we assume that there is only one time variable $t$ ($l=1$).
+In this section, we formally define numerical schemes and introduce the necessary tools from difference algebra for our algorithm.
+Throughout this paper, we assume that there is only one time variable $t$ ($l=1$).
 \begin{definition}
@@
 However, our algorithms and our code handle parameters.
-Hence, one can generalize our results to schemes with other step sizes.
+Hence, our results can be generalized to schemes with other step sizes.
 \end{remark}
+\begin{example}
+	\label{ee}
+Consider the heat equation	$u_t-u_{xx}=0$ \cite{smoller94}.  A possible scheme is
+\[	
+		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t)).
+\]
+A solution of the corresponding numerical approximation is a function $u\in \mathcal{P}(\Ii\times\Nn,\Rr),$ that satisfies
+\[
+		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t))=0
+\]
+for all $n\in \Ii$ and $t\in \Nn$. 
+\end{example}
+
 \begin{definition}
@@
 \begin{example}\label{timeexplicitheateq}
-	Consider the forward-difference scheme for the heat equation $u_t-u_{xx}=0$ \cite{smoller94}
-	\begin{equation*}
-		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t)).
-	\end{equation*}
+	Consider the forward-difference scheme for the heat equation in Example \ref{ee}. 
 	This scheme is in time-explicit form with right-hand side
@@
 \end{definition}
-Now we construct the difference ideal containing the scheme:
+We now construct the difference ideal containing the scheme.
 \begin{definition}
@@
 \end{definition}
-A solution of a numerical scheme is a function $u$, that makes all translations of the scheme vanish. Hence, every element of the difference ideal generated by the scheme vanishes under $u$.
-Algorithm \ref{alg: discreteconservedqimplicit} seeks to determine if the polynomial associated with the discrete time derivative $\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$ belongs to the ideal $\langle H^1,\ldots,H^i\rangle$.
-To examine, if $\mathcal{T}\in\langle H^1,\ldots,H^i\rangle$, Gerdt proposes the notion of a standard (Gr{\"o}bner) basis for the ideal $\langle H^1,\ldots,H^i\rangle$.
-This standard basis may not be finite.
-We overcome this problem by considering polynomial instead of difference algebra and using a smaller polynomial ideal, that is contained in the difference ideal and admits a finite standard basis.
-However, there may be functions $\mathcal{J}$, whose sums add to zero, even though they may fail to belong to the difference ideal.
-Hence, we combine the discrete partial variational derivative with polynomial ideals.
+
+A solution of a numerical scheme is a function, $u$, that makes all translations of the scheme vanish.
+Hence, every element of the difference ideal generated by the scheme vanishes under $u$.
+As a result, every element of the difference ideal generated by the scheme vanishes under $u$.
+If the polynomial associated with the discrete time derivative  $\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$
+belongs to the ideal $\langle H^1,\ldots,H^i\rangle$, the corresponding functional vanishes. 
+ Gerdt introduced the concept of a standard  Gr{\"o}bner basis for the ideal $\langle H^1,\ldots,H^i\rangle$, which might not be finite.
+To address this issue, we use polynomial algebra instead of difference algebra and a smaller polynomial ideal contained within the difference ideal that has a finite standard basis.
+
+ However, ideal membership is not  a necessary condition for conservation. Some functions, $\mathcal{J}$,
+ may have sums that add to zero even if they do not belong to the difference ideal.
+Thus, we combine the discrete partial variational derivative with polynomial ideals.
+This idea is implemented in Algorithm \ref{alg: discreteconservedqimplicit} in Section \ref{secgen}.
+
+\section{Parametric Groebner basis}
+\label{pgb}
+
+TBD
+
 \section{The time-explicit case}
 \label{time-exp}
-In this section, we present Algorithm \ref{alg: discreteconservedqexplicit}  for checking the conservation of a quantity under the right-hand side of a scheme in time-explicit form.
-We discuss our algorithm for general schemes in the next section.
+In this section, we introduce Algorithm \ref{alg: discreteconservedqexplicit}  for checking the conservation of a quantity for time-explicit schemes.
+We discuss the algorithm for general schemes in the next section.
 \begin{algorithm}
@@
 \end{algorithm}
-Note, that the result of Step 2 exists only at time $t$ because all instances of $t+1$ have been replaced. Thus, we do not need any computations of difference ideals or Gr{\"o}bner basis.
+The result of Step 2 exists only at time $t$ because all instances of $t+1$ have been replaced. Thus, we do not need any computations of difference ideals or Gr{\"o}bner basis.
 \begin{example}
 	\label{timeexplicitheateqalgo}
-	We check for conservation of $\sum_nu(n,t)$ under the scheme
+	We examine the conservation of $\sum_nu(n,t)$ under the following scheme.
 	\begin{equation*}
@@
 \end{example}
+
+{\color{red}ADD HERE TIME EXPLICIT EXAMPLE WITH PARAMETERS - DISCUSS WITH RICARDO THE ISSUE}
+
+{\color{red} DG HERE}
+
 \section{The general case}
@@
 In this section, we present Algorithm \ref{alg: discreteconservedqimplicit} that deals with general, not necessarily time-explicit, schemes.
-We explain its steps in detail below and demonstrate the algorithm in Example \ref{heateqalgoimplicit}.\\
-%In this section we explain our algorithm, that checks if a given quantity is preserved in time under a given scheme in detail.
+We explain its steps in detail below and demonstrate the algorithm in Example \ref{heateqalgoimplicit}.
 \begin{algorithm}
@@
 \end{algorithm}
-{\bf Step 1: Build the discrete time derivative.}
-We build the discrete time derivative by subtracting $\mathcal{J}[u](t)$ from $\mathcal{J}[u](t+1)$. This task is done in our code by the {\sc TimeDifference}.
+
+\noindent{\bf Step 1: Build the discrete time derivative.}
+We build the discrete time derivative by subtracting $\mathcal{J}[u](t)$ from $\mathcal{J}[u](t+1)$. This task is performed in our code using {\sc TimeDifference}.
 \begin{example}
@@
 \end{example}
-%	Our code allows us to check not only for conservation in time but also if higher time derivatives are zero under the scheme by specifying a {\sc timederivativeorder} in the {\sc variables}.	
-{\bf Step 2: Translate the scheme.}
-%	Now we compute the translations of the scheme.
-As mentioned above, difference algebra does not guarantee termination of the algorithm.
-Hence, we need to find a finite set of translations of the scheme, to guarantee termination of the algorithm.
-We translate the scheme such, that the range of its stencil stays exactly inside the range of the stencil of the discrete time derivative.
-This guarantees, that every instance of $u_j(n+e,t+l)$, occuring in $\mathcal{T}$, also occurs in the translated scheme.
-We now make the step from difference algebra, with an infinite number of independent variables, to polynomial algebra, where we only have a finite number of variables to guarantee finiteness of the Gr{\"o}bner basis.
-Hence, we need to compute a finite number of translations of the scheme and then consider every instance $u_j(n+e,t+l)$, that appears either in the translated scheme or in the discrete time derivative as independent polynomial variable.
+
+\noindent{\bf Step 2: Translate the scheme.}
+Now, we transition from difference algebra with infinite independent variables to polynomial algebra with finite variables to ensure the finiteness of the Gr{\"o}bner basis. 
+Hence, we need to compute a finite number of translations of the scheme and consider every instance of $u_j(n+e,t+l)$ that appears in either the translated scheme or the discrete time derivative as an independent polynomial variable.
 \begin{example}
@@
 	where $k_0=0$, $x^{k_l +}_i = \max_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$,
-	and\\ $x^{k_l -}_i = \min_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$. Here, $[a,b]$ denotes the discrete interval in $\Zz$, i.e. $[0,2]=\{0,1,2\}$.
-	\end{definition}
-If for any $[a,b]$ involved $b<a$, we write $[a,b]=\emptyset$ with the convention that $\emptyset-\emptyset=\emptyset$.
+	and $x^{k_l -}_i = \min_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$. Here, $[a,b]$ denotes the discrete interval in $\Zz$, i.e. $[0,2]=\{0,1,2\}$.
+\end{definition}
+If for any $[a,b]$ involved we have $b<a$, we write $[a,b]=\emptyset$ with the convention that $\emptyset-\emptyset=\emptyset$.
 Further, $\emptyset$ in the translations results in the use of the respective entry of the original scheme without translations.
@@
 \end{example}
-{\bf Step 3 and 4: Compute the Gr{\"o}bner basis and reduce the discrete time derivative.}\\
+
+\noindent{\bf Step 3 and 4: Compute Gr{\"o}bner basis and reduce the discrete time derivative.}\\
 \emph{Polynomial ideals and Gr{\"o}bner bases.}\label{polid}
-	We found a finite set of polynomials (the translated scheme $\{H^1,\ldots,H^r\}$ and the discrete time derivative) with a finite number of instances of $u_j(n+e,t+l)$ in the previous step.
-Now we reduce $\mathcal{T}$ as an element of a polynomial ring by multivariate polynomial division with respect to the translated scheme.
-Here, we adapt the definitions and theorems from \cite{hibi13} to our setting.
+In the previous step, we identified a finite set of polynomials comprising the translated scheme ${H^1,\ldots,H^r}$ and the discrete time derivative, each containing a finite number of instances of $u_j(n+e,t+l)$.
+We now proceed to reduce $\mathcal{T}$ by treating it as an element of a polynomial ring and performing multivariate polynomial division with respect to the translated scheme.
+To accomplish this, we adapt the definitions and theorems presented in \cite{hibi13} to suit our specific context.
 \begin{definition}
@@
 Contrary to the standard basis for the difference ideal that Gerdt's algorithm computes, a Gr{\"o}bner basis for the polynomial ideal always exists and is finite.
-	A Gr{\"o}bner basis is defined up to the order of the monomials involved, so, depending on the order, we get different remainders of the polynomial division.\\
+A Gr{\"o}bner basis is defined up to the order of the monomials involved, so, depending on the order, we get different remainders of the polynomial division.
+
 \emph{Polynomial reduction.}
-	In Algorithm \ref{alg: discreteconservedqimplicit}, we consider two monomial orders: the (a) lexicographic order and an experimental (b) elimination order.
-	To reduce $\mathcal{T}$, we compute the Gr{\"o}bner basis of $\langle H^1,\ldots,H^r\rangle$ with respect to the monomial order and then calculate the remainder of the polynomial division of $\mathcal{T}$ with respect to this Gr{\"o}bner basis.
-	The (a) lexicographic order is induced by the time-explicit ordering of the instances of $u_j(n+e,t+l)$, which gives more weight to instances at later times than to instances at earlier times.
-	For the (b) elimination order, we first order all instances of $u_j(n+e,t+l)$ according to a pre-defined ordering.
-Our code uses by default the time-explicit ordering, but other orderings, like an implicit ordering or an user-defined ordering, are also possible.
-Then, we try to successively eliminate the instance with the highest ordering from the discrete time derivative, using the elimination order induced by a weight-matrix.
-	This elimination might not always be possible, hence the reduction may leave the discrete time derivative unchanged.
-Then we repeat this process for the instance at the second latest time and continue until all instances of $u_j(n+e,t+l)$ have been eliminated.
-	However, the elimination order is rather experimental, as its weight-matrix does not have a full rank and hence does not induce a total order on the set of monomials.\\
-	Out of the two remainders from (a) and (b), we choose the result with the least number of different instances of time.\\
-{\bf Step 5 to 7: Compute the discrete partial variational derivative and reduce again.}
+In Algorithm \ref{alg: discreteconservedqimplicit}, we employ two monomial orders: (a) the lexicographic order and (b) an experimental elimination order.
+To reduce $\mathcal{T}$, we first compute the Gr\"obner basis of $\langle H^1,\ldots,H^r\rangle$ with respect to the chosen monomial order and subsequently calculate the remainder resulting from the polynomial division of $\mathcal{T}$ with respect to this Gr\"obner basis.
+
+The lexicographic order (a) is induced by the time-explicit ordering of the instances of $u_j(n+e,t+l)$, giving greater importance to instances at later times over those at earlier times.
+For the elimination order (b), we initially order all instances of $u_j(n+e,t+l)$ based on a predefined ordering.
+By default, our code employs the time-explicit ordering, but other options, such as implicit or user-defined orderings, can also be used.
+We then attempt to sequentially eliminate the instance with the highest ordering from the discrete time derivative, utilizing the elimination order induced by a weight-matrix.
+This elimination may not always be feasible, resulting in an unchanged discrete time derivative.
+In such cases, we repeat the process for the instance at the second latest time and continue until all instances of $u_j(n+e,t+l)$ have been eliminated.
+
+It should be noted that the elimination order (b) is experimental in nature, as its weight-matrix lacks full rank and therefore does not induce a total order on the set of monomials.
+Finally, we select the result with the fewest distinct time instances from the two remainders produced by the lexicographic order (a) and the elimination order (b).
+
+\noindent {\bf Step 5 to 7: Compute the discrete partial variational derivative and reduce again.}
 In the next step, we take the resulting expression and compute its discrete partial variational derivative. Then, we repeat Step 2 to Step 4 applied to the discrete partial variational derivative.
@@
 	\label{heateqalgoimplicit}
-		To demonstrate Algorithm \ref{alg: discreteconservedqimplicit}, we use the setting from Example \ref{timeexplicitheateqalgo}.
+	To illustrate Algorithm \ref{alg: discreteconservedqimplicit}, we use the setting of Example \ref{timeexplicitheateqalgo}.
 	We compute the discrete time derivative (Step 1)
@@
 \begin{example}
-		We verify our result from Example \ref{heateqalgoimplicit}, using {\sc DiscreteConservedQ}, which detects automatically if we are in the time-explicit or the general setting.
+	We verify our result from Example \ref{heateqalgoimplicit}, using {\sc DiscreteConservedQ}, which detects automatically if the scheme is time-explicit or if
+	 the general setting should be applied.
 	\begin{EXE}
@@
 \begin{remark}\label{remark}
-	The algorithm can only detect conservation, but can not check if something is not conserved for sure.
-So a resulting {\sc False} should be understood as our algorithm not detecting conservation, not as there being no conservation at all.
+The algorithm is designed to identify conservation, but it cannot definitively confirm the absence of conservation.
+Therefore, a {\sc False} result should be interpreted as the algorithm being unable to detect conservation, rather than an absolute assertion of non-conservation.
+The reason for this is that we are working with a Gr\"obner basis that does not generate the full difference ideal.
 \end{remark}
-%\begin{remark}
-%	Our algorithm is also able to check for detection of time derivatives of higher order.
-To check for example the conservation of the first time derivative of the solution of the scheme, one can %specify {\sc timederivativeorder $\rightarrow$ 2} when calling {\sc DiscreteConservedQOperator}.
-In that case the algorithm will build the discrete forward second order time derivative of the %given quantity.
-%\end{remark}
+
+{\color{red} add an example with parameters}
+
 \section{A basis for conserved quantities}
@@
 	\DontPrintSemicolon
-	\KwIn{right-hand side $\{H^1,\ldots,H^m\}\subset S(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ of time-explicit scheme, list of {\sc generators}, {\sc degree} of polynomials }
+	\KwIn{right-hand side set of polynomials $\{H^1,\ldots,H^m\}\subset S(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ ofa time-explicit scheme, list of {\sc generators}, {\sc degree} of polynomials }
 	\KwOut{basis of conserved quantities}
 
-	\label{fstep1} Build all monomials combined from the {\sc generators} up to the total {\sc degree}\;
-
-	\label{fstep2} Linearly combine all monomials with undetermined coefficients\;
-
-	\label{fstep3} Compute the discrete time derivative and plug in the appropriate right-hand side(s) of the scheme\;
+	\label{fstep1} Construct all monomials combined from the {\sc generators} up to the total {\sc degree}\;
+
+	\label{fstep2} Form linear combinations of all monomials with undetermined coefficients\:
+
+	\label{fstep3} Compute the discrete time derivative and substitute in the appropriate right-hand side(s) polynomials of the scheme\;
 
@@
 
-	\label{fstep4} Simplify and use the undetermined coefficients method to determine coefficients that make the result vanish\;
+	\label{fstep4} Simplify the expression and apply the undetermined coefficients method to find coefficients that make the result vanish\;
 
@@
 \end{algorithm}
-We have implemented this algorithm in the {\sc FindDiscreteConservedQuantityBasis}.
-\begin{example}
-	We find a basis for conserved quantities that are generated by $u(n,t)$ and $n$ and that have at most degree 3, admitted by the scheme for the heat equation from Example \ref{heatex1}.
+We have implemented this algorithm in {\sc FindDiscreteConservedQuantityBasis}.
+\begin{example}
+In the context of the heat equation from Example \ref{heatex1}, we aim to find a basis for conserved quantities generated by $u(n,t)$ and $n$ with a maximum degree of 3.
 	\begin{EXE}
@@
 	\end{small}
+The first conserved quantity corresponds to mass conservation and the second to center of mass conservation. 
 \end{example}
@@
 \label{eaa}
-We illustrate our code with the inviscid Burgers equation and a system of mean-field games.
-These only have first-order time-derivatives, but the algorithm handles PDEs with all orders of derivatives.\\
-{\bf Burgers equation.}
+We demonstrate our code's capabilities using two examples: the inviscid Burgers equation and a mean-field game system. 
+Through these examples, we examine different discretizations, evaluate their conservation properties, and identify conserved quantities.
+Although the examples presented feature only first-order time derivatives, the algorithm can handle PDEs with derivatives of any order.
+
+\noindent{\bf Burgers equation.}
+In the first example, we focus on the inviscid Burgers equation, a well-known PDE with the form $u_t + uu_x = 0$ \cite{smoller94}. 
+This PDE preserves the total mass $\int u\ dx$.
+Our first example demonstrates how the code verifies mass conservation by a forward-difference discretization.
+Next, we demonstrate the algorithm's ability to find a mass-preserving scheme by investigating conservation properties under a class of schemes involving a parameter.
+Finally, we explore time-implicit versions of the schemes and their conservation properties.
+
 The inviscid Burgers equation is the PDE $u_t+uu_x=0$ \cite{smoller94} in one space dimension. Any function of $u$ is a conserved quantity. We search for discretizations preserving $\int u\ dx$.
@@
 \begin{example}
-	We discretize $u_x$ by a three-point stencil with a parameter $a$.
+	We discretize $u_x$ with a three-point stencil with a parameter $a$.
 	\begin{EXE}
@@
 	\end{small}
-	Conservation holds for $a=\frac{1}{2}$, the central difference case.
+The conservation property holds when $a=\frac{1}{2}$, which corresponds to the central difference case.
 \end{example}
@@
 \end{example}
-	Our results also hold for the viscous Burgers equation $u_t+uu_x-u_{xx}=0$, if we use the second-order central difference for $u_{xx}$.\\
+%Our results also hold for the viscous Burgers equation $u_t+uu_x-u_{xx}=0$, if we use the second-order central difference for $u_{xx}$.\\
+
 {\bf Conserved quantities for a mean-field game.}
-We study conserved quantities admitted by the discretization of a PDE system.
+We analyze the conserved quantities preserved by the discretization of a PDE system in the context of mean-field games.
+The examples showcase our algorithm's effectiveness in identifying conserved quantities for various discretization schemes, including time-explicit, time-implicit, and mixed schemes.
 \begin{example}
@@
 	\end{small}
-	We reproduce this result, noting that we have a time-explicit scheme:
+We replicate this result, observing that the scheme used here is time-explicit:
 	\begin{EXE}
@@
 	\end{small}
-	Those are the only conserved quantities for this scheme, that are polynomials up to degree 4 in $v$ and $m$.
+	Those are the only conserved quantities for this scheme, which are polynomials up to degree 4 in $v$ and $m$.
 	\begin{EXE}
@@
 	This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly in time for $m$ and implicitly for $v$.
-Our code can handle this setting, if we specify an order for the elimination of the variables, telling the code to use an explicit monomial order for $m$ and an implicit one for $v$:
+	Our code can handle this setting by specifying an order for variable elimination, using an explicit monomial order for $m$ and an implicit order for $v$:
 	\begin{EXE}
@@
 We use our implementation of the algorithm to analyze the conservation properties of several schemes for PDEs that arise in applications.
-
-%\citestyle{acmnumeric}
-\bibliographystyle{ACM-Reference-Format}
+{\color{red} add remark on parametric groebner basis}
+
+\bibliographystyle{splncs04}
 \bibliography{references}

### 2023-05-07 Papers/Paper CASC format/paper.tex
various

@@
 
-{\color{red}ADD HERE TIME EXPLICIT EXAMPLE WITH PARAMETERS - DISCUSS WITH RICARDO THE ISSUE}
-
-{\color{red} DG HERE}
+{\color{red}ADD HERE TIME EXPLICIT EXAMPLE WITH PARAMETERS}
 
@@
 
-	\label{step3} Reduce $\mathcal{T}$ by using the Gr{\"o}bner basis generated by the translated scheme in two different ways
+	\label{step3} Reduce $\mathcal{T}$ by using {\color{red} the Gr{\"o}bner basis} generated by the translated scheme in two different ways
 	\label{step3a} (a) Using the lexicographic order

### 2023-05-08 Papers/Paper CASC format/paper.tex
paper

@@
 \newcommand{\Mm}{\mathcal{M}}
+\newcommand{\Kk}{\mathcal{K}}
 
@@
 %\label{intro}
+{\color{blue} add 
+Comprehensive Gr{\"o}bner Systems 
+}
 In many partial differential equations (PDEs), conserved integral quantities play a crucial role, as they often represent properties of physical significance, such as energy or mass.
@@
 
-\section{Parametric Groebner basis}
-\label{pgb}
-
-TBD
-
+ \newpage 
+%just for now 
+\section{Parametric Groebner Systems}
+\label{pgs}
+
+% This section covers key concepts for studying parametric 
+% polynomial ideals, including parametric rings and specialization 
+% ideals and comprehensive Gr{\"o}bner systems for parametric polynomials that generalize 
+% Gr{\'o}bner basis in section \ref{4sec1}. Section 2 explains why comprehensive Gr{\"o}bner systems are advantageous 
+% and some problems that arise without CGS. Section 3 
+% extends the ideal membership problem to parametric ideals.
+In this section, we delve into essential concepts for 
+parametric polynomial ideals. These include parametric rings, 
+specialized ideals, and comprehensive Gr{\"o}bner systems, which are built upon 
+Gr{\"o}bner basis theory. 
+We discuss the benefits of using comprehensive Gr{\"o}bner systems 
+and how they help address problems that occur when not using them. 
+Furthermore, we expand on the ideal membership problem to encompass parametric ideals.
+
+% \subsection{Parametric Polynomials and Comprehensive Groebner Systems}
+% \label{4sec1}
+
+% This section introduces the general setting of parametric polynomial rings and a ring homomorphism that 
+% represents the choice of parameters. Moreover, we define comprehensive Gr{\"o}bner systems (CGS),
+% and show some examples of (CGS) using our implementation of a parametric Buchberger's algorithm.
+
+\begin{definition}
+    Let $\mathcal{K}$ be a field. The ring of parametric polynomials
+    $\mathcal{K}[a_1,..,a_s][x_1,...,x_n]$ is a ring of polynomials in the variables $x_1,...,x_n$ with coefficients that are polynomials
+    in the ring $\mathcal{K}[a_1,..,a_s]$. 
+    For simplicity we denote this ring by $\mathcal{K}[A][X]$.
+\end{definition}
+
+\begin{definition}
+    Let $L$ be an algebraically closed field extension of $\mathcal{K}$ and let $\sigma:\mathcal{K}[A][X]\rightarrow L[X]$ be a ring homomorphism.
+$\sigma$ is a specialization given by a polynomial evaluation on the parameters and an identity on the variables, i.e.,
+    for $P\in \mathcal{K}[A][X]$
+    $\sigma$ is given by $P\rightarrow P(a_{1_0},..,a_{s_0})$, for some $(a_{1_0},..,a_{s_0})\in L^s$.
+
+\end{definition}
+% For example let $P=a x+by \in \Rr[a,b][x,y]$. 
+% a specialization $\sigma$ is defined by $P\rightarrow P(2,5)$ so that $$\sigma (P)=\sigma(a x+by)$$
+% and since $\sigma $ is a ring homomorphism we have,
+% $$\sigma(a x+by)= \sigma(a) \sigma(x)+\sigma(b)\sigma(y)$$
+% $$= 2x+5y.$$
+It is easier to think about specializations as a choice of parameters. \\
+
+Comprehensive Gr{\"o}bner systems extend Gr{\"o}bner basis to 
+parametric polynomials and have diverse practical applications in 
+algebraic geometry, combinatorics, coding theory, computer-aided design, and robotics.
+\begin{definition}
+    Let $\Ii$ be a finitely generated, nonzero ideal in the ring $\mathcal{K}[A][X]$ with a fixed monomial order $\prec$ on the variables. 
+    Let $\{A_1,\dots,A_k\}$ be a family of locally closed sets that forms a partition of $L^m$ 
+    , such that the disjoint union $\bigcup\limits_{i=1}^k A_i=L^m$, and $\{G_1,\dots,G_k\}$ 
+    be a family of subsets in $\mathcal{K}[A][X]$.
+    The system, 
+    \[   \left\{
+    \begin{array}{ll}
+        G_1 & A_1 \\
+        G_2  & A_2\\
+        \vdots  & \vdots \\
+        G_k & A_k \\
+    \end{array} 
+    \right. \]
+
+    is called a comprehensive Gr{\"o}bner system (CGS) of $I$
+     if for all specializations $\sigma_i \in A_i, \sigma_i (G_i)$ 
+     is a Gr{\"o}bner basis of the specialized ideal $\sigma_i (I)\subset \mathcal{K}[X]$.
+     Where each pair $\{G_i,A_i\}$ is called a branch. 
+\end{definition}
+Here, the notation $\sigma_i (\Ff)$, with $\Ff=\{f_1,f_2,\dots ,f_k\}$, is the specialization $\sigma_i$
+of each $f_j$ such that  $\sigma_i (\Ff)=\{\sigma_i(f_j)~|~ f_j\in \Ff\}$
+and  $"\sigma_i \in A_i"$ stands for a specialization given by $P\rightarrow P(a_{1_0},\dots,a_{m_0})$ 
+for $(a_{1_0},\dots,a_{m_0})\in A_i$.\\
+
+In this paper, we choose $A_i$ to be of the form 
+$$V(p_1,\dots,p_s)\backslash V(q_1,\dots,q_t) \subset L^m.$$
+Where $p_1,\dots,p_s,q_1,\dots,q_t\in \mathcal{K}[A]$ and $V(p_1,\dots,p_s)$ is a Variety.
+
+A proof of existence and the first algorithm to compute the (CGS) were introduced by Volker Weispfenning.
+\begin{theorem}
+    Let $\mathcal{I}\subseteq\mathcal{K}[A][X].$ The corresponding comprehensive Gr{\"o}bner system exists.\cite{weispfenning}
+\end{theorem}
+We use our implementation of a parametric Buchberger's algorithm to compute (CGS). 
+This is illustrated in the following examples.
+\begin{example}
+    Let $\Ii\subset \Rr[a,b][x,y]$ be the ideal generated by 
+    $$\Ff=\left\{a x^2+\left(a^2-2 a b\right) x y, b^2 y^2-x\right\}.$$
+    The comprehensive Gr{\"o}bner system of $\Ii$ with a Lexicographic order $y\prec x$ is:
+    {\color{red} we use our code here}
+	% $$
+    % % \begin{cases}\left\{a^2 b^2 y^3+a b^4 y^4-2 a b^3 y^3,x-b^2
+    % %     y^2\right\} &  a b \neq 0 \\
+    % % \{x\} &  b=0 \\
+    % % \left\{b^2 y^2-x\right\} &  a=0 \land b \neq 0.
+    % % \end{cases}
+    % $$
+\end{example}
+The idea of a comprehensive Gr{\"o}bner system is to provide a "parameter sensitive Gr{\"o}bner basis".
+It branches the cases of the parameters that affect leading terms. For instance, in the previous
+example if $a=0$ then the leading term $a x^2$ vanishes which leads to a different Gr{\"o}bner basis. Therefore,
+the (CGS) considers cases of the leading coefficients parameters Vanishing.
+
+\begin{example}
+    Let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
+    $$\Ff=\left\{a^2 x-y,-a+b^2 x
+    y-y^2\right\}.$$
+    The comprehensive Gr{\"o}bner system of $\mathcal{I}$ with a Lexicographic order $y\prec x$ is:
+
+    {\color{red} we use our code here}
+\end{example}
+
+\begin{remark}
+    If a polynomial ideal contains a nonzero constant polynomial, then it is called a unit ideal $\langle 1\rangle$.
+   Unit ideals can generate all polynomials in the ring, so $\langle 1\rangle=\mathcal{K}[A][X]$.
+    As a convention the Gr{\"o}bner basis of such ideals is $\{1\}$.
+\end{remark}
+
+Issues arise when computing the Gr{\"o}bner basis of $\Ff$ without considering the parameters.
+In practice, to compute the Gr{\"o}bner basis, we assume non-zero parameters by default.
+This assumption guarantees a Gr{\"o}bner basis only for one set of specializations $\Rr^2\backslash V(a,b)$. i.e., 
+$\sigma(\Gg)$ is a Gr{\"o}bner basis of the specialization ideal $\sigma(\Ii)$ only for $\sigma \in \Rr^2\backslash V(a,b)$.
+ The impact of this assumption is seen in (i) and (ii):
+\begin{enumerate}
+    \item In general $\sigma\left(\Gg\right)$ is not necessarily a Gr{\"o}bner basis.
+    \begin{example}
+        \label{ex431}
+        Let $\Ii=\langle 
+           {a x^2 - y, -2 + b x y} \rangle$. The corresponding Gr{\"o}bner basis 
+           is $\Gg=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$. Consider 
+           the specialization $\sigma \in \Rr^2$ given by $P\rightarrow P(0,1)$.
+           Now, We claim that the set $\sigma(\Gg)=\left\{y^3,-y^2,x y-2\right\}$ is not a Gr{\"o}bner basis.
+           To show that we find the S-polynomials of each pair in the set and we find that: $S(-y^2, -2 + x y)=-2 y$.
+           We divide $S(-y^2, -2 + x y)$ by $\sigma(\Gg)$ and find that the remainder is $-2y$. 
+           Therefore by the Buchberger Criterion, $\sigma(\Gg)$ is not a Gr{\"o}bner basis.
+    \end{example}
+    \item Consequently, the algorithm utilized to resolve the ideal membership 
+    problem in \ref{sec2.1} can not be used for all specializations.
+
+    \begin{example}
+        \label{ex433}
+        Consider the ideal $\Ii$ in the above example. The corresponding Gr{\"o}bner basis computed under the assumption of non-zero parameters
+        is $$\sigma_0(\Gg)=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$$ where $\sigma_0 \in \Rr^2\backslash V(a,b)$. 
+        Now, it is possible to check if a polynomial $f\in \Rr[x,y]$ belongs to $\sigma(\Ii)$ only for $\sigma_0 $     
+        because we know that $\sigma_0(\Gg)$ is a Gr{\"o}bner basis. For other specializations such as 
+        the specialization $\sigma$ given by $P\rightarrow P(0,1)$, $\sigma(\Gg)$ is not a Gr{\"o}bner basis as shown in the above example. 
+        Therefore, we can not proceed with the division algorithm as we can not guarantee unique remainder if we used $\sigma(\Gg)$.
+
+    \end{example}
+
+\end{enumerate}
+
+\newpage
 \section{The time-explicit case}

### 2023-05-09 Papers/Paper CASC format/MLaTeX/CLpaper.tex
...

--- 2023-05-08 Papers/Paper CASC format/paper.tex
@@
 % This is samplepaper.tex, a sample chapter demonstrating the
-% LLNCS macro package for Springer Computer Science proceedings;
-% Version 2.21 of 2022/01/12
-%
 \documentclass[runningheads]{llncs}
-%
 \usepackage[T1]{fontenc}
-% T1 fonts will be used to generate the final print and online PDFs,
-% so please use T1 fonts in your manuscript whenever possible.
-% Other font encondings may result in incorrect characters.
-%
 \usepackage{graphicx}
-% Used for displaying a sample figure. If possible, figure files should
-% be included in EPS format.
-%
-% If you use the hyperref package, please uncomment the following two lines
-% to display URLs in blue roman font according to Springer's eBook style:
-%\usepackage{color}
-%\renewcommand\UrlFont{\color{blue}\rmfamily}
-%
 
@@
 \usepackage{mmacells}
-\usepackage[invisible]{MLaTeX}
 
@@
 \begin{document}
-%
 \title{Algorithmic detection of conserved quantities for finite-difference schemes\thanks{Supported by King Abdullah University of Science and Technology (KAUST) baseline funds and KAUST OSR-CRG2021-4674.}}
-%
-%\titlerunning{Abbreviated paper title}
-%
 \author{Diogo A. Gomes\inst{1}\orcidID{0000-0002-3129-3956} \and
@@
 	Ricardo de Lima Ribeiro\inst{1}\orcidID{0000-0002-2707-2210}}
-%
 \authorrunning{D.A. Gomes et al.}
-%
 \institute{King Abdullah University of Science and Technology (KAUST), AMCS/CEMSE Division, Thuwal 23955-6900, Saudi Arabia\
 	\email{{diogo.gomes,friedemann.krannich,bashayer.majrashi, ricardo.ribeiro}@kaust.edu.sa}}
-%
 \maketitle
-%
 \begin{abstract}
@@
 \section{Introduction}
-%\label{intro}
 {\color{blue} add 
-Comprehensive Gr{\"o}bner Systems 
+Comprehensive Gr{\"o}bner Systems 2222
 }
@@
 Here, all quantities, whose conservation in time we check, are functions $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$. 
-%We restrict $G$ to the set of polynomials in $k$ variables, because our algorithm below can only treat polynomial quantities.
 Let $\tilde{n}:=(n_1,\dots,n_d,s_1,\dots,s_l)\in\Ii\times\Nn^l$. Using the Kronecker delta $\delta$, we rewrite
@@
 	In our {\sc Mathematica} implementation, Example \ref{dpvd1} is computed by
-	\begin{EXE}
-		LABEL="dpvd1"
-		(*CODE*)
-		variables = Association["indVars"->{n,t},"depVars"->{u},"timeVars"->{t}]
-		expression = u[n,t+1] - u[n,t]
-		PartialDVarDOperator[variables][expression]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}	
+
 	\begin{small}
-		\input{executables/dpvd1ctt.ed.tex}
-		\input{executables/dpvd1ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, expression, PartialDVarD}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n,t\},"depVars"\(\pmb{\to}\)\{u\},
+  "timeVars"\(\pmb{\to}\)\{t\}|>
+  PartialDVarD[variables][u[n,t+1]-u[n,t]]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  \{-KroneckerDelta[t]+KroneckerDelta[1+t]\}
+\end{mmaCell}
 	\end{small}
@@
  \newpage 
-%just for now 
 \section{Parametric Groebner Systems}
@@
 
-% This section covers key concepts for studying parametric 
-% polynomial ideals, including parametric rings and specialization 
-% ideals and comprehensive Gr{\"o}bner systems for parametric polynomials that generalize 
-% Gr{\'o}bner basis in section \ref{4sec1}. Section 2 explains why comprehensive Gr{\"o}bner systems are advantageous 
-% and some problems that arise without CGS. Section 3 
-% extends the ideal membership problem to parametric ideals.
 In this section, we delve into essential concepts for 
@@
 
-% \subsection{Parametric Polynomials and Comprehensive Groebner Systems}
-% \label{4sec1}
-
-% This section introduces the general setting of parametric polynomial rings and a ring homomorphism that 
-% represents the choice of parameters. Moreover, we define comprehensive Gr{\"o}bner systems (CGS),
-% and show some examples of (CGS) using our implementation of a parametric Buchberger's algorithm.
-
 \begin{definition}
@@
 \end{definition}
-% For example let $P=a x+by \in \Rr[a,b][x,y]$. 
-% a specialization $\sigma$ is defined by $P\rightarrow P(2,5)$ so that $$\sigma (P)=\sigma(a x+by)$$
-% and since $\sigma $ is a ring homomorphism we have,
-% $$\sigma(a x+by)= \sigma(a) \sigma(x)+\sigma(b)\sigma(y)$$
-% $$= 2x+5y.$$
 It is easier to think about specializations as a choice of parameters. \\
@@
     The system, 
-    \[   \left\{
+	\begin{equation}\label{CGS}
+	\left\{
 		\begin{array}{ll}
 			G_1 & A_1 \\
-        G_2  & A_2\\
 			\vdots  & \vdots \\ 
@@
 		\end{array}
-    \right. \]
+		\right. 
+	\end{equation}
 
@@
     {\color{red} we use our code here}
-	% $$
-    % % \begin{cases}\left\{a^2 b^2 y^3+a b^4 y^4-2 a b^3 y^3,x-b^2
-    % %     y^2\right\} &  a b \neq 0 \\
-    % % \{x\} &  b=0 \\
-    % % \left\{b^2 y^2-x\right\} &  a=0 \land b \neq 0.
-    % % \end{cases}
-    % $$
 \end{example}
@@
 Issues arise when computing the Gr{\"o}bner basis of $\Ff$ without considering the parameters.
-In practice, to compute the Gr{\"o}bner basis, we assume non-zero parameters by default.
+In practice, to compute the Gr{\"o}bner basis without considering the parameters, we assume non-zero parameters by default.
 This assumption guarantees a Gr{\"o}bner basis only for one set of specializations $\Rr^2\backslash V(a,b)$. i.e., 
-$\sigma(\Gg)$ is a Gr{\"o}bner basis of the specialization ideal $\sigma(\Ii)$ only for $\sigma \in \Rr^2\backslash V(a,b)$.
- The impact of this assumption is seen in (i) and (ii):
+$\sigma(\Gg)$ is a Gr{\"o}bner basis of the specialized ideal $\sigma(\Ii)$ only for $\sigma \in \Rr^2\backslash V(a,b)$.
+Next, we show the impact of this assumption.
 \begin{enumerate}
@@
            is $\Gg=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$. Consider 
-           the specialization $\sigma \in \Rr^2$ given by $P\rightarrow P(0,1)$.
-           Now, We claim that the set $\sigma(\Gg)=\left\{y^3,-y^2,x y-2\right\}$ is not a Gr{\"o}bner basis.
-           To show that we find the S-polynomials of each pair in the set and we find that: $S(-y^2, -2 + x y)=-2 y$.
-           We divide $S(-y^2, -2 + x y)$ by $\sigma(\Gg)$ and find that the remainder is $-2y$. 
-           Therefore by the Buchberger Criterion, $\sigma(\Gg)$ is not a Gr{\"o}bner basis.
+           the specialization $\sigma$ given by $P\rightarrow P(0,1)$.
+           The set $\sigma(\Gg)=\left\{y^3,-y^2,x y-2\right\}$ is not a Gr{\"o}bner basis.
+		   This can be shown by the Buchberger Criterion (see chapter 2 in\cite{Cox}).
+
     \end{example}
     \item Consequently, the algorithm utilized to resolve the ideal membership 
-    problem in \ref{sec2.1} can not be used for all specializations.
+    problem in chapter 2 of \cite{Cox} can not be used for all specializations.
 
@@
         Therefore, we can not proceed with the division algorithm as we can not guarantee unique remainder if we used $\sigma(\Gg)$.  
-
     \end{example}
@@
 \end{enumerate}
+
+Using CGS can resolve these problems. Problem 1 is fixed by the definition of (CGS), as we have $\sigma_i(G_i)$ a 
+ Gröbner basis for any specialization $\sigma_i$. Moreover, 
+ to discuses problem 2 we need to look at the parametric ideal membership first.
+
+{\color{red} Is this a definition?}
+ Consider, $f\in \Kk[A][X]$ we say that $f\in\Ii$ under $\sigma$, if there exists a specialization 
+ $\sigma\in L^m$ such that $\sigma(f)\in\sigma(\Ii)$.\\
+
+To determine if $f\in\Ii$ we consider the (CGS) of $\Ii$ given \ref{CGS}. 
+We check if $f\in\sigma_i(\Ii)$ for each $i\in\{1,..,k\}$ and $\sigma_i\in A_i$.
+For each $\sigma_i(\Ii)$ we have the corresponding Gröbner basis $\sigma_i(G_i)$
+so we divide $f$ by the $\sigma_i(G_i)$ if the remainder is zero for some $i\in\{1,..,k\}$,
+then $f\in\sigma_i(\Ii)$. Therefore, $f\in\Ii$. However, if the remainder is non-zero for all $i$ we conclude that 
+$f\notin\Ii$.
+
+\begin{example}
+	Let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
+    $$\Ff=\{a^2 x - y, -b^2 x y^2 - x\}.$$
+	We will determine if $f(x,y)=-x+y^2\in\mathcal{I}$.
+	We compute the corresponding (CGS)
+	{\color{red} add example 3}\\
+	First let $A_1=\{(a,b)~|~ a\ne 0, b\ne 0\}$ and $A_2=\{(a,b)~|~ ab=0\}$. 
+	We divide $f$ by $\{\frac{y}{b^2}+y^3,x-\frac{y}{a^2}\}$ and we obtain the remainder $-\frac{y}{a^2}+y^2$ and 
+	since the exists no specialization in $A_1$ such that this remainder is zero we conclude that $f\notin \sigma_1(\Ii)$ for
+	$\sigma_1\in A_1$. Now we repeat the reduction of $f$ but with the second Gröbner basis $\{x,y\}$. We obtain a zero remainder, which imply that 
+	$f\in \sigma_2(\Ii)$ for $\sigma_2\in A_2$. Therefore, $f\in\Ii$ under a specialization $\sigma_2\in A_2$.
+\end{example}
 
@@
 	We verify the result from Example \ref{timeexplicitheateqalgo}, using our implementation in {\sc Mathematica} of Algorithm \ref{alg: discreteconservedqexplicit}.
-	\begin{EXE}
-		LABEL="rhs1"
-		(*CODE*)
-		variables = Association["indVars"->{n},"depVars"->{u},"rhs"->{u[n]+(u[n+1]-2*u[n]+u[n-1])}]
-		expression = u[n]
-		DiscreteConservedQOperator[variables][expression]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
+
 	\begin{small}
-		\input{executables/rhs1ctt.ed.tex}
-		\input{executables/rhs1ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, expression, DiscreteConservedQ}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n\},"depVars"\(\pmb{\to}\)\{u\},
+  "rhs"\(\pmb{\to}\)\{u[n]+(u[n+1]-2u[n]+u[n-1])\}|>
+  DiscreteConservedQ[variables][u[n]]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  True
+\end{mmaCell}
 	\end{small}
@@
 	The discrete time derivative for $\sum_n u(n,t)$ is
-	\begin{EXE}
-		LABEL="heateq2"
-		(*CODE*)
-		variables = <|"indVars"->{n,t},"depVars"->{u}|>
-		expression = u[n,t]
-		TimeDifferenceOperator[variables][expression]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
 
 	\begin{small}
-		\input{executables/heateq2ctt.ed.tex}
-		\input{executables/heateq2ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, expression, TimeDifference}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n,t\},"depVars"\(\pmb{\to}\)\{u\}|>
+  TimeDifference[variables][u[n,t]]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  <|exp\(\to\)-u[n,t]+u[n,1+t]|>
+\end{mmaCell}
 	\end{small}
@@
 	 the general setting should be applied.
-	\begin{EXE}
-		LABEL="heateq10"
-		(*CODE*)
-		variables = Association["indVars"->{n,t},"depVars"->{u},"scheme"->{u[n,t+1]-u[n,t]-(u[n+1,t]-2*u[n,t]+u[n-1,t])}]
-		expression = u[n,t]
-		DiscreteConservedQOperator[variables][expression]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
+
 	\begin{small}
-		\input{executables/heateq10ctt.ed.tex}
-		\input{executables/heateq10ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, expression, DiscreteConservedQ}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n,t\},"depVars"\(\pmb{\to}\)\{u\},
+  "scheme"\(\pmb{\to}\)\{u[n,t+1]-u[n,t]
+  -(u[n+1,t]-2u[n,t]+u[n-1,t])\}|>
+  DiscreteConservedQ[variables][u[n,t]]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  True
+\end{mmaCell}
 	\end{small}
@@
 In the context of the heat equation from Example \ref{heatex1}, we aim to find a basis for conserved quantities generated by $u(n,t)$ and $n$ with a maximum degree of 3.
-	\begin{EXE}
-		LABEL="heateq11"
-		(*CODE*)
-		variables = <|"indVars"->{n},"depVars"->{u},"rhs"->{u[n]+u[n+1]-2*u[n]+u[n-1]}|>
-		FindDiscreteConservedQuantityBasisOperator[variables][<|"degree"->3,"generators"->{u[n],n}|>]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
+
 	\begin{small}
-		\input{executables/heateq11ctt.ed.tex}
-		\input{executables/heateq11ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, FindDiscreteConservedQuantityBasis}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n\},"depVars"\(\pmb{\to}\)\{u\},
+  "rhs"\(\pmb{\to}\)\{u[n]+u[n+1]-2u[n]+u[n-1]\}|>
+  FindDiscreteConservedQuantityBasis[variables]
+  [<|"degree"\(\pmb{\to}\)3,"generators"\(\pmb{\to}\)\{u[n],n\}|>]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  \{u[n],nu[n]\}
+\end{mmaCell}
 	\end{small}
@@
 	We check for conservation of mass using a forward-difference discretization:
-	\begin{EXE}
-		LABEL="burgers1"
-		(*CODE*)
-		variables = Association["indVars"->{n,t},"depVars"->{u},"scheme"->{u[n,t+1]-u[n,t]-u[n,t]*(u[n+1,t]-u[n,t])}];
-		expression = u[n,t];
-		DiscreteConservedQOperator[variables][expression]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
+
 	\begin{small}
-		\input{executables/burgers1ctt.ed.tex}
-		\input{executables/burgers1ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, expression, DiscreteConservedQ}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n,t\},"depVars"\(\pmb{\to}\)\{u\},
+  "scheme"\(\pmb{\to}\)\{u[n,t+1]-u[n,t]
+  -u[n,t](u[n+1,t]-u[n,t])\}|>
+  DiscreteConservedQ[variables][u[n,t]]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  False
+\end{mmaCell}
 	\end{small}
@@
 	We discretize $u_x$ with a three-point stencil with a parameter $a$.
-	\begin{EXE}
-		LABEL="burgers2"
-		(*CODE*)
-		variables = Association["indVars"->{n,t}, "depVars"->{u}, "pars"->{a}, "scheme"->{u[n,t+1]-u[n,t]-u[n,t]*(a*(u[n+1,t]-u[n,t])+(1-a)*(u[n,t]-u[n-1,t]))}];
-		expression = u[n,t];
-		DiscreteConservedQOperator[variables][expression]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
+
 	\begin{small}
-		\input{executables/burgers2ctt.ed.tex}
-		\input{executables/burgers2ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, expression, DiscreteConservedQ}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n,t\},"depVars"\(\pmb{\to}\)\{u\},
+  "pars"\(\pmb{\to}\)\{a\},"scheme"\(\pmb{\to}\)\{u[n,t+1]-u[n,t]
+  -u[n,t](a(u[n+1,t]-u[n,t])
+  +(1-a)(u[n,t]-u[n-1,t]))\}|>
+  DiscreteConservedQ[variables][u[n,t]]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  False		a==0||a==1||-1+2a\(\neq\)0
+  True		a==\mmaFrac{1}{2}
+
+\end{mmaCell}
 	\end{small}
@@
 	We consider the scheme from the previous example but in the time-implicit version.
-	\begin{EXE}
-		LABEL="burgers3"
-		(*CODE*)
-		variables = Association["indVars"->{n,t}, "depVars"->{u}, "pars"->{a},"elimOrder"->"implicit", "scheme"->{u[n,t+1]-u[n,t]-u[n,t+1]*(a*(u[n+1,t+1]-u[n,t+1])+(1-a)*(u[n,t+1]-u[n-1,t+1]))}]
-		DiscreteConservedQOperator[variables][u[n,t]]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
+
 	\begin{small}
-		\input{executables/burgers3ctt.ed.tex}
-		\input{executables/burgers3ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, DiscreteConservedQ}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n,t\},"depVars"\(\pmb{\to}\)\{u\},
+  "pars"\(\pmb{\to}\)\{a\},"elimOrder"\(\pmb{\to}\)"implicit",
+  "scheme"\(\pmb{\to}\)\{u[n,t+1]-u[n,t]-u[n,t+1]
+  (a(u[n+1,t+1]-u[n,t+1])
+  +(1-a)(u[n,t+1]-u[n-1,t+1]))\}|>
+  DiscreteConservedQ[variables][u[n,t]]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  False		a==0||a==1||-1+2 a\(\neq\)0
+  True		a==\mmaFrac{1}{2}
+
+\end{mmaCell}
 	\end{small}
 \end{example}
-%Our results also hold for the viscous Burgers equation $u_t+uu_x-u_{xx}=0$, if we use the second-order central difference for $u_{xx}$.\\
 
@@
 We check with our code if the scheme admits the same preserved quantities as the continuous system:
-	\begin{EXE}
-		LABEL="mfg1"
-		(*CODE*)
-		variables = Association[
-		"indVars" -> {n, t},
-		"depVars" -> {v, m},
-		"display result" -> True,
-		"scheme" -> {v[n, t + 1] - 
-			v[n, 
-			t] + (v[n + 1, t]^2/2 - m[n + 1, t]^2/2) - (v[n, t]^2/2 - 
-			m[n, t]^2/2), 
-			m[n, t + 1] - 
-			m[n, t] - (m[n + 1, t]*v[n + 1, t]) + (m[n, t]*v[n, t])}
-		];
-		DiscreteConservedQOperator[variables][{v[n, t], m[n, t]}]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
+
 	\begin{small}
-		\input{executables/mfg1ctt.ed.tex}
-		\input{executables/mfg1ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, DiscreteConservedQ}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n,t\},"depVars"\(\pmb{\to}\)\{v,m\},
+  "scheme"\(\pmb{\to}\)\{v[n,t+1]-v[n,t]
+  +(\mmaSup{v[n+1,t]}{2}-\mmaSup{m[n+1,t]}{2})/2-(\mmaSup{v[n,t]}{2}-\mmaSup{m[n,t]}{2})/2,
+  m[n,t+1]-m[n,t]-m[n+1,t]v[n+1,t]+m[n,t]v[n,t]\}|>
+  DiscreteConservedQ[variables][\{v[n,t],m[n,t]\}]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  True
+\end{mmaCell}
 	\end{small}
 We replicate this result, observing that the scheme used here is time-explicit:
-	\begin{EXE}
-		LABEL="mfg2"
-		(*CODE*)
-		variables = Association[
-		"indVars" -> {n},
-		"depVars" -> {v, m},
-		"rhs" -> { 
-			v[n] -(v[n + 1]^2/2 - m[n + 1]^2/2) + (v[n]^2/2 - 
-			m[n]^2/2), 
-			m[n] + (m[n + 1]*v[n + 1]) - (m[n]*v[n])}
-		];
-		DiscreteConservedQOperator[variables][{v[n], m[n]}]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
+
 	\begin{small}
-		\input{executables/mfg2ctt.ed.tex}
-		\input{executables/mfg2ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, DiscreteConservedQ}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n\},"depVars"\(\pmb{\to}\)\{v,m\},
+  "rhs"\(\pmb{\to}\)\{v[n]-(\mmaSup{v[n+1]}{2}-\mmaSup{m[n+1]}{2})/2
+  +(\mmaSup{v[n]}{2}-\mmaSup{m[n]}{2})/2,m[n]+m[n+1]v[n+1]-m[n]v[n]\}|>
+  DiscreteConservedQ[variables][\{v[n],m[n]\}]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  True
+\end{mmaCell}
 	\end{small}
 	Those are the only conserved quantities for this scheme, which are polynomials up to degree 4 in $v$ and $m$.
-	\begin{EXE}
-		LABEL="mfg3"
-		(*CODE*)
-		variables = Association[
-		"indVars" -> {n, t},
-		"depVars" -> {v, m},
-		"VarDOperator"->PartialDVarDOperator,
-		"timeVars"->{t},
-		"eqRhs" -> { 
-			(v[n + 1, t]^2/2 - m[n + 1, t]^2/2) - (v[n, t]^2/2 - 
-			m[n, t]^2/2), 
-			(m[n + 1, t]*v[n + 1, t]) - (m[n, t]*v[n, t])}
-		];
-		FindConservedQuantityBasisOperator[variables][Association["degree"->4,"generators"->{v[n,t],m[n,t]}]]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
+
 	\begin{small}
-		\input{executables/mfg3ctt.ed.tex}
-		\input{executables/mfg3ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, PartialDVarD, FindConservedQuantityBasis}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n,t\},"depVars"\(\pmb{\to}\)\{v,m\},
+  "eqRhs"\(\pmb{\to}\)\{0.5(\mmaSup{v[n+1,t]}{2}-\mmaSup{m[n+1,t]}{2})
+  -0.5(\mmaSup{v[n,t]}{2}-\mmaSup{m[n,t]}{2}),
+  m[n+1,t]v[n+1,t]-m[n,t]v[n,t]\}|>
+  FindConservedQuantityBasis[variables]
+  [<|"degree"\(\pmb{\to}\)4,"generators"\(\pmb{\to}\)\{v[n,t],m[n,t]\}|>]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  \{m[n,t],v[n,t]\}
+\end{mmaCell}
 	\end{small}
@@
 	Our code can handle this setting by specifying an order for variable elimination, using an explicit monomial order for $m$ and an implicit order for $v$:
-	\begin{EXE}
-		LABEL="mfg4"
-		(*CODE*)
-		variables = Association[
-		"indVars" -> {n, t},
-		"depVars" -> {m, v},
-		"elimOrder" -> "explicitimplicit",
-		"scheme" -> {-(v[n, t + 1] - 
-			v[n, t]) + (v[n + 1, t+1]^2/2 - m[n + 1, t+1]^2/2) - (v[n, t+1]^2/2 - 
-			m[n, t+1]^2/2), 
-			m[n, t + 1] - 
-			m[n, t] - (m[n + 1, t]*v[n + 1, t]) + (m[n, t]*v[n, t])}
-		];
-		DiscreteConservedQOperator[variables][{v[n, t], m[n, t]}]
-		(*CODE*)
-		PUNCT={,.}
-	\end{EXE}
+
 	\begin{small}
-		\input{executables/mfg4ctt.ed.tex}
-		\input{executables/mfg4ctt.result.ed.tex}
+		\begin{mmaCell}[moredefined={variables, DiscreteConservedQ}]{Input}
+  variables=<|"indVars"\(\pmb{\to}\)\{n,t\},"depVars"\(\pmb{\to}\)\{m,v\},
+  "elimOrder"\(\pmb{\to}\)"explicitimplicit",
+  "scheme"\(\pmb{\to}\)\{-(v[n,t+1]-v[n,t])
+  +(\mmaSup{v[n+1,t+1]}{2}-\mmaSup{m[n+1,t+1]}{2})/2
+  -(\mmaSup{v[n,t+1]}{2}-\mmaSup{m[n,t+1]}{2})/2,
+  m[n,t+1]-m[n,t]-m[n+1,t]v[n+1,t]+m[n,t]v[n,t]\}|>
+  DiscreteConservedQ[variables][\{v[n,t],m[n,t]\}]
+
+\end{mmaCell}
+		\begin{mmaCell}{Output}
+  True
+\end{mmaCell}
 	\end{small}

### 2023-05-09 Papers/Paper CASC format/paper.tex
...

@@
 {\color{blue} add 
-Comprehensive Gr{\"o}bner Systems 
+Comprehensive Gr{\"o}bner Systems 2222
 }
@@
     in the ring $\mathcal{K}[a_1,..,a_s]$. 
-    For simplicity we denote this ring by $\mathcal{K}[A][X]$.
+    For simplicity, we denote this ring by $\mathcal{K}[A][X]$.
 \end{definition}
@@
 \begin{definition}
-    Let $\Ii$ be a finitely generated, nonzero ideal in the ring $\mathcal{K}[A][X]$ with a fixed monomial order $\prec$ on the variables. 
+    Let $\Ii$ be a finitely generated, non-zero ideal in the ring $\mathcal{K}[A][X]$ with a fixed monomial order $\prec$ on the variables. 
     Let $\{A_1,\dots,A_k\}$ be a family of locally closed sets that forms a partition of $L^m$ 
@@
      is a Gr{\"o}bner basis of the specialized ideal $\sigma_i (I)\subset \mathcal{K}[X]$.
-     Where each pair $\{G_i,A_i\}$ is called a branch. 
+     Where each pair $\{G_i,A_i\}$ is called a branch. \cite{Kapur}
 \end{definition}
@@
 It branches the cases of the parameters that affect leading terms. For instance, in the previous
-example if $a=0$ then the leading term $a x^2$ vanishes which leads to a different Gr{\"o}bner basis. Therefore,
+example, if $a=0$, then the leading term $a x^2$ vanishes which leads to a different Gr{\"o}bner basis. Therefore,
 the (CGS) considers cases of the leading coefficients parameters Vanishing.
@@
 \begin{remark}
-    If a polynomial ideal contains a nonzero constant polynomial, then it is called a unit ideal $\langle 1\rangle$.
+    If a polynomial ideal contains a non-zero constant polynomial, then it is called a unit ideal $\langle 1\rangle$.
    Unit ideals can generate all polynomials in the ring, so $\langle 1\rangle=\mathcal{K}[A][X]$.
-    As a convention the Gr{\"o}bner basis of such ideals is $\{1\}$.
+    As a convention, the Gr{\"o}bner basis of such ideals is $\{1\}$.
 \end{remark}
@@
            is $\Gg=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$. Consider 
-           the specialization $\sigma \in \Rr^2$ given by $P\rightarrow P(0,1)$.
-           Now, We claim that the set $\sigma(\Gg)=\left\{y^3,-y^2,x y-2\right\}$ is not a Gr{\"o}bner basis.
-           To show that we find the S-polynomials of each pair in the set and we find that: $S(-y^2, -2 + x y)=-2 y$.
-           We divide $S(-y^2, -2 + x y)$ by $\sigma(\Gg)$ and find that the remainder is $-2y$. 
-           Therefore by the Buchberger Criterion, $\sigma(\Gg)$ is not a Gr{\"o}bner basis.
+           the specialization $\sigma$ given by $P\rightarrow P(0,1)$.
+           The set $\sigma(\Gg)=\left\{y^3,-y^2,x y-2\right\}$ is not a Gr{\"o}bner basis.
+		   This can be shown by the Buchberger Criterion (see Chapter 2 in\cite{Cox}).
+
+        %    To show that we find the S-polynomials of each pair in the set and we find that: $S(-y^2, -2 + x y)=-2 y$.
+        %    We divide $S(-y^2, -2 + x y)$ by $\sigma(\Gg)$ and find that the remainder is $-2y$. 
+        %    Therefore by the Buchberger Criterion, $\sigma(\Gg)$ is not a Gr{\"o}bner basis.
     \end{example}
     \item Consequently, the algorithm utilized to resolve the ideal membership 
-    problem in \ref{sec2.1} can not be used for all specializations.
+    problem in chapter 2 of \cite{Cox} can not be used for all specializations.
 
@@
         the specialization $\sigma$ given by $P\rightarrow P(0,1)$, $\sigma(\Gg)$ is not a Gr{\"o}bner basis as shown in the above example. 
-        Therefore, we can not proceed with the division algorithm as we can not guarantee unique remainder if we used $\sigma(\Gg)$.
-
+        Therefore, we can not proceed with the division algorithm as we can not guarantee a unique remainder if we use $\sigma(\Gg)$.  
     \end{example}
@@
 \end{enumerate}
+
+Using CGS can resolve these problems. Problem 1 is fixed by the definition of (CGS), as we have $\sigma_i(G_i)$ a 
+ Gr{\"o}bner basis for any specialization $\sigma_i$. Moreover, 
+ to discuss problem 2, we need to look at the parametric ideal membership first.
+
+{\color{red} Is this a definition?}
+ Consider, $f\in \Kk[A][X]$ we say that $f\in\Ii$ under $\sigma$, if there exists a specialization 
+ $\sigma\in L^m$ such that $\sigma(f)\in\sigma(\Ii)$.\\
+
+To determine if $f\in\Ii$ we consider the (CGS) of $\Ii$ given \ref{CGS}. 
+We check if $f\in\sigma_i(\Ii)$ for each $i\in\{1,..,k\}$ and $\sigma_i\in A_i$.
+For each $\sigma_i(\Ii)$ we have the corresponding Gr{\"o}bner basis $\sigma_i(G_i)$
+so we divide $f$ by the $\sigma_i(G_i)$ if the remainder is zero for some $i\in\{1,..,k\}$,
+then $f\in\sigma_i(\Ii)$. Therefore, $f\in\Ii$. However, if the remainder is non-zero for all $i$ we conclude that 
+$f\notin\Ii$.
+
+\begin{example}
+	Let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
+    $$\Ff=\{a^2 x - y, -b^2 x y^2 - x\}.$$
+	We will determine if $f(x,y)=-x+y^2\in\mathcal{I}$.
+	We compute the corresponding (CGS)
+	{\color{red} add example 3}\\
+	First let $A_1=\{(a,b)~|~ a\ne 0, b\ne 0\}$ and $A_2=\{(a,b)~|~ ab=0\}$. 
+	We divide $f$ by $\{\frac{y}{b^2}+y^3,x-\frac{y}{a^2}\}$ and we obtain the remainder $-\frac{y}{a^2}+y^2$ and 
+	since the exists no specialization in $A_1$ such that this remainder is zero we conclude that $f\notin \sigma_1(\Ii)$ for
+	$\sigma_1\in A_1$. Now we repeat the reduction of $f$ but with the second Gr{\"o}bner basis $\{x,y\}$. We obtain a zero remainder, which implies that 
+	$f\in \sigma_2(\Ii)$ for $\sigma_2\in A_2$. Therefore, $f\in\Ii$ under a specialization $\sigma_2\in A_2$.
+\end{example}
 

### 2023-05-09 Papers/Paper CASC format/tmp/ANTpaper.tex
...

StringJoin["\\documentclass{amsart}\n\\usepackage{color}\n\\begin{document}\n\\section{Basic Text Statistics}\n \\[\\begin{array}{ll}\n \\text{Language = } & \\text{LanguageIdentify}[\\text{StringReplace}[\\text{StringReplace}[\\text{StringReplace}[\\text{$\\$$Failed},\\{\\text{$\\backslash $r$\\backslash $n}\\to \\text{},\\text{$\\backslash $n}\\to \\text{},\\text{Whitespace}\\to  \\}],\\{,\\to \\text{},\\text{[1]}\\to \\text{}\\}],\\{\\text{Whitespace}\\text{..}\\sim \\sim .\\to .,\\text{Whitespace}\\text{..}\\to  \\}]] \\\\\n \\text{Sentiment = } & \\text{ClassifierFunction}\\left[\\right](\\text{StringReplace}[\\text{StringReplace}[\\text{StringReplace}[\\text{$\\$$Failed},\\{\\text{$\\backslash $r$\\backslash $n}\\to \\text{},\\text{$\\backslash $n}\\to \\text{},\\text{Whitespace}\\to  \\}],\\{,\\to \\text{},\\text{[1]}\\to \\text{}\\}],\\{\\text{Whitespace}\\text{..}\\sim \\sim .\\to .,\\text{Whitespace}\\text{..}\\to  \\}],\\text{AcceptanceThreshold}\\to \\text{Automatic},\\text{AnomalyDetector}\\to \\text{None},\\text{BatchProcessing}\\to \\text{Automatic},\\text{ClassPriors}\\to \\text{Automatic},\\text{IndeterminateThreshold}\\to \\text{Automatic},\\text{MissingValueSynthesis}\\to \\text{Automatic},\\text{PerformanceGoal}\\to \\text{Automatic},\\text{RandomSeeding}\\to 1234,\\text{RecalibrationFunction}\\to \\text{Automatic},\\text{TargetDevice}\\to \\text{CPU},\\text{TieBreakerFunction}\\to \\text{RandomChoice},\\text{UtilityFunction}\\to \\text{Automatic}) \\\\\n \\text{$\\#$ Characters = } & \\text{Total}[\\text{StringLength}[\\text{TextWords}[\\text{StringReplace}[\\text{StringReplace}[\\text{StringReplace}[\\text{$\\$$Failed},\\{\\text{$\\backslash $r$\\backslash $n}\\to \\text{},\\text{$\\backslash $n}\\to \\text{},\\text{Whitespace}\\to  \\}],\\{,\\to \\text{},\\text{[1]}\\to \\text{}\\}],\\{\\text{Whitespace}\\text{..}\\sim \\sim .\\to .,\\text{Whitespace}\\text{..}\\to  \\}]]]] \\\\\n \\text{$\\#$ Words = } & 1 \\\\\n \\text{$\\#$ Sentences = } & 1 \\\\\n \\text{$\\#$ Words/Sentence = } & 1. \\\\\n \\text{$\\#$ Distinct words = } & 1 \\\\\n \\text{$\\#$ Distinct words (without stop words) = } & 1 \\\\\n \\text{Automated Readibility index = } & 4.71 \\text{Total}[\\text{StringLength}[\\text{TextWords}[\\text{StringReplace}[\\text{StringReplace}[\\text{StringReplace}[\\text{$\\$$Failed},\\{\\text{$\\backslash $r$\\backslash $n}\\to \\text{},\\text{$\\backslash $n}\\to \\text{},\\text{Whitespace}\\to  \\}],\\{,\\to \\text{},\\text{[1]}\\to \\text{}\\}],\\{\\text{Whitespace}\\text{..}\\sim \\sim .\\to .,\\text{Whitespace}\\text{..}\\to  \\}]]]]-20.93 \\\\\n \\text{Coleman-Liau index = } & 5.88 \\text{Total}[\\text{StringLength}[\\text{TextWords}[\\text{StringReplace}[\\text{StringReplace}[\\text{StringReplace}[\\text{$\\$$Failed},\\{\\text{$\\backslash $r$\\backslash $n}\\to \\text{},\\text{$\\backslash $n}\\to \\text{},\\text{Whitespace}\\to  \\}],\\{,\\to \\text{},\\text{[1]}\\to \\text{}\\}],\\{\\text{Whitespace}\\text{..}\\sim \\sim .\\to .,\\text{Whitespace}\\text{..}\\to  \\}]]]]-45.4 \\\\\n \\text{SMOG grade = } & 3.1291 \\\\\n \\text{Glue index = } & 0. \\\\\n \\text{Word repetition index = } & 0. \\\\\n\\end{array}\\]\n\n\\section{Sentence Analysis}", StringReplace[TextSentences["\n\n\\noindent {\\color{green} ********************}\n\n\n\\noindent {\\color{red} "~~If[-20.93 + 4.71*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]] > Quantile[TextSentences[{-20.93 + 4.71*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], -45.4 + 5.88*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], 3.1291, 0.}], 93/100] || -45.4 + 5.88*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]] > Transpose[Quantile[TextSentences[{-20.93 + 4.71*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], -45.4 + 5.88*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], 3.1291, 0.}], 93/100]][[2]] || 3.1291 > Transpose[Quantile[TextSentences[{-20.93 + 4.71*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], -45.4 + 5.88*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], 3.1291, 0.}], 93/100]][[3]], " {\\color{blue} (Low readability) } ", ""]~~If[0. > Transpose[Quantile[TextSentences[{-20.93 + 4.71*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], -45.4 + 5.88*StringLength[StringReplace[ToLowerCase[StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]], "[1]" -> ""]], 3.1291, 0.}], 93/100]][[4]], " {\\color{blue} (High-glue sentence) } ", ""]~~" {\\color{blue} (Style problems: }  {\\color{green}  {\\bf Hedging and Vague}: almost, a lot of, and so on, apparently, appears, appears to be, appear to be, appropriate, approximately, are sure, arguably, assume, believe, be sure, comparatively, conceivably, could be the case that, doubt, etc, fairly, feel, frequently, generally, in general, has been generally agreed that, help, helps, hope, hopeful, hopefully, if true, indicate, indicates, in part, is generally agreed that, is important to develop, is sure, is useful to study, like, likely, look, looks, may, may be possible, may be possible to obtain, may be the case that, might, might be suggested that, my view, nearly, occasionally, often, our view, partially, perhaps, possibility, possible, possibly, practically, predominantly, presumably, probable, rather, reasonable, reckon, relatively, reportedly, roughly, seem, seemingly, seems, should, sometimes, somewhat, sort of, so to speak, stuff, suggest, suggests, suitable, tend, tends, there is every hope that, there is every hope that, thing, things like that, things like this, think, this kind, to a certain degree, to our knowledge, to some extent, type of, unlikely, usually, vague, virtually, whatever, whenever, wherever, whichever, whoever, would, would argue;  {\\bf Intensifiers}: completely, extremely, highly, novel, quite, really, totally, very;  {\\bf Cliches}: bread and butter, change the subject, changing the subject, for the first time, game changer, give and take, give or take, look at the facts, looking at the facts, novel, the long and the short;  {\\bf Abstract names}: framework, issues, level, model, perspective, prospects, strategy;  {\\bf Wordiness}: are allowed, is allowed, absolute, actually, already, alternative, as a matter of fact, as the ability to, at such time as, at that point in time, at the present time, at this point in time, based on the fact that, basically, brought about the organization of, call your attention to the fact that, certain, complete, completely, continue to, despite the fact that, due partly to the fact that, due to the fact that, during the course of, empty, equal, eternal, extremely, fact of the matter, for the purpose of, has the potential to, have a facilitative impact, have the ability to, I might add that, in light of the fact that, in order, in order to, in spite of the fact that, in the event that, in the final analysis, in the matter of, in the vicinity of, in view of the fact that, is able to, is capable of being, is dedicated to providing, it is imperative that we, it is noteworthy, it is often the case that, it is significant that, it should be noted, it should be pointed out that, kind of, less, make an appearance with, more, on a daily basis, on a weekly basis, or the same, owing to the fact that, perfect, practically, previously, quite, really, relating to the subject of, relatively, significantly expedite the process of, somewhat, still, take under consideration, the course of, the fact that, the presence of, the question as to whether, the reason is, the reason why, there is no doubt but that, this is a subject that, to be of the opinion, to make reference to, unique, until such a time as, very, were in great need of, which is, which was, who is, who was, widely observed that; } {\\color{blue} ) }  {\\color{blue} (No verb) }\n\n\\noindent "~~StringReplace[StringReplace[$Failed, {"\r\n" -> "", "\n" -> "", Whitespace -> " "}], {Whitespace..~~"." -> ".", Whitespace.. -> " "}]~~"}\n\n\\noindent {\\color{green} ********************}\n\n\\noindent "], "." -> ".\n\n\\noindent "], "\\end{document}"]

### 2023-05-09 Papers/Paper CASC format/tmp/CPpaper.tex
...

--- 2023-05-09 Papers/Paper CASC format/MLaTeX/CLpaper.tex
@@
 \usepackage{mmacells}
+\usepackage[invisible]{MLaTeX}
 
@@
 	The discrete time derivative for $\sum_n u(n,t)$ is
+	\begin{EXE}
+		LABEL="heateq2"
+		(*CODE*)
+		variables = <|"indVars"->{n,t},"depVars"->{u}|>
+		expression = u[n,t]
+		TimeDifferenceOperator[variables][expression]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 
@@
 	 the general setting should be applied.
-
+	\begin{EXE}
+		LABEL="heateq10"
+		(*CODE*)
+		variables = Association["indVars"->{n,t},"depVars"->{u},"scheme"->{u[n,t+1]-u[n,t]-(u[n+1,t]-2*u[n,t]+u[n-1,t])}]
+		expression = u[n,t]
+		DiscreteConservedQOperator[variables][expression]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 In the context of the heat equation from Example \ref{heatex1}, we aim to find a basis for conserved quantities generated by $u(n,t)$ and $n$ with a maximum degree of 3.
-
+	\begin{EXE}
+		LABEL="heateq11"
+		(*CODE*)
+		variables = <|"indVars"->{n},"depVars"->{u},"rhs"->{u[n]+u[n+1]-2*u[n]+u[n-1]}|>
+		FindDiscreteConservedQuantityBasisOperator[variables][<|"degree"->3,"generators"->{u[n],n}|>]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 	We discretize $u_x$ with a three-point stencil with a parameter $a$.
-
+	\begin{EXE}
+		LABEL="burgers2"
+		(*CODE*)
+		variables = Association["indVars"->{n,t}, "depVars"->{u}, "pars"->{a}, "scheme"->{u[n,t+1]-u[n,t]-u[n,t]*(a*(u[n+1,t]-u[n,t])+(1-a)*(u[n,t]-u[n-1,t]))}];
+		expression = u[n,t];
+		DiscreteConservedQOperator[variables][expression]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 We replicate this result, observing that the scheme used here is time-explicit:
-
+	\begin{EXE}
+		LABEL="mfg2"
+		(*CODE*)
+		variables = Association[
+		"indVars" -> {n},
+		"depVars" -> {v, m},
+		"rhs" -> { 
+			v[n] -(v[n + 1]^2/2 - m[n + 1]^2/2) + (v[n]^2/2 - 
+			m[n]^2/2), 
+			m[n] + (m[n + 1]*v[n + 1]) - (m[n]*v[n])}
+		];
+		DiscreteConservedQOperator[variables][{v[n], m[n]}]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 	Those are the only conserved quantities for this scheme, which are polynomials up to degree 4 in $v$ and $m$.
-
+	\begin{EXE}
+		LABEL="mfg3"
+		(*CODE*)
+		variables = Association[
+		"indVars" -> {n, t},
+		"depVars" -> {v, m},
+		"VarDOperator"->PartialDVarDOperator,
+		"timeVars"->{t},
+		"eqRhs" -> { 
+			(v[n + 1, t]^2/2 - m[n + 1, t]^2/2) - (v[n, t]^2/2 - 
+			m[n, t]^2/2), 
+			(m[n + 1, t]*v[n + 1, t]) - (m[n, t]*v[n, t])}
+		];
+		FindConservedQuantityBasisOperator[variables][Association["degree"->4,"generators"->{v[n,t],m[n,t]}]]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}
@@
 	Our code can handle this setting by specifying an order for variable elimination, using an explicit monomial order for $m$ and an implicit order for $v$:
-
+	\begin{EXE}
+		LABEL="mfg4"
+		(*CODE*)
+		variables = Association[
+		"indVars" -> {n, t},
+		"depVars" -> {m, v},
+		"elimOrder" -> "explicitimplicit",
+		"scheme" -> {-(v[n, t + 1] - 
+			v[n, t]) + (v[n + 1, t+1]^2/2 - m[n + 1, t+1]^2/2) - (v[n, t+1]^2/2 - 
+			m[n, t+1]^2/2), 
+			m[n, t + 1] - 
+			m[n, t] - (m[n + 1, t]*v[n + 1, t]) + (m[n, t]*v[n, t])}
+		];
+		DiscreteConservedQOperator[variables][{v[n, t], m[n, t]}]
+		(*CODE*)
+		PUNCT={,.}
+	\end{EXE}
 	\begin{small}

### 2023-05-09 Papers/Paper CASC format/tmp/FCpaper.tex
...

--- 2023-05-09 Papers/Paper CASC format/MLaTeX/CLpaper.tex
@@
 
-\begin{document}
+\usepackage{color}\begin{document}
+
 \title{Algorithmic detection of conserved quantities for finite-difference schemes\thanks{Supported by King Abdullah University of Science and Technology (KAUST) baseline funds and KAUST OSR-CRG2021-4674.}}
@@
 
-{\color{red} $\Phi$ and $\Psi$ need domains and ranges, also explain that divergence is only in $x$}
+{\color{red} \(\Phi\)
+{\color{blue} New variable(s): \(\Phi\),  }
+ and \(\Psi\)
+{\color{blue} New variable(s): \(\Psi\),  }
+ need domains and ranges, also explain that divergence is only in \(x\)
+{\color{blue} New variable(s): \(x\),  }
+}
 
@@
 \begin{gather*}
-	\frac{\partial }{\partial t}\Phi[u]+\div \Psi[u]=0
-\end{gather*}
-holding for all solutions $u$ of the considered PDE,
-that induces the conserved quantity $\int \Phi[u]\ dx$ as
+	\frac{\partial }{\partial t}\Phi[u]+\operatorname{div} \Psi[u]=0
+\end{gather*}
+holding for all solutions \(u\)
+{\color{blue} New variable(s): \(u\),  }
+ of the considered PDE,
+that induces the conserved quantity \(\int \Phi[u]\ dx\)
+{\color{blue} New variable(s): \(d\),  }
+ as
 \begin{gather*}
@@
 \end{gather*}
-assuming periodic boundary conditions in $x$.
+assuming periodic boundary conditions in \(x\).
 
@@
 In this paper, we present an algorithm that verifies if a quantity is conserved over time under a given scheme.
-Our approach diverges from previous methods, as we do not attempt to construct discrete $\Phi$ and $\Psi$ using the scheme, but rather check if a given discrete $\Phi$ is conserved over time under the scheme.
+Our approach diverges from previous methods, as we do not attempt to construct discrete \(\Phi\) and \(\Psi\) using the scheme, but rather check if a given discrete \(\Phi\) is conserved over time under the scheme.
 Typically, conserved quantities involve integral expressions or their discrete analogs, such as sums.
@@
 proposed algorithms for simplifying sums \cite{gomes20} and developed algorithms for detecting conserved quantities in PDEs and semi-discretized schemes \cite{gomes20}.
-We review these techniques in Section \ref{assp}.
-To generalize these methods for cases without summing over all arguments of the involved functions, we introduce the discrete partial variational derivative (Section \ref{pvard}).
+We review these techniques in Section \ref{assp}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{assp}\end{verbatim}.}
+.
+To generalize these methods for cases without summing over all arguments of the involved functions, we introduce the discrete partial variational derivative (Section \ref{pvard}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{pvard}\end{verbatim}.}
+).
 Our main contribution is an algorithm for verifying the conservation of a quantity under a numerical scheme.
@@
 Furthermore, Gerdt's algorithm may not always terminate, as the Gr{\"o}bner basis for the difference ideal might be infinite.
-We address these issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) having a finite Gr{\"o}bner basis (Section \ref{algodetec}).
+We address these issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) having a finite Gr{\"o}bner basis (Section \ref{algodetec}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{algodetec}\end{verbatim}.}
+).
 Our algorithm is applicable to schemes that are explicit and implicit in time, can handle schemes with multiple higher-order time and space derivatives, multiple space dimensions, systems of equations, and schemes with parameters.
@@
 We employ parametric Gr{\"o}bner  bases, which are crucial for the proper treatment of parameters in our algorithm, see Section \ref{pgb}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{pgb}\end{verbatim}.}
+
  We have implemented this algorithm as a part of a {\sc Mathematica} package \cite{wolfram21}\footnote{The code is available upon request.
 For access, contact ricardo.ribeiro@kaust.edu.sa.}.
-Our examples include Burgers equation and a PDE system related to mean-field games (Section \ref{eaa}).
-In these examples,
+Our examples include Burgers equation and a PDE system related to mean-field games (Section \ref{eaa}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{eaa}\end{verbatim}.}
+). In these examples, 
  we show that our code detects conserved quantities and can be used to select suitable schemes for time-implicit and time-explicit discretizations.
@@
 In this paper, subscripts denote coordinate or tuple indices, and superscripts denote sequence indices.
-To avoid boundary terms, we work with periodic functions in $\Zz^d$.
-
-When discretizing PDEs using finite difference schemes, we need to work with functions defined on discrete sets.
-For instance, periodic functions in $\Rr^d$ can be replaced by functions defined on the discrete torus, which will be introduced in the following discussion.
+To avoid boundary terms, we work with periodic functions in \({\mathbb{Z}}^d\)
+{\color{blue} New variable(s): \(Z\),  }
+.
+
+When discretizing PDEs using finite difference schemes, we need to work with functions defined on discrete sets. For instance, periodic functions in \(\mathbb{R}^d\)
+{\color{blue} New variable(s): \(R\),  }
+ can be replaced by functions defined on the discrete torus, which will be introduced in the following discussion.
 \begin{definition}
 	\label{d1}
-	Let $N$, $d$ and $m$ be positive integers. The \emph{discrete torus} is the set $\Ii:=\{0,\ldots,N-1\}^d\subset \Zz^d$. Define the space
-	\begin{gather*}
-		\mathcal{P}(\Ii,\Rr^m):=\{u\ |\ u:\Ii\rightarrow\Rr^m\}.
-	\end{gather*}
-Functions in $\mathcal{P}(\Ii,\Rr^m)$ extend periodically to $\Zz^d$.
+	Let \(N\)
+{\color{blue} New variable(s): \(N\),  }
+, \(d\) and \(m\)
+{\color{blue} New variable(s): \(m\),  }
+ be positive integers. The \emph{discrete torus} is the set \(\mathcal{I}:=\{0,\ldots,N-1\}^d\subset {\mathbb{Z}}^d\). Define the space
+	\begin{gather*}
+		\mathcal{P}(\mathcal{I},\mathbb{R}^m):=\{u\ |\ u:\mathcal{I}\rightarrow\mathbb{R}^m\}.
+	\end{gather*}
+Functions in \(\mathcal{P}(\mathcal{I},\mathbb{R}^m)\)
+{\color{blue} New variable(s): \(P\),  }
+ extend periodically to \({\mathbb{Z}}^d\).
 \end{definition}
@@
 	\label{d2}
-	The space of functionals $\mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$ (not necessarily linear) on $\mathcal{P}(\Ii,\Rr^m)$ is
-	\begin{gather*}
-		\mathcal{F}(\mathcal{P}(\Ii,\Rr^m)):= \Bigg\{\mathcal{J}:\mathcal{P}(\Ii,\Rr^m)\to \Rr\ \Bigg|\ \mathcal{J}[u]=\sum_{n\in\Ii}F_n[u],\\
+	The space of functionals \(\mathcal{F}(\mathcal{P}(\mathcal{I},\mathbb{R}^m))\)
+{\color{blue} New variable(s): \(F\),  }
+ (not necessarily linear) on \(\mathcal{P}(\mathcal{I},\mathbb{R}^m)\) is
+	\begin{gather*}
+		\mathcal{F}(\mathcal{P}(\mathcal{I},\mathbb{R}^m)):= \Bigg\{\mathcal{J}:\mathcal{P}(\mathcal{I},\mathbb{R}^m)\to \mathbb{R}\ \Bigg|\ \mathcal{J}[u]=\sum_{n\in\mathcal{I}}F_n[u],\\
 		\text{ where }
 		F_n[u]=G(u(n+e^1),u(n+e^2),\ldots,u(n+e^k))\\
-		\text{ and }\{e^1,\ldots,e^k\}\subset\Zz^d,\ \text{G is smooth}
+		\text{ and }\{e^1,\ldots,e^k\}\subset{\mathbb{Z}}^d,\ \text{G is smooth}
 		\Bigg\}.
@@
 \end{definition}
-In the previous definition, $e^1,\ldots,e^k$ are not necessarily unit vectors and may vary between functionals.
+In the previous definition, \(e^1,\ldots,e^k\)
+{\color{blue} New variable(s): \(k\),  }
+ are not necessarily unit vectors and may vary between functionals.
 \begin{example}
 	\label{functional1}
-	Let $d=m=1$ and
-	$	\mathcal{J}[u]:=\sum_{n}(u(n+1)-u(n))^2$.
+	Let \(d=m=1\) and
+	\(	\mathcal{J}[u]:=\sum_{n}(u(n+1)-u(n))^2\)
+{\color{blue} New variable(s): \(J\), \(n\),  }
+.
 \end{example}
 The discrete variational derivative is a useful tool for simplifying sums.
-Let $u,v\in\mathcal{P}(\Ii,\Rr^m)$, $\epsilon 
-\in\Rr$ and $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$. Define
+Let \(u,v\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\)
+{\color{blue} New variable(s): \(v\),  }
+, \(\epsilon 
+\in\mathbb{R}\)
+{\color{blue} New variable(s): \(\epsilon\),  }
+ and \(\mathcal{J}\in\mathcal{F}(\mathcal{P}(\mathcal{I},\mathbb{R}^m))\). Define
 \begin{gather*}
@@
 \end{gather*}
-In \eqref{periodicity}, we shift the indices because of the periodicity of $u$ and $v$,
+In \eqref{periodicity}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\eqref{periodicity}\end{verbatim}.}
+, we shift the indices because of the periodicity of \(u\) and \(v\),
 \begin{gather}
-	\notag	D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\Ii}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
-	\notag	=\sum_{n\in\Ii}\frac{d}{d\epsilon}G(u(n+e^1)+\epsilon v(n+e^1),\ldots,u(n+e^k)+\epsilon v(n+e^k))\bigg|_{\epsilon=0}\\
-	\notag	=\sum_{n\in\Ii}D_1G(u(n+e^1),\ldots,u(n+e^k))v(n+e^1)\\
+	\notag	D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\mathcal{I}}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
+	\notag	=\sum_{n\in\mathcal{I}}\frac{d}{d\epsilon}G(u(n+e^1)+\epsilon v(n+e^1),\ldots,u(n+e^k)+\epsilon v(n+e^k))\bigg|_{\epsilon=0}\\
+	\notag	=\sum_{n\in\mathcal{I}}D_1G(u(n+e^1),\ldots,u(n+e^k))v(n+e^1)\\
 	\notag	+\ldots+D_kG(u(n+e^1),\ldots,u(n+e^k))v(n+e^k)\\
-	\label{periodicity}	=\sum_{n\in\Ii}D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))v(n)\\
+	\label{periodicity}	=\sum_{n\in\mathcal{I}}D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))v(n)\\
 	\notag	+D_2G(u(n+e^1-e^2),u(n),\ldots,u(n+e^k-e^2))v(n)\\
 	\notag	+\ldots+D_kG(u(n+e^1-e^k),u(n+e^2-e^k),\ldots,u(n))v(n)\\
-	\notag	=\sum_{n\in\Ii}\Big(D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))\\
+	\notag	=\sum_{n\in\mathcal{I}}\Big(D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))\\
 	\notag	+D_2G(u(n+e^1-e^2),u(n),\ldots,u(n+e^k-e^2))\\
 	\notag	+\ldots+D_kG(u(n+e^1-e^k),u(n+e^2-e^k),\ldots,u(n))\Big)v(n)\\
-	\notag	=:\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n).	
+	\notag	=:\sum_{n\in\mathcal{I}}\mathcal{V}(\mathcal{J})[u](n)v(n).	
 \end{gather}
-The operator $\mathcal{V}(\mathcal{J})$ can be viewed as a representation of the derivative $D\mathcal{J}u$ with respect to the $L^2$ inner product $\langle w, v\rangle=\sum_{n\in \Ii} w_n v_n$, leading to the following definition.
-\begin{definition}
-	$\mathcal{V}(\mathcal{J})$ is the \emph{discrete variational derivative} of $\mathcal{J}$.
+The operator \(\mathcal{V}(\mathcal{J})\)
+{\color{blue} New variable(s): \(V\),  }
+ can be viewed as a representation of the derivative \(D\mathcal{J}u\)
+{\color{blue} New variable(s): \(D\),  }
+ with respect to the \(L^2\)
+{\color{blue} New variable(s): \(L\),  }
+ inner product \(\langle w, v\rangle=\sum_{n\in \mathcal{I}} w_n v_n\)
+{\color{blue} New variable(s): \(w\), \(v_n\), \(w_n\),  }
+, leading to the following definition.
+\begin{definition}
+	\(\mathcal{V}(\mathcal{J})\) is the \emph{discrete variational derivative} of \(\mathcal{J}\).
 \end{definition}
@@
 	\begin{gather*}
-		D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\Ii}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
-		=\sum_{n\in\Ii}\frac{d}{d\epsilon}G[u(n+1)+\epsilon v(n+1),u(n)+\epsilon v(n)]\bigg|_{\epsilon=0}\\
-		=\sum_{n\in\Ii}2(u(n+1)-u(n))v(n+1)+(-2)(u(n+1)-u(n))v(n)\\
-		=\sum_{n\in\Ii}2(u(n)-u(n-1))v(n)+(-2)(u(n+1)-u(n))v(n)\\
-		=\sum_{n\in\Ii}\Big(2(u(n)-u(n-1))+(-2)(u(n+1)-u(n))\Big)v(n)\\
-		=\sum_{n\in\Ii}\Big(-2u(n+1)+4u(n)-2u(n-1)\Big)v(n)\\
-		=\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n).
-	\end{gather*}
-Thus, $\mathcal{V}(\mathcal{J})[u](n)=-2u(n+1)+4u(n)-2u(n-1)$.
+		D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\mathcal{I}}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
+		=\sum_{n\in\mathcal{I}}\frac{d}{d\epsilon}G[u(n+1)+\epsilon v(n+1),u(n)+\epsilon v(n)]\bigg|_{\epsilon=0}\\
+		=\sum_{n\in\mathcal{I}}2(u(n+1)-u(n))v(n+1)+(-2)(u(n+1)-u(n))v(n)\\
+		=\sum_{n\in\mathcal{I}}2(u(n)-u(n-1))v(n)+(-2)(u(n+1)-u(n))v(n)\\
+		=\sum_{n\in\mathcal{I}}\Big(2(u(n)-u(n-1))+(-2)(u(n+1)-u(n))\Big)v(n)\\
+		=\sum_{n\in\mathcal{I}}\Big(-2u(n+1)+4u(n)-2u(n-1)\Big)v(n)\\
+		=\sum_{n\in\mathcal{I}}\mathcal{V}(\mathcal{J})[u](n)v(n).
+	\end{gather*}
+Thus, \(\mathcal{V}(\mathcal{J})[u](n)=-2u(n+1)+4u(n)-2u(n-1)\).
 \end{example}
@@
 \begin{theorem}\label{theo1}
-	Let $\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$.
-If $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[v]=0$ for all $v\in\mathcal{P}(\Ii,\Rr^m)$ and if there exists $u_0\in\mathcal{P}(\Ii,\Rr^m)$ such that $\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]$, then
-	$\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$
-	for all $u\in\mathcal{P}(\Ii,\Rr^m)$.
+	Let \(\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\mathcal{I},\mathbb{R}^m))\).
+If \(\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[v]=0\) for all \(v\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\) and if there exists \(u_0\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\)
+{\color{blue} New variable(s): \(u_0\),  }
+ such that \(\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]\), then
+	\(\mathcal{J}[u]=\tilde{\mathcal{J}}[u]\)
+	for all \(u\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\).
 	Conversely,
-	$\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u]=0$
-	for all $u\in\mathcal{P}(\Ii,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
+	\(\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u]=0\)
+	for all \(u\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\), if \(\mathcal{J}[u]=\tilde{\mathcal{J}}[u]\).
 \end{theorem}
@@
 \begin{example}
-	Let $d=m=1$ and consider the functionals
-	\begin{gather*}
-		\mathcal{J}[u]:=\sum_{n\in\Ii}u(n)u(n+1)\text{ and }
-		\tilde{\mathcal{J}}[u]:=\sum_{n\in\Ii}u(n-2)u(n-1).
-	\end{gather*}
-	It is clear that $\mathcal{J}$ and $\tilde{\mathcal{J}}$ represent the same quantity (by shifting $n$). We confirm this, using the discrete variational derivative:
+	Let \(d=m=1\) and consider the functionals
+	\begin{gather*}
+		\mathcal{J}[u]:=\sum_{n\in\mathcal{I}}u(n)u(n+1)\text{ and }
+		\tilde{\mathcal{J}}[u]:=\sum_{n\in\mathcal{I}}u(n-2)u(n-1).
+	\end{gather*}
+	It is clear that \(\mathcal{J}\) and \(\tilde{\mathcal{J}}\) represent the same quantity (by shifting \(n\)). We confirm this, using the discrete variational derivative:
 	\begin{gather*}
 		D(\mathcal{J}-\tilde{\mathcal{J}})[u](v)=\\
-		=\sum_{n\in\Ii}(u(n+1)+u(n-1)-u(n+1)-u(n-1))v(n)=0.
-	\end{gather*}
-	Hence, $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})=0$ and $\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]$. Therefore, both functionals represent the same quantity.
-\end{example}
-
-\begin{example}
-	Let $d=2$ and $m=1$. It may not be obvious, that
+		=\sum_{n\in\mathcal{I}}(u(n+1)+u(n-1)-u(n+1)-u(n-1))v(n)=0.
+	\end{gather*}
+	Hence, \(\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})=0\) and \(\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]\). Therefore, both functionals represent the same quantity.
+\end{example}
+
+\begin{example}
+	Let \(d=2\) and \(m=1\). It may not be obvious, that
 	\begin{gather*}
@@
 	\end{gather*}
-	We confirm this by computing the discrete variational derivative and noticing that the expression is $0$ for $u=0$.
+	We confirm this by computing the discrete variational derivative and noticing that the expression is \(0\)
+{\color{blue} New variable(s): \(\text{Null}\),  }
+ for \(u=0\).
 \end{example}
@@
 This generalization provides a more flexible approach to handling various problems in conservation laws for finite-difference schemes, addressing challenges and limitations of the traditional discrete variational derivative.
-The main result of this section is Theorem \ref{theo2}, which gives a criterion for equality of two functionals based on their discrete partial variational derivatives."
-
-Let $n=(n_1,\ldots,n_d)$ be the variables for the functions in $\mathcal{P}(\Ii,\Rr^m)$.
-We call $n$ the space variables.
-Let $l\in\Nn$ and call $t:=(t_1,\ldots,t_l)\in\Nn^l$ the time variables.
-For generality, we consider the vectorial time case, while later in this paper, we focus specifically on the case when $l=1$.
+The main result of this section is Theorem \ref{theo2}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{theo2}\end{verbatim}.}
+, which gives a criterion for equality of two functionals based on their discrete partial variational derivatives."
+
+Let \(n=(n_1,\ldots,n_d)\)
+{\color{blue} New variable(s): \(n_1\), \(n_d\),  }
+ be the variables for the functions in \(\mathcal{P}(\mathcal{I},\mathbb{R}^m)\). We call \(n\) the space variables. Let \(l\in\mathbb{N}\)
+{\color{blue} New variable(s): \(l\),  }
+ and call \(t:=(t_1,\ldots,t_l)\in\mathbb{N}^l\)
+{\color{blue} New variable(s): \(t\), \(t_1\), \(t_l\),  }
+ the time variables. For generality, we consider the vectorial time case, while later in this paper, we focus specifically on the case when \(l=1\).
 The analog to Definition \ref{d1} is the following. 
 \begin{definition}
-	Let $\mathcal{P}(\Ii\times\Nn^l,\Rr^m):=\{u\ |\ u(\cdot,t)\in\mathcal{P}(\Ii,\Rr^m)\ \forall t\in\Nn^l \}$.
-	Functions in $\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ extend to $\Zz^d\times\Nn^l$ through  periodicity in the space variables.
-	Let $\mathcal{D}(\Nn^l):=\{u\ |\ u:\Nn^l\rightarrow\Rr\}$.
-\end{definition}
-We can regard $\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ as the set of functions depending in space and time and $\mathcal{D}(\Nn^l)$ as the functions depending only on time.
+	Let {\color{red} \(\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m):=\{u\ |\ u(\cdot,t)\in\mathcal{P}(\mathcal{I},\mathbb{R}^m)\ \forall t\in\mathbb{N}^l \}\)
+
+  Equation errors and warnings:
+
+ Unbalanced $|\hdots|$}
+
+.
+	Functions in \(\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m)\) extend to \({\mathbb{Z}}^d\times\mathbb{N}^l\) through  periodicity in the space variables.
+	Let {\color{red} \(\mathcal{D}(\mathbb{N}^l):=\{u\ |\ u:\mathbb{N}^l\rightarrow\mathbb{R}\}\)
+
+  Equation errors and warnings:
+
+ Unbalanced $|\hdots|$}
+
+.
+\end{definition}
+We can regard \(\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m)\) as the set of functions depending in space and time and \(\mathcal{D}(\mathbb{N}^l)\) as the functions depending only on time.
 
@@
 	\label{quantityfunc2}
-	The space of functions $\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$ is
-	\begin{gather*}
-		\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l)):=\\	
-		\Bigg\{\mathcal{J}:\mathcal{P}(\Ii\times\Nn^l)\rightarrow\mathcal{D}(\Nn^l) \Bigg|\ \mathcal{J}[u](t)=\sum_{n\in\Ii}F_{(n,t)}[u],\\
+	The space of functions \(\mathcal{F}(\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m),\mathcal{D}(\mathbb{N}^l))\) is
+	\begin{gather*}
+		\mathcal{F}(\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m),\mathcal{D}(\mathbb{N}^l)):=\\	
+		\Bigg\{\mathcal{J}:\mathcal{P}(\mathcal{I}\times\mathbb{N}^l)\rightarrow\mathcal{D}(\mathbb{N}^l) \Bigg|\ \mathcal{J}[u](t)=\sum_{n\in\mathcal{I}}F_{(n,t)}[u],\\
 		\text{where } F_{(n,t)}[u]=G(u((n,t)+e^1),u((n,t)+e^2),\ldots,u((n,t)+e^k)),\\
 		\text{and }
-		\{e^1,\ldots,e^k\}\subset\Zz^{d+l},\text{G is a polynomial in k variables}\Bigg\}.
-	\end{gather*}
-	Here, $(n,t):=(n_1,\ldots,n_d,t_1,\ldots,t_l)\in\Ii\times\Nn^l$.
-\end{definition}
-
-Here, all quantities, whose conservation in time we check, are functions $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$. 
-Let $\tilde{n}:=(n_1,\dots,n_d,s_1,\dots,s_l)\in\Ii\times\Nn^l$. Using the Kronecker delta $\delta$, we rewrite
+		\{e^1,\ldots,e^k\}\subset{\mathbb{Z}}^{d+l},\text{G is a polynomial in k variables}\Bigg\}.
+	\end{gather*}
+	Here, \((n,t):=(n_1,\ldots,n_d,t_1,\ldots,t_l)\in\mathcal{I}\times\mathbb{N}^l\).
+\end{definition}
+
+Here, all quantities, whose conservation in time we check, are functions \(\mathcal{J}\in\mathcal{F}(\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m),\mathcal{D}(\mathbb{N}^l))\). 
+Let \(\tilde{n}:=(n_1,\dots,n_d,s_1,\dots,s_l)\in\mathcal{I}\times\mathbb{N}^l\)
+{\color{blue} New variable(s): \(s_1\), \(s_l\),  }
+. Using the Kronecker delta \(\delta\)
+{\color{blue} New variable(s): \(\delta\),  }
+, we rewrite
 \begin{gather*}
 	\mathcal{J}[u](t_1,\ldots,t_{l})
-	=\sum_{\tilde{n}\in\Ii\times\Nn^l}F_{\tilde{n}}[u]\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l).
-\end{gather*}
-Let $u,v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$, then
+	=\sum_{\tilde{n}\in\mathcal{I}\times\mathbb{N}^l}F_{\tilde{n}}[u]\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l).
+\end{gather*}
+Let \(u,v\in\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m)\), then
 \footnotesize
@@
 \begin{definition}
-$\mathcal{V}^l(\mathcal{J})$ is the \emph{discrete partial variational derivative} of $\mathcal{J}$.
-	Note that $\mathcal{V}^0=\mathcal{V}$.
+\(\mathcal{V}^l(\mathcal{J})\) is the \emph{discrete partial variational derivative} of \(\mathcal{J}\).
+	Note that \(\mathcal{V}^0=\mathcal{V}\).
 \end{definition}
@@
 	\end{gather*}
-	is $\delta_{t+1}(s)-\delta_t(s)$, because
+	is \(\delta_{t+1}(s)-\delta_t(s)\)
+{\color{blue} New variable(s): \(s\), \(\delta _t\),  }
+, because
 	\small
@@
 \begin{theorem}\label{theo2}
-	Let $\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$.
-If $\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[v]=0$ for all $v\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ and if there exists $u_0\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$ such that $\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]$, then
-	$\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$
-	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$.
+	Let \(\mathcal{J},\tilde{\mathcal{J}}\in \mathcal{F}(\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m),\mathcal{D}(\mathbb{N}^l))\).
+If \(\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[v]=0\) for all \(v\in\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m)\) and if there exists \(u_0\in\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m)\) such that \(\mathcal{J}[u_0]=\tilde{\mathcal{J}}[u_0]\), then
+	\(\mathcal{J}[u]=\tilde{\mathcal{J}}[u]\)
+	for all \(u\in\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m)\).
 	Conversely,
-	$\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[u]=0$
-	for all $u\in\mathcal{P}(\Ii\times\Nn^l,\Rr^m)$, if $\mathcal{J}[u]=\tilde{\mathcal{J}}[u]$.
+	\(\mathcal{V}^l(\mathcal{J}-\tilde{\mathcal{J}})[u]=0\)
+	for all \(u\in\mathcal{P}(\mathcal{I}\times\mathbb{N}^l,\mathbb{R}^m)\), if \(\mathcal{J}[u]=\tilde{\mathcal{J}}[u]\).
 \end{theorem}
@@
 In this section, we formally define numerical schemes and introduce the necessary tools from difference algebra for our algorithm.
-Throughout this paper, we assume that there is only one time variable $t$ ($l=1$).
+Throughout this paper, we assume that there is only one time variable \(t\) (\(l=1\)).
 \begin{definition}
@@
 	\begin{gather*}
-		\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr)):=\\
-		\Bigg\{H:\mathcal{P}(\Ii\times\Nn,\Rr^m)\rightarrow\mathcal{P}(\Ii\times\Nn,\Rr)\ \Bigg|\\
+		\mathcal{S}(\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m),\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R})):=\\
+		\Bigg\{H:\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m)\rightarrow\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R})\ \Bigg|\\
 		H[u](n,t)=H(u(n+\tilde{e}^1,t+\tilde{l}^1),\ldots,u(n+\tilde{e}^r,t+\tilde{l}^r)),\\
 		\text{where }
-		\{\tilde{e}^1,\ldots,\tilde{e}^r\}\subset\Zz^d,\ \{\tilde{l}^1,\ldots,\tilde{l}^r\}\subset\Zz,\\
+		\{\tilde{e}^1,\ldots,\tilde{e}^r\}\subset{\mathbb{Z}}^d,\ \{\tilde{l}^1,\ldots,\tilde{l}^r\}\subset{\mathbb{Z}},\\
 		\text{H is a polynomial in r variables}
@@
 	\begin{gather*}
-		\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))
-	\end{gather*}
-	that represent the equations $\{H^1[u]=0,\ldots,H^i[u]=0\}$ for $u\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$, holding pointwise for all points in $\Ii\times\Nn$.
-\end{definition}
-If a scheme contains a single function $H$, we call $H$ the scheme. Sometimes, we also call the expression $H[u]$ the scheme.
+		\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m),\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}))
+	\end{gather*}
+	that represent the equations \(\{H^1[u]=0,\ldots,H^i[u]=0\}\)
+{\color{blue} New variable(s): \(H\), \(i\),  }
+ for \(u\in\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m)\), holding pointwise for all points in \(\mathcal{I}\times\mathbb{N}\).
+\end{definition}
+If a scheme contains a single function \(H\), we call \(H\) the scheme. Sometimes, we also call the expression \(H[u]\) the scheme.
 \begin{remark}
-	In our examples, we only consider finite-difference schemes with fixed step sizes $\Delta x=\Delta t=1$.
-However, our algorithms and our code handle parameters.
-Hence, our results can be generalized to schemes with other step sizes.
+	In our examples, we only consider finite-difference schemes with fixed step sizes \(\Delta x=\Delta t=1\)
+{\color{blue} New variable(s): \(\Delta\),  }
+. However, our algorithms and our code handle parameters. Hence, our results can be generalized to schemes with other step sizes.
 \end{remark}
@@
 	\label{ee}
-Consider the heat equation	$u_t-u_{xx}=0$ \cite{smoller94}.  A possible scheme is
+Consider the heat equation	\(u_t-u_{xx}=0\)
+{\color{blue} New variable(s): \(u_t\),  }
+ \cite{smoller94}.  A possible scheme is
 \[	
@@
 \]
-A solution of the corresponding numerical approximation is a function $u\in \mathcal{P}(\Ii\times\Nn,\Rr),$ that satisfies
+A solution of the corresponding numerical approximation is a function \(u\in \mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}),\) that satisfies
 \[
@@
 \]
-for all $n\in \Ii$ and $t\in \Nn$. 
-\end{example}
-
-\begin{definition}
-	A scheme $\{H^1,\ldots,H^m\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ given by
+for all \(n\in \mathcal{I}\) and \(t\in \mathbb{N}\). 
+\end{example}
+
+\begin{definition}
+	A scheme \(\{H^1,\ldots,H^m\}\subset\mathcal{S}(\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m),\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}))\)
+{\color{blue} New variable(s): \(S\),  }
+ given by
 	\begin{equation*}
@@
 	\end{equation*}
+{\color{blue} New variable(s): \(e\), \(j\), \(r\),  }
+
 	is in \emph{time-explicit form} if we can rewrite the previous equation (eventually translating the scheme) as
@@
 	\end{equation*}
-	for $1\leq j\leq m$
-	with $\tilde{H}^j\in\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$. We call $\{\tilde{H}^1,\ldots,\tilde{H}^m\}$ the right-hand side of $\{H^1,\ldots,H^m\}$.
+{\color{blue} New variable(s): \(u_j\),  }
+
+	for \(1\leq j\leq m\)
+	with \(\tilde{H}^j\in\mathcal{S}(\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m),\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}))\).
+We call \(\{\tilde{H}^1,\ldots,\tilde{H}^m\}\) the right-hand side of \(\{H^1,\ldots,H^m\}\).
 \end{definition}
@@
 \begin{definition}
-	Let $u=(u_1,\dots,u_m)\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$. The \emph{shift} in the $i$-th coordinate for $1\leq i\leq d+1$ by $k\in\Zz$ is
+	Let \(u=(u_1,\dots,u_m)\in\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m)\)
+{\color{blue} New variable(s): \(u_1\), \(u_m\),  }
+. The \emph{shift} in the \(i\)-th coordinate for \(1\leq i\leq d+1\) by \(k\in{\mathbb{Z}}\) is
 	\begin{gather*}
@@
 	\end{gather*}
-	understanding that $\sigma_{d+1}^k$ shifts the time variable.
-\end{definition}
-\begin{definition}
-	The \emph{set of all possible shifts} $\Theta$ is
-	\begin{gather*}
-		\Theta:=\{\sigma_1^{k_1}\circ\cdots\circ\sigma_{d+1}^{k_{d+1}}\ |\ k_1,\ldots,k_{d+1}\in\Zz\}.
+	understanding that \(\sigma_{d+1}^k\)
+{\color{blue} New variable(s): \(\sigma\),  }
+ shifts the time variable.
+\end{definition}
+\begin{definition}
+	The \emph{set of all possible shifts} \(\Theta\)
+{\color{blue} New variable(s): \(\Theta\),  }
+ is
+	\begin{gather*}
+		\Theta:=\{\sigma_1^{k_1}\circ\cdots\circ\sigma_{d+1}^{k_{d+1}}\ |\ k_1,\ldots,k_{d+1}\in{\mathbb{Z}}\}.
 	\end{gather*}
@@
 	\label{diffideal}
-	Let $\ \mathcal{K}$ be the field generated by $\ \Rr$ and the variables $\{n_1,\ldots,n_d,t\}$.
-Let $\ \tilde{\mathcal{R}}$ be the polynomial ring over the field $\ \mathcal{K}$ and the variables $\theta\circ u_j(n,t)$ for $\theta\in\Theta$ and $1\leq j\leq m$.
-	A set $\tilde{I}\subseteq\tilde{\mathcal{R}}$ is a \emph{difference ideal} if
-	$p_1,p_2\in\tilde{I}$ implies $p_1+p_2\in\tilde{I}$;
-	$p_1\in\tilde{I},p_2\in\tilde{\mathcal{R}}$ implies $p_1\cdot p_2\in\tilde{I}$ and
-	$p_1\in\tilde{I},\theta\in\Theta$ implies $\theta\circ p_1\in\tilde{I}$.
-\end{definition}
-\begin{definition}
-	$\langle H^1,\ldots,H^i\rangle$ is the smallest difference ideal containing the scheme $\{H^1,\ldots,H^i\}$.
-\end{definition}
-
-A solution of a numerical scheme is a function, $u$, that makes all translations of the scheme vanish.
-Hence, every element of the difference ideal generated by the scheme vanishes under $u$.
-As a result, every element of the difference ideal generated by the scheme vanishes under $u$.
-If the polynomial associated with the discrete time derivative  $\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$
-belongs to the ideal $\langle H^1,\ldots,H^i\rangle$, the corresponding functional vanishes. 
- Gerdt introduced the concept of a standard  Gr{\"o}bner basis for the ideal $\langle H^1,\ldots,H^i\rangle$, which might not be finite.
+	Let \(\ \mathcal{K}\)
+{\color{blue} New variable(s): \(K\),  }
+ be the field generated by \(\ \mathbb{R}\) and the variables \(\{n_1,\ldots,n_d,t\}\).
+Let \(\ \tilde{\mathcal{R}}\) be the polynomial ring over the field \(\ \mathcal{K}\) and the variables \(\theta\circ u_j(n,t)\)
+{\color{blue} New variable(s): \(\theta\),  }
+ for \(\theta\in\Theta\) and \(1\leq j\leq m\).
+	A set \(\tilde{I}\subseteq\tilde{\mathcal{R}}\) is a \emph{difference ideal} if
+	\(p_1,p_2\in\tilde{I}\)
+{\color{blue} New variable(s): \(p_1\), \(p_2\),  }
+ implies \(p_1+p_2\in\tilde{I}\);
+	\(p_1\in\tilde{I},p_2\in\tilde{\mathcal{R}}\) implies \(p_1\cdot p_2\in\tilde{I}\) and
+	\(p_1\in\tilde{I},\theta\in\Theta\) implies \(\theta\circ p_1\in\tilde{I}\).
+\end{definition}
+\begin{definition}
+	\(\langle H^1,\ldots,H^i\rangle\) is the smallest difference ideal containing the scheme \(\{H^1,\ldots,H^i\}\).
+\end{definition}
+
+A solution of a numerical scheme is a function, \(u\), that makes all translations of the scheme vanish.
+Hence, every element of the difference ideal generated by the scheme vanishes under \(u\).
+As a result, every element of the difference ideal generated by the scheme vanishes under \(u\).
+If the polynomial associated with the discrete time derivative  \(\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)\)
+{\color{blue} New variable(s): \(T\),  }
+
+belongs to the ideal \(\langle H^1,\ldots,H^i\rangle\), the corresponding functional vanishes. 
+ Gerdt introduced the concept of a standard  Gr{\"o}bner basis for the ideal \(\langle H^1,\ldots,H^i\rangle\), which might not be finite.
 To address this issue, we use polynomial algebra instead of difference algebra and a smaller polynomial ideal contained within the difference ideal that has a finite standard basis.
 
- However, ideal membership is not  a necessary condition for conservation. Some functions, $\mathcal{J}$,
+ However, ideal membership is not  a necessary condition for conservation. Some functions, \(\mathcal{J}\),
  may have sums that add to zero even if they do not belong to the difference ideal.
 Thus, we combine the discrete partial variational derivative with polynomial ideals.
-This idea is implemented in Algorithm \ref{alg: discreteconservedqimplicit} in Section \ref{secgen}.
+This idea is implemented in Algorithm \ref{alg: discreteconservedqimplicit}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{alg: discreteconservedqimplicit}\end{verbatim}.}
+ in Section \ref{secgen}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{secgen}\end{verbatim}.}
+.
 
@@
 \begin{definition}
-    Let $\mathcal{K}$ be a field. The ring of parametric polynomials
-    $\mathcal{K}[a_1,..,a_s][x_1,...,x_n]$ is a ring of polynomials in the variables $x_1,...,x_n$ with coefficients that are polynomials
-    in the ring $\mathcal{K}[a_1,..,a_s]$. 
-    For simplicity we denote this ring by $\mathcal{K}[A][X]$.
-\end{definition}
-
-\begin{definition}
-    Let $L$ be an algebraically closed field extension of $\mathcal{K}$ and let $\sigma:\mathcal{K}[A][X]\rightarrow L[X]$ be a ring homomorphism.
-$\sigma$ is a specialization given by a polynomial evaluation on the parameters and an identity on the variables, i.e.,
-    for $P\in \mathcal{K}[A][X]$
-    $\sigma$ is given by $P\rightarrow P(a_{1_0},..,a_{s_0})$, for some $(a_{1_0},..,a_{s_0})\in L^s$.
+    Let \(\mathcal{K}\) be a field. The ring of parametric polynomials
+    \(\mathcal{K}[a_1,..,a_s][x_1,...,x_n]\)
+{\color{blue} New variable(s): \(a_1\), \(a_s\), \(x_1\), \(x_n\),  }
+ is a ring of polynomials in the variables \(x_1,...,x_n\) with coefficients that are polynomials
+    in the ring \(\mathcal{K}[a_1,..,a_s]\). 
+    For simplicity we denote this ring by \(\mathcal{K}[A][X]\)
+{\color{blue} New variable(s): \(A\), \(X\),  }
+.
+\end{definition}
+
+\begin{definition}
+    Let \(L\) be an algebraically closed field extension of \(\mathcal{K}\) and let \(\sigma:\mathcal{K}[A][X]\rightarrow L[X]\) be a ring homomorphism.
+\(\sigma\) is a specialization given by a polynomial evaluation on the parameters and an identity on the variables, i.e.,
+    for \(P\in \mathcal{K}[A][X]\)
+    \(\sigma\) is given by \(P\rightarrow P(a_{1_0},..,a_{s_0})\)
+{\color{blue} New variable(s): \(a\), \(1_0\), \(s_0\),  }
+, for some \((a_{1_0},..,a_{s_0})\in L^s\).
 
@@
 \begin{definition}
-    Let $\Ii$ be a finitely generated, nonzero ideal in the ring $\mathcal{K}[A][X]$ with a fixed monomial order $\prec$ on the variables. 
-    Let $\{A_1,\dots,A_k\}$ be a family of locally closed sets that forms a partition of $L^m$ 
-    , such that the disjoint union $\bigcup\limits_{i=1}^k A_i=L^m$, and $\{G_1,\dots,G_k\}$ 
-    be a family of subsets in $\mathcal{K}[A][X]$.
+    Let \(\mathcal{I}\) be a finitely generated, nonzero ideal in the ring \(\mathcal{K}[A][X]\) with a fixed monomial order \(\prec\) on the variables. 
+    Let \(\{A_1,\dots,A_k\}\)
+{\color{blue} New variable(s): \(A_1\), \(A_k\),  }
+ be a family of locally closed sets that forms a partition of \(L^m\) 
+    , such that the disjoint union \(\bigcup\limits_{i=1}^k A_i=L^m\)
+{\color{blue} New variable(s): \(A_i\),  }
+, and \(\{G_1,\dots,G_k\}\)
+{\color{blue} New variable(s): \(G_1\), \(G_k\),  }
+
+    be a family of subsets in \(\mathcal{K}[A][X]\).
     The system, 
@@
 
-    is called a comprehensive Gr{\"o}bner system (CGS) of $I$
-     if for all specializations $\sigma_i \in A_i, \sigma_i (G_i)$ 
-     is a Gr{\"o}bner basis of the specialized ideal $\sigma_i (I)\subset \mathcal{K}[X]$.
-     Where each pair $\{G_i,A_i\}$ is called a branch. 
-\end{definition}
-Here, the notation $\sigma_i (\Ff)$, with $\Ff=\{f_1,f_2,\dots ,f_k\}$, is the specialization $\sigma_i$
-of each $f_j$ such that  $\sigma_i (\Ff)=\{\sigma_i(f_j)~|~ f_j\in \Ff\}$
-and  $"\sigma_i \in A_i"$ stands for a specialization given by $P\rightarrow P(a_{1_0},\dots,a_{m_0})$ 
-for $(a_{1_0},\dots,a_{m_0})\in A_i$.\\
-
-In this paper, we choose $A_i$ to be of the form 
-$$V(p_1,\dots,p_s)\backslash V(q_1,\dots,q_t) \subset L^m.$$
-Where $p_1,\dots,p_s,q_1,\dots,q_t\in \mathcal{K}[A]$ and $V(p_1,\dots,p_s)$ is a Variety.
+    is called a comprehensive Gr{\"o}bner system (CGS) of \(I\)
+     if for all specializations \(\sigma_i \in A_i, \sigma_i (G_i)\)
+{\color{blue} New variable(s): \(G_i\), \(\sigma _i\),  }
+
+     is a Gr{\"o}bner basis of the specialized ideal \(\sigma_i (I)\subset \mathcal{K}[X]\).
+     Where each pair \(\{G_i,A_i\}\) is called a branch. 
+\end{definition}
+Here, the notation \(\sigma_i (\mathcal{F})\), with \(\mathcal{F}=\{f_1,f_2,\dots ,f_k\}\)
+{\color{blue} New variable(s): \(f_1\), \(f_2\), \(f_k\),  }
+, is the specialization \(\sigma_i\)
+of each \(f_j\)
+{\color{blue} New variable(s): \(f_j\),  }
+ such that  {\color{red} \(\sigma_i (\mathcal{F})=\{\sigma_i(f_j)~|~ f_j\in \mathcal{F}\}\)
+
+  Equation errors and warnings:
+
+ Unbalanced $|\hdots|$}
+
+and  \("\sigma_i \in A_i"\) stands for a specialization given by \(P\rightarrow P(a_{1_0},\dots,a_{m_0})\)
+{\color{blue} New variable(s): \(m_0\),  }
+
+for \((a_{1_0},\dots,a_{m_0})\in A_i\).\\
+
+In this paper, we choose \(A_i\) to be of the form 
+\[V(p_1,\dots,p_s)\backslash V(q_1,\dots,q_t) \subset L^m.\]
+{\color{blue} New variable(s): \(p_s\), \(q_1\), \(q_t\),  }
+
+Where \(p_1,\dots,p_s,q_1,\dots,q_t\in \mathcal{K}[A]\) and \(V(p_1,\dots,p_s)\) is a Variety.
 
@@
 \begin{theorem}
-    Let $\mathcal{I}\subseteq\mathcal{K}[A][X].$ The corresponding comprehensive Gr{\"o}bner system exists.\cite{weispfenning}
+    Let \(\mathcal{I}\subseteq\mathcal{K}[A][X].\) The corresponding comprehensive Gr{\"o}bner system exists.\cite{weispfenning}
 \end{theorem}
@@
 \begin{example}
-    Let $\Ii\subset \Rr[a,b][x,y]$ be the ideal generated by 
-    $$\Ff=\left\{a x^2+\left(a^2-2 a b\right) x y, b^2 y^2-x\right\}.$$
-    The comprehensive Gr{\"o}bner system of $\Ii$ with a Lexicographic order $y\prec x$ is:
+    Let \(\mathcal{I}\subset \mathbb{R}[a,b][x,y]\)
+{\color{blue} New variable(s): \(b\), \(y\),  }
+ be the ideal generated by 
+    \[\mathcal{F}=\left\{a x^2+\left(a^2-2 a b\right) x y, b^2 y^2-x\right\}.\]
+    The comprehensive Gr{\"o}bner system of \(\mathcal{I}\) with a Lexicographic order \(y\prec x\) is:
     {\color{red} we use our code here}
@@
 It branches the cases of the parameters that affect leading terms. For instance, in the previous
-example if $a=0$ then the leading term $a x^2$ vanishes which leads to a different Gr{\"o}bner basis. Therefore,
+example if \(a=0\) then the leading term \(a x^2\) vanishes which leads to a different Gr{\"o}bner basis. Therefore,
 the (CGS) considers cases of the leading coefficients parameters Vanishing.
@@
 \begin{example}
-    Let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
-    $$\Ff=\left\{a^2 x-y,-a+b^2 x
-    y-y^2\right\}.$$
-    The comprehensive Gr{\"o}bner system of $\mathcal{I}$ with a Lexicographic order $y\prec x$ is:
+    Let \(\mathcal{I}\subset \mathbb{R}[a,b][x,y]\) be the ideal generated by 
+    \[\mathcal{F}=\left\{a^2 x-y,-a+b^2 x
+    y-y^2\right\}.\]
+    The comprehensive Gr{\"o}bner system of \(\mathcal{I}\) with a Lexicographic order \(y\prec x\) is:
 
@@
 \begin{remark}
-    If a polynomial ideal contains a nonzero constant polynomial, then it is called a unit ideal $\langle 1\rangle$.
-   Unit ideals can generate all polynomials in the ring, so $\langle 1\rangle=\mathcal{K}[A][X]$.
-    As a convention the Gr{\"o}bner basis of such ideals is $\{1\}$.
+    If a polynomial ideal contains a nonzero constant polynomial, then it is called a unit ideal \(\langle 1\rangle\).
+   Unit ideals can generate all polynomials in the ring, so \(\langle 1\rangle=\mathcal{K}[A][X]\).
+    As a convention the Gr{\"o}bner basis of such ideals is \(\{1\}\).
 \end{remark}
 
-Issues arise when computing the Gr{\"o}bner basis of $\Ff$ without considering the parameters.
+Issues arise when computing the Gr{\"o}bner basis of \(\mathcal{F}\) without considering the parameters.
 In practice, to compute the Gr{\"o}bner basis without considering the parameters, we assume non-zero parameters by default.
-This assumption guarantees a Gr{\"o}bner basis only for one set of specializations $\Rr^2\backslash V(a,b)$. i.e., 
-$\sigma(\Gg)$ is a Gr{\"o}bner basis of the specialized ideal $\sigma(\Ii)$ only for $\sigma \in \Rr^2\backslash V(a,b)$.
+This assumption guarantees a Gr{\"o}bner basis only for one set of specializations \(\mathbb{R}^2\backslash V(a,b)\). i.e., 
+\(\sigma(\mathcal{G})\)
+{\color{blue} New variable(s): \(G\),  }
+ is a Gr{\"o}bner basis of the specialized ideal \(\sigma(\mathcal{I})\) only for \(\sigma \in \mathbb{R}^2\backslash V(a,b)\).
 Next, we show the impact of this assumption.
 \begin{enumerate}
-    \item In general $\sigma\left(\Gg\right)$ is not necessarily a Gr{\"o}bner basis.
+    \item In general \(\sigma\left(\mathcal{G}\right)\) is not necessarily a Gr{\"o}bner basis.
     \begin{example}
         \label{ex431}
-        Let $\Ii=\langle 
-           {a x^2 - y, -2 + b x y} \rangle$. The corresponding Gr{\"o}bner basis 
-           is $\Gg=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$. Consider 
-           the specialization $\sigma$ given by $P\rightarrow P(0,1)$.
-           The set $\sigma(\Gg)=\left\{y^3,-y^2,x y-2\right\}$ is not a Gr{\"o}bner basis.
+        Let \(\mathcal{I}=\langle 
+           {a x^2 - y, -2 + b x y} \rangle\). The corresponding Gr{\"o}bner basis 
+           is \(\mathcal{G}=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}\). Consider 
+           the specialization \(\sigma\) given by \(P\rightarrow P(0,1)\).
+           The set \(\sigma(\mathcal{G})=\left\{y^3,-y^2,x y-2\right\}\) is not a Gr{\"o}bner basis.
 		   This can be shown by the Buchberger Criterion (see chapter 2 in\cite{Cox}).
@@
         \label{ex433}
-        Consider the ideal $\Ii$ in the above example. The corresponding Gr{\"o}bner basis computed under the assumption of non-zero parameters
-        is $$\sigma_0(\Gg)=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$$ where $\sigma_0 \in \Rr^2\backslash V(a,b)$. 
-        Now, it is possible to check if a polynomial $f\in \Rr[x,y]$ belongs to $\sigma(\Ii)$ only for $\sigma_0 $     
-        because we know that $\sigma_0(\Gg)$ is a Gr{\"o}bner basis. For other specializations such as 
-        the specialization $\sigma$ given by $P\rightarrow P(0,1)$, $\sigma(\Gg)$ is not a Gr{\"o}bner basis as shown in the above example. 
-        Therefore, we can not proceed with the division algorithm as we can not guarantee unique remainder if we used $\sigma(\Gg)$.  
+        Consider the ideal \(\mathcal{I}\) in the above example. The corresponding Gr{\"o}bner basis computed under the assumption of non-zero parameters
+        is \[\sigma_0(\mathcal{G})=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}\]
+{\color{blue} New variable(s): \(\sigma _0\),  }
+ where \(\sigma_0 \in \mathbb{R}^2\backslash V(a,b)\). 
+        Now, it is possible to check if a polynomial \(f\in \mathbb{R}[x,y]\)
+{\color{blue} New variable(s): \(f\),  }
+ belongs to \(\sigma(\mathcal{I})\) only for \(\sigma_0 \)     
+        because we know that \(\sigma_0(\mathcal{G})\) is a Gr{\"o}bner basis. For other specializations such as 
+        the specialization \(\sigma\) given by \(P\rightarrow P(0,1)\), \(\sigma(\mathcal{G})\) is not a Gr{\"o}bner basis as shown in the above example. 
+        Therefore, we can not proceed with the division algorithm as we can not guarantee unique remainder if we used \(\sigma(\mathcal{G})\).  
     \end{example}
@@
 
-Using CGS can resolve these problems. Problem 1 is fixed by the definition of (CGS), as we have $\sigma_i(G_i)$ a 
- Gröbner basis for any specialization $\sigma_i$. Moreover, 
+Using CGS can resolve these problems. Problem 1 is fixed by the definition of (CGS), as we have \(\sigma_i(G_i)\) a 
+ GrA¶bner basis for any specialization \(\sigma_i\). Moreover, 
  to discuses problem 2 we need to look at the parametric ideal membership first.
@@
 {\color{red} Is this a definition?}
- Consider, $f\in \Kk[A][X]$ we say that $f\in\Ii$ under $\sigma$, if there exists a specialization 
- $\sigma\in L^m$ such that $\sigma(f)\in\sigma(\Ii)$.\\
-
-To determine if $f\in\Ii$ we consider the (CGS) of $\Ii$ given \ref{CGS}. 
-We check if $f\in\sigma_i(\Ii)$ for each $i\in\{1,..,k\}$ and $\sigma_i\in A_i$.
-For each $\sigma_i(\Ii)$ we have the corresponding Gröbner basis $\sigma_i(G_i)$
-so we divide $f$ by the $\sigma_i(G_i)$ if the remainder is zero for some $i\in\{1,..,k\}$,
-then $f\in\sigma_i(\Ii)$. Therefore, $f\in\Ii$. However, if the remainder is non-zero for all $i$ we conclude that 
-$f\notin\Ii$.
-
-\begin{example}
-	Let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
-    $$\Ff=\{a^2 x - y, -b^2 x y^2 - x\}.$$
-	We will determine if $f(x,y)=-x+y^2\in\mathcal{I}$.
+ Consider, \(f\in \mathcal{K}[A][X]\) we say that \(f\in\mathcal{I}\) under \(\sigma\), if there exists a specialization 
+ \(\sigma\in L^m\) such that \(\sigma(f)\in\sigma(\mathcal{I})\).\\
+
+To determine if \(f\in\mathcal{I}\) we consider the (CGS) of \(\mathcal{I}\) given \ref{CGS}. 
+We check if \(f\in\sigma_i(\mathcal{I})\) for each \(i\in\{1,..,k\}\) and \(\sigma_i\in A_i\).
+For each \(\sigma_i(\mathcal{I})\) we have the corresponding GrA¶bner basis \(\sigma_i(G_i)\)
+so we divide \(f\) by the \(\sigma_i(G_i)\) if the remainder is zero for some \(i\in\{1,..,k\}\),
+then \(f\in\sigma_i(\mathcal{I})\). Therefore, \(f\in\mathcal{I}\). However, if the remainder is non-zero for all \(i\) we conclude that 
+\(f\notin\mathcal{I}\).
+
+\begin{example}
+	Let \(\mathcal{I}\subset \mathbb{R}[a,b][x,y]\) be the ideal generated by 
+    \[\mathcal{F}=\{a^2 x - y, -b^2 x y^2 - x\}.\]
+	We will determine if \(f(x,y)=-x+y^2\in\mathcal{I}\).
 	We compute the corresponding (CGS)
 	{\color{red} add example 3}\\
-	First let $A_1=\{(a,b)~|~ a\ne 0, b\ne 0\}$ and $A_2=\{(a,b)~|~ ab=0\}$. 
-	We divide $f$ by $\{\frac{y}{b^2}+y^3,x-\frac{y}{a^2}\}$ and we obtain the remainder $-\frac{y}{a^2}+y^2$ and 
-	since the exists no specialization in $A_1$ such that this remainder is zero we conclude that $f\notin \sigma_1(\Ii)$ for
-	$\sigma_1\in A_1$. Now we repeat the reduction of $f$ but with the second Gröbner basis $\{x,y\}$. We obtain a zero remainder, which imply that 
-	$f\in \sigma_2(\Ii)$ for $\sigma_2\in A_2$. Therefore, $f\in\Ii$ under a specialization $\sigma_2\in A_2$.
+	First let {\color{red} \(A_1=\{(a,b)~|~ a\ne 0, b\ne 0\}\)
+
+  Equation errors and warnings:
+
+ Unbalanced $|\hdots|$}
+
+ and {\color{red} \(A_2=\{(a,b)~|~ ab=0\}\)
+
+  Equation errors and warnings:
+
+ Unbalanced $|\hdots|$}
+
+{\color{blue} New variable(s): \(A_2\),  }
+. 
+	We divide \(f\) by \(\{\frac{y}{b^2}+y^3,x-\frac{y}{a^2}\}\) and we obtain the remainder \(-\frac{y}{a^2}+y^2\) and 
+	since the exists no specialization in \(A_1\) such that this remainder is zero we conclude that \(f\notin \sigma_1(\mathcal{I})\)
+{\color{blue} New variable(s): \(\sigma _1\),  }
+ for
+	\(\sigma_1\in A_1\). Now we repeat the reduction of \(f\) but with the second GrA¶bner basis \(\{x,y\}\). We obtain a zero remainder, which imply that 
+	\(f\in \sigma_2(\mathcal{I})\)
+{\color{blue} New variable(s): \(\sigma _2\),  }
+ for \(\sigma_2\in A_2\). Therefore, \(f\in\mathcal{I}\) under a specialization \(\sigma_2\in A_2\).
 \end{example}
@@
 \label{time-exp}
-In this section, we introduce Algorithm \ref{alg: discreteconservedqexplicit}  for checking the conservation of a quantity for time-explicit schemes.
-We discuss the algorithm for general schemes in the next section.
+In this section, we introduce Algorithm \ref{alg: discreteconservedqexplicit}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{alg: discreteconservedqexplicit}\end{verbatim}.}
+  for checking the conservation of a quantity for time-explicit schemes. We discuss the algorithm for general schemes in the next section.
 \begin{algorithm}
 	\DontPrintSemicolon
-	\KwIn{$\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{D}(\Nn))$ and $\{H^1,\ldots,H^m\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ right-hand side of a time-explicit scheme}
-	\KwOut{ {\sc True}, if $\mathcal{J}$ is conserved in time, {\sc False} otherwise }
-
-	\label{estep1} Build the discrete time derivative $\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$\;
-
-	\label{estep2} Replace all instances of $u_j(n+e,t+1)$ (for $e\in\Zz^{d}$) by an appropriate translation of $\{H^1,\ldots,H^m\}$\;
+	\KwIn{\(\mathcal{J}\in\mathcal{F}(\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m),\mathcal{D}(\mathbb{N}))\) and \(\{H^1,\ldots,H^m\}\subset\mathcal{S}(\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m),\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}))\) right-hand side of a time-explicit scheme}
+	\KwOut{ {\sc True}, if \(\mathcal{J}\) is conserved in time, {\sc False} otherwise }
+
+	\label{estep1} Build the discrete time derivative \(\mathcal{J}[u](t+1)-\mathcal{J}[u](t)\)\;
+
+	\label{estep2} Replace all instances of \(u_j(n+e,t+1)\) (for \(e\in{\mathbb{Z}}^{d}\)) by an appropriate translation of \(\{H^1,\ldots,H^m\}\)\;
 
@@
 \end{algorithm}
-The result of Step 2 exists only at time $t$ because all instances of $t+1$ have been replaced. Thus, we do not need any computations of difference ideals or Gr{\"o}bner basis.
+The result of Step 2 exists only at time \(t\) because all instances of \(t+1\) have been replaced. Thus, we do not need any computations of difference ideals or Gr{\"o}bner basis.
 \begin{example}
 	\label{timeexplicitheateqalgo}
-	We examine the conservation of $\sum_nu(n,t)$ under the following scheme.
+	We examine the conservation of \(\sum_nu(n,t)\) under the following scheme.
 	\begin{equation*}
@@
 	\begin{gather*}
-		\sum_{n\in\Ii}u(n,t+1)-u(n,t).
-	\end{gather*}
-	Replace $u(n,t+1)$ by the right-hand side (Step 2) to get
-	\begin{gather*}
-		\sum_{n\in\Ii}u(n,t)+(u(n+1,t)-2u(n,t)+u(n-1,t))-u(n,t).
+		\sum_{n\in\mathcal{I}}u(n,t+1)-u(n,t).
+	\end{gather*}
+	Replace \(u(n,t+1)\) by the right-hand side (Step 2) to get
+	\begin{gather*}
+		\sum_{n\in\mathcal{I}}u(n,t)+(u(n+1,t)-2u(n,t)+u(n-1,t))-u(n,t).
 	\end{gather*}
@@
 \label{secgen}
-In this section, we present Algorithm \ref{alg: discreteconservedqimplicit} that deals with general, not necessarily time-explicit, schemes.
-We explain its steps in detail below and demonstrate the algorithm in Example \ref{heateqalgoimplicit}.
+In this section, we present Algorithm \ref{alg: discreteconservedqimplicit}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{alg: discreteconservedqimplicit}\end{verbatim}.}
+ that deals with general, not necessarily time-explicit, schemes. We explain its steps in detail below and demonstrate the algorithm in Example \ref{heateqalgoimplicit}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{heateqalgoimplicit}\end{verbatim}.}
+.
 \begin{algorithm}
 	\DontPrintSemicolon
-	\KwIn{$\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{D}(\Nn))$ and $\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ }
-	\KwOut{ {\sc True}, if $\mathcal{J}$ is conserved in time, {\sc False} otherwise }
-
-	\label{step1} Build the discrete time derivative $\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)$\;
-
-	\label{step2}Translate $\{H^1,\ldots,H^i\}$ by subtracting the range of the stencil of the scheme from the range of the stencil of $\mathcal{T}$\;
-
-	\label{step3} Reduce $\mathcal{T}$ by using {\color{red} the Gr{\"o}bner basis} generated by the translated scheme in two different ways
+	\KwIn{\(\mathcal{J}\in\mathcal{F}(\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m),\mathcal{D}(\mathbb{N}))\) and \(\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m),\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}))\) }
+	\KwOut{ {\sc True}, if \(\mathcal{J}\) is conserved in time, {\sc False} otherwise }
+
+	\label{step1} Build the discrete time derivative \(\mathcal{T}:=\mathcal{J}[u](t+1)-\mathcal{J}[u](t)\)\;
+
+	\label{step2}Translate \(\{H^1,\ldots,H^i\}\) by subtracting the range of the stencil of the scheme from the range of the stencil of \(\mathcal{T}\)\;
+
+	\label{step3} Reduce \(\mathcal{T}\) by using {\color{red} the Gr{\"o}bner basis} generated by the translated scheme in two different ways
 	\label{step3a} (a) Using the lexicographic order
@@
 \noindent{\bf Step 1: Build the discrete time derivative.}
-We build the discrete time derivative by subtracting $\mathcal{J}[u](t)$ from $\mathcal{J}[u](t+1)$. This task is performed in our code using {\sc TimeDifference}.
-\begin{example}
-	The discrete time derivative for $\sum_n u(n,t)$ is
+We build the discrete time derivative by subtracting \(\mathcal{J}[u](t)\) from \(\mathcal{J}[u](t+1)\). This task is performed in our code using {\sc TimeDifference}.
+\begin{example}
+	The discrete time derivative for \(\sum_n u(n,t)\) is
 
@@
 Now, we transition from difference algebra with infinite independent variables to polynomial algebra with finite variables to ensure the finiteness of the Gr{\"o}bner basis. 
-Hence, we need to compute a finite number of translations of the scheme and consider every instance of $u_j(n+e,t+l)$ that appears in either the translated scheme or the discrete time derivative as an independent polynomial variable.
-\begin{example}
-	The algorithm treats the discrete time derivative $\sum_{n\in\Ii}u(n+1,t+1)-u(n+1,t)$
-	as the polynomial $z_1-z_2$.
+Hence, we need to compute a finite number of translations of the scheme and consider every instance of \(u_j(n+e,t+l)\) that appears in either the translated scheme or the discrete time derivative as an independent polynomial variable.
+\begin{example}
+	The algorithm treats the discrete time derivative \(\sum_{n\in\mathcal{I}}u(n+1,t+1)-u(n+1,t)\)
+	as the polynomial \(z_1-z_2\)
+{\color{blue} New variable(s): \(z_1\), \(z_2\),  }
+.
 \end{example}
 There are several possiblities which translations of the scheme to take as basis for the polynomial ideal.
-However, an obvious idea is to compute the minimal number of translations of the scheme, such that all instances of $u_j(n+e,t+l)$, that appear in the range of the stencil of the discrete time derivative, also appear in the translated scheme.
+However, an obvious idea is to compute the minimal number of translations of the scheme, such that all instances of \(u_j(n+e,t+l)\), that appear in the range of the stencil of the discrete time derivative, also appear in the translated scheme.
 This is done by elementwise subtracting the range of the stencil of the scheme from the range of the stencil of the discrete time derivative.
@@
 	\end{gather*}
-	which we want to reduce by the scheme $u(n+1)-u(n)=0$. We compute the translations as described above as
+	which we want to reduce by the scheme \(u(n+1)-u(n)=0\). We compute the translations as described above as
 	\begin{gather*}
@@
 	\end{gather*}
-	and see, that the expression indeed reduces to $0$ under the scheme.
-\end{example}
-\begin{definition}
-	Let $u=(u_1,\ldots,u_m)\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$ and
+	and see, that the expression indeed reduces to \(0\) under the scheme.
+\end{example}
+\begin{definition}
+	Let \(u=(u_1,\ldots,u_m)\in\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m)\) and
 	\begin{gather*}
 		F:
-		\mathcal{P}(\Ii\times\Nn,\Rr^m)\rightarrow\mathcal{P}(\Ii\times\Nn,\Rr),\\
+		\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m)\rightarrow\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}),\\
 		F[u](n,t)=F(u_1((n,t)+x^1),\ldots,u_1((n,t)+x^{k_1}),\\
@@
 	\normalsize
-	with $\{x^1,\ldots,x^{k_1},\ldots,x^{k_m}\}\subset\Zz^{d+1}$.
-	The \emph{stencil} of $F$ is the $m$-tuple of sets of vectors
+	with \(\{x^1,\ldots,x^{k_1},\ldots,x^{k_m}\}\subset{\mathbb{Z}}^{d+1}\)
+{\color{blue} New variable(s): \(k_1\), \(k_m\),  }
+.
+	The \emph{stencil} of \(F\) is the \(m\)-tuple of sets of vectors
 	\begin{gather*}
@@
 	\end{gather*}
-	The \emph{range} of the stencil of $F$ is
+	The \emph{range} of the stencil of \(F\) is
 	\begin{gather*}
@@
 	\end{gather*}
-	where $k_0=0$, $x^{k_l +}_i = \max_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$,
-	and $x^{k_l -}_i = \min_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}$. Here, $[a,b]$ denotes the discrete interval in $\Zz$, i.e. $[0,2]=\{0,1,2\}$.
-\end{definition}
-If for any $[a,b]$ involved we have $b<a$, we write $[a,b]=\emptyset$ with the convention that $\emptyset-\emptyset=\emptyset$.
-Further, $\emptyset$ in the translations results in the use of the respective entry of the original scheme without translations.
+	where \(k_0=0\)
+{\color{blue} New variable(s): \(k_0\),  }
+, \(x^{k_l +}_i = \max_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}\)
+{\color{blue} New variable(s): \(k_l\), \(x_i\),  }
+,
+	and \(x^{k_l -}_i = \min_{k_{l-1}+1\leq k\leq k_l}\{x_i^k\}\). Here, \([a,b]\) denotes the discrete interval in \({\mathbb{Z}}\), i.e. \([0,2]=\{0,1,2\}\).
+\end{definition}
+If for any \([a,b]\) involved we have \(b<a\), we write \([a,b]=\emptyset\) with the convention that \(\emptyset-\emptyset=\emptyset\).
+Further, \(\emptyset\) in the translations results in the use of the respective entry of the original scheme without translations.
 \begin{example}
@@
 	\footnotesize
-	$
+	\(
 	\{u_1(n+1,t+1)-u_1(n,t),u_2(n,t+1)-u_2(n,t)\}.
-	$
+	\)
+{\color{blue} New variable(s): \(u_2\),  }
+
 	\normalsize
@@
 	\footnotesize
-	$
+	\(
 	\Big([0,2]\times[0,1],[0,0]\times [0,1]\Big).
-	$
+	\)
 	\normalsize
@@
 	\footnotesize
-	$
+	\(
 	\Big\{\Big([0,1]\times[0,1],\emptyset\Big),\Big(\emptyset,[0,0]\times [0,1]\Big)\Big\}.
-	$
+	\)
 	\normalsize
@@
 	\normalsize
-	Therefore, we translate the first equation of the scheme by $(0,0)$ and $(1,0)$ and the second one by $(0,0)$ to get the translated scheme
+	Therefore, we translate the first equation of the scheme by \((0,0)\) and \((1,0)\) and the second one by \((0,0)\) to get the translated scheme
 	\footnotesize
@@
 \emph{Polynomial ideals and Gr{\"o}bner bases.}\label{polid}
-In the previous step, we identified a finite set of polynomials comprising the translated scheme ${H^1,\ldots,H^r}$ and the discrete time derivative, each containing a finite number of instances of $u_j(n+e,t+l)$.
-We now proceed to reduce $\mathcal{T}$ by treating it as an element of a polynomial ring and performing multivariate polynomial division with respect to the translated scheme.
+In the previous step, we identified a finite set of polynomials comprising the translated scheme \({H^1,\ldots,H^r}\) and the discrete time derivative, each containing a finite number of instances of \(u_j(n+e,t+l)\).
+We now proceed to reduce \(\mathcal{T}\) by treating it as an element of a polynomial ring and performing multivariate polynomial division with respect to the translated scheme.
 To accomplish this, we adapt the definitions and theorems presented in \cite{hibi13} to suit our specific context.
 \begin{definition}
-	Let $\mathcal{K}$ be as in Definition \ref{diffideal}. Let $\mathcal{R}$ be the polynomial ring generated by $\mathcal{K}$ and all instances of $u_j(n+e,t+l)$ that occur in $\{H^1,\ldots,H^r\}$.
-	We call a set $ I\subseteq \mathcal{R}$ a \emph{(polynomial) ideal} if
-	$p_1,p_2\in I$ implies $p_1+p_2\in I$ and
-	$p_1\in I,p_2\in\mathcal{R}$ implies $p_1p_2\in I$.
-\end{definition}
-\begin{definition}
-	We denote by $\langle H^1,\ldots,H^r\rangle$ the smallest (polynomial) ideal containing $\{H^1,\ldots,H^r\}$.
-\end{definition}
-Given the ideal $\langle H^1,\ldots,H^r\rangle$ and the discrete time derivative $\mathcal{T}\in\mathcal{R}$, we want to determine if $\mathcal{T}\in\langle H^1,\ldots,H^r\rangle$ or if we can write $\mathcal{T}$ in a simpler form, using $\{H^1,\ldots,H^r\}$.
-Hence, using multivariate polynomial division, we search for $p^1,\ldots,p^r,p^{r+1}\in\mathcal{R}$ such that
+	Let \(\mathcal{K}\) be as in Definition \ref{diffideal}.
+Let \(\mathcal{R}\) be the polynomial ring generated by \(\mathcal{K}\) and all instances of \(u_j(n+e,t+l)\) that occur in \(\{H^1,\ldots,H^r\}\).
+	We call a set \( I\subseteq \mathcal{R}\) a \emph{(polynomial) ideal} if
+	\(p_1,p_2\in I\) implies \(p_1+p_2\in I\) and
+	\(p_1\in I,p_2\in\mathcal{R}\) implies \(p_1p_2\in I\).
+\end{definition}
+\begin{definition}
+	We denote by \(\langle H^1,\ldots,H^r\rangle\) the smallest (polynomial) ideal containing \(\{H^1,\ldots,H^r\}\).
+\end{definition}
+Given the ideal \(\langle H^1,\ldots,H^r\rangle\) and the discrete time derivative \(\mathcal{T}\in\mathcal{R}\), we want to determine if \(\mathcal{T}\in\langle H^1,\ldots,H^r\rangle\) or if we can write \(\mathcal{T}\) in a simpler form, using \(\{H^1,\ldots,H^r\}\).
+Hence, using multivariate polynomial division, we search for \(p^1,\ldots,p^r,p^{r+1}\in\mathcal{R}\)
+{\color{blue} New variable(s): \(p\),  }
+ such that
 \begin{gather*}
@@
 \end{gather*}
-Unfortunately, the resulting remainder $p^{r+1}$ may not be unique \cite[page 14, Example 1.2.3]{hibi13}, i.e. non-zero, although $\mathcal{T}$ belongs to $\langle H^1,\ldots,H^r \rangle$. 
-Replacing $\{H^1,\ldots,H^r\}$ by a Gr{\"o}bner basis for the ideal $\langle H^1,\ldots,H^r \rangle$ guarantees the uniqueness of the remainder of the polynomial division. 
+Unfortunately, the resulting remainder \(p^{r+1}\) may not be unique \cite[page 14, Example 1.2.3]{hibi13}, i.e. non-zero, although \(\mathcal{T}\) belongs to \(\langle H^1,\ldots,H^r \rangle\). 
+Replacing \(\{H^1,\ldots,H^r\}\) by a Gr{\"o}bner basis for the ideal \(\langle H^1,\ldots,H^r \rangle\) guarantees the uniqueness of the remainder of the polynomial division. 
 Contrary to the standard basis for the difference ideal that Gerdt's algorithm computes, a Gr{\"o}bner basis for the polynomial ideal always exists and is finite.
@@
 In Algorithm \ref{alg: discreteconservedqimplicit}, we employ two monomial orders: (a) the lexicographic order and (b) an experimental elimination order.
-To reduce $\mathcal{T}$, we first compute the Gr\"obner basis of $\langle H^1,\ldots,H^r\rangle$ with respect to the chosen monomial order and subsequently calculate the remainder resulting from the polynomial division of $\mathcal{T}$ with respect to this Gr\"obner basis.
-
-The lexicographic order (a) is induced by the time-explicit ordering of the instances of $u_j(n+e,t+l)$, giving greater importance to instances at later times over those at earlier times.
-For the elimination order (b), we initially order all instances of $u_j(n+e,t+l)$ based on a predefined ordering.
+To reduce \(\mathcal{T}\), we first compute the Gr\"obner basis of \(\langle H^1,\ldots,H^r\rangle\) with respect to the chosen monomial order and subsequently calculate the remainder resulting from the polynomial division of \(\mathcal{T}\) with respect to this Gr\"obner basis.
+
+The lexicographic order (a) is induced by the time-explicit ordering of the instances of \(u_j(n+e,t+l)\), giving greater importance to instances at later times over those at earlier times.
+For the elimination order (b), we initially order all instances of \(u_j(n+e,t+l)\) based on a predefined ordering.
 By default, our code employs the time-explicit ordering, but other options, such as implicit or user-defined orderings, can also be used.
@@
 This elimination may not always be feasible, resulting in an unchanged discrete time derivative.
-In such cases, we repeat the process for the instance at the second latest time and continue until all instances of $u_j(n+e,t+l)$ have been eliminated.
+In such cases, we repeat the process for the instance at the second latest time and continue until all instances of \(u_j(n+e,t+l)\) have been eliminated.
 
@@
 	\begin{gather*}
-		\mathcal{T}=\sum_{n\in\Ii} u(n,t+1)-u(n,t).
+		\mathcal{T}=\sum_{n\in\mathcal{I}} u(n,t+1)-u(n,t).
 	\end{gather*}
@@
 	is
-	$
+	\(
 	\Big(\big\{(0,1),(0,0),(1,0),(-1,0)\big\}\Big)
-	$
+	\)
 	\normalsize
-	with range \small$\Big([-1,1]\times[0,1]\Big)$\normalsize.
-	The discrete time derivative has stencil \small$\Big(\big\{(0,1),(0,0)\big\}\Big)$\normalsize 
-	and range \small$\Big([0,0]\times[0,1]\Big)$\normalsize.
+	with range \small\(\Big([-1,1]\times[0,1]\Big)\)\normalsize.
+	The discrete time derivative has stencil \small\(\Big(\big\{(0,1),(0,0)\big\}\Big)\)\normalsize 
+	and range \small\(\Big([0,0]\times[0,1]\Big)\)\normalsize.
 	The translations are
@@
 	\end{gather*}
-	For the reduction using the elimination order, we first eliminate $u(n,t+1)$ and then the remaining instances of $u$.
+	For the reduction using the elimination order, we first eliminate \(u(n,t+1)\) and then the remaining instances of \(u\).
 	The elimination order yields the same remainder as the lexicographic order. Hence, we do not check for the number of instances of time (Step 4).
@@
 	\begin{gather*}
-		\sum_{n\in\Ii}u(n-1,t)-2u(n,t)+u(n+1,t)
-	\end{gather*}
-	that equals $0$.
+		\sum_{n\in\mathcal{I}}u(n-1,t)-2u(n,t)+u(n+1,t)
+	\end{gather*}
+	that equals \(0\).
 	Hence, repeating the above procedure (Step 6) becomes unnecessary.
@@
 So far, we have discussed how to check if a quantity is preserved in time under a scheme. But it is also desirable to have a systematic way to find conserved quantities, given a scheme.
-Algorithm \ref{alg: findconservedquantitybasis} finds, for a time-explicit scheme, a basis for conserved quantities that are generated by monomials up to a degree. 
+Algorithm \ref{alg: findconservedquantitybasis}
+ {\color{red} Forward or undefined reference: \begin{verbatim}\ref{alg: findconservedquantitybasis}\end{verbatim}.}
+ finds, for a time-explicit scheme, a basis for conserved quantities that are generated by monomials up to a degree. 
 \begin{algorithm}
 	\DontPrintSemicolon
-	\KwIn{right-hand side set of polynomials $\{H^1,\ldots,H^m\}\subset S(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ ofa time-explicit scheme, list of {\sc generators}, {\sc degree} of polynomials }
+	\KwIn{right-hand side set of polynomials \(\{H^1,\ldots,H^m\}\subset S(\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}^m),\mathcal{P}(\mathcal{I}\times\mathbb{N},\mathbb{R}))\) ofa time-explicit scheme, list of {\sc generators}, {\sc degree} of polynomials }
 	\KwOut{basis of conserved quantities}
@@
 \begin{example}
-In the context of the heat equation from Example \ref{heatex1}, we aim to find a basis for conserved quantities generated by $u(n,t)$ and $n$ with a maximum degree of 3.
+In the context of the heat equation from Example \ref{heatex1}, we aim to find a basis for conserved quantities generated by \(u(n,t)\) and \(n\) with a maximum degree of 3.
 
@@
 \noindent{\bf Burgers equation.}
-In the first example, we focus on the inviscid Burgers equation, a well-known PDE with the form $u_t + uu_x = 0$ \cite{smoller94}. 
-This PDE preserves the total mass $\int u\ dx$.
+In the first example, we focus on the inviscid Burgers equation, a well-known PDE with the form \(u_t + uu_x = 0\)
+{\color{blue} New variable(s): \(u_x\),  }
+ \cite{smoller94}. 
+This PDE preserves the total mass \(\int u\ dx\).
 Our first example demonstrates how the code verifies mass conservation by a forward-difference discretization.
@@
 
-The inviscid Burgers equation is the PDE $u_t+uu_x=0$ \cite{smoller94} in one space dimension. Any function of $u$ is a conserved quantity. We search for discretizations preserving $\int u\ dx$.
+The inviscid Burgers equation is the PDE \(u_t+uu_x=0\) \cite{smoller94} in one space dimension. Any function of \(u\) is a conserved quantity. We search for discretizations preserving \(\int u\ dx\).
 \begin{example}
@@
 \begin{example}
-	We discretize $u_x$ with a three-point stencil with a parameter $a$.
+	We discretize \(u_x\) with a three-point stencil with a parameter \(a\).
 
@@
 	\end{small}
-The conservation property holds when $a=\frac{1}{2}$, which corresponds to the central difference case.
+The conservation property holds when \(a=\frac{1}{2}\), which corresponds to the central difference case.
 \end{example}
@@
 	\end{small}
-	Those are the only conserved quantities for this scheme, which are polynomials up to degree 4 in $v$ and $m$.
+	Those are the only conserved quantities for this scheme, which are polynomials up to degree 4 in \(v\) and \(m\).
 
@@
 	\end{gather*}
-	This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly in time for $m$ and implicitly for $v$.
-	Our code can handle this setting by specifying an order for variable elimination, using an explicit monomial order for $m$ and an implicit order for $v$:
+	This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly in time for \(m\) and implicitly for \(v\).
+	Our code can handle this setting by specifying an order for variable elimination, using an explicit monomial order for \(m\) and an implicit order for \(v\):
 
@@
 An experimental approach, that gives a hint, if a quantity is conserved, and that can refute conservation, is the following:
-We fix $d$ and $N$.
-Then, we generate experimental initial data by randomly sampling $N\cdot d\cdot m$ values, with every value corresponding to an instance of $u_j(n,t)$ for $n\in\Ii,1\leq j\leq m$.
-Afterward, we calculate $u_j(n,t+1)$ for all $n\in\Ii,1\leq j\leq m$, using the scheme.
+We fix \(d\) and \(N\).
+Then, we generate experimental initial data by randomly sampling \(N\cdot d\cdot m\) values, with every value corresponding to an instance of \(u_j(n,t)\) for \(n\in\mathcal{I},1\leq j\leq m\).
+Afterward, we calculate \(u_j(n,t+1)\) for all \(n\in\mathcal{I},1\leq j\leq m\), using the scheme.
 With this values, we explicitly calculate the value of the discrete time derivative.

### 2023-05-09 Papers/Paper CASC format/tmp/GCpaper.tex
...

\documentclass{article}
\begin{document}

{Algorithmic detection of conserved quantities for finite-difference schemes{Supported by King Abdullah University of Science and Technology (KAUST) baseline funds and KAUST OSR-CRG2021-4674.}}
{Diogo A. Gomes{1}{0000-0002-3129-3956} 
	Friedemann Krannich{1}{1111-2222-3333-4444} 
	Bashayer Majrashi{1}{0009-0008-3677-8064 } 
	Ricardo de Lima Ribeiro{1}{0000-0002-2707-2210}}
{D.A. Gomes et al.}
{King Abdullah University of Science and Technology (KAUST), AMCS/CEMSE Division, Thuwal 23955-6900, Saudi Arabia	{{diogo.gomes,friedemann.krannich,bashayer.majrashi, ricardo.ribeirokaust.edu.sa}}

	Many partial differential equations (PDEs) admit conserved quantities like mass or energy. These quantities are often essential for establishing well-posedness results. When approximating a PDE with a finite-difference scheme, it is crucial to determine whether related discretized quantities remain conserved over time by the scheme. Such conservation may ensure the stability of the numerical scheme. We present an algorithm for verifying the preservation of a polynomial quantity under a polynomial finite-difference scheme. Our schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables. Additionally, we introduce an algorithm for finding conserved quantities, given a scheme. We illustrate our algorithm with several finite-difference schemes. Our approach incorporates parametric G bases to handle parameters, ensuring accurate computation of conserved quantities.

	{Symbolic computations  Finite-difference schemes  Discrete variational derivative  Discrete partial variational derivative  Conserved quantities  Implicit schemes  Explicit schemes  Parametric Groebner basis.}

Introduction.

{  add 
Comprehensive Gr{}bner Systems 2222
}
In many partial differential equations (PDEs), conserved integral quantities play a crucial role, as they often represent properties of physical significance, such as energy or mass. For example, the advection equation preserves energy and the heat equation conserves mass. 
Since closed-form solutions for these PDEs are often unavailable, finite-difference schemes serve as a practical numerical alternative. Hence, it becomes imperative to determine whether the chosen scheme preserves these conserved quantities. This information is not only valuable for gauging the accuracy of the approximation, but also vital for evaluating its stability. Automating conservation computations in computer algebra systems streamlines the process, saves time, enhances accuracy, and may create new research opportunities.

{  [1] and [1] need domains and ranges, also explain that divergence is only in [1]}

Conserved quantities correspond to conservation laws, admitted by the PDE. A conservation law [1] is an equation of the form

	{ }{ t}[u]+{div} [u]=0

holding for all solutions [1] of the considered PDE,
that induces the conserved quantity [1] as

	{d}{dt} [u]= D_t[u]=- D_x[u]=0

assuming periodic boundary conditions in [1].

There is a similar formulation for
for conservation laws for finite difference schemes of a PDE  where the derivatives in the preceding identity are replaced by shifts [1].  . As a result, standard approaches for constructing schemes with conservation involve either discretizing the continuous conservation law or finding multipliers for the scheme [1].
Alternatively, 
to verify that a particular scheme admits conservation laws, we can use the  discrete Euler operator [1]. Kupershmidt [II. Theorem 31]{kupershmidt85} discovered that an equation represents a discrete conservation law if and only if it is in the kernel of the discrete Euler operator
This strategy has been employed in recent studies, such as by Dorodnitsyn et al. for developing a shallow water equation scheme that preserves energy [1], and by Cheviakov et al. for creating linear and nonlinear wave equation schemes that admit discrete analogs of continuous conservation laws [1].
These strategies also apply to continuous or semi-discretized PDEs. Algorithms using multipliers and the Euler operator for PDEs were developed and implemented by Cheviakov [1]. Hereman et al. [1] proposed an algorithm for computing conserved densities for semi-discretized PDE schemes with first-order time derivatives, using scaling symmetries and the discrete Euler operator. Gao et al. extended this algorithm to first-order time-explicit [1] and time-implicit schemes [1].
Bihlo et al. [1] used another approach and derived finite-difference schemes for the shallow water equations, that are symmetry-invariant on an adaptive mesh and then analyzed them regarding conserved quantities.

In this paper, we present an algorithm that verifies if a quantity is conserved over time under a given scheme. Our approach diverges from previous methods, as we do not attempt to construct discrete [1] and [1] using the scheme, but rather check if a given discrete [1] is conserved over time under the scheme. Typically, conserved quantities involve integral expressions or their discrete analogs, such as sums. Gomes et al. proposed algorithms for simplifying sums [1] and developed algorithms for detecting conserved quantities in PDEs and semi-discretized schemes [1]. We review these techniques in Section [1].
To generalize these methods for cases without summing over all arguments of the involved functions, we introduce the discrete partial variational derivative (Section [1]). Our main contribution is an algorithm for verifying the conservation of a quantity under a numerical scheme. Gerdt [1] demonstrated that the quantity is conserved if its discrete time derivative belongs to the difference ideal generated by the scheme. However, some quantities may add only to a constant and be trivially preserved without being part of the difference ideal. Furthermore, Gerdt's algorithm may not always terminate, as the Gr{}bner basis for the difference ideal might be infinite.
We address these issues by combining the discrete partial variational derivative with a polynomial ideal (instead of a difference ideal) having a finite Gr{}bner basis (Section [1]). Our algorithm is applicable to schemes that are explicit and implicit in time, can handle schemes with multiple higher-order time and space derivatives, multiple space dimensions, systems of equations, and schemes with parameters.
As the process of finding conserved quantities involves polynomial division, special care must be taken to ensure that the division is performed correctly with respect to the parameters. We employ parametric Gr{}bner  bases, which are crucial for the proper treatment of parameters in our algorithm, see Section [1]
 We have implemented this algorithm as a part of a {Mathematica} package [1]{The code is available upon request. For access, contact ricardo.ribeirkaust.edu.sa.}. Our examples include Burgers equation and a PDE system related to mean-field games (Section [1]). In these examples, 
 we show that our code detects conserved quantities and can be used to select suitable schemes for time-implicit and time-explicit discretizations.

Preliminaries.

In this section, we introduce the concept of discrete torus and discrete variational derivatives, and discuss their properties and applications in simplifying sums. 
These definitions are essential for the analysis of methods presented in the rest of the paper.
Here, we 
discuss the connection between discrete functionals and discrete variational derivatives and present examples that illustrate how to simplify sums using this approach.
We follow the ideas in [1].
In this paper, subscripts denote coordinate or tuple indices, and superscripts denote sequence indices.
To avoid boundary terms, we work with periodic functions in [1].

When discretizing PDEs using finite difference schemes, we need to work with functions defined on discrete sets. For instance, periodic functions in [1] can be replaced by functions defined on the discrete torus, which will be introduced in the following discussion.

	Let [1], [1] and [1] be positive integers. The {discrete torus} is the set [1]. Define the space

		{P}({I},{R}^m):=\ |:{I}\mathbb{R}^m.
Functions in [1] extend periodically to [1].

We introduce discrete functionals in the previous function space. These functionals are the analog of integral functionals for discrete approximations of functions.

	The space of functionals [1] (not necessarily linear) on [1] is

		{F}({P}({I},{R}^m)):= \{{J}:{P}({I},{R}^m) {R}\Bigg|\mathcal{J}[u]=_{n\mathcal{I}}F_n[u],
		{ where }
		F_n[u]=G(u(n+e^1),u(n+e^2),,u(n+e^k))
		{ and }^1,,e^k\subset{{Z}}^d,\text{G is smooth}
		\}.

In the previous definition, [1] are not necessarily unit vectors and may vary between functionals.

	Let [1] and
	[1].

The discrete variational derivative is a useful tool for simplifying sums.
Let [1], [1] and [1]. Define

	D{J}[u](v):={d}{d}{J}[u+ v]|_{=0}.
In [1], we shift the indices because of the periodicity of [1] and [1],

		D{J}[u](v)={d}{d}{J}[u+ v]|_{=0}={d}{d}_{n\mathcal{I}}F_n[u+ v]|_{=0}
		=_{n\mathcal{I}}{d}{d}G(u(n+e^1)+ v(n+e^1),,u(n+e^k)+ v(n+e^k))|_{=0}
		=_{n\mathcal{I}}D_1G(u(n+e^1),,u(n+e^k))v(n+e^1)
		++D_kG(u(n+e^1),,u(n+e^k))v(n+e^k)
		=_{n\mathcal{I}}D_1G(u(n),u(n+e^2-e^1),,u(n+e^k-e^1))v(n)
		+D_2G(u(n+e^1-e^2),u(n),,u(n+e^k-e^2))v(n)
		++D_kG(u(n+e^1-e^k),u(n+e^2-e^k),,u(n))v(n)
		=_{n\mathcal{I}}(D_1G(u(n),u(n+e^2-e^1),,u(n+e^k-e^1))
		+D_2G(u(n+e^1-e^2),u(n),,u(n+e^k-e^2))
		++D_kG(u(n+e^1-e^k),u(n+e^2-e^k),,u(n)))v(n)
		=:_{n\mathcal{I}}{V}({J})[u](n)v(n).
The operator [1] can be viewed as a representation of the derivative [1] with respect to the [1] inner product [1], leading to the following definition.

	[1] is the {discrete variational derivative} of [1].

Because we assume all related functions are smooth, the discrete variational derivative always exists.

Now, we return to Example [1] and compute

		D{J}[u](v)={d}{d}{J}[u+ v]|_{=0}={d}{d}_{n\mathcal{I}}F_n[u+ v]|_{=0}
		=_{n\mathcal{I}}{d}{d}G[u(n+1)+ v(n+1),u(n)+ v(n)]|_{=0}
		=_{n\mathcal{I}}2(u(n+1)-u(n))v(n+1)+(-2)(u(n+1)-u(n))v(n)
		=_{n\mathcal{I}}2(u(n)-u(n-1))v(n)+(-2)(u(n+1)-u(n))v(n)
		=_{n\mathcal{I}}(2(u(n)-u(n-1))+(-2)(u(n+1)-u(n)))v(n)
		=_{n\mathcal{I}}(-2u(n+1)+4u(n)-2u(n-1))v(n)
		=_{n\mathcal{I}}{V}({J})[u](n)v(n).
Thus, [1].

The algorithms for the simplification of sums presented in [1] relies on the following result, which 
provides a criterion to compare functionals and evaluate if they represent the same quantity.

	Let [1]. If [1] for all [1] and if there exists [1] such that [1], then
	[1]
	for all [1].
	Conversely,
	[1]
	for all [1], if [1].

	We have

		({J}-{{J}})[u]=({J}-{{J}})[u]-({J}-{{J}})[u_0]
		=_{0}^{1}{d}{d}({J}-{{J}})[u_0+(u-u_0)]\lambda
		=_{0}^{1}{V}({J}-{{J}})[u_0+(u-u_0)](u-u_0)\lambda=0

	and hence [1].

We use Theorem [1] to determine whether different sums represent the same quantity.

	Let [1] and consider the functionals

		{J}[u]:=_{n\mathcal{I}}u(n)u(n+1){ and }
		{{J}}[u]:=_{n\mathcal{I}}u(n-2)u(n-1).
	It is clear that [1] and [1] represent the same quantity (by shifting [1]). We confirm this, using the discrete variational derivative:

		D({J}-{{J}})[u](v)=
		=_{n\mathcal{I}}(u(n+1)+u(n-1)-u(n+1)-u(n-1))v(n)=0.
	Hence, [1] and [1]. Therefore, both functionals represent the same quantity.

	Let [1] and [1]. It may not be obvious, that

		_{n_1,n_2}u(n_1,n_2-2)^4 - 3 u(n_1,n_2-2)^3 u(n_1+1,n_2-2) + 
		u(n_1,n_2) u(n_1+1,n_2)^3 + 3 u(n_1,n_2-2)^2 u(n_1+1,n_2-2)^2 -
		u(n_1,n_2-2) u(n_1+1,n_2-2)^3
		- u(n_1, n_2)^4 +
		3 u(n_1,n_2)^3 u(n_1+1,n_2)- 3 u(n_1,n_2)^2 u(n_1+1,n_2)
		-3 u(n_1,n_2-2)^2 u(n_1+1,n_2-2)^2+ 3 u(n_1,n_2)^2 u(n_1+1,n_2)=0.
	We confirm this by computing the discrete variational derivative and noticing that the expression is [1] for [1].

The discrete partial variational derivative.

In this section, we introduce the concept of the discrete partial variational derivative, a generalization of the discrete variational derivative for cases where one or several arguments of the indices are kept constant. This generalization provides a more flexible approach to handling various problems in conservation laws for finite-difference schemes, addressing challenges and limitations of the traditional discrete variational derivative. The main result of this section is Theorem [1], which gives a criterion for equality of two functionals based on their discrete partial variational derivatives."

Let [1] be the variables for the functions in [1]. We call [1] the space variables. Let [1] and call [1] the time variables. For generality, we consider the vectorial time case, while later in this paper, we focus specifically on the case when [1].
The analog to Definition [1] is the following.

	Let [1].
	Functions in [1] extend to [1] through  periodicity in the space variables.
	Let [1].

We can regard [1] as the set of functions depending in space and time and [1] as the functions depending only on time.

Next, we examine functions that involve summation only over spatial variables, extending the concept presented in Definition [1].

	The space of functions [1] is

		{F}({P}({I}\mathbb{N}^l,{R}^m),{D}({N}^l)):=	
		\{{J}:{P}({I}\mathbb{N}^l)\mathcal{D}({N}^l) |\mathcal{J}[u](t)=_{n\mathcal{I}}F_{(n,t)}[u],
		{where } F_{(n,t)}[u]=G(u((n,t)+e^1),u((n,t)+e^2),,u((n,t)+e^k)),
		{and }
		^1,,e^k\subset{{Z}}^{d+l},{G is a polynomial in k variables}\}.
	Here, [1].

Here, all quantities, whose conservation in time we check, are functions [1]. 
Let [1]. Using the Kronecker delta [1], we rewrite

	{J}[u](t_1,,t_{l})
	=_{{n}\mathcal{I}\mathbb{N}^l}F_{{n}}[u]_{(t_1,,t_{l})}(s_1,,s_l).
Let [1], then

	D^l{J}[u](v)(t_1,,t_{l}):={d}{d}{J}[u+ v]|_{=0}(t_1,,t_{l})
	=_{{n}}D_1G(u({n}+e^1),,u({n}+e^k))_{(t_1,,t_{l})}(s_1,,s_l)v({n}+e^1)
	+D_2G(u({n}+e^1),,u({n}+e^k))_{(t_1,,t_{l})}(s_1,,s_l)v({n}+e^2)+
	+D_kG(u({n}+e^1),,u({n}+e^k))_{(t_1,,t_{l})}(s_1,,s_l)v({n}+e^k)
	=_{{n}}D_1G(u({n}),,u({n}+e^k-e^1))_{(t_1,,t_{l})}(s_1-e^1_{d+1},,s_l-e^1_{d+l})v({n})
	+D_2G(u({n}+e^1-e^2),,u({n}+e^k-e^2))_{(t_1,,t_{l})}(s_1-e^2_{d+1},,s_l-e^2_{d+l})v({n})+
	+D_kG(u({n}+e^1-e^k),,u({n}))_{(t_1,,t_{l})}(s_1-e^k_{d+1},,s_l-e^k_{d+l})v({n})
	=_{{n}}(D_1G(u({n}),,u({n}+e^k-e^1))_{(t_1+e^1_{d+1},,t_{l}+e^1_{d+l})}(s_1,,s_l)
	+D_2G(u({n}+e^1-e^2),,u({n}+e^k-e^2))_{(t_1+e^2_{d+1},,t_{l}+e^2_{d+l})}(s_1,,s_l)+
	+D_kG(u({n}+e^1-e^k),,u({n}))_{(t_1+e^k_{d+1},,t_{l}+e^k_{d+l})}(s_1,,s_l))v({n})
	=:_{{n}}{V}^l({J})[u]({n},t)v({n}).

[1] is the {discrete partial variational derivative} of [1].
	Note that [1].

	The discrete partial variational derivative of

		{J}[u](t)=_nu(n,t+1)-u(n,t)=_{n,s}(u(n,s+1)-u(n,s))_t(s)

	is [1], because

		D^1({J})[u](v)(t)
		={d}{d} _{n,s}((u(n,s+1)+ v(n,s+1))-(u(n,s)+ v(n,s)))_t(s)
		=_{n,s}v(n,s+1)_t(s)-v(n,s)_t(s)=_{n,s}v(n,s)_t(s-1)-v(n,s)_t(s)
		=_{n,s}(_{t+1}(s)-_t(s))v(n,s)=_{n,s}{V}^1({J})(n,s,t)v(n,s).

	In our {Mathematica} implementation, Example [1] is computed by

		[moredefined={variables, expression, PartialDVarD}]{Input}
  variables=<|"indVars"[1],t,"depVars"[1]\},
  "timeVars"[1]\}|>
  PartialDVarD[variables][u[n,t+1]-u[n,t]]

		{Output}
  -KroneckerDelta[t]+KroneckerDelta[1+t]

We have a result similar to Theorem [1] for the discrete partial variational derivative:

	Let [1]. If [1] for all [1] and if there exists [1] such that [1], then
	[1]
	for all [1].
	Conversely,
	[1]
	for all [1], if [1].

The proof is similar to the proof of Theorem [1].

Difference schemes and algebra.

In this section, we formally define numerical schemes and introduce the necessary tools from difference algebra for our algorithm.
Throughout this paper, we assume that there is only one time variable [1] ([1]).

	The {space of schemes} is

		{S}({P}({I}\mathbb{N},{R}^m),{P}({I}\mathbb{N},{R})):=
		\{H:{P}({I}\mathbb{N},{R}^m)\mathcal{P}({I}\mathbb{N},{R})\Bigg|
		H[u](n,t)=H(u(n+{e}^1,t+{l}^1),,u(n+{e}^r,t+{l}^r)),
		{where }
		\tilde{e}^1,,{e}^r\subset{{Z}}^d,\{{l}^1,,{l}^r\subset{{Z}},
		{H is a polynomial in r variables}
		\}.
	A {scheme} is a set of functions

		^1,,H^i\subset{S}({P}({I}\mathbb{N},{R}^m),{P}({I}\mathbb{N},{R}))

	that represent the equations [1] for [1], holding pointwise for all points in [1].

If a scheme contains a single function [1], we call [1] the scheme. Sometimes, we also call the expression [1] the scheme.

	In our examples, we only consider finite-difference schemes with fixed step sizes [1]. However, our algorithms and our code handle parameters. Hence, our results can be generalized to schemes with other step sizes.

Consider the heat equation	[1] [1].  A possible scheme is
[1].
A solution of the corresponding numerical approximation is a function [1], that satisfies
[1]
for all [1] and [1].

	A scheme [1] given by
	[1]
	is in {time-explicit form} if we can rewrite the previous equation (eventually translating the scheme) as
	[1]
	for [1]
	with [1]. We call [1] the right-hand side of [1].

	Consider the forward-difference scheme for the heat equation in Example [1]. 
	This scheme is in time-explicit form with right-hand side
	[1].

{Difference ideals.}
Following Gerdt [1], we now introduce difference ideals and how we use them in our algorithm.

	Let [1]. The {shift} in the [1]-th coordinate for [1] by [1] is

		_i^k u_j(n,t):=u_j(n_1,,n_{i-1},n_i+k,n_{i+1},,t)

	understanding that [1] shifts the time variable.

	The {set of all possible shifts} [1] is

		:=\sigma_1^{k_1}\cdots\sigma_{d+1}^{k_{d+1}}|_1,,k_{d+1}{{Z}}.

We now construct the difference ideal containing the scheme.

	Let [1] be the field generated by [1] and the variables [1]. Let [1] be the polynomial ring over the field [1] and the variables [1] for [1] and [1].
	A set [1] is a {difference ideal} if
	[1] implies [1];
	[1] implies [1] and
	[1] implies [1].

	[1] is the smallest difference ideal containing the scheme [1].

A solution of a numerical scheme is a function, [1], that makes all translations of the scheme vanish. Hence, every element of the difference ideal generated by the scheme vanishes under [1]. As a result, every element of the difference ideal generated by the scheme vanishes under [1]. If the polynomial associated with the discrete time derivative  [1]
belongs to the ideal [1], the corresponding functional vanishes. 
 Gerdt introduced the concept of a standard  Gr{}bner basis for the ideal [1], which might not be finite. To address this issue, we use polynomial algebra instead of difference algebra and a smaller polynomial ideal contained within the difference ideal that has a finite standard basis.

 However, ideal membership is not  a necessary condition for conservation. Some functions, [1],
 may have sums that add to zero even if they do not belong to the difference ideal. Thus, we combine the discrete partial variational derivative with polynomial ideals. This idea is implemented in Algorithm [1] in Section [1].

Parametric Groebner Systems.

In this section, we delve into essential concepts for 
parametric polynomial ideals. These include parametric rings, 
specialized ideals, and comprehensive Gr{}bner systems, which are built upon 
Gr{}bner basis theory. 
We discuss the benefits of using comprehensive Gr{}bner systems 
and how they help address problems that occur when not using them. 
Furthermore, we expand on the ideal membership problem to encompass parametric ideals.

    Let [1] be a field. The ring of parametric polynomials
    [1] is a ring of polynomials in the variables [1] with coefficients that are polynomials
    in the ring [1]. 
    For simplicity we denote this ring by [1].

    Let [1] be an algebraically closed field extension of [1] and let [1] be a ring homomorphism. [1] is a specialization given by a polynomial evaluation on the parameters and an identity on the variables, i.e., 
    for [1]
    [1] is given by [1], for some [1].

It is easier to think about specializations as a choice of parameters.

Comprehensive Gr{}bner systems extend Gr{}bner basis to 
parametric polynomials and have diverse practical applications in 
algebraic geometry, combinatorics, coding theory, computer-aided design, and robotics.

    Let [1] be a finitely generated, nonzero ideal in the ring [1] with a fixed monomial order [1] on the variables. 
    Let [1] be a family of locally closed sets that forms a partition of [1] 
    , such that the disjoint union [1], and [1] 
    be a family of subsets in [1].
    The system, 
	[1].

    is called a comprehensive Gr{}bner system (CGS) of [1]
     if for all specializations [1] 
     is a Gr{}bner basis of the specialized ideal [1].
     Where each pair [1] is called a branch.

Here, the notation [1], with [1], is the specialization [1]
of each [1] such that  [1]
and  [1] stands for a specialization given by [1] 
for [1].

In this paper, we choose [1] to be of the form 
[1].
Where [1] and [1] is a Variety.

A proof of existence and the first algorithm to compute the (CGS) were introduced by Volker Weispfenning.

    Let [1]. The corresponding comprehensive Gr{}bner system exists.[1]

We use our implementation of a parametric Buchberger's algorithm to compute (CGS). 
This is illustrated in the following examples.

    Let [1] be the ideal generated by 
    [1].
    The comprehensive Gr{}bner system of [1] with a Lexicographic order [1] is:
    {  we use our code here}

The idea of a comprehensive Gr{}bner system is to provide a "parameter sensitive Gr{}bner basis".
It branches the cases of the parameters that affect leading terms. For instance, in the previous
example if [1] then the leading term [1] vanishes which leads to a different Gr{}bner basis. Therefore,
the (CGS) considers cases of the leading coefficients parameters Vanishing.

    Let [1] be the ideal generated by 
    [1].
    The comprehensive Gr{}bner system of [1] with a Lexicographic order [1] is:

    {  we use our code here}

    If a polynomial ideal contains a nonzero constant polynomial, then it is called a unit ideal [1].
   Unit ideals can generate all polynomials in the ring, so [1].
    As a convention the Gr{}bner basis of such ideals is [1].

Issues arise when computing the Gr{}bner basis of [1] without considering the parameters.
In practice, to compute the Gr{}bner basis without considering the parameters, we assume non-zero parameters by default.
This assumption guarantees a Gr{}bner basis only for one set of specializations [1]. i.e., 
[1] is a Gr{}bner basis of the specialized ideal [1] only for [1].
Next, we show the impact of this assumption.

     In general [1] is not necessarily a Gr{}bner basis.

        Let [1]. The corresponding Gr{}bner basis 
           is [1]. Consider 
           the specialization [1] given by [1].
           The set [1] is not a Gr{}bner basis.
		   This can be shown by the Buchberger Criterion (see chapter 2 in[1]).

     Consequently, the algorithm utilized to resolve the ideal membership 
    problem in chapter 2 of [1] can not be used for all specializations.

        Consider the ideal [1] in the above example. The corresponding Gr{}bner basis computed under the assumption of non-zero parameters
        is [1] where [1]. 
        Now, it is possible to check if a polynomial [1] belongs to [1] only for [1]     
        because we know that [1] is a Gr{}bner basis. For other specializations such as 
        the specialization [1] given by [1], [1] is not a Gr{}bner basis as shown in the above example. 
        Therefore, we can not proceed with the division algorithm as we can not guarantee unique remainder if we used [1].

Using CGS can resolve these problems. Problem 1 is fixed by the definition of (CGS), as we have [1] a 
 GrA¶bner basis for any specialization [1]. Moreover, 
 to discuses problem 2 we need to look at the parametric ideal membership first.

{  Is this a definition?}
 Consider, [1] we say that [1] under [1], if there exists a specialization 
 [1] such that [1].

To determine if [1] we consider the (CGS) of [1] given [1]. 
We check if [1] for each [1] and [1].
For each [1] we have the corresponding GrA¶bner basis [1]
so we divide [1] by the [1] if the remainder is zero for some [1],
then [1]. Therefore, [1]. However, if the remainder is non-zero for all [1] we conclude that 
[1].

	Let [1] be the ideal generated by 
    [1].
	We will determine if [1].
	We compute the corresponding (CGS)
	{  add example 3}
	First let [1] and [1]. 
	We divide [1] by [1] and we obtain the remainder [1] and 
	since the exists no specialization in [1] such that this remainder is zero we conclude that [1] for
	[1]. Now we repeat the reduction of [1] but with the second GrA¶bner basis [1]. We obtain a zero remainder, which imply that 
	[1] for [1]. Therefore, [1] under a specialization [1].

The time-explicit case.

In this section, we introduce Algorithm [1]  for checking the conservation of a quantity for time-explicit schemes. We discuss the algorithm for general schemes in the next section.

	{[1] and [1] right-hand side of a time-explicit scheme}
	{ {True}, if [1] is conserved in time, {False} otherwise }

	 Build the discrete time derivative [1]

	 Replace all instances of [1] (for [1]) by an appropriate translation of [1]

	 Compute the discrete partial variational derivative

	 If the result is zero, conservation is {True}, else {False}

	{{time-explicit DiscreteConservedQ}}

The result of Step 2 exists only at time [1] because all instances of [1] have been replaced. Thus, we do not need any computations of difference ideals or Gr{}bner basis.

	We examine the conservation of [1] under the following scheme.
	[1].
	The discrete time derivative (Step 1) is

		_{n\mathcal{I}}u(n,t+1)-u(n,t).
	Replace [1] by the right-hand side (Step 2) to get

		_{n\mathcal{I}}u(n,t)+(u(n+1,t)-2u(n,t)+u(n-1,t))-u(n,t).
	Then, we compute the discrete partial variational derivative (Step 3), which equals zero. Hence (Step 4), we have conservation.

	We verify the result from Example [1], using our implementation in {Mathematica} of Algorithm [1].

		[moredefined={variables, expression, DiscreteConservedQ}]{Input}
  variables=<|"indVars"[1]\},"depVars"[1]\},
  "rhs"[1][n]+(u[n+1]-2u[n]+u[n-1])|>
  DiscreteConservedQ[variables][u[n]]

		{Output}
  True

{ ADD HERE TIME EXPLICIT EXAMPLE WITH PARAMETERS}

The general case.

In this section, we present Algorithm [1] that deals with general, not necessarily time-explicit, schemes. We explain its steps in detail below and demonstrate the algorithm in Example [1].

	{[1] and [1] }
	{ {True}, if [1] is conserved in time, {False} otherwise }

	 Build the discrete time derivative [1]

	Translate [1] by subtracting the range of the stencil of the scheme from the range of the stencil of [1]

	 Reduce [1] by using {  the Gr{}bner basis} generated by the translated scheme in two different ways
	 (a) Using the lexicographic order
	 (b) Using the explicit elimination order

	 Choose the result that admits the least different instances of time

	 Compute the discrete partial variational derivative

	 Apply Steps 2 to 4 to the result

	 If the result is zero, conservation is {True}, else {False}

	{{General DiscreteConservedQ}}

{Step 1: Build the discrete time derivative.}
We build the discrete time derivative by subtracting [1] from [1]. This task is performed in our code using {TimeDifference}.

	The discrete time derivative for [1] is

		[moredefined={variables, expression, TimeDifference}]{Input}
  variables=<|"indVars"[1],t,"depVars"[1]\}|>
  TimeDifference[variables][u[n,t]]

		{Output}
  <|exp[1]-u[n,t]+u[n,1+t]|>

{Step 2: Translate the scheme.}
Now, we transition from difference algebra with infinite independent variables to polynomial algebra with finite variables to ensure the finiteness of the Gr{}bner basis. 
Hence, we need to compute a finite number of translations of the scheme and consider every instance of [1] that appears in either the translated scheme or the discrete time derivative as an independent polynomial variable.

	The algorithm treats the discrete time derivative [1]
	as the polynomial [1].

There are several possiblities which translations of the scheme to take as basis for the polynomial ideal. However, an obvious idea is to compute the minimal number of translations of the scheme, such that all instances of [1], that appear in the range of the stencil of the discrete time derivative, also appear in the translated scheme.
This is done by elementwise subtracting the range of the stencil of the scheme from the range of the stencil of the discrete time derivative. Before we define the above translations formally, we give a simple example to illustrate the idea of computing the translations:

	Consider the expression

		u(n+3)-u(n-2)

	which we want to reduce by the scheme [1]. We compute the translations as described above as

		(n+3)-u(n+2),,u(n-1)-u(n-2).
	and see, that the expression indeed reduces to [1] under the scheme.

	Let [1] and

		F:
		{P}({I}\mathbb{N},{R}^m)\mathcal{P}({I}\mathbb{N},{R}),
		F[u](n,t)=F(u_1((n,t)+x^1),,u_1((n,t)+x^{k_1}),
		u_2((n,t)+x^{k_1+1}),
		,u_2((n,t)+x^{k_2}),
		,
		u_m((n,t)+x^{k_{m-1}+1}),,u_m((n,t)+x^{k_m}))

	with [1].
	The {stencil} of [1] is the [1]-tuple of sets of vectors

		(\{x^1,,x^{k_1}\},
		\{x^{k_1+1},,x^{k_2}\},
		,
		\{x^{k_{m-1}+1},,x^{k_m}\}).
	The {range} of the stencil of [1] is

		([x_1^{k_1 -},x_1^{k_1 +}]\cdots[x_{d+1}^{k_1 -},x_{d+1}^{k_1 +}],
		[x_1^{k_2 -},x_1^{k_2 +}]\cdots[x_{d+1}^{k_2 -},x_{d+1}^{k_2 +}],
			,
		[x_1^{k_m -},x_1^{k_m +}]\cdots[x_{d+1}^{k_m -},x_{d+1}^{k_m +}]
		),
	where [1], [1],
	and [1]. Here, [1] denotes the discrete interval in [1], i.e. [1].

If for any [1] involved we have [1], we write [1] with the convention that [1].  Further, [1] in the translations results in the use of the respective entry of the original scheme without translations.

	Consider the discrete time derivative

	[1]

	and the scheme

	[1].

	Then, the stencil of the discrete time derivative equals

	[1]

	with range

	[1].

	The stencil of the scheme equals the set of stencils of the equations in the scheme, i.e.

	[1]

	with range

	[1].

	Hence, we get the translations for the first equation of the scheme by

		([0,2][0,1],[0,0] [0,1])-([0,1][0,1],\Big)=([0,1][0,0],\Big)

	and for the second equation of the scheme by

		([0,2][0,1],[0,0] [0,1])-(,[0,0] [0,1])=(,[0,0] [0,0]).

	Therefore, we translate the first equation of the scheme by [1] and [1] and the second one by [1] to get the translated scheme

		_1(n+1,t+1)-u_1(n,t),
		u_1(n+2,t+1)-u_1(n+1,t),
		u_2(n,t+1)-u_2(n,t)
		.

{Step 3 and 4: Compute Gr{}bner basis and reduce the discrete time derivative.}
{Polynomial ideals and Gr{}bner bases.}
In the previous step, we identified a finite set of polynomials comprising the translated scheme [1] and the discrete time derivative, each containing a finite number of instances of [1]. We now proceed to reduce [1] by treating it as an element of a polynomial ring and performing multivariate polynomial division with respect to the translated scheme. To accomplish this, we adapt the definitions and theorems presented in [1] to suit our specific context.

	Let [1] be as in Definition [1]. Let [1] be the polynomial ring generated by [1] and all instances of [1] that occur in [1].
	We call a set [1] a {(polynomial) ideal} if
	[1] implies [1] and
	[1] implies [1].

	We denote by [1] the smallest (polynomial) ideal containing [1].

Given the ideal [1] and the discrete time derivative [1], we want to determine if [1] or if we can write [1] in a simpler form, using [1].
Hence, using multivariate polynomial division, we search for [1] such that

	{T}=p^1H^1++p^rH^r+p^{r+1}.
Unfortunately, the resulting remainder [1] may not be unique [page 14, Example 1.2.3]{hibi13}, i.e. non-zero, although [1] belongs to [1]. 
Replacing [1] by a Gr{}bner basis for the ideal [1] guarantees the uniqueness of the remainder of the polynomial division. 
Contrary to the standard basis for the difference ideal that Gerdt's algorithm computes, a Gr{}bner basis for the polynomial ideal always exists and is finite.
A Gr{}bner basis is defined up to the order of the monomials involved, so, depending on the order, we get different remainders of the polynomial division.

{Polynomial reduction.}
In Algorithm [1], we employ two monomial orders: (a) the lexicographic order and (b) an experimental elimination order. To reduce [1], we first compute the Gr basis of [1] with respect to the chosen monomial order and subsequently calculate the remainder resulting from the polynomial division of [1] with respect to this Gr basis.

The lexicographic order (a) is induced by the time-explicit ordering of the instances of [1], giving greater importance to instances at later times over those at earlier times. For the elimination order (b), we initially order all instances of [1] based on a predefined ordering. By default, our code employs the time-explicit ordering, but other options, such as implicit or user-defined orderings, can also be used. We then attempt to sequentially eliminate the instance with the highest ordering from the discrete time derivative, utilizing the elimination order induced by a weight-matrix. This elimination may not always be feasible, resulting in an unchanged discrete time derivative. In such cases, we repeat the process for the instance at the second latest time and continue until all instances of [1] have been eliminated.

It should be noted that the elimination order (b) is experimental in nature, as its weight-matrix lacks full rank and therefore does not induce a total order on the set of monomials. Finally, we select the result with the fewest distinct time instances from the two remainders produced by the lexicographic order (a) and the elimination order (b).

 {Step 5 to 7: Compute the discrete partial variational derivative and reduce again.}
In the next step, we take the resulting expression and compute its discrete partial variational derivative. Then, we repeat Step 2 to Step 4 applied to the discrete partial variational derivative.

	To illustrate Algorithm [1], we use the setting of Example [1].
	We compute the discrete time derivative (Step 1)

		{T}=_{n\mathcal{I}} u(n,t+1)-u(n,t).
	For the translations (Step 2), the stencil of the scheme
	[1]

	is
	[1]

	with range [1].
	The discrete time derivative has stencil [1] 
	and range [1].
	The translations are

		([0,0][0,1])-([-1,1][0,1])=([1,-1][0,0])=(\times[0,0])

	which results in no translations, as the range of the stencil of the scheme is greater than the range of the stencil of the discrete time derivative.
	Hence, the Gr{}bner basis (Step 3) coincides with the original scheme for both monomial orders.
	The reduction using the lexicographic monomial order yields as remainder

		u(n-1,t)-2u(n,t)+u(n+1,t).
	For the reduction using the elimination order, we first eliminate [1] and then the remaining instances of [1].
	The elimination order yields the same remainder as the lexicographic order. Hence, we do not check for the number of instances of time (Step 4).
	We calculate the discrete partial variational derivative (Step 5) of

		_{n\mathcal{I}}u(n-1,t)-2u(n,t)+u(n+1,t)

	that equals [1].
	Hence, repeating the above procedure (Step 6) becomes unnecessary.
	Therefore, (Step 7) we see that the scheme conserves the quantity.

	We verify our result from Example [1], using {DiscreteConservedQ}, which detects automatically if the scheme is time-explicit or if
	 the general setting should be applied.

		[moredefined={variables, expression, DiscreteConservedQ}]{Input}
  variables=<|"indVars"[1],t,"depVars"[1]\},
  "scheme"[1][n,t+1]-u[n,t]
  -(u[n+1,t]-2u[n,t]+u[n-1,t])|>
  DiscreteConservedQ[variables][u[n,t]]

		{Output}
  True

The algorithm is designed to identify conservation, but it cannot definitively confirm the absence of conservation. Therefore, a {False} result should be interpreted as the algorithm being unable to detect conservation, rather than an absolute assertion of non-conservation. The reason for this is that we are working with a Gr basis that does not generate the full difference ideal.

{  add an example with parameters}

A basis for conserved quantities.

So far, we have discussed how to check if a quantity is preserved in time under a scheme. But it is also desirable to have a systematic way to find conserved quantities, given a scheme.
Algorithm [1] finds, for a time-explicit scheme, a basis for conserved quantities that are generated by monomials up to a degree.

	{right-hand side set of polynomials [1] ofa time-explicit scheme, list of {generators}, {degree} of polynomials }
	{basis of conserved quantities}

	 Construct all monomials combined from the {generators} up to the total {degree}

	 Form linear combinations of all monomials with undetermined coefficients

	 Compute the discrete time derivative and substitute in the appropriate right-hand side(s) polynomials of the scheme

	Compute the discrete partial variational derivative

	 Simplify the expression and apply the undetermined coefficients method to find coefficients that make the result vanish

	{{FindDiscreteConservedQuantityBasis}}

We have implemented this algorithm in {FindDiscreteConservedQuantityBasis}.

In the context of the heat equation from Example [1], we aim to find a basis for conserved quantities generated by [1] and [1] with a maximum degree of 3.

		[moredefined={variables, FindDiscreteConservedQuantityBasis}]{Input}
  variables=<|"indVars"[1]\},"depVars"[1]\},
  "rhs"[1][n]+u[n+1]-2u[n]+u[n-1]|>
  FindDiscreteConservedQuantityBasis[variables]
  [<|"degree"[1]3,"generators"[1][n],n|>]

		{Output}
  [n],nu[n]

The first conserved quantity corresponds to mass conservation and the second to center of mass conservation.

Examples and applications.

We demonstrate our code's capabilities using two examples: the inviscid Burgers equation and a mean-field game system. 
Through these examples, we examine different discretizations, evaluate their conservation properties, and identify conserved quantities.
Although the examples presented feature only first-order time derivatives, the algorithm can handle PDEs with derivatives of any order.

{Burgers equation.}
In the first example, we focus on the inviscid Burgers equation, a well-known PDE with the form [1] [1]. 
This PDE preserves the total mass [1].
Our first example demonstrates how the code verifies mass conservation by a forward-difference discretization.
Next, we demonstrate the algorithm's ability to find a mass-preserving scheme by investigating conservation properties under a class of schemes involving a parameter.
Finally, we explore time-implicit versions of the schemes and their conservation properties.

The inviscid Burgers equation is the PDE [1] [1] in one space dimension. Any function of [1] is a conserved quantity. We search for discretizations preserving [1].

	We check for conservation of mass using a forward-difference discretization:

		[moredefined={variables, expression, DiscreteConservedQ}]{Input}
  variables=<|"indVars"[1],t,"depVars"[1]\},
  "scheme"[1][n,t+1]-u[n,t]
  -u[n,t](u[n+1,t]-u[n,t])|>
  DiscreteConservedQ[variables][u[n,t]]

		{Output}
  False

For finding a scheme that preserves mass, we check for conservation under a class of schemes with a parameter.

	We discretize [1] with a three-point stencil with a parameter [1].

		[moredefined={variables, expression, DiscreteConservedQ}]{Input}
  variables=<|"indVars"[1],t,"depVars"[1]\},
  "pars"[1]\},"scheme"[1][n,t+1]-u[n,t]
  -u[n,t](a(u[n+1,t]-u[n,t])
  +(1-a)(u[n,t]-u[n-1,t]))|>
  DiscreteConservedQ[variables][u[n,t]]

		{Output}
  False		a==0||a==1||-1+2a[1]0
  True		a=={1}{2}

The conservation property holds when [1], which corresponds to the central difference case.

So far, we have only seen schemes that are in time-explicit form, but our algorithm allows also for general schemes:

	We consider the scheme from the previous example but in the time-implicit version.

		[moredefined={variables, DiscreteConservedQ}]{Input}
  variables=<|"indVars"[1],t,"depVars"[1]\},
  "pars"[1]\},"elimOrder"[1]"implicit",
  "scheme"[1][n,t+1]-u[n,t]-u[n,t+1]
  (a(u[n+1,t+1]-u[n,t+1])
  +(1-a)(u[n,t+1]-u[n-1,t+1]))|>
  DiscreteConservedQ[variables][u[n,t]]

		{Output}
  False		a==0||a==1||-1+2 a[1]0
  True		a=={1}{2}

{Conserved quantities for a mean-field game.}
We analyze the conserved quantities preserved by the discretization of a PDE system in the context of mean-field games.
The examples showcase our algorithm's effectiveness in identifying conserved quantities for various discretization schemes, including time-explicit, time-implicit, and mixed schemes.

	In [1], Gomes et al. derived the following system

		v_t+({v^2}{2}-{m^2}{2})_x=0,
		m_t-(vm)_x=0.
	This system admits the conserved quantities [1] and [1].
	Because this system was derived from a forward-forward mean-field game, we discretize it forward in time. We check with our code if the scheme admits the same preserved quantities as the continuous system:

		[moredefined={variables, DiscreteConservedQ}]{Input}
  variables=<|"indVars"[1],t,"depVars"[1],m,
  "scheme"[1][n,t+1]-v[n,t]
  +({v[n+1,t]}{2}-{m[n+1,t]}{2})/2-({v[n,t]}{2}-{m[n,t]}{2})/2,
  m[n,t+1]-m[n,t]-m[n+1,t]v[n+1,t]+m[n,t]v[n,t]|>
  DiscreteConservedQ[variables][[n,t],m[n,t]]

		{Output}
  True

We replicate this result, observing that the scheme used here is time-explicit:

		[moredefined={variables, DiscreteConservedQ}]{Input}
  variables=<|"indVars"[1]\},"depVars"[1],m,
  "rhs"[1][n]-({v[n+1]}{2}-{m[n+1]}{2})/2
  +({v[n]}{2}-{m[n]}{2})/2,m[n]+m[n+1]v[n+1]-m[n]v[n]|>
  DiscreteConservedQ[variables][[n],m[n]]

		{Output}
  True

	Those are the only conserved quantities for this scheme, which are polynomials up to degree 4 in [1] and [1].

		[moredefined={variables, PartialDVarD, FindConservedQuantityBasis}]{Input}
  variables=<|"indVars"[1],t,"depVars"[1],m,
  "eqRhs"[1]0.5({v[n+1,t]}{2}-{m[n+1,t]}{2})
  -0.5({v[n,t]}{2}-{m[n,t]}{2}),
  m[n+1,t]v[n+1,t]-m[n,t]v[n,t]|>
  FindConservedQuantityBasis[variables]
  [<|"degree"[1]4,"generators"[1][n,t],m[n,t]|>]

		{Output}
  [n,t],v[n,t]

	The related backward-forward system reads

		-v_t+({v^2}{2}-{m^2}{2})_x=0,
		m_t-(vm)_x=0.
	This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly in time for [1] and implicitly for [1].
	Our code can handle this setting by specifying an order for variable elimination, using an explicit monomial order for [1] and an implicit order for [1]:

		[moredefined={variables, DiscreteConservedQ}]{Input}
  variables=<|"indVars"[1],t,"depVars"[1],v,
  "elimOrder"[1]"explicitimplicit",
  "scheme"[1]-(v[n,t+1]-v[n,t])
  +({v[n+1,t+1]}{2}-{m[n+1,t+1]}{2})/2
  -({v[n,t+1]}{2}-{m[n,t+1]}{2})/2,
  m[n,t+1]-m[n,t]-m[n+1,t]v[n+1,t]+m[n,t]v[n,t]|>
  DiscreteConservedQ[variables][[n,t],m[n,t]]

		{Output}
  True

Possible extensions and concluding remarks.

 Polynomial treatment of non-polynomial expressions.

It may be possible to extend our methods to non-polynomial PDEs and schemes. In this case, non-polynomial expressions can be handled by writing them in polynomial form.
{Translate the scheme more accurately.}
In our elimination procedure, we work with a specific polynomial ideal that is a subset of the difference ideal generated by the scheme. However, it could be necessary to translate the scheme more than we did in our algorithm to get cancellation. Therefore, flexible methods to determine the translations of the scheme are desirable.
Experimental check of conservation.

An experimental approach, that gives a hint, if a quantity is conserved, and that can refute conservation, is the following:
We fix [1] and [1]. Then, we generate experimental initial data by randomly sampling [1] values, with every value corresponding to an instance of [1] for [1]. Afterward, we calculate [1] for all [1], using the scheme. With this values, we explicitly calculate the value of the discrete time derivative.
If the discrete time derivative is zero for several different samples of initial data, we might have conservation on the discrete level. Otherwise, if we find initial data such that the discrete time derivative is not zero, we can refute conservation.
Discrete conserved quantities that are not conserved by the PDE.

It may arise the question if we can find conserved quantities, that are not preserved by the original PDE. Gerdt et al. [1] define the notion of s-consistency, which guarantees that all discrete quantities are also preserved in the continuous setting. One possible extension of the code would be to check if s-consistency holds for the translated scheme.
Concluding remarks.

We present an algorithm for checking the conservation of a quantity under a finite-difference scheme. Our algorithm allows for systems of equations, arbitrary time and space derivatives, and both explicit and implicit schemes. Also, we implemented a function for finding conserved quantities admitted by a scheme. We use our implementation of the algorithm to analyze the conservation properties of several schemes for PDEs that arise in applications. {  add remark on parametric groebner basis}

{splncs04}
{references}

\end{document}

### 2023-05-09 Papers/Paper CASC format/paper.tex
working on page 12/13

@@
 
+{\color{blue} Just include the parameters in the variables in a way that they are less expensive then the original variables. Consider $\mathcal{K}[X,A]$.}
+
 Issues arise when computing the Gr{\"o}bner basis of $\Ff$ without considering the parameters.
-In practice, to compute the Gr{\"o}bner basis without considering the parameters, we assume non-zero parameters by default.
-This assumption guarantees a Gr{\"o}bner basis only for one set of specializations $\Rr^2\backslash V(a,b)$. i.e., 
+In practice, not considering the parameters means they are assumed non-zero by default.
+With this assumption, only for one set of specializations  $\Rr^2\backslash V(a,b)$ do we get a Gr{\"o}bner basis, i.e. 
 $\sigma(\Gg)$ is a Gr{\"o}bner basis of the specialized ideal $\sigma(\Ii)$ only for $\sigma \in \Rr^2\backslash V(a,b)$.
-Next, we show the impact of this assumption.
+Next, we discuss the impact of such an assumption:
 \begin{enumerate}
@@
         Let $\Ii=\langle 
-           {a x^2 - y, -2 + b x y} \rangle$. The corresponding Gr{\"o}bner basis 
+           {a x^2 - y, -2 + b x y} \rangle$. The corresponding Gr{\"o}bner basis, with the variables $x>y>a>b$ with the Lexicographic monomial ordering, 
            is $\Gg=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$. Consider 
@@
     \end{example}
-    \item Consequently, the algorithm utilized to resolve the ideal membership 
+    \item The algorithm utilized to resolve the ideal membership 
     problem in chapter 2 of \cite{Cox} can not be used for all specializations.
@@
         because we know that $\sigma_0(\Gg)$ is a Gr{\"o}bner basis. For other specializations such as 
-        the specialization $\sigma$ given by $P\rightarrow P(0,1)$, $\sigma(\Gg)$ is not a Gr{\"o}bner basis as shown in the above example. 
+        the specialization $\sigma$ given by $P\rightarrow P(0,1)$, $\sigma(\Gg)$ is not a Gr{\"o}bner basis as discussed in the above example. 
         Therefore, we can not proceed with the division algorithm as we can not guarantee a unique remainder if we use $\sigma(\Gg)$.  
@@
 
+	{\color{blue} For this specialization (0,1), we have that the GB is \{1\}. But pseudo-division of 1 by the "not"-GB obtained does not have 0 remainder.}
+
 \end{enumerate}
 
-Using CGS can resolve these problems. Problem 1 is fixed by the definition of (CGS), as we have $\sigma_i(G_i)$ a 
- Gr{\"o}bner basis for any specialization $\sigma_i$. Moreover, 
- to discuss problem 2, we need to look at the parametric ideal membership first.
+Using CGS resolves these problems. Problem 1 is {\color{red}fixed} by the definition of (CGS), as $\sigma_i(G_i)$ is a 
+ Gr{\"o}bner basis for any specialization $\sigma_i\in A_i$. Moreover, 
+ to discuss problem 2, we need to look/define at the parametric ideal membership first.
 

### 2023-05-09 Papers/Paper CASC format/papercopy.tex
paper

--- 2023-05-09 Papers/Paper CASC format/paper.tex
@@
 \author{Diogo A. Gomes\inst{1}\orcidID{0000-0002-3129-3956} \and
-	Friedemann Krannich\inst{1}\orcidID{1111-2222-3333-4444} \and
+	Friedemann Krannich\inst{1}\orcidID{0000-0002-8642-4319} \and
 	Bashayer Majrashi\inst{1}\orcidID{0009-0008-3677-8064 } \and
@@
 	Many partial differential equations (PDEs) admit conserved quantities like mass or energy.
-These quantities are often essential for establishing well-posedness results.
-When approximating a PDE with a finite-difference scheme, it is crucial to determine whether related discretized quantities remain conserved over time by the scheme.
+These are often essential for establishing well-posedness results.
+When approximating a PDE with a finite-difference scheme, it is crucial to determine whether related discretized quantities remain conserved by the scheme.
 Such conservation may ensure the stability of the numerical scheme.
 We present an algorithm for verifying the preservation of a polynomial quantity under a polynomial finite-difference scheme.
-Our schemes can be explicit or implicit, have higher-order time and space derivatives, and an arbitrary number of variables.
-Additionally, we introduce an algorithm for finding conserved quantities, given a scheme.
-We illustrate our algorithm with several finite-difference schemes.
+Our schemes can be explicit or implicit, have higher-order time and space derivatives, and  arbitrary number of variables.
+Additionally, we introduce an algorithm for finding conserved quantities.
+We illustrate our algorithm in several finite-difference schemes.
 Our approach incorporates parametric G\"obner bases to handle parameters, ensuring accurate computation of conserved quantities.
 
-	\keywords{Symbolic computations \and Finite-difference schemes \and Discrete variational derivative \and Discrete partial variational derivative \and Conserved quantities \and Implicit schemes \and Explicit schemes \and Parametric Groebner basis.}
+	\keywords{Finite-difference schemes \and Discrete variational derivative \and Discrete partial variational derivative \and Conserved quantities \and Implicit schemes \and Explicit schemes \and Parametric Groebner systems.}
 \end{abstract}
@@
 This information is not only valuable for gauging the accuracy of the approximation, but also vital for evaluating its stability.
-Automating conservation computations in computer algebra systems streamlines the process, saves time, enhances accuracy, and may create new research opportunities.
+Automating conservation computations in computer algebra systems streamlines the process, saves time and enhances accuracy.
 
@@
 Conserved quantities correspond to conservation laws, admitted by the PDE. A conservation law \cite{cheviakov20} is an equation of the form
-\begin{gather*}
+$
 	\frac{\partial }{\partial t}\Phi[u]+\div \Psi[u]=0
-\end{gather*}
+$
 holding for all solutions $u$ of the considered PDE,
 that induces the conserved quantity $\int \Phi[u]\ dx$ as
-\begin{gather*}
+$
 	\frac{d}{dt}\int \Phi[u]\ dx=\int D_t\Phi[u]\ dx=-\int D_x\Psi[u]\ dx=0
-\end{gather*}
+$
 assuming periodic boundary conditions in $x$.
@@
 As the process of finding conserved quantities involves polynomial division, special care must be taken to ensure that the division is performed correctly with respect to the parameters.
-We employ parametric Gr{\"o}bner  bases, which are crucial for the proper treatment of parameters in our algorithm, see Section \ref{pgb}
+We employ parametric Gr{\"o}bner  bases, which are crucial for the proper treatment of parameters in our algorithm, see Section \ref{pgs}.
  We have implemented this algorithm as a part of a {\sc Mathematica} package \cite{wolfram21}\footnote{The code is available upon request.
@@
 	\label{d1}
-	Let $N$, $d$ and $m$ be positive integers. The \emph{discrete torus} is the set $\Ii:=\{0,\ldots,N-1\}^d\subset \Zz^d$. Define the space
-	\begin{gather*}
+	Let $N$, $d$ and $m$ be positive integers. The \emph{discrete torus} is the set $\Ii:=\{0,\ldots,N-1\}^d\subset \Zz^d$. Let
+$
 		\mathcal{P}(\Ii,\Rr^m):=\{u\ |\ u:\Ii\rightarrow\Rr^m\}.
-	\end{gather*}
+$
 Functions in $\mathcal{P}(\Ii,\Rr^m)$ extend periodically to $\Zz^d$.
@@
 	\begin{gather*}
-		\mathcal{F}(\mathcal{P}(\Ii,\Rr^m)):= \Bigg\{\mathcal{J}:\mathcal{P}(\Ii,\Rr^m)\to \Rr\ \Bigg|\ \mathcal{J}[u]=\sum_{n\in\Ii}F_n[u],\\
-		\text{ where }
-		F_n[u]=G(u(n+e^1),u(n+e^2),\ldots,u(n+e^k))\\
-		\text{ and }\{e^1,\ldots,e^k\}\subset\Zz^d,\ \text{G is smooth}
-		\Bigg\}.
+		\mathcal{F}(\mathcal{P}(\Ii,\Rr^m)):= \Bigg\{\mathcal{J}:\mathcal{P}(\Ii,\Rr^m)\to \Rr\ \Bigg|\ \mathcal{J}[u]=\sum_{n\in\Ii}F_n[u],\text{ where for} G \ \text{ smooth} \\
+		F_n[u]=G(u(n+e^1),u(n+e^2),\ldots,u(n+e^k))
+		\text{ and }\{e^1,\ldots,e^k\}\subset\Zz^d
+		\Bigg\}, 
 	\end{gather*}
-\end{definition}
-In the previous definition, $e^1,\ldots,e^k$ are not necessarily unit vectors and may vary between functionals.
+where $e^1,\ldots,e^k$ are may not be unit vectors and may vary between functionals.
+\end{definition}
 \begin{example}
@@
 \in\Rr$ and $\mathcal{J}\in\mathcal{F}(\mathcal{P}(\Ii,\Rr^m))$. Define
-\begin{gather*}
+$
 	D\mathcal{J}[u](v):=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}.
-\end{gather*}
+$
 In \eqref{periodicity}, we shift the indices because of the periodicity of $u$ and $v$,
@@
 	\notag	=\sum_{n\in\Ii}\frac{d}{d\epsilon}G(u(n+e^1)+\epsilon v(n+e^1),\ldots,u(n+e^k)+\epsilon v(n+e^k))\bigg|_{\epsilon=0}\\
-	\notag	=\sum_{n\in\Ii}D_1G(u(n+e^1),\ldots,u(n+e^k))v(n+e^1)\\
-	\notag	+\ldots+D_kG(u(n+e^1),\ldots,u(n+e^k))v(n+e^k)\\
-	\label{periodicity}	=\sum_{n\in\Ii}D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))v(n)\\
-	\notag	+D_2G(u(n+e^1-e^2),u(n),\ldots,u(n+e^k-e^2))v(n)\\
-	\notag	+\ldots+D_kG(u(n+e^1-e^k),u(n+e^2-e^k),\ldots,u(n))v(n)\\
+%	\notag	=\sum_{n\in\Ii}D_1G(u(n+e^1),\ldots,u(n+e^k))v(n+e^1)\\
+%	\notag	+\ldots+D_kG(u(n+e^1),\ldots,u(n+e^k))v(n+e^k)\\
+	\label{periodicity}	%=\sum_{n\in\Ii}D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))v(n)\\
+%	\notag	+D_2G(u(n+e^1-e^2),u(n),\ldots,u(n+e^k-e^2))v(n)\\
+	%\notag	+\ldots+D_kG(u(n+e^1-e^k),u(n+e^2-e^k),\ldots,u(n))v(n)\\
 	\notag	=\sum_{n\in\Ii}\Big(D_1G(u(n),u(n+e^2-e^1),\ldots,u(n+e^k-e^1))\\
-	\notag	+D_2G(u(n+e^1-e^2),u(n),\ldots,u(n+e^k-e^2))\\
-	\notag	+\ldots+D_kG(u(n+e^1-e^k),u(n+e^2-e^k),\ldots,u(n))\Big)v(n)\\
-	\notag	=:\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n).	
+%	\notag	+D_2G(u(n+e^1-e^2),u(n),\ldots,u(n+e^k-e^2))\\
+	\notag	\ldots+D_kG(u(n+e^1-e^k),\ldots,u(n))\Big)v(n)
+=:\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n).	
 \end{gather}
-The operator $\mathcal{V}(\mathcal{J})$ can be viewed as a representation of the derivative $D\mathcal{J}u$ with respect to the $L^2$ inner product $\langle w, v\rangle=\sum_{n\in \Ii} w_n v_n$, leading to the following definition.
+The operator $\mathcal{V}(\mathcal{J})$ represents the derivative $D\mathcal{J}u$ with respect to the $L^2$ inner product $\langle w, v\rangle=\sum_{n\in \Ii} w_n v_n$, leading to the following definition.
 \begin{definition}
@@
 \end{definition}
-Because we assume all related functions are smooth, the discrete variational derivative always exists.
+Because here all functions are smooth, the discrete variational derivative exists.
 \begin{example}
@@
 		D\mathcal{J}[u](v)=\frac{d}{d\epsilon}\mathcal{J}[u+\epsilon v]\bigg|_{\epsilon=0}=\frac{d}{d\epsilon}\sum_{n\in\Ii}F_n[u+\epsilon v]\bigg|_{\epsilon=0}\\
-		=\sum_{n\in\Ii}\frac{d}{d\epsilon}G[u(n+1)+\epsilon v(n+1),u(n)+\epsilon v(n)]\bigg|_{\epsilon=0}\\
+	%	=\sum_{n\in\Ii}\frac{d}{d\epsilon}G[u(n+1)+\epsilon v(n+1),u(n)+\epsilon v(n)]\bigg|_{\epsilon=0}\\
 		=\sum_{n\in\Ii}2(u(n+1)-u(n))v(n+1)+(-2)(u(n+1)-u(n))v(n)\\
-		=\sum_{n\in\Ii}2(u(n)-u(n-1))v(n)+(-2)(u(n+1)-u(n))v(n)\\
-		=\sum_{n\in\Ii}\Big(2(u(n)-u(n-1))+(-2)(u(n+1)-u(n))\Big)v(n)\\
-		=\sum_{n\in\Ii}\Big(-2u(n+1)+4u(n)-2u(n-1)\Big)v(n)\\
+	%	=\sum_{n\in\Ii}2(u(n)-u(n-1))v(n)+(-2)(u(n+1)-u(n))v(n)\\
+	%	=\sum_{n\in\Ii}\Big(2(u(n)-u(n-1))+(-2)(u(n+1)-u(n))\Big)v(n)\\
+	%	=\sum_{n\in\Ii}\Big(-2u(n+1)+4u(n)-2u(n-1)\Big)v(n)
 		=\sum_{n\in\Ii}\mathcal{V}(\mathcal{J})[u](n)v(n).
@@
 \end{example}
-The algorithms for the simplification of sums presented in \cite{gomes20} relies on the following result, which 
-provides a criterion to compare functionals and evaluate if they represent the same quantity.
+The algorithms for simplification of sums in \cite{gomes20} rely on the following result, which 
+gives a criterion to decide if two functionals represent the same quantity.
 
@@
 	\begin{gather*}
-		(\mathcal{J}-\tilde{\mathcal{J}})[u]=(\mathcal{J}-\tilde{\mathcal{J}})[u]-(\mathcal{J}-\tilde{\mathcal{J}})[u_0]\\
-		=\int_{0}^{1}\frac{d}{d\lambda}(\mathcal{J}-\tilde{\mathcal{J}})[u_0+\lambda(u-u_0)]\ d\lambda\\
+		(\mathcal{J}-\tilde{\mathcal{J}})[u]=(\mathcal{J}-\tilde{\mathcal{J}})[u]-(\mathcal{J}-\tilde{\mathcal{J}})[u_0]=
+		\int_{0}^{1}\frac{d}{d\lambda}(\mathcal{J}-\tilde{\mathcal{J}})[u_0+\lambda(u-u_0)]\ d\lambda\\
 		=\int_{0}^{1}\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})[u_0+\lambda(u-u_0)](u-u_0)\ d\lambda=0
@@
 	Let $d=m=1$ and consider the functionals
-	\begin{gather*}
-		\mathcal{J}[u]:=\sum_{n\in\Ii}u(n)u(n+1)\text{ and }
+$
+		\mathcal{J}[u]:=\sum_{n\in\Ii}u(n)u(n+1)$ and $
 		\tilde{\mathcal{J}}[u]:=\sum_{n\in\Ii}u(n-2)u(n-1).
-	\end{gather*}
-	It is clear that $\mathcal{J}$ and $\tilde{\mathcal{J}}$ represent the same quantity (by shifting $n$). We confirm this, using the discrete variational derivative:
+$
+By shifting $n$, we see that	$\mathcal{J}$ and $\tilde{\mathcal{J}}$ represent the same quantity .
+Accordingly, $\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]$, and the discrete variational derivative must vanish
 	\begin{gather*}
@@
 	\end{gather*}
-	Hence, $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})=0$ and $\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]$. Therefore, both functionals represent the same quantity.
+%	Hence, $\mathcal{V}(\mathcal{J}-\tilde{\mathcal{J}})=0$
+%	and $\mathcal{J}[0]=0=\tilde{\mathcal{J}}[0]$. Thus, the functionals agree.
+	%	both functionals represent the same quantity.
 \end{example}
@@
 
-Let $n=(n_1,\ldots,n_d)$ be the variables for the functions in $\mathcal{P}(\Ii,\Rr^m)$.
-We call $n$ the space variables.
+Let $n=(n_1,\ldots,n_d)$, the space variables, be the variables for the functions in $\mathcal{P}(\Ii,\Rr^m)$.
 Let $l\in\Nn$ and call $t:=(t_1,\ldots,t_l)\in\Nn^l$ the time variables.
-For generality, we consider the vectorial time case, while later in this paper, we focus specifically on the case when $l=1$.
+For generality, we consider the vectorial time case, while later in this paper, we focus on the case when $l=1$.
 The analog to Definition \ref{d1} is the following. 
@@
 	\label{quantityfunc2}
-	The space of functions $\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$ is
+	The space of functions $\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l))$ is the set
 	\begin{gather*}
-		\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l)):=\\	
+	%	\mathcal{F}(\mathcal{P}(\Ii\times\Nn^l,\Rr^m),\mathcal{D}(\Nn^l)):=\\	
 		\Bigg\{\mathcal{J}:\mathcal{P}(\Ii\times\Nn^l)\rightarrow\mathcal{D}(\Nn^l) \Bigg|\ \mathcal{J}[u](t)=\sum_{n\in\Ii}F_{(n,t)}[u],\\
@@
 	=\sum_{\tilde{n}}D_1G(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^1)\\
-	+D_2G(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^2)\\+\ldots
+%	+D_2G(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^2)\\
++\ldots
 	+D_kG(u(\tilde{n}+e^1),\ldots,u(\tilde{n}+e^k))\delta_{(t_1,\ldots,t_{l})}(s_1,\ldots,s_l)v(\tilde{n}+e^k)\\
-	=\sum_{\tilde{n}}D_1G(u(\tilde{n}),\ldots,u(\tilde{n}+e^k-e^1))\delta_{(t_1,\ldots,t_{l})}(s_1-e^1_{d+1},\ldots,s_l-e^1_{d+l})v(\tilde{n})\\
-	+D_2G(u(\tilde{n}+e^1-e^2),\ldots,u(\tilde{n}+e^k-e^2))\delta_{(t_1,\ldots,t_{l})}(s_1-e^2_{d+1},\ldots,s_l-e^2_{d+l})v(\tilde{n})\\+\ldots
-	+D_kG(u(\tilde{n}+e^1-e^k),\ldots,u(\tilde{n}))\delta_{(t_1,\ldots,t_{l})}(s_1-e^k_{d+1},\ldots,s_l-e^k_{d+l})v(\tilde{n})\\
+%	=\sum_{\tilde{n}}D_1G(u(\tilde{n}),\ldots,u(\tilde{n}+e^k-e^1))\delta_{(t_1,\ldots,t_{l})}(s_1-e^1_{d+1},\ldots,s_l-e^1_{d+l})v(\tilde{n})\\
+%	+D_2G(u(\tilde{n}+e^1-e^2),\ldots,u(\tilde{n}+e^k-e^2))\delta_{(t_1,\ldots,t_{l})}(s_1-e^2_{d+1},\ldots,s_l-e^2_{d+l})v(\tilde{n})\\+\ldots
+%	+D_kG(u(\tilde{n}+e^1-e^k),\ldots,u(\tilde{n}))\delta_{(t_1,\ldots,t_{l})}(s_1-e^k_{d+1},\ldots,s_l-e^k_{d+l})v(\tilde{n})\\
 	=\sum_{\tilde{n}}\Big(D_1G(u(\tilde{n}),\ldots,u(\tilde{n}+e^k-e^1))\delta_{(t_1+e^1_{d+1},\ldots,t_{l}+e^1_{d+l})}(s_1,\ldots,s_l)\\
-	+D_2G(u(\tilde{n}+e^1-e^2),\ldots,u(\tilde{n}+e^k-e^2))\delta_{(t_1+e^2_{d+1},\ldots,t_{l}+e^2_{d+l})}(s_1,\ldots,s_l)\\+\ldots
+%	+D_2G(u(\tilde{n}+e^1-e^2),\ldots,u(\tilde{n}+e^k-e^2))\delta_{(t_1+e^2_{d+1},\ldots,t_{l}+e^2_{d+l})}(s_1,\ldots,s_l)\\+\ldots
 	+D_kG(u(\tilde{n}+e^1-e^k),\ldots,u(\tilde{n}))\delta_{(t_1+e^k_{d+1},\ldots,t_{l}+e^k_{d+l})}(s_1,\ldots,s_l)\Big)v(\tilde{n})\\
@@
 	\begin{gather*}
-		D^1(\mathcal{J})[u](v)(t)\\
-		=\frac{d}{d\epsilon} \sum_{n,s}((u(n,s+1)+\epsilon v(n,s+1))-(u(n,s)+\epsilon v(n,s)))\delta_t(s)\\
-		=\sum_{n,s}v(n,s+1)\delta_t(s)-v(n,s)\delta_t(s)=\sum_{n,s}v(n,s)\delta_t(s-1)-v(n,s)\delta_t(s)\\
-		=\sum_{n,s}(\delta_{t+1}(s)-\delta_t(s))v(n,s)=\sum_{n,s}\mathcal{V}^1(\mathcal{J})(n,s,t)v(n,s).
+		D^1(\mathcal{J})[u](v)(t)
+		=\frac{d}{d\epsilon} \sum_{n,s}((u(n,s+1)+\epsilon v(n,s+1))-(u(n,s)+\epsilon v(n,s)))\delta_t(s)=\\
+		\sum_{n,s}v(n,s+1)\delta_t(s)-v(n,s)\delta_t(s)%=\sum_{n,s}v(n,s)\delta_t(s-1)-v(n,s)\delta_t(s)\\
+		=\sum_{n,s}(\delta_{t+1}(s)-\delta_t(s))v(n,s).%=\sum_{n,s}\mathcal{V}^1(\mathcal{J})(n,s,t)v(n,s).
 	\end{gather*}
@@
 \end{example}
-We have a result similar to Theorem \ref{theo1} for the discrete partial variational derivative:
+The analog to Theorem \ref{theo1} for the discrete partial variational derivative is:
 \begin{theorem}\label{theo2}
@@
 \label{algodetec}
-In this section, we formally define numerical schemes and introduce the necessary tools from difference algebra for our algorithm.
+In this section, we formally define numerical schemes and introduce the tools from difference algebra for our algorithm.
 Throughout this paper, we assume that there is only one time variable $t$ ($l=1$).
 \begin{definition}
-	The \emph{space of schemes} is
+	The \emph{space of schemes}, $	\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ is the set
 	\begin{gather*}
-		\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr)):=\\
 		\Bigg\{H:\mathcal{P}(\Ii\times\Nn,\Rr^m)\rightarrow\mathcal{P}(\Ii\times\Nn,\Rr)\ \Bigg|\\
@@
 		\text{where }
-		\{\tilde{e}^1,\ldots,\tilde{e}^r\}\subset\Zz^d,\ \{\tilde{l}^1,\ldots,\tilde{l}^r\}\subset\Zz,\\
-		\text{H is a polynomial in r variables}
+		\{\tilde{e}^1,\ldots,\tilde{e}^r\}\subset\Zz^d,\ \{\tilde{l}^1,\ldots,\tilde{l}^r\}\subset\Zz,
+		\text{H is polynomial}
 		\Bigg\}.
@@
 	A \emph{scheme} is a set of functions
-	\begin{gather*}
+$
 		\{H^1,\ldots,H^i\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))
-	\end{gather*}
+$
 	that represent the equations $\{H^1[u]=0,\ldots,H^i[u]=0\}$ for $u\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$, holding pointwise for all points in $\Ii\times\Nn$.
@@
 If a scheme contains a single function $H$, we call $H$ the scheme. Sometimes, we also call the expression $H[u]$ the scheme.
-\begin{remark}
+%\begin{remark}
 	In our examples, we only consider finite-difference schemes with fixed step sizes $\Delta x=\Delta t=1$.
@@
 Hence, our results can be generalized to schemes with other step sizes.
-\end{remark}
+%\end{remark}
 \begin{example}
@@
 Consider the heat equation	$u_t-u_{xx}=0$ \cite{smoller94}.  A possible scheme is
-\[	
+$
 		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t)).
-\]
+$
 A solution of the corresponding numerical approximation is a function $u\in \mathcal{P}(\Ii\times\Nn,\Rr),$ that satisfies
-\[
+$
 		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t))=0
-\]
+$
 for all $n\in \Ii$ and $t\in \Nn$. 
@@
 	A scheme $\{H^1,\ldots,H^m\}\subset\mathcal{S}(\mathcal{P}(\Ii\times\Nn,\Rr^m),\mathcal{P}(\Ii\times\Nn,\Rr))$ given by
-	\begin{equation*}
+	$
 		H^j(u(n+\tilde{e}_j^1,t+\tilde{l}_j^1),\ldots,u(n+\tilde{e}_j^r,t+\tilde{l}_j^r))
-	\end{equation*}
+	$
 	is in \emph{time-explicit form} if we can rewrite the previous equation (eventually translating the scheme) as
-	\begin{equation*}
+$
 		u_j(n,t+1)-\tilde{H}^j(u(n+\tilde{\tilde{e}}_j^1,t),u(n+\tilde{\tilde{e}}_j^2,t),\ldots,u(n+\tilde{\tilde{e}}_j^r,t))
-	\end{equation*}
+$
 	for $1\leq j\leq m$
@@
 	This scheme is in time-explicit form with right-hand side
-	\begin{equation*}
+	$
 		u(n,t)+(u(n+1,t)-2u(n,t)+u(n-1,t)).
-	\end{equation*}
+$
 \end{example}
@@
 	Let $u=(u_1,\dots,u_m)\in\mathcal{P}(\Ii\times\Nn,\Rr^m)$. The \emph{shift} in the $i$-th coordinate for $1\leq i\leq d+1$ by $k\in\Zz$ is
-	\begin{gather*}
+$
 		\sigma_i^k\circ u_j(n,t):=u_j(n_1,\ldots,n_{i-1},n_i+k,n_{i+1},\ldots,t)
-	\end{gather*}
+$
 	understanding that $\sigma_{d+1}^k$ shifts the time variable.
@@
 	The \emph{set of all possible shifts} $\Theta$ is
-	\begin{gather*}
+$
 		\Theta:=\{\sigma_1^{k_1}\circ\cdots\circ\sigma_{d+1}^{k_{d+1}}\ |\ k_1,\ldots,k_{d+1}\in\Zz\}.
-	\end{gather*}
+$
 \end{definition}
@@
 
- \newpage 
+ %\newpage 
 %just for now 
@@
 % extends the ideal membership problem to parametric ideals.
-In this section, we delve into essential concepts for 
+Here, we review essential concepts for 
 parametric polynomial ideals. These include parametric rings, 
@@
 Gr{\"o}bner basis theory. 
-We discuss the benefits of using comprehensive Gr{\"o}bner systems 
-and how they help address problems that occur when not using them. 
+We discuss the advantages of comprehensive Gr{\"o}bner systems 
+compared with standard Gr\"obner basis. 
 Furthermore, we expand on the ideal membership problem to encompass parametric ideals.
 
-% \subsection{Parametric Polynomials and Comprehensive Groebner Systems}
-% \label{4sec1}
-
-% This section introduces the general setting of parametric polynomial rings and a ring homomorphism that 
-% represents the choice of parameters. Moreover, we define comprehensive Gr{\"o}bner systems (CGS),
-% and show some examples of (CGS) using our implementation of a parametric Buchberger's algorithm.
-
-\begin{definition}
-    Let $\mathcal{K}$ be a field. The ring of parametric polynomials
-    $\mathcal{K}[a_1,..,a_s][x_1,...,x_n]$ is a ring of polynomials in the variables $x_1,...,x_n$ with coefficients that are polynomials
+\begin{definition}
+The ring of parametric polynomials
+    $\mathcal{K}[a_1,..,a_s][x_1,...,x_n]$ in a field $\mathcal{K}$ is a ring of polynomials in the variables $x_1,...,x_n$ with coefficients that are polynomials
     in the ring $\mathcal{K}[a_1,..,a_s]$. 
-    For simplicity, we denote this ring by $\mathcal{K}[A][X]$.
+    To simplify, we denote this ring by $\mathcal{K}[A][X]$.
 \end{definition}
@@
 \end{definition}
-% For example let $P=a x+by \in \Rr[a,b][x,y]$. 
-% a specialization $\sigma$ is defined by $P\rightarrow P(2,5)$ so that $$\sigma (P)=\sigma(a x+by)$$
-% and since $\sigma $ is a ring homomorphism we have,
-% $$\sigma(a x+by)= \sigma(a) \sigma(x)+\sigma(b)\sigma(y)$$
-% $$= 2x+5y.$$
-It is easier to think about specializations as a choice of parameters. \\
-
+It is easier to think about specializations as a choice of parameters. 
 Comprehensive Gr{\"o}bner systems extend Gr{\"o}bner basis to 
-parametric polynomials and have diverse practical applications in 
-algebraic geometry, combinatorics, coding theory, computer-aided design, and robotics.
+parametric polynomials.
+% and have diverse applications in 
+%algebraic geometry, combinatorics, coding theory, computer-aided design, and robotics.
 \begin{definition}
@@
 	\left\{
-		\begin{array}{ll}
-			G_1 & A_1 \\
-			\vdots  & \vdots \\ 
-			G_k & A_k \\
+		\begin{array}{lll}
+			G_1 &\quad& A_1 \\
+			\vdots  &\quad& \vdots \\ 
+			G_k &\quad & A_k \\
 		\end{array}
@@
 	\end{equation}
-
-    is called a comprehensive Gr{\"o}bner system (CGS) of $I$
+    is a {\it comprehensive Gr{\"o}bner system} (CGS) of $I$
      if for all specializations $\sigma_i \in A_i, \sigma_i (G_i)$ 
@@
 and  $"\sigma_i \in A_i"$ stands for a specialization given by $P\rightarrow P(a_{1_0},\dots,a_{m_0})$ 
-for $(a_{1_0},\dots,a_{m_0})\in A_i$.\\
-
-In this paper, we choose $A_i$ to be of the form 
-$$V(p_1,\dots,p_s)\backslash V(q_1,\dots,q_t) \subset L^m.$$
-Where $p_1,\dots,p_s,q_1,\dots,q_t\in \mathcal{K}[A]$ and $V(p_1,\dots,p_s)$ is a Variety.
-
-A proof of existence and the first algorithm to compute the (CGS) were introduced by Volker Weispfenning.
-\begin{theorem}
-    Let $\mathcal{I}\subseteq\mathcal{K}[A][X].$ The corresponding comprehensive Gr{\"o}bner system exists.\cite{weispfenning}
+for $(a_{1_0},\dots,a_{m_0})\in A_i$.
+In this paper, $A_i$ is of the form 
+$V(p_1,\dots,p_s)\backslash V(q_1,\dots,q_t) \subset L^m$.
+Where $p_1,\dots,p_s,q_1,\dots,q_t\in \mathcal{K}[A]$ and $V(p_1,\dots,p_s)$ is a variety.
+
+A proof of existence and an algorithm to compute CGS appeared first in \cite{weispfenning}.
+\begin{theorem}[\cite{weispfenning}]
+    Let $\mathcal{I}\subseteq\mathcal{K}[A][X]$. Then, there exists a CGS . 
 \end{theorem}
-We use our implementation of a parametric Buchberger's algorithm to compute (CGS). 
-This is illustrated in the following examples.
-\begin{example}
-    Let $\Ii\subset \Rr[a,b][x,y]$ be the ideal generated by 
-    $$\Ff=\left\{a x^2+\left(a^2-2 a b\right) x y, b^2 y^2-x\right\}.$$
+Here, to compute CGS, we use a parametric implementation of 
+Buchberger's algorithm which is illustrated in the following examples.
+
+{\color{red} I think we need to explain briefly the parametric implementation. 
+Basically we use Buchberger algorithm but branch every time we divide by a coefficient of a polynomial, considering the cases where this coefficient vanishes and when it does not. }
+
+\begin{example}
+    Let  $\Ff=\left\{a x^2+\left(a^2-2 a b\right) x y, b^2 y^2-x\right\}$, 
+    and     
+    $\Ii\subset \Rr[a,b][x,y]$ be the ideal generated by $\Ff$. 
     The comprehensive Gr{\"o}bner system of $\Ii$ with a Lexicographic order $y\prec x$ is:
     {\color{red} we use our code here}
+
+    {\color{blue} I had issues running the code, will check again once online}
+
 	% $$
@@
 example, if $a=0$, then the leading term $a x^2$ vanishes which leads to a different Gr{\"o}bner basis. Therefore,
-the (CGS) considers cases of the leading coefficients parameters Vanishing.
-
-\begin{example}
-    Let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
-    $$\Ff=\left\{a^2 x-y,-a+b^2 x
-    y-y^2\right\}.$$
+the (CGS) considers cases of the leading coefficients parameters vanishing.
+
+\begin{example}
+	Let $\Ff=\left\{a^2 x-y,-a+b^2 x
+	y-y^2\right\}$, and 
+ $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by $\Ff$.
     The comprehensive Gr{\"o}bner system of $\mathcal{I}$ with a Lexicographic order $y\prec x$ is:
@@
 
-Issues arise when computing the Gr{\"o}bner basis of $\Ff$ without considering the parameters.
-In practice, to compute the Gr{\"o}bner basis without considering the parameters, we assume non-zero parameters by default.
-This assumption guarantees a Gr{\"o}bner basis only for one set of specializations $\Rr^2\backslash V(a,b)$. i.e., 
+Issues arise when computing the Gr{\"o}bner basis of $\Ff$ without considering the parameters; that is, 
+computing Gr{\"o}bner basis assuming the non-zero coefficients.
+This assumption guarantees a Gr{\"o}bner basis only for one set of specializations $\Rr^2\backslash V(a,b)$, i.e., 
 $\sigma(\Gg)$ is a Gr{\"o}bner basis of the specialized ideal $\sigma(\Ii)$ only for $\sigma \in \Rr^2\backslash V(a,b)$.
-Next, we show the impact of this assumption.
-\begin{enumerate}
-    \item In general $\sigma\left(\Gg\right)$ is not necessarily a Gr{\"o}bner basis.
+%Next, we show the impact of this assumption.
+In general $\sigma\left(\Gg\right)$ is not necessarily a Gr{\"o}bner basis.
     \begin{example}
@@
     \end{example}
-    \item Consequently, the algorithm utilized to resolve the ideal membership 
-    problem in chapter 2 of \cite{Cox} can not be used for all specializations.
+
+Consequently, the algorithm utilized to resolve the ideal membership 
+    problem in Chapter 2 of \cite{Cox} can not be used for all specializations.
 
@@
         \label{ex433}
-        Consider the ideal $\Ii$ in the above example. The corresponding Gr{\"o}bner basis computed under the assumption of non-zero parameters
+        Consider the ideal $\Ii$ in the above example. The corresponding Gr{\"o}bner basis computed under the assumption of non-zero coefficients
         is $$\sigma_0(\Gg)=\left\{b^2 y^3-4 a,2 a x-b y^2,b x y-2\right\}$$ where $\sigma_0 \in \Rr^2\backslash V(a,b)$. 
@@
         the specialization $\sigma$ given by $P\rightarrow P(0,1)$, $\sigma(\Gg)$ is not a Gr{\"o}bner basis as shown in the above example. 
-        Therefore, we can not proceed with the division algorithm as we can not guarantee a unique remainder if we use $\sigma(\Gg)$.  
-    \end{example}
-
-\end{enumerate}
+        Therefore, we can not proceed with the division algorithm as we can not guarantee a unique remainder.% if we use $\sigma(\Gg)$.  
+    \end{example}
 
@@
 
-{\color{red} Is this a definition?}
- Consider, $f\in \Kk[A][X]$ we say that $f\in\Ii$ under $\sigma$, if there exists a specialization 
- $\sigma\in L^m$ such that $\sigma(f)\in\sigma(\Ii)$.\\
-
-To determine if $f\in\Ii$ we consider the (CGS) of $\Ii$ given \ref{CGS}. 
+{\color{red} Is this a definition? DG: I think now it is better }
+Let $f\in \Kk[A][X]$. We say that $f\in\Ii$ under 
+ $\sigma\in L^m$ if  $\sigma(f)\in\sigma(\Ii)$.\\
+
+To determine  $f\in\Ii$, we consider the (CGS) of $\Ii$ given \ref{CGS}. 
 We check if $f\in\sigma_i(\Ii)$ for each $i\in\{1,..,k\}$ and $\sigma_i\in A_i$.
-For each $\sigma_i(\Ii)$ we have the corresponding Gr{\"o}bner basis $\sigma_i(G_i)$
-so we divide $f$ by the $\sigma_i(G_i)$ if the remainder is zero for some $i\in\{1,..,k\}$,
-then $f\in\sigma_i(\Ii)$. Therefore, $f\in\Ii$. However, if the remainder is non-zero for all $i$ we conclude that 
+For each $\sigma_i(\Ii)$, we have the corresponding Gr{\"o}bner basis $\sigma_i(G_i)$. Thus, 
+we divide $f$ by the $\sigma_i(G_i)$. If the remainder is zero for some $i\in\{1,..,k\}$,
+then $f\in\sigma_i(\Ii)$. Therefore, $f\in\Ii$. However, if the remainder is non-zero for all $i$, we conclude that 
 $f\notin\Ii$.
@@
 \begin{example}
-	Let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
-    $$\Ff=\{a^2 x - y, -b^2 x y^2 - x\}.$$
+	Let $\Ff=\{a^2 x - y, -b^2 x y^2 - x\}$ and let $\mathcal{I}\subset \Rr[a,b][x,y]$ be the ideal generated by 
+    $\Ff$.
 	We will determine if $f(x,y)=-x+y^2\in\mathcal{I}$.
@@
 
-\newpage
 \section{The time-explicit case}
@@
 	We examine the conservation of $\sum_nu(n,t)$ under the following scheme.
-	\begin{equation*}
+	$
 		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t)).
-	\end{equation*}
+	$
 	The discrete time derivative (Step 1) is
-	\begin{gather*}
+	$
 		\sum_{n\in\Ii}u(n,t+1)-u(n,t).
-	\end{gather*}
+$
 	Replace $u(n,t+1)$ by the right-hand side (Step 2) to get
-	\begin{gather*}
+	$
 		\sum_{n\in\Ii}u(n,t)+(u(n+1,t)-2u(n,t)+u(n-1,t))-u(n,t).
-	\end{gather*}
+	$
 	Then, we compute the discrete partial variational derivative (Step 3), which equals zero. Hence (Step 4), we have conservation.
@@
 Hence, we need to compute a finite number of translations of the scheme and consider every instance of $u_j(n+e,t+l)$ that appears in either the translated scheme or the discrete time derivative as an independent polynomial variable.
-\begin{example}
-	The algorithm treats the discrete time derivative $\sum_{n\in\Ii}u(n+1,t+1)-u(n+1,t)$
-	as the polynomial $z_1-z_2$.
-\end{example}
+%\begin{example}
+%	The algorithm treats the discrete time derivative $\sum_{n\in\Ii}u(n+1,t+1)-u(n+1,t)$
+%	as the polynomial $z_1-z_2$.
+%\end{example}
 There are several possiblities which translations of the scheme to take as basis for the polynomial ideal.
@@
 	Consider the expression
-	\begin{gather*}
+$
 		u(n+3)-u(n-2)
-	\end{gather*}
+$
 	which we want to reduce by the scheme $u(n+1)-u(n)=0$. We compute the translations as described above as
-	\begin{gather*}
+	$
 		\{u(n+3)-u(n+2),\dots,u(n-1)-u(n-2)\}.
-	\end{gather*}
+$
 	and see, that the expression indeed reduces to $0$ under the scheme.
@@
 	The \emph{stencil} of $F$ is the $m$-tuple of sets of vectors
-	\begin{gather*}
+$
 		\Big(\big\{x^1,\ldots,x^{k_1}\big\},
@@
 		\big\{x^{k_{m-1}+1},\ldots,x^{k_m}\big\}\Big).
-	\end{gather*}
+$
 	The \emph{range} of the stencil of $F$ is
@@
 	Consider the discrete time derivative
-	\footnotesize
-	\begin{equation*}
+$
 		u_1(n+2,t+1)+u_1(n,t+1)+u_2(n,t+1)-u_1(n+2,t)-u_1(n,t)-u_2(n,t)
-	\end{equation*}
-	\normalsize
+$
 	and the scheme
-	\footnotesize
 	$
@@
 	$
-	\normalsize
 	Then, the stencil of the discrete time derivative equals
-	\footnotesize
-	\begin{equation*}
+$
 		\Big( \big\{ (2,1),(0,1),(2,0),(0,0) \big\},\big\{(0,1),(0,0)\big\} \Big)
-	\end{equation*}
-	\normalsize
+$
 	with range
-	\footnotesize
 	$
@@
 	$
-	\normalsize
 	The stencil of the scheme equals the set of stencils of the equations in the scheme, i.e.
-	\footnotesize
-	\begin{equation*}
+$
 		\Big\{\Big(\big\{(1,1),(0,0)\big\},\emptyset\Big),\Big(\emptyset,\big\{(0,1),(0,0)\big\}\Big)\Big\}
-	\end{equation*}
-	\normalsize
+$
 	with range
-	\footnotesize
 	$
@@
 	$
-	\normalsize
 	Hence, we get the translations for the first equation of the scheme by
-	\footnotesize
-	\begin{gather*}
+$
 		\Big([0,2]\times[0,1],[0,0]\times [0,1]\Big)-\Big([0,1]\times[0,1],\emptyset\Big)=\Big([0,1]\times[0,0],\emptyset\Big)
-	\end{gather*}
-	\normalsize
+$
 	and for the second equation of the scheme by
-	\footnotesize
-	\begin{gather*}
+$
 		\Big([0,2]\times[0,1],[0,0]\times [0,1]\Big)-\Big(\emptyset,[0,0]\times [0,1]\Big)=\Big(\emptyset,[0,0]\times [0,0]\Big).
-	\end{gather*}
-	\normalsize
+$
 	Therefore, we translate the first equation of the scheme by $(0,0)$ and $(1,0)$ and the second one by $(0,0)$ to get the translated scheme
-	\footnotesize
-	\begin{gather*}
+$
 		\{u_1(n+1,t+1)-u_1(n,t),
@@
 		\}.
-	\end{gather*}
-	\normalsize
+$
 \end{example}
@@
 Hence, using multivariate polynomial division, we search for $p^1,\ldots,p^r,p^{r+1}\in\mathcal{R}$ such that
-\begin{gather*}
+$
 	\mathcal{T}=p^1H^1+\ldots+p^rH^r+p^{r+1}.
-\end{gather*}
+$
 Unfortunately, the resulting remainder $p^{r+1}$ may not be unique \cite[page 14, Example 1.2.3]{hibi13}, i.e. non-zero, although $\mathcal{T}$ belongs to $\langle H^1,\ldots,H^r \rangle$. 
@@
 	We compute the discrete time derivative (Step 1)
-	\begin{gather*}
+$
 		\mathcal{T}=\sum_{n\in\Ii} u(n,t+1)-u(n,t).
-	\end{gather*}
+$
 	For the translations (Step 2), the stencil of the scheme
-	\begin{equation*}
+$
 		u(n,t+1)-u(n,t)-(u(n+1,t)-2u(n,t)+u(n-1,t))
-	\end{equation*}
-	\small
+$
 	is
@@
 	$
-	\normalsize
 	with range \small$\Big([-1,1]\times[0,1]\Big)$\normalsize.
@@
 	The translations are
-	\small
-	\begin{gather*}
+$
 		\Big([0,0]\times[0,1]\Big)-\Big([-1,1]\times[0,1]\Big)=\Big([1,-1]\times[0,0]\Big)=\Big(\emptyset\times[0,0]\Big)
-	\end{gather*}
-	\normalsize
+$
 	which results in no translations, as the range of the stencil of the scheme is greater than the range of the stencil of the discrete time derivative.
@@
 	The reduction using the lexicographic monomial order yields as remainder
-	\begin{gather*}
+$
 		u(n-1,t)-2u(n,t)+u(n+1,t).
-	\end{gather*}
+$
 	For the reduction using the elimination order, we first eliminate $u(n,t+1)$ and then the remaining instances of $u$.
@@
 	We calculate the discrete partial variational derivative (Step 5) of
-	\begin{gather*}
+$
 		\sum_{n\in\Ii}u(n-1,t)-2u(n,t)+u(n+1,t)
-	\end{gather*}
+$
 	that equals $0$.
@@
 \begin{example}
-	We verify our result from Example \ref{heateqalgoimplicit}, using {\sc DiscreteConservedQ}, which detects automatically if the scheme is time-explicit or if
-	 the general setting should be applied.
+	We verify our result from Example \ref{heateqalgoimplicit}, using {\sc DiscreteConservedQ}, which detects automatically if the scheme is time-explicit or not.
 	\begin{EXE}
@@
 \begin{example}
-In the context of the heat equation from Example \ref{heatex1}, we aim to find a basis for conserved quantities generated by $u(n,t)$ and $n$ with a maximum degree of 3.
+In the context of the heat equation from Example \ref{heatex1}, we aim to find a basis for conserved quantities generated by $u$ and $n$ with degree up to 3.
 	\begin{EXE}
@@
 
-The inviscid Burgers equation is the PDE $u_t+uu_x=0$ \cite{smoller94} in one space dimension. Any function of $u$ is a conserved quantity. We search for discretizations preserving $\int u\ dx$.
+The inviscid Burgers equation is $u_t+uu_x=0$ \cite{smoller94}. Any function of $u$ is a conserved quantity. We search for discretizations preserving $\int u\ dx$.
 \begin{example}
@@
 	\end{small}
-The conservation property holds when $a=\frac{1}{2}$, which corresponds to the central difference case.
+The conservation property holds when $a=\frac{1}{2}$, the central difference case.
 \end{example}
@@
 	In \cite{gomes18}, Gomes et al. derived the following system
-	\begin{gather*}
-		v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0,\quad
+$
+		v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0,$ and $
 		m_t-(vm)_x=0.
-	\end{gather*}
+$
 	This system admits the conserved quantities $\int v\ dx$ and $\int m\ dx$.
@@
 \begin{example}
-	The related backward-forward system reads
-	\begin{gather*}
-		-v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0,\quad
+	The related backward-forward system is
+$
+		-v_t+\Big(\frac{v^2}{2}-\frac{m^2}{2}\Big)_x=0$ and $
 		m_t-(vm)_x=0.
-	\end{gather*}
+$
 	This system admits the same conserved quantities as the forward-forward system, but we approximate it explicitly in time for $m$ and implicitly for $v$.
@@
 \end{example}
+\begin{comment}
 \section{Possible extensions and concluding remarks}
@@
 One possible extension of the code would be to check if s-consistency holds for the translated scheme.\\
-\subsection{Concluding remarks}
+\end{comment}
+\section{Concluding remarks}
 We present an algorithm for checking the conservation of a quantity under a finite-difference scheme.

### 2023-05-10 PDESymbolics2022/Discrete.m
small improvements in the code

@@ -157,7 +157,7 @@ DiscreteConservedQ[variables_Association][schemeexpression_] :=
     schexp = 
      EchoLabel["DiscreteConservedQ: schexp"]@VariationalTimeDifferenceOperator[KeyDrop[var, "listop"]][
      	EchoLabel["DiscreteConservedQ: schexp: input to VariationalTimeDifferenceOperator"]@schexp];
-    schexp = PiecewiseMap[Lookup["exp"] (*/@ # &*), schexp];
+    schexp = EchoLabel["DiscreteConservedQ: schexp[exp]"]@PiecewiseMap[Lookup["exp"] (*/@ # &*), schexp];
     generators = PiecewiseExtractGeneratorsOperator[var][schexp];
     var = Append[var, "generators" -> generators];
     If[Lookup[var, "display result", False], Print[schexp]];

### 2023-05-10 PDESymbolics2022/ParametricGroebnerBasis.m

@@ -43,6 +43,10 @@ SPolynomialOperator::usage = "";
 Division::usage = "";
 
 NotPiecewise::usage = "";
+
+MonicOperator::usage = "";
+
+AllCasesOperator::usage = "";
 Begin["`Private`"]
 InferGeneratorsOperator[variables_][xplist_List] :=
     With[ {glc = First@EchoLabel["InferGenerators: gLc"]@GenericLinearCombinationOperator[variables][EchoLabel["InferGenerators: input"]@xplist]},
@@ -128,7 +132,7 @@ LeadingCoefficient[variables_Association][xp_?NotPiecewise] :=
     ];
 
 LeadingCoefficientOperator[vars_][xp_] :=
-    KleisliListable[LeadingCoefficient][vars][xp]
+    KleisliListable[LeadingCoefficient][vars][xp]//QuietEcho
 
 Division[variables_][a_?NotPiecewise,b_?NotPiecewise] :=
     If[ a === $Failed|| b === $Failed,
@@ -267,18 +271,21 @@ PiecewiseApplyConditionOperator[variables_][px_Piecewise] :=
         facts = Lookup[variables, "facts", True];
         If[ Reduce[facts] === False,
             $Failed,
-            newPx = 
+            newPx = EchoLabel["PACO: newPx"][
              SplitOr @@@ 
               PiecewiseLastCaseClean[
-               EchoLabel["PACO: first beautify"]@PiecewiseBeautifyOperator[variables][px]];
-            generators = EchoLabel["PACO: generators"]@Lookup[variables, "generators", InferGeneratorsOperator[variables][Piecewise@newPx]];
+               (*EchoLabel["PACO: first beautify"]@*)PiecewiseBeautifyOperator[variables][px]]];
+            generators = (*EchoLabel["PACO: generators"]@*)Lookup[variables, "generators", InferGeneratorsOperator[variables][Piecewise@newPx]];
             If[ AnyTrue[First@newPx,Head[#]===List&],
+            	(*Print["1"];*)
                 cleanList = Function[dd,
-                Assuming[dd[[2]], {Simplify[dd[[1]]], dd[[2]]}]] /@ newPx,
+                Assuming[dd[[2]], {(*EchoLabel["PACO: simplifyed this"]@*)
[…]

### 2023-05-10 PDESymbolics2022/PiecewiseTools.m

@@ -92,7 +92,7 @@ PiecewiseLastCaseClean[xp_Piecewise] :=
     ];
 
 PiecewiseMap[F_, XPO_] :=
-    With[ {XP = PiecewiseExpand[XPO]},
+    With[ {XP = PiecewiseExpand[EchoLabel["PiecewiseMap: input"]@XPO]},
         Which[
             XP === $Failed, 
             $Failed,
@@ -111,7 +111,7 @@ PiecewiseMap[F_, XPO_] :=
             True,
             F[XP]
         ]
-    ];
+    ]//QuietEcho;
 
 PiecewiseOperatorMap[Operator_, variables_, XPO_] :=
     With[ {XP = PiecewiseExpand[XPO],reduce = Lookup[variables,"reduce",Reduce]},
@@ -158,7 +158,7 @@ ConditionalPiecewiseMap[F_,XP_] :=
     ]
 
 Clear[PiecewiseBeautify];
-Options[PiecewiseBeautify] = {"domain"->Complex, "reduce"->Resolve};
+Options[PiecewiseBeautify] = {"domain"->Complex, "reduce"->Reduce};
 
 
 (*keeping to check one alteration
@@ -180,7 +180,7 @@ PiecewiseBeautify[P_,OptionsPattern[]] :=
         Head[P] === Piecewise,
         With[ {pp = PiecewiseLastCaseClean[P],reduce = OptionValue["reduce"]},
             Module[ {qq},
-                qq = Simplify@EchoLabel["PiecewiseBeautify:  qq"]@MapThread[BooleanConvert@reduce[!#1&&#2,OptionValue["domain"]]&, {FoldList[Or,False, pp[[All,2]]][[;;-2]],pp[[All,2]]}];
+                qq = Simplify@(*EchoLabel["PiecewiseBeautify:  qq"]@*)MapThread[BooleanConvert@reduce[!#1&&#2,OptionValue["domain"]]&, {FoldList[Or,False, pp[[All,2]]][[;;-2]],pp[[All,2]]}];
                 Piecewise[Select[Transpose[{pp[[All,1]], qq}], #[[1]]=!=$Failed &], $Failed]
             ]
         ],

### 2023-05-10 Papers/Paper CASC format/paper.tex
paper changes...
no need for under sigma in the definition of membership to the parametric ideal

@@
 {\color{red} Is this a definition?}
- Consider, $f\in \Kk[A][X]$ we say that $f\in\Ii$ under $\sigma$, if there exists a specialization 
+ Consider, $f\in \Kk[A][X]$ we say that $f\in\Ii$ if there exists a specialization 
  $\sigma\in L^m$ such that $\sigma(f)\in\sigma(\Ii)$.\\

### 2023-05-11 PDESymbolics2022/ParametricGroebnerBasis.m
Update ParametricGroebnerBasis.m
Beautifying after done with GrobOp

@@ -292,7 +292,7 @@ PiecewiseApplyConditionOperator[variables_][px_Piecewise] :=
 Clear[GrobOp];
 
 ComprehensiveGroebnerSystemOperator[variables_][ideal_]:=
-Kleisli[GrobOp][variables][ideal];
+Kleisli[GrobOp][variables][ideal]//PiecewiseBeautifyOperator[variables];
 
 hasFailed[x_] :=
     MemberQ[$Failed][x];

### 2023-05-18 Papers/Paper CASC format/paper.tex
finished paper for CASC

--- 2023-05-09 Papers/Paper CASC format/paper.tex
@@
 and  $"\sigma_i \in A_i"$ stands for a specialization given by $P\rightarrow P(a_{1_0},\dots,a_{m_0})$ 
-for $(a_{1_0},\dots,a_{m_0})\in A_i$.\\
-
-In this paper, we choose $A_i$ to be of the form 
-$$V(p_1,\dots,p_s)\backslash V(q_1,\dots,q_t) \subset L^m.$$
-Where $p_1,\dots,p_s,q_1,\dots,q_t\in \mathcal{K}[A]$ and $V(p_1,\dots,p_s)$ is a Variety.
-
-A proof of existence and the first algorithm to compute the (CGS) were introduced by Volker Weispfenning.
-\begin{theorem}
-    Let $\mathcal{I}\subseteq\mathcal{K}[A][X].$ The corresponding comprehensive Gr{\"o}bner system exists.\cite{weispfenning}
+for $(a_{1_0},\dots,a_{m_0})\in A_i$.
+In this paper, $A_i$ is of the form 
+$V(p_1,\dots,p_s)\backslash V(q_1,\dots,q_t) \subset L^m$.
+Where $p_1,\dots,p_s,q_1,\dots,q_t\in \mathcal{K}[A]$ and $V(p_1,\dots,p_s)$ is a variety.
+
+A proof of existence and an algorithm to compute CGS appeared first in \cite{weispfenning}.
+\begin{theorem}[\cite{weispfenning}]
+    Let $\mathcal{I}\subseteq\mathcal{K}[A][X]$. Then, there exists a CGS . 
 \end{theorem}
-We use our implementation of a parametric Buchberger's algorithm to compute (CGS). 
-This is illustrated in the following examples.
-\begin{example}
-    Let $\Ii\subset \Rr[a,b][x,y]$ be the ideal generated by 
-    $$\Ff=\left\{a x^2+\left(a^2-2 a b\right) x y, b^2 y^2-x\right\}.$$
+Here, to compute CGS, we use a parametric implementation of 
+Buchberger's algorithm which is illustrated in the following examples.
+
+{\color{red} I think we need to explain briefly the parametric implementation. 
+Basically we use Buchberger algorithm but branch every time we divide by a coefficient of a polynomial, considering the cases where this coefficient vanishes and when it does not. }
+
+\begin{example}
+    Let  $\Ff=\left\{a x^2+\left(a^2-2 a b\right) x y, b^2 y^2-x\right\}$, 
+    and     
+    $\Ii\subset \Rr[a,b][x,y]$ be the ideal generated by $\Ff$. 
     The comprehensive Gr{\"o}bner system of $\Ii$ with a Lexicographic order $y\prec x$ is:
-    {\color{red} we use our code here}
 	% $$
